---
title: "Create VMS Representation Index"
author: "Owen Liu"
date: "10/21/2021"
output: html_document
---

## Purpose

Create a relative VMS representation index by year, month, state, vessel pot limit tier, and vessel length. This key will be used as a weighting factor in the vertical line model. NOTE: this is an update to a script that was used in exploration of the matched fish ticket-VMS data for the O. Liu's Dungenesss fishing behavioral types project.

```{r pkgs, include=FALSE,message=F,warning=FALSE}
library(tidyverse)
library(knitr)
library(extrafont)
library(ggsci)
library(lubridate)
library(sf)
library(geosphere)
library(magrittr)
library(here)

# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=10,color="black"),
        legend.text = element_text(size=10),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
```

We need fish ticket, VMS, and tier info

```{r import data}
# Load Rdata files

fishtix <- read_rds(here::here('data','processed','fish tickets','fish_tickets_w_dayofseason.rds'))
fishtix %<>% filter(ticket_day_of_season>=0)

# vms_rep <- read_rds(here('fishing behavior','vms_proportional_landings_key.rds'))
vms_tickets <- read_rds(here('fishing behavior','fish_tickets_with_vms.rds'))
recid_crabseason <- fishtix %>% dplyr::select(Rec_ID,crab_season) %>% distinct()

# combined tier information
tier_key <- read_rds(here::here('vertical line model','tier_information_year_month_allstates.rds'))
tier_key %<>% mutate(month=month.name[month]) %>% dplyr::select(-crabyr)
```

# Total Landings, Trips, and Number of Vessels over Time

In this section, we summarize the data on broad scales to look at total landings over time and representation of the fleet through VMS records.

For the tier level, WE ASSUME THAT IF NUMBER OF PERMITTED TRAPS IS NA, WE DEFAULT TO A LIMIT OF 500

```{r}
fishtix <- fishtix %>% 
  left_join(vms_tickets) %>% 
  mutate(has_vms=replace_na(has_vms,0)) %>% 
  left_join(tier_key) %>% 
  mutate(ntraps_permitted=coalesce(ntraps_permitted,500))
glimpse(fishtix)
```

## Total Dungeness Vessels

*  How many vessels recorded landings in each month in our data?
*  What proportion of vessels have associated VMS data in each month?

```{r count vessels}
nvessels_by_month <- fishtix %>% 
  group_by(year,month,has_vms) %>% 
  summarise(nvessels=n_distinct(drvid)) %>% 
  ungroup() %>% 
  mutate(day=1) %>% 
  unite(date,year,month,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%B-%d",tz="America/Los_Angeles"))

nvessels_by_month %>% 
  ggplot(aes(date,nvessels,fill=factor(has_vms),col=factor(has_vms)))+
  geom_bar(stat='identity')+
  scale_fill_locuszoom(name="",labels=c("No VMS","VMS"))+
  scale_color_locuszoom(name="",labels=c("No VMS","VMS"))+
  scale_x_date(date_breaks= "6 months",date_labels="%b-%Y",expand=c(0,0))+
  labs(x="Date",y="Vessels Recording Landings",title="Number of Vessels by Month")+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())
```


## Total Landings

*  What are the total Dungeness landings represented in the fish ticket data over time?
*  What proportion of those landings are represented in our VMS records?

```{r}
landings_by_month <- fishtix %>%
  group_by(year,month,has_vms) %>% 
  summarise(DCRB_lbs=sum(DCRB_lbs)) %>%
  mutate(DCRB_mt=DCRB_lbs/2204.62) %>% 
  ungroup() %>% 
  mutate(day=1) %>% 
  unite(date,year,month,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%B-%d",tz="America/Los_Angeles"))

landings_by_month %>% 
  ggplot(aes(date,DCRB_lbs/1e6,fill=factor(has_vms),col=factor(has_vms)))+
  geom_bar(stat='identity')+
  scale_fill_locuszoom(labels=c("No VMS","VMS"),name="")+
  scale_color_locuszoom(labels=c("No VMS","VMS"),name="")+
  scale_x_date(date_breaks= "6 months",date_labels="%b-%Y",expand=c(0,0))+
  labs(x="Date",y="DCRB Landings (Million Lbs)",title="DCRB Landings by Month, all Tickets")+
  theme(legend.position = c(0.2,0.9))+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())

landings_by_year <- fishtix %>%
  group_by(year,has_vms) %>% 
  summarise(DCRB_lbs=sum(DCRB_lbs)) %>%
  mutate(DCRB_mt=DCRB_lbs/2204.62) %>% 
  ungroup()

landings_by_year %>% 
  ggplot(aes(year,DCRB_lbs/1e6,fill=factor(has_vms),col=factor(has_vms)))+
  geom_bar(stat='identity')+
  scale_fill_locuszoom(labels=c("No VMS","VMS"),name="")+
  scale_color_locuszoom(labels=c("No VMS","VMS"),name="")+
  labs(x="Year",y="DCRB Landings (Million Lbs)",title="DCRB Landings by Year, all Tickets")+
  theme(legend.position = c(0.2,0.9))
```
As a percentage

```{r}
prop <- landings_by_year %>% 
  dplyr::select(-DCRB_lbs) %>% 
  group_by(year) %>% 
  pivot_wider(names_from=has_vms,values_from=DCRB_mt,names_prefix="vms") %>% 
  mutate(prop=vms1/sum(vms0,vms1))
mean(prop$prop)
```

## By State

```{r}
landings_by_month_state <- fishtix %>%
  group_by(year,month,agency_code,has_vms) %>% 
  summarise(DCRB_lbs=sum(DCRB_lbs)) %>%
  mutate(DCRB_mt=DCRB_lbs/2204.62) %>% 
  ungroup() %>% 
  mutate(day=1) %>% 
  unite(date,year,month,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%B-%d",tz="America/Los_Angeles"))

landings_by_month_state %>% 
  ggplot(aes(date,DCRB_lbs/1e6,fill=factor(has_vms),col=factor(has_vms)))+
  geom_bar(stat='identity')+
  facet_wrap(~agency_code,nrow=3)+
  scale_fill_locuszoom(labels=c("No VMS","VMS"),name="")+
  scale_color_locuszoom(labels=c("No VMS","VMS"),name="")+
  scale_x_date(date_breaks= "6 months",date_labels="%b-%Y",expand=c(0,0))+
  labs(x="Date",y="DCRB Landings (Million Lbs)",title="DCRB Landings by Month, all Tickets")+
  theme(legend.position = c(0.2,0.95))+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())

revenue_by_month_state <- fishtix %>%
  group_by(year,month,agency_code,has_vms) %>% 
  summarise(DCRB_revenue=sum(DCRB_revenue)) %>%
  mutate(DCRB_revenue=DCRB_revenue/1000) %>% 
  ungroup() %>% 
  mutate(day=1) %>% 
  unite(date,year,month,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%B-%d",tz="America/Los_Angeles"))

revenue_by_month_state %>% 
  ggplot(aes(date,DCRB_revenue,fill=factor(has_vms),col=factor(has_vms)))+
  geom_bar(stat='identity')+
  facet_wrap(~agency_code,nrow=3)+
  scale_fill_locuszoom(labels=c("No VMS","VMS"),name="")+
  scale_color_locuszoom(labels=c("No VMS","VMS"),name="")+
  scale_x_date(date_breaks= "6 months",date_labels="%b-%Y",expand=c(0,0))+
  labs(x="Date",y="Revenue ($1000s)",title="DCRB Landings by Month, all Tickets")+
  theme(legend.position = c(0.2,0.95))+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())

relative_revenue <- revenue_by_month_state %>% 
  mutate(has_vms=ifelse(has_vms==1,"VMS","nonVMS")) %>% 
  pivot_wider(names_from=has_vms,values_from = DCRB_revenue) %>% 
  drop_na() %>%
  mutate(vms_relative=VMS/(VMS+nonVMS)) %>% 
  filter(!(month(date) %in% c(7,8,9,10)))
  
```

## By State and Vessel Length

```{r}
# calc landings of VMS vs. non-VMS tix
vms_prop_landings <- fishtix %>% 
  
  # add vessel size designation
  mutate(vessel_size_category=ifelse(FINAL_LENGTH<40,'small','large')) %>% 
  # group and summarise
  group_by(year,month,agency_code,ntraps_permitted,vessel_size_category,has_vms) %>% 
  summarise(totland=sum(DCRB_lbs,na.rm=T),tottix=n()) %>% 
  ungroup() %>% 
  group_by(year,month,agency_code,ntraps_permitted,vessel_size_category) %>% 
  mutate(sumland=sum(totland),
         sumland_thousand_mt=sumland/1000/2204.64,
         propland=totland/sumland,
         sumtix=sum(tottix),
         proptix=tottix/sumtix) %>% 
  ungroup() %>% 
  # only save the proportion of tickets and landings with VMS
  filter(has_vms==1)
```

```{r,fig.width=8,fig.height=8}
vms_representation_index <- vms_prop_landings %>% 
  filter(!is.na(vessel_size_category)) %>% 
  dplyr::select(year:vessel_size_category,propland)

# California
vms_representation_index %>%
  filter(agency_code=="C") %>% 
  mutate(day=1) %>% 
  unite(date,year,month,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%B-%d",tz="America/Los_Angeles")) %>% 
  ggplot(aes(date,1/propland,col=factor(vessel_size_category)))+
  # geom_bar(stat='identity',position = 'dodge')+
  geom_line()+
  facet_wrap(~ntraps_permitted,nrow=7,scales='free_y')+
  scale_fill_locuszoom(name="",na.value='black')+
  scale_color_locuszoom(name="",na.value='black')+
  scale_x_date(date_breaks= "6 months",date_labels="%b-%Y",expand=c(0,0))+
  labs(x="Date",y="Landings Weight",title="Inverse of Proportional Landings Represented in VMS\nCalifornia")+
  theme(legend.position = c(0.1,0.95))+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())

# Oregon
vms_representation_index %>%
  filter(agency_code=="O") %>% 
  mutate(day=1) %>% 
  unite(date,year,month,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%B-%d",tz="America/Los_Angeles")) %>% 
  ggplot(aes(date,1/propland,col=factor(vessel_size_category)))+
  # geom_bar(stat='identity',position = 'dodge')+
  geom_line()+
  facet_wrap(~ntraps_permitted,nrow=7,scales='free_y')+
  scale_fill_locuszoom(name="",na.value='black')+
  scale_color_locuszoom(name="",na.value='black')+
  scale_x_date(date_breaks= "6 months",date_labels="%b-%Y",expand=c(0,0))+
  labs(x="Date",y="Landings Weight",title="Inverse of Proportional Landings Represented in VMS\nOregon")+
  theme(legend.position = c(0.1,0.95))+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())

# Washington
vms_representation_index %>%
  filter(agency_code=="W") %>% 
  mutate(day=1) %>% 
  unite(date,year,month,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%B-%d",tz="America/Los_Angeles")) %>% 
  ggplot(aes(date,1/propland,col=factor(vessel_size_category)))+
  # geom_bar(stat='identity',position = 'dodge')+
  geom_line()+
  facet_wrap(~ntraps_permitted,nrow=7,scales='free_y')+
  scale_fill_locuszoom(name="",na.value='black')+
  scale_color_locuszoom(name="",na.value='black')+
  scale_x_date(date_breaks= "6 months",date_labels="%b-%Y",expand=c(0,0))+
  labs(x="Date",y="Landings Weight",title="Inverse of Proportional Landings Represented in VMS\nWashington")+
  theme(legend.position = c(0.1,0.95))+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())
```

## Save

```{r}
write_rds(vms_representation_index,here::here('vertical line model','vms_representation_index.rds'))
```

