---
title: "Oregon Logbook Comparison"
author: "Owen Liu"
date: "3/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pkgs, include=FALSE,message=F,warning=FALSE}
library(raster)
library(tidyverse)
library(lubridate)
library(sf)
library(magrittr)
library(rnaturalearth)
library(fasterize)
library(viridis)
library(here)

# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=12,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
```

## Purpose

The purpose of this analysis is to compare Oregon logbook data on Dungeness fishing to the vertical line model estimated from VMS data and PacFIN fish tickets.

We will compare them in two ways-- trip by trip and in regards to overall density. In other words, we will investigate how the two data sources align, on both a small and large scale.

## Import Data

```{r}
# Oregon logbook data
permits <- read_csv(here::here('data','oregon','oregon_permits.csv'))
logs <- read_csv(here::here('data','oregon','oregon_logs.csv'))

# fish tickets
fishtix_matched_all <- read_rds("C:/Users/Owen.Liu/Documents/github/VMS-NWFSC-2/data/processed/fish tickets/fish_tickets_w_dayofseason.rds")
# filter for out of season?
fishtix_matched_all %<>% filter(ticket_day_of_season>=0)

# 3 keys for matching Oregon VLM
drvid_tier_match <- read_rds(here('data','oregon_tier_drvid_matching_key.rds'))
traps_month_key <- read_rds(here('data','oregon_traps_month_key.rds'))
RecID_tier_weights <- read_rds(here('data','oregon_tier_vms_weights.rds'))

# a coastline, for plotting
# coastline for plotting
coaststates <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('Oregon'))
coastline <- ne_coastline(scale='medium',returnclass = 'sf') %>% 
  st_crop(st_bbox(coaststates))

# Load if not already loaded
simulated_traps <- read_rds(here::here('data','processed',"sim_traps_speed_depth_filters.rds")) %>% ungroup()

```

## Organize Oregon Data

Clean up the data such that it is Oregon-specific

### VLM

```{r}
# traps locations only (discard VMS data for now), attached to fish ticket data
traplocs <- simulated_traps %>% dplyr::select(Rec_ID,traps) %>% 
  left_join(fishtix_matched_all,by="Rec_ID") %>% 
  dplyr::select(Rec_ID,date,year,month,ticket_day_of_season,pacfin_port_code,port_group_code,agency_code,drvid,FINAL_LENGTH,traps) %>% 
  distinct(Rec_ID,.keep_all = T) %>% 
  # count number of traps associated with each trip
  mutate(ntraps=purrr::map_int(traps,nrow))

# join the permits key by year/month/vesselID
OR_traplocs <- traplocs %>% filter(agency_code=="O") %>% left_join(drvid_tier_match,by=c('year','month','drvid'))

# subsample traps down to trap limit (hacky...)
OR_traplocs <- OR_traplocs %>% 
  mutate(traps=purrr::map2(traps,ntraps_permitted,function(x,y){
    if(y<nrow(x)&!is.na(y)) {sample_n(x,y)} else{x}
  }))
```

### Logbooks

Organize the permit and logbook data such that we can use it to compare to vertical line model.

Here's the metadata about the Oregon logbook records

```{r}
logs_meta <- read_csv(here::here('data','oregon','oregon_logs_metadata.csv'))

knitr::kable(logs_meta)
```

Pull out the important variables, rename them to match our other data. Also, remove any records for whom the spatial information is unreasonable (variable `SpatialFlag==T`)

```{r}
glimpse(logs)

# discard and rename some variables
logs %<>% 
  filter(!SpatialFlag) %>% 
  select(FishTicket,VessID,Date,Pots,AdjLbs,BegLat,Beglon,EndLat,EndLon) %>% 
  rename(FTID=FishTicket,drvid=VessID,date=Date,n_pots=Pots,adj_lbs=AdjLbs,startlat=BegLat,startlon=Beglon,endlat=EndLat,endlon=EndLon)

# Reduce logs to only those years for which we have VMS data
logs %<>% 
  mutate(date=dmy(date)) %>%
  filter(year(date)>2008)
# reduce logs data to only those vessels for which we have VMS data
# Unique vessels
# oregon_vessels <- unique(logs$drvid)
# vms_vessels <- unique(traplocs$drvid)
# vessel_list <- intersect(oregon_vessels,vms_vessels)

# logs %<>%
#   filter(drvid %in% vessel_list) %>% 
#   mutate(date=dmy(date)) %>% 
#   filter(year(date)>2008)

```

Some of the records are associated with multiple fish tickets (around 7 percent of records). For now, we discard them, but we should come back to this later and figure out how to handle those records.

```{r}
# mult_tix <- logs %>% slice(str_which(logs$Rec_ID,';')) %>% 
#   mutate(num_tix=str_count(Rec_ID,";")+1)
# # make a new row record for each ticket
# expanded_tix <- mult_tix %>% 
#   uncount(num_tix) %>% 
#   group_by(Rec_ID) %>% 
#   mutate(which_obs=row_number())
  
# a list of all the fish tickets and vessels associated with Oregon logbook data
# logs %<>%
#   mutate(num_tix=str_count(Rec_ID,";")+1) %>% 
#   # filter out records with more than 1 fish ticket
#   filter(num_tix==1) %>% 
#   mutate(Rec_ID=as.numeric(Rec_ID))
# 
```

## Create Oregon Logbooks Trap Density

Make density rasters from Oregon logbook data.

*  Use spatial information (start and end points of strings) to assign traps to the 5km grid
*  If the start and end points of the string are in different 5km grid cells, split them evenly between those cells

```{r}
string_starts <- logs %>% 
  #lat and lons cannot be NA
  filter(!is.na(startlat),!is.na(startlon),!is.na(endlat),!is.na(endlon)) %>% 
  select(startlat,startlon) %>% 
  mutate(startlon=startlon*-1) %>% 
  st_as_sf(coords=c('startlon','startlat'),crs=4326) %>% 
  st_transform(st_crs(grd))
# example map
string_starts %>% slice(1:1000) %>% ggplot()+geom_sf()+geom_sf(data=coaststates)

string_ends <- logs %>%
  #lat and lons cannot be NA
  filter(!is.na(startlat),!is.na(startlon),!is.na(endlat),!is.na(endlon)) %>% 
  select(endlat,endlon) %>% 
  mutate(endlon=endlon*-1) %>% 
  st_as_sf(coords=c('endlon','endlat'),crs=4326)%>% 
  st_transform(st_crs(grd))
# example map
string_ends %>% slice(1:1000) %>% ggplot()+geom_sf()+geom_sf(data=coaststates)

# Match to the grid
string_starts %<>% st_join(grd) %>% select(GRID5KM_ID) %>% rename(startgrd=GRID5KM_ID)
string_ends %<>% st_join(grd) %>% select(GRID5KM_ID) %>% rename(endgrd=GRID5KM_ID)

# collect grid matches and join to logbook data
logs_grd <- logs %>% 
  #lat and lons cannot be NA
  filter(!is.na(startlat),!is.na(startlon),!is.na(endlat),!is.na(endlon)) %>%
  mutate(startgrd=string_starts$startgrd,endgrd=string_ends$endgrd) %>% 
  filter(!is.na(startgrd),!is.na(endgrd)) %>% 
  ungroup() %>% 
  # add a new identifier for unique fished strings
  mutate(newID=row_number())
  
logs_trap_density <- logs_grd %>% 
  select(newID,FTID,date,n_pots,startgrd,endgrd) %>% 
  pivot_longer(startgrd:endgrd,names_to='type',values_to='GRID5KM_ID') %>% 
  group_by(newID) %>% 
  # see whether start and end grids are the same
  mutate(n_grd=n_distinct(GRID5KM_ID)) %>% 
  # if grids are not the same, split pots evenly amongst start and end grids
  mutate(n_pots_adj=ifelse(n_grd==1,n_pots,n_pots/2)) %>% 
  ungroup() %>% 
  distinct(newID,FTID,date,GRID5KM_ID,n_grd,n_pots_adj) %>% 
  filter(!is.na(n_pots_adj)) %>% 
  # finally, organize by grid cell and date, and sum the number of traps
  group_by(GRID5KM_ID,date) %>% 
  summarise(n_pots=sum(n_pots_adj)) %>% 
  ungroup() %>% 
  # convert to a density
  mutate(pot_density=n_pots/25)

# by month
logs_trap_density_month <- logs_trap_density %>% 
  mutate(year=year(date),month=month(date)) %>% 
  # MAX by grid cell/year/month
  group_by(GRID5KM_ID,year,month) %>% 
  summarise(max_pot_density=max(pot_density,na.rm=T))
  
```

Distribution of Trap Densities, all years, according to Oregon logbook data.
```{r}
logs_trap_density %>% 
  mutate(yr=year(date)) %>% 
  ggplot(aes(pot_density))+
  geom_density(alpha=0.5,fill=viridis_pal(begin=0.2,end=0.8)(1))+
  labs(x="Traps per sq. km",y="Kernel Density")

# and by month across all years

logs_trap_density %>% 
  mutate(mth=month(date),mth_nm=month.name[mth],mth_nm=factor(mth_nm,levels=month.name)) %>%
  ggplot(aes(pot_density))+
  geom_density(alpha=0.5,fill=viridis_pal(begin=0.2,end=0.8)(1))+
  facet_wrap(~mth_nm)+
  labs(x="Traps per sq. km",y="Kernel Density")
```

# Join Grid to VLM

We make a grid representation of the simulated traps data for Oregon, assigning a 5km grid cell to each observation. Then, we produce both an unweighted (no correction for tier weights) and weighted (tier-representation-adjusted) version of monthly rasters.

First, attach the appropriate grid cell to each simulated trap

```{r grid fxn,warning=FALSE,fig.height=8,fig.width=4}
# 5x5 grid shapefile
# grd <- read_sf(here::here("Input_Data","5x5 Grid","five_km_grid_polys_geo.shp"))
grd <- read_sf(here::here("data","grid","regions_master_final_lamb.shp"))
names(grd)

# raster representation of grid.
# transform to a projection that uses meters
grd_rast <- fasterize(grd,raster = raster(grd,res=5000,crs=crs(grd)),field="GRID5KM_ID")
# set -99999 to NA
grd_rast[grd_rast==-99999] <- NA

grd_xy <- rasterToPoints(grd_rast) %>% as_tibble() %>% set_colnames(c("x","y","GRID5KM_ID")) %>%
  st_as_sf(coords=c('x','y'),crs=st_crs(grd)) %>% st_transform(4326)
grd_xy <- grd_xy %>% st_coordinates() %>% as_tibble() %>% mutate(GRID5KM_ID=grd_xy$GRID5KM_ID) %>% 
  set_colnames(c("grd_x","grd_y","GRID5KM_ID"))

# unlist the traps and attach grid cells using nearest neighbors
OR_traps_long <- OR_traplocs %>% 
  mutate(traps_xy=purrr::map(traps,function(x)st_coordinates(x) %>% set_colnames(c('x','y')) %>% as_tibble())) %>%
  select(-traps) %>% 
  unnest(c(traps_xy))

# nearest neighbors
traps_nn <- RANN::nn2(grd_xy[,1:2],OR_traps_long[,c('x','y')],k=1)[[1]]
# attach to observations
OR_traps_long <- OR_traps_long %>% mutate(GRID5KM_ID = grd_xy$GRID5KM_ID[traps_nn])
# clean up variable names
OR_traps_long %<>%
  rename(ntraps_trip=ntraps)
```

Make weighted and unweighted rasters of trap counts. 

```{r}
# add tier weights to traps list-column
RecID_tier_weights %>% ungroup() %>%  select(Rec_ID,tot_vessels:tier_wt) %>% right_join(OR_traps_long)->OR_traps_long

OR_traps_long <- OR_traps_long %>% left_join(grd_xy,by="GRID5KM_ID")

# add total Oregon trap numbers by month
OR_traps_long <- OR_traps_long %>% 
  left_join(traps_month_key) %>% 
  group_by(year,month) %>% 
  # if no recorded trap limits, the number of traps is just the sum of all simulated traps
  mutate(max_traps_region=replace_na(max_traps_region,n()))
```

```{r}
count_traps <- function(yr,mth,use_weights=F,norm_index=F){
  mth_name <- month.name[mth]
  samp <- OR_traps_long %>%
    ungroup() %>%
    filter(year==yr,month==mth_name,!is.na(GRID5KM_ID))
  if(nrow(samp)==0){
    print(paste(mth_name,yr,"had zero observations."))
    zero_rast <- setValues(grd_rast,NA)
    names(zero_rast) <- paste0(mth_name,yr)
    return(zero_rast)
  } else {
    samp_sf <- samp %>% 
      mutate(norm=norm_index) %>% 
      group_by(norm,GRID5KM_ID,grd_x,grd_y,max_traps_region) %>% 
      # if using weights, the counted trap value is equal to the sum of the relative tier weights
      summarise(tot_traps=ifelse(use_weights,sum(tier_wt),n())) %>%
      ungroup() %>%
      # the relative index is the total traps in each grid cell, divided by the sum of all simulated traps in the region
      mutate(traps_relative=tot_traps/sum(tot_traps)) %>% 
      ungroup() %>% 
      # if normalizing by region, leave it as is. If looking for total traps, multiply by total regional traps (from fish tickets)
      mutate(ntraps_calc=traps_relative*max_traps_region) %>% 
      mutate(ntraps=ifelse(norm,traps_relative,ntraps_calc)) %>% 
      ungroup() %>% 
      sf::st_as_sf(coords=c('grd_x','grd_y'),crs=4326) %>% 
      st_transform(st_crs(grd)) %>%
      dplyr::select(ntraps) %>% 
      as_Spatial()
  out_rast <- raster::rasterize(samp_sf,grd_rast,field='ntraps',fun='first')
  names(out_rast) <- paste0(mth_name,yr)
  }
  
  print(paste(mth_name,yr,"done."))

  return(out_rast)
}


# sample map
p <- proc.time()
dec2013 <- count_traps(2013,12,use_weights = T,norm_index=F)
proc.time()-p
dec2013
plot(dec2013)
# data.table version (dunno why i can't get something like this to work in tidy) 
#   traps_sf<-setDT( samp )[, rbindlist( traps ) ] %>% 
#     sf::st_as_sf()
```
Apply to all years/months
```{r,echo=F,message=F,warning=F}
p <- proc.time()
years_months <- crossing(yr=2009:2018,mth=1:12)
trap_density_unweighted_norm <- purrr::map2(years_months$yr,years_months$mth,count_traps,use_weights=F,norm_index=T)
trap_density_weighted_norm <- purrr::map2(years_months$yr,years_months$mth,count_traps,use_weights=T,norm_index=T)
trap_density_unweighted <- purrr::map2(years_months$yr,years_months$mth,count_traps,use_weights=F,norm_index=F)
trap_density_weighted <- purrr::map2(years_months$yr,years_months$mth,count_traps,use_weights=T,norm_index=F)
proc.time()-p

trap_density_unweighted_stack <- raster::stack(trap_density_unweighted)
trap_density_weighted_stack <- raster::stack(trap_density_weighted)
trap_density_unweighted_norm_stack <- raster::stack(trap_density_unweighted_norm)
trap_density_weighted_norm_stack <- raster::stack(trap_density_weighted_norm)

write_rds(trap_density_unweighted_stack,here::here('data','processed',"OR_trap_counts_unweighted.rds"))
write_rds(trap_density_weighted_stack,here::here('data','processed',"OR_trap_counts_weighted.rds"))
write_rds(trap_density_unweighted_norm_stack,here::here('data','processed',"OR_trap_counts_unweighted_norm.rds"))
write_rds(trap_density_weighted_norm_stack,here::here('data','processed',"OR_trap_counts_weighted_norm.rds"))
```

## Compare to Vertical Lines Density by Grid Cell

Convert raster data to tibble of trap densities by year/month/grid cell

```{r}
grd_template <- grd_rast %>% rasterToPoints() %>% as_tibble() %>% set_names(c('x','y','GRID5KM_ID'))

# RasterStack to long form xy
all_obs_xy_weighted_norm <- purrr::map(as.list(trap_density_weighted_norm_stack),function(x){
  mth <- names(x)
  y<-rasterToPoints(x) %>% as_tibble() %>% set_names(c('x','y','value'))
  z <-y %>% mutate(month=mth)
  z
}) %>% bind_rows() %>% 
  mutate(year=str_sub(month,-4,-1) %>% as.numeric(),month=str_remove_all(month,"[:digit:]"),
         month=match(month,month.name)) %>% 
  left_join(grd_template)

all_obs_xy_unweighted <- purrr::map(as.list(trap_density_unweighted_stack),function(x){
  mth <- names(x)
  y<-rasterToPoints(x) %>% as_tibble() %>% set_names(c('x','y','value'))
  z <-y %>% mutate(month=mth)
  z
}) %>% bind_rows() %>% 
  mutate(year=str_sub(month,-4,-1) %>% as.numeric(),month=str_remove_all(month,"[:digit:]"),
         month=match(month,month.name)) %>% 
  left_join(grd_template)

all_obs_xy_weighted <- purrr::map(as.list(trap_density_weighted_stack),function(x){
  mth <- names(x)
  y<-rasterToPoints(x) %>% as_tibble() %>% set_names(c('x','y','value'))
  z <-y %>% mutate(month=mth)
  z
}) %>% bind_rows() %>% 
  mutate(year=str_sub(month,-4,-1) %>% as.numeric(),month=str_remove_all(month,"[:digit:]"),
         month=match(month,month.name)) %>% 
  left_join(grd_template)

# vert_lines_density <- purrr::map(as.list(trap_density),function(x){
#   mth <- names(x)
#   y<-rasterToPoints(x) %>% as_tibble() %>% set_names(c('x','y','density'))
#   z <-y %>% mutate(month=mth) %>% left_join(grd_template,by=c('x','y'))
#   z
# }) %>% bind_rows() %>%
#   select(GRID5KM_ID,month,density) %>% 
#   filter(!is.na(GRID5KM_ID)) %>% 
#   mutate(yr=str_extract(month,"[:digit:]+") %>% as.integer(),mth=str_extract(month,"[:alpha:]+")) %>% 
#   rowwise() %>% 
#   mutate(mth=which(month.name==mth)) %>% 
#   select(yr,mth,GRID5KM_ID,density) %>% 
#   rename(vlm_density=density)

# vert_lines_density %>% 
#   filter(GRID5KM_ID %in% unique(logs_trap_density$GRID5KM_ID)) %>% 
#   ggplot(aes(vlm_density))+
#   geom_density(alpha=0.5,fill=viridis_pal(begin=0.2,end=0.8)(1))+
#   facet_wrap(~mth)
```

Join and compare. First, using VMS-weighted, normalized density

```{r}
trap_dens_both <- logs_trap_density_month %>%
  left_join(all_obs_xy_weighted_norm) %>% 
  rename(vlm_density=value) %>% 
  # rename(vlm_trap_count=value) %>% 
  # mutate(vlm_density=vlm_trap_count/25) %>% 
  ungroup() %>% 
  distinct() %>% 
  mutate(logs_z=(max_pot_density-mean(max_pot_density,na.rm=T))/sd(max_pot_density,na.rm=T),
         vlm_z=(vlm_density-mean(vlm_density,na.rm=T))/sd(vlm_density,na.rm=T))

# ggplot(trap_dens_both,aes(max_pot_density,vlm_density))+
  # geom_point(size=0.5)+
  # geom_smooth(method='lm',formula='y~x-1')
trap_density_comparison_long <- trap_dens_both %>% 
  pivot_longer(contains("_density"),names_to = 'type',values_to = 'density')

trap_density_comparison <- trap_density_comparison_long %>% 
  ggplot(aes(density,fill=type))+
  geom_density(alpha=0.7,col=NA)+
  xlim(c(0,90))+
  scale_fill_manual(values=viridis_pal(begin=0.2,end=0.8)(2),labels=c('Logbook','VLM'))+
  labs(x='Trap Density',y="",fill='')
ggsave(here::here('plots','OR_trap_density_comparison.png'),trap_density_comparison,w=7,h=5)
```

```{r}
library(cowplot)
make_comparison_map <- function(yr,mth){
  dat <-trap_dens_both %>%
    filter(year==yr,month==mth) %>% 
    distinct(GRID5KM_ID,vlm_z,logs_z)
  if(nrow(dat)==0) stop('no observations for that month')
  dat %<>% left_join(grd,by="GRID5KM_ID") %>% st_as_sf()
  
  mnth_name<-month.name[mth]
  
  bbox <- st_bbox(dat)
  plt_vlm <- dat %>% 
    filter(vlm_z<=quantile(vlm_z,0.98,na.rm=T)) %>% 
    ggplot()+
    geom_sf(aes(fill=vlm_z),col=NA)+
    geom_sf(data=coaststates,fill='gray70')+
    scale_fill_viridis(na.value='white',limits=c(-1,1),breaks=seq(-1,1,by=0.5))+
    coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]),datum=NA)+
    labs(title=paste('VLM',mnth_name,yr),fill='Trap Density (Z-score)')+
    theme(axis.text=element_blank())
  plt_logs <- dat %>% 
    filter(logs_z<=quantile(logs_z,0.98,na.rm=T)) %>% 
    ggplot()+
    geom_sf(aes(fill=logs_z),col=NA)+
    geom_sf(data=coaststates,fill='gray70')+
    scale_fill_viridis(na.value='white',limits=c(-1,1),breaks=seq(-1,1,by=0.5))+
    coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]),datum=NA)+
    labs(title=paste('Logbooks',mnth_name,yr),fill='Trap Density (Z-score)')+
    theme(axis.text=element_blank())
  plot_grid(plt_vlm,plt_logs)
}
jan2014 <- make_comparison_map(yr=2014,mth=1)
jan2014
ggsave(here::here('plots','jan2014_zscore_comparison.png'),dec2013)
make_comparison_map(yr=2013,mth=12)

```

## Total Traps Timeseries Comparison

Compare the total number of estimated traps between logbooks and VLM over time

```{r}
# Total logbook traps
# have to figure this out
# logs_timeseries <- logs_trap_density_month %>% 
#   distinct() %>% 
#   select(date,n_pots) %>%
#   mutate(year=year(date),month=month(date)) %>% 
#   group_by(year,month) %>% 
#   summarise(tot_pots_logs=sum(n_pots,na.rm=T)) %>% 
#   ungroup()

# VLM timeseries
vlm_timeseries <- all_obs_xy_weighted %>% 
  group_by(year,month) %>% 
  summarise(tot_pots_vlm=sum(value,na.rm=T)) %>% 
  ungroup() %>% 
  mutate(date=paste(year,month,"01",sep="_") %>% as_date())

vlm_ts <-vlm_timeseries %>% 
  # filter(month %in% c(1:8,12)) %>% 
  ggplot(aes(date,tot_pots_vlm/1000))+
  geom_point()+geom_line()+
  labs(x="Date",y="Thousands of Traps in Oregon")
ggsave(here('plots','OR_vlm_ts.png'),vlm_ts,w=7,h=5)
```


# OLD

## Investigate Fish Ticket Numbering Issue

The fish tickets don't seem to line up to the numbers in the VMS-matched data. Let's investigate

```{r}
oregon_tix <- unique(logs$FTID)
vms_tix <- unique(traplocs$Rec_ID)
fishtix_tix <- fishtix_matched_all %>% .$Rec_ID %>% unique() %>% as.numeric()
# and tickets that match our vertical line model
tix_match <- intersect(fishtix_tix,oregon_tix)

# these don't match

## Raw, raw fish tickets with FTID column
tixfiles <- list.files("C:/Users/Owen.Liu/Documents/github/VMS-NWFSC-2/data/raw/fish tickets",full.names = T)

rawtest <- read_csv(tixfiles[4])
rawdat <- purrr::map_df(tixfiles,function(fl){
  read_csv(fl,col_types= cols_only(
    FISH_TICKET_ID = col_double(), 
    FTID = col_character(),
    PACFIN_PORT_CODE= col_character(), 
    PACFIN_GROUP_PORT_CODE= col_character(), 
    VESSEL_NUM= col_character(), 
    AGENCY_CODE= col_character(), 
    GEAR_CODE= col_double(),
    GEAR_NAME= col_character(), 
    PACFIN_GROUP_GEAR_CODE= col_character(), 
    REMOVAL_TYPE_CODE= col_character(), 
    REMOVAL_TYPE_NAME= col_character(), 
    LANDING_DATE= col_character(),
    LANDING_MONTH= col_double(), 
    LANDING_YEAR= col_double(), 
    PACFIN_SPECIES_CODE= col_character(), 
    LANDED_WEIGHT_LBS= col_double(), 
    EXVESSEL_REVENUE= col_double()))
})
raw_fishtix_FTID <- unique(rawdat$FTID)
tix_match2 <- intersect(raw_fishtix_FTID,oregon_tix)

# okay, these match, so let's make a matching key to match PacFIN Rec_IDs to FTIDs for Oregon
ftid_recid <- rawdat %>% select(FISH_TICKET_ID,FTID) %>% distinct() %>% 
  filter(FTID %in% tix_match2)

# join to logs data
logs %<>%
  left_join(ftid_recid,by='FTID') %>% 
  rename(Rec_ID=FISH_TICKET_ID)
```

## Match Individual Trips

Now that we can match logbook data to individual Oregon trips, we can explore the vertical line model on a trip-by-trip basis.

Note that in the logbook data, there are often more than one string of pots associated with each fish ticket.

```{r}
# logs_nest <- logs %>% 
#   filter(!is.na(Rec_ID)) %>% 
#   ungroup() %>% 
#   nest(logbook_traps=c(n_pots,adj_lbs,startlat,startlon,endlat,endlon))

# inner join to only retain trips/fish tickets that are associated with both VMS-simulated traps and logbooks
traplocs_sim_logbook <- logs %>% inner_join(traplocs)

# sample trip
# samptrip <- unique(traplocs_sim_logbook$Rec_ID) %>% sample(1)
samptrip <- 449325098
sample_logs <- traplocs_sim_logbook %>% filter(Rec_ID==samptrip)

lns <- purrr::map(1:nrow(sample_logs),function(i){
  ln <- sample_logs[i,c(7,6,9,8)] %>% as.numeric() %>%  matrix(nrow=2,byrow = T)
  ln[,1] <- -ln[,1]
  ln
}) %>%  st_multilinestring() %>% st_sfc() %>% st_sf() %>% st_set_crs(4326) %>%  st_transform(st_crs(grd))
simpots <- sample_logs %>% slice(1) %>% pluck('traps',1) %>% st_set_crs(4326) %>% st_transform(st_crs(grd))
totlogpots <- sum(sample_logs$n_pots)
totsimpots <- unique(sample_logs$ntraps)
bbox1 <- st_bbox(lns) %>% st_as_sfc()
bbox2 <- st_bbox(simpots) %>% st_as_sfc()
bbox <- st_union(bbox1,bbox2) %>% st_bbox()
ggplot()+
  geom_sf(data=coaststates,fill='gray70')+
  geom_sf(data=simpots)+
  geom_sf(data=lns,col='red',size=2)+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  labs(title=paste0("Simulated versus Logged Traps\nTotal Traps Logged: ",totlogpots,"\nTotal Traps Simulated: ",totsimpots))
```

We can map out individual tracks to compare

```{r}
trip_compare <- function(recid) {
  sample_logs <- traplocs_sim_logbook %>% filter(Rec_ID==recid) 
  lns <- purrr::map(1:nrow(sample_logs),function(i){
    ln <- sample_logs[i,c(7,6,9,8)] %>% as.numeric() %>%  matrix(nrow=2,byrow = T)
    ln[,1] <- -ln[,1]
    ln
  }) %>%  st_multilinestring() %>% st_sfc() %>% st_set_crs(4326) %>%  st_transform(st_crs(grd))
  simpots <- sample_logs %>% slice(1) %>% pluck('traps',1) %>% st_set_crs(4326) %>% st_transform(st_crs(grd))
  totlogpots <- sum(sample_logs$n_pots)
  totsimpots <- unique(sample_logs$ntraps)
  bbox1 <- st_bbox(lns) %>% st_as_sfc()
  bbox2 <- st_bbox(simpots) %>% st_as_sfc()
  bbox <- st_union(bbox1,bbox2) %>% st_bbox()
  ggplot()+
    geom_sf(data=coaststates,fill='gray70')+
    geom_sf(data=simpots)+
    geom_sf(data=lns,col='red',size=2)+
    coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
    theme(axis.text.x = element_text(angle=90,vjust=0.5))+
    labs(title=paste0("Simulated versus Logged Traps\nTotal Traps Logged: ",totlogpots,"\nTotal Traps Simulated: ",totsimpots))
}
```

```{r,fig.height=10,fig.width=10}
library(cowplot)
# samp_recids  <- unique(traplocs_sim_logbook$Rec_ID) %>% sample(6)
samp_recids <- c(50073739, 449325098, 427580126, 449333293, 427598582, 48070529)
samptrips <- purrr::map(samp_recids,function(x){
  trip_compare(x)
})
sample_trip_plots <- plot_grid(plotlist=samptrips,nrow=3)
sample_trip_plots
ggsave(here::here('plots','sample_comparison_plots.png'),sample_trip_plots,h=10,w=8)
```


