---
title: "Summarise VLM Outputs"
author: "Owen Liu"
date: "10/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster)
library(data.table)
library(tidyverse)
library(lubridate)
library(magrittr)
library(sf)
library(rnaturalearth)
library(fasterize)
library(tictoc)

library(furrr)
library(future)
```

# Purpose

The goal of this script is to take the output of the vertical line model simulation and apply functions that can summarise the total number and/or density of simulated crab pots by 1, 2, or month-long intervals, either expanded (accounting for relative VMS representation by vessel size class, tier level, and month) or unexpanded.

# Import Data

Simulated traps

```{r}
simtraps <- read_rds(here::here('vertical line model',"sim_traps_speed_depth_filters.rds"))
```

5x5km grid
```{r}
grd <- read_sf(here::here("data","raw","grid","regions_master_final_lamb.shp"))
names(grd)
# raster representation of grid.
# transform to a projection that uses meters instead of lat/long
grd_rast <- fasterize(grd,raster = raster(grd,res=5000,crs=crs(grd)),field="GRID5KM_ID")
# set -99999 to NA
grd_rast[grd_rast==-99999] <- NA
```

Fish tickets
```{r}
fishtix_matched_all <- read_rds(here::here('data','processed','fish tickets','fish_tickets_w_dayofseason.rds'))
# filter for out of season
fishtix_matched_all %<>% filter(ticket_day_of_season>=0)
```

Coastline (just for plotting)
```{r}
# a coastline, for plotting
# coastline for plotting
coaststates <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('California','Oregon','Washington','Nevada'))
coastline <- ne_coastline(scale='medium',returnclass = 'sf') %>% 
  st_crop(st_bbox(coaststates))
```

Trap locations, discarding VMS data
```{r}
# traps locations only (discard VMS data for now), attached to fish ticket data
traplocs <- simtraps %>% dplyr::select(Rec_ID,traps) %>% 
  left_join(fishtix_matched_all,by="Rec_ID") %>% 
  dplyr::select(Rec_ID,date,year,month,ticket_day_of_season,pacfin_port_code,port_group_code,agency_code,drvid,FINAL_LENGTH,traps) %>% 
  distinct(Rec_ID,.keep_all = T) %>% 
  # count number of traps associated with each trip
  mutate(ntraps=purrr::map_int(traps,nrow))
```

Vessel tier info

```{r}
tier_key <- read_rds(here::here('vertical line model','tier_information_year_month_allstates.rds'))
tier_key %<>% mutate(month=month.name[month]) %>% dplyr::select(-crabyr)
```

VMS representation index, for weighting of simulated traps
```{r}
vms_representation_index <- read_rds(here::here('vertical line model','vms_representation_index.rds'))
```


# Summarising Traps on the 5km Grid

We want to write a function that takes the full simulated traps dataset, filters it by 1 week, 2 week, or 1 month intervals, then gets counts of total traps per 5km grid cell

## Create Periods

First, we need to create identifiers for 1 week, 2 week, and 1 month periods.

```{r}
traplocs <- traplocs %>% 
  filter(!is.na(date)) %>% 
  # week of year
  mutate(period_1wk=week(date)) %>% 
  # 2-week segment of year
  mutate(period_2wk=round(period_1wk/2)+1)
```

## Add Tier Info to Each Fish Ticket

Using the 3 states' tier info, add a pot limit to the trap locations. IF POT LIMIT IS NA, ASSUME 500!

```{r}
traplocs <- traplocs %>% 
  left_join(tier_key,by=c('year','month','agency_code','drvid'))
```

Add the 2 different weighting factors- one for the total number of allowable traps, and one for relative VMS representation

```{r}
traplocs_wts <- traplocs %>% 

  group_by(drvid,period_1wk) %>% 
  mutate(total_traps_period1wk=sum(ntraps)) %>%
  
  # total traps weighting
  mutate(ntraps_permitted=coalesce(ntraps_permitted,500)) %>% 
  mutate(ntraps_wt=ifelse(total_traps_period1wk<ntraps_permitted,1,ntraps_permitted/total_traps_period1wk)) %>% 

  # add vessel size designation
  mutate(vessel_size_category=ifelse(FINAL_LENGTH<40,'small','large')) %>% 
  # join vms representation key
  left_join(vms_representation_index,by=c('year','month','agency_code','ntraps_permitted','vessel_size_category')) %>% 
  # relative vms representation weighting
  mutate(vms_representation_wt=1/propland) %>% 
  ungroup() %>% 
  dplyr::select(-propland)
```


Write the function to count traps, per grid cell, by period.

The `expand` option determines whether traps are up-weighted by the relative representation in VMS of each vessel's length/tier level in each year/month in the fish ticket data.

```{r}
count_traps <- function(yr,per,expand=F){
  samp <- traplocs_wts %>%
    ungroup() %>%
    filter(year==yr,period_1wk==per)
  
  # expand with VMS representation weight or no?
  samp <- samp %>% 
    mutate(trapwt=ifelse(expand,ntraps_wt*vms_representation_wt,ntraps_wt))
  
  # add as a "weight" for density calcs
  samp <- samp %>% 
    mutate(traps=purrr::map2(traps,trapwt,function(df,wt) mutate(df,wt=wt)))

  if(nrow(samp)==0){
    print(paste('Week',per,'in',yr,"had zero observations."))
    zero_rast <- setValues(grd_rast,NA)
    names(zero_rast) <- paste0('week'," ",per," ",yr)
    return(zero_rast)
  } else{
    samp_sf<-setDT( samp )[, rbindlist( traps ) ] %>%
    sf::st_as_sf(crs=4326) %>% 
    # transform to match the grid
    st_transform(st_crs(grd)) %>% 
    dplyr::select(wt) %>% 
    as_Spatial()
    
    # use raster::rasterize() to count the number of traps in each cell
  out_rast <- raster::rasterize(samp_sf,grd_rast,field='wt',fun='sum')
  names(out_rast) <- paste0('week'," ",per," ",yr)

  print(paste0('week'," ",per," ",yr," done."))
  return(out_rast)
  }

}
```

```{r}
# sample map
tic()
week35_2013 <- count_traps(2013,35)
week35_2013_expanded <- count_traps(2013,35,expand=T)
toc()
week35_2013
week35_2013_expanded
plot(week35_2013)
plot(week35_2013_expanded)

```

Apply to all years/months' data.

Non-expanded

```{r apply fxn}
tic("1 Week Interval Pot Counts, non-Expanded")
years_periods <- traplocs %>% ungroup() %>% distinct(year,period_1wk) %>% rename(yr=year,per=period_1wk)
future::plan(multisession)
trap_density_all_periods_no_expansion <- furrr::future_map2(years_periods$yr,years_periods$per,count_traps,expand=F)
toc()
trap_density_stack_no_expansion <- raster::stack(trap_density_all_periods_no_expansion)
write_rds(trap_density_stack_no_expansion,here::here('vertical line model',"trap_density_all_1week_periods_non_expanded.rds"))

```

Expanded
```{r}
tic("1 Week Interval Pot Counts, Expanded")
trap_density_all_periods_expanded <- furrr::future_map2(years_periods$yr,years_periods$per,count_traps,expand=T)
toc()
trap_density_stack_expanded <- raster::stack(trap_density_all_periods_expanded)
write_rds(trap_density_stack_expanded,here::here('vertical line model',"trap_density_all_1week_periods_expanded.rds"))
```

```{r}
future::plan(sequential)
```

