---
title: "Washington VLM Maps"
author: "Owen Liu"
date: "3/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pkgs, include=FALSE,message=F,warning=FALSE}
library(raster)
library(tidyverse)
library(knitr)
library(extrafont)
library(ggsci)
library(lubridate)
library(sf)
library(geosphere)
library(magrittr)
library(rnaturalearth)
library(gstat)
# library(data.table)
library(fasterize)
library(gganimate)
library(here)
library(RANN)

# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=12,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
options(dplyr.summarise.inform = FALSE)
```

# Introduction

The purpose of this document is to use the outputs from a vertical line/crab pot simulation estimation to investigate preliminary spatial and temporal patterns in the estimated number of traps in the water. We will do this by using a spatial grid to aggregate total crab traps across space and time.

```{r import data}
# fp <- "C:/Users/Owen.Liu/Documents/NWFSC Research/Dungeness Fleet Dynamics/VMS DATA CONFIDENTIAL/Processed Data/VMS/"
# fp_tix <- "C:/Users/Owen.Liu/Documents/NWFSC Research/Dungeness Fleet Dynamics/VMS DATA CONFIDENTIAL/Processed Data/Fish tickets/"

# fish tickets
# load(file=paste0(fp_tix,"FishTix_all_2009_2018.RData"))
fishtix_matched_all <- read_rds("C:/Users/Owen.Liu/Documents/github/VMS-NWFSC-2/data/processed/fish tickets/fish_tickets_w_dayofseason.rds")
# filter for out of season?
fishtix_matched_all %<>% filter(ticket_day_of_season>=0)

# a coastline, for plotting
# coastline for plotting
coaststates <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('California','Oregon','Washington','Nevada'))
coastline <- ne_coastline(scale='medium',returnclass = 'sf') %>% 
  st_crop(st_bbox(coaststates))

# Load if not already loaded
simulated_traps <- read_rds(here::here('data','processed',"sim_traps_speed_depth_filters.rds")) %>% ungroup()

# traps locations only (discard VMS data for now), attached to fish ticket data
traplocs <- simulated_traps %>% dplyr::select(Rec_ID,traps) %>% 
  left_join(fishtix_matched_all,by="Rec_ID") %>% 
  dplyr::select(Rec_ID,date,year,month,ticket_day_of_season,pacfin_port_code,port_group_code,agency_code,drvid,FINAL_LENGTH,traps) %>% 
  distinct(Rec_ID,.keep_all = T) %>% 
  # count number of traps associated with each trip
  mutate(ntraps=purrr::map_int(traps,nrow))

# 5x5 grid shapefile
# grd <- read_sf(here::here("Input_Data","5x5 Grid","five_km_grid_polys_geo.shp"))
grd <- read_sf(here::here("data","grid","regions_master_final_lamb.shp"))
names(grd)

meta <- readxl::read_xlsx(here::here("data","grid","regions_master_final_lamb_metadata.xlsx")) 

meta %>% kable()
```

```{r}
# WA only traps
wa_traps <- traplocs %>% filter(agency_code=="W",ticket_day_of_season>=0)
```


# Join Grid

We make a grid representation of the simulated traps data, assigning a 5km grid cell to each observation. Then, we produce both an unweighted version of monthly rasters for WA.

First, attach the appropriate grid cell to each simulated trap.

```{r grid fxn,warning=FALSE,fig.height=8,fig.width=4}
# Function to build a grid from a given year/month's simulated traps data

# raster representation of grid.
# transform to a projection that uses meters
grd_rast <- fasterize(grd,raster = raster(grd,res=5000,crs=crs(grd)),field="GRID5KM_ID")
# set -99999 to NA
grd_rast[grd_rast==-99999] <- NA
# wa/or bbox
wa_extent <- grd %>% 
  filter(STATE %in% c('OR','WA')) %>% 
  st_bbox() %>% extent()

grd_xy <- rasterToPoints(grd_rast) %>% as_tibble() %>% set_colnames(c("x","y","GRID5KM_ID")) %>%
  st_as_sf(coords=c('x','y'),crs=st_crs(grd)) %>% st_transform(4326)
grd_xy <- grd_xy %>% st_coordinates() %>% as_tibble() %>% mutate(GRID5KM_ID=grd_xy$GRID5KM_ID) %>% 
  set_colnames(c("grd_x","grd_y","GRID5KM_ID"))

# unlist the traps and attach grid cells using nearest neighbors
wa_traps_long <- wa_traps %>% 
  mutate(traps_xy=purrr::map(traps,function(x)st_coordinates(x) %>% set_colnames(c('x','y')) %>% as_tibble())) %>%
  select(-traps) %>% 
  unnest(c(traps_xy))

# nearest neighbors
traps_nn <- RANN::nn2(grd_xy[,1:2],wa_traps_long[,c('x','y')],k=1)[[1]]
# attach to observations
wa_traps_long <- wa_traps_long %>% 
  mutate(GRID5KM_ID = grd_xy$GRID5KM_ID[traps_nn]) %>% 
  left_join(grd_xy,by="GRID5KM_ID")
# clean up variable names
wa_traps_long %<>%
  rename(ntraps_trip=ntraps)
```

Make unweighted rasters of trap density.

```{r}
count_traps <- function(yr,mth,norm_index=F){
  mth_name <- month.name[mth]
  samp <- wa_traps_long %>%
    ungroup() %>%
    filter(year==yr,month==mth_name,!is.na(GRID5KM_ID))
  if(nrow(samp)==0){
    print(paste(mth_name,yr,"had zero observations."))
    zero_rast <- setValues(grd_rast,NA)
    names(zero_rast) <- paste0(mth_name,yr)
    return(zero_rast)
  } else {
    samp_sf <- samp %>% 
      mutate(norm=norm_index) %>% 
      group_by(norm,GRID5KM_ID,grd_x,grd_y) %>% 
      # count all the traps
      summarise(tot_traps=n()) %>%
      ungroup() %>%
      # the relative index is the total traps in each grid cell, divided by the sum of all simulated traps in the region
      mutate(traps_relative=tot_traps/sum(tot_traps)) %>%
      # if normalizing by region, leave it as is. If looking for total traps, multiply by total regional traps (from fish tickets)
      # mutate(ntraps_calc=traps_relative*max_traps_region) %>% 
      mutate(ntraps=ifelse(norm,traps_relative,tot_traps)) %>% 
      ungroup() %>% 
      sf::st_as_sf(coords=c('grd_x','grd_y'),crs=4326) %>% 
      st_transform(st_crs(grd)) %>%
      dplyr::select(ntraps) %>% 
      as_Spatial()
  out_rast <- raster::rasterize(samp_sf,grd_rast,field='ntraps',fun='first')
    #crop to WA grid
    # crop(wa_extent)
  names(out_rast) <- paste0(mth_name,yr)
  }
  
  print(paste(mth_name,yr,"done."))

  return(out_rast)
}


# sample map
p <- proc.time()
feb2013 <- count_traps(2013,2,norm_index=F)
proc.time()-p
feb2013
plot(feb2013)
# data.table version (dunno why i can't get something like this to work in tidy) 
#   traps_sf<-setDT( samp )[, rbindlist( traps ) ] %>% 
#     sf::st_as_sf()
```

Apply to all years/months' data.

```{r apply fxn}
p <- proc.time()
years_months <- crossing(yr=2009:2018,mth=1:12)
trap_density_unweighted_norm <- purrr::map2(years_months$yr,years_months$mth,count_traps,norm_index=T)
trap_density_unweighted <- purrr::map2(years_months$yr,years_months$mth,count_traps,norm_index=F)

trap_density_unweighted_stack <- raster::stack(trap_density_unweighted)
trap_density_unweighted_norm_stack <- raster::stack(trap_density_unweighted_norm)

write_rds(trap_density_unweighted_stack,here::here('data','processed',"WA_trap_counts_unweighted.rds"))
write_rds(trap_density_unweighted_norm_stack,here::here('data','processed',"WA_trap_counts_unweighted_norm.rds"))
```


# Animate output

We can then represent the output in an animation

```{r interpolate output,warning=FALSE,fig.height=8,fig.width=4}
# RasterStack to long form xy
all_obs_xy_unweighted_norm <- purrr::map(as.list(trap_density_unweighted_norm_stack),function(x){
  mth <- names(x)
  y<-rasterToPoints(x) %>% as_tibble() %>% set_names(c('x','y','value'))
  z <-y %>% mutate(month=mth)
  z
}) %>% bind_rows()

all_obs_xy_unweighted <- purrr::map(as.list(trap_density_unweighted_stack),function(x){
  mth <- names(x)
  y<-rasterToPoints(x) %>% as_tibble() %>% set_names(c('x','y','value'))
  z <-y %>% mutate(month=mth)
  z
}) %>% bind_rows()

#bounding box for animation

# bbox=grd %>% filter(STATE %in% c("WA","OR")) %>% st_bbox()
bbox = c(750000,1500000,1013103,2045068)

# test plot
library(viridis)
coaststates %<>% st_transform(st_crs(grd))
test<-all_obs_xy_unweighted %>% 
  filter(month=="February2013")

test %>% ggplot()+
    geom_tile(aes(x,y,fill=value/25),na.rm=T,alpha=0.8,interpolate=FALSE)+
    geom_sf(data=coaststates,col=NA,fill='gray50')+
    scale_fill_viridis(na.value=NA,option="C",name="Relative Density\nIndex")+
    coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
    labs(x='',y='',fill='Traps per\nsq. km',title='{current_frame}')+
    theme(axis.text.x.bottom = element_text(angle=45),
          legend.position = c(0.8,0.3),
          title=element_text(size=16),
          legend.title = element_text(size=10))

# ordering factor for frames of the animation
all_obs_xy_unweighted %<>% mutate(labeler=factor(month,levels=unique(month)))
# # produce the animation
# library(viridis)
# library(gifski)
# traps.gif <- all_obs_xy_unweighted %>%
#   ggplot()+
#     geom_tile(aes(x,y,fill=value/25),na.rm=T,alpha=0.8,interpolate=TRUE)+
#     geom_sf(data=coaststates,col=NA,fill='gray50')+
#     scale_fill_viridis(na.value=NA,option="C",name="Traps per\nsq. km")+
#     coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
#     labs(x='',y='',fill='Traps per\nsq. km',title='{current_frame}')+
#     theme(axis.text.x.bottom = element_text(angle=45),
#           legend.position = c(0.7,0.6),
#           title=element_text(size=16))+
#     transition_manual(labeler)
# animate(traps.gif,fps=1,width=500,height=1000,renderer = gifski_renderer())
# 
# #save the animation
# # fp_plots <- "C:/Users/Owen.Liu/Documents/NWFSC Research/Dungeness Fleet Dynamics/VMS DATA CONFIDENTIAL/Processed Data/owen processed/plots/"
# anim_save(here::here('plots',"wa_trap_density_all_years_unweighted.gif"))
```

# Average Distribution of Pots

Average distribution, all years and months, using the normalized data.

```{r}
wa_trap_distribution<-all_obs_xy_unweighted_norm %>%  
  group_by(x,y) %>% 
  summarise(meannorm=mean(value)) %>% 
  ungroup() %>% 
  # hmm how to summarise? Normalize? divide by max?
  mutate(rescale=(meannorm-mean(meannorm,na.rm=T))/sd(meannorm,na.rm=T))

wa_trap_dist_map <- wa_trap_distribution %>% 
  ggplot()+
    geom_raster(aes(x,y,fill=meannorm),na.rm=T,alpha=0.8,interpolate=F)+
    geom_sf(data=coaststates,col=NA,fill='gray50')+
    scale_fill_viridis(na.value=NA,option="C",name="Relative Density\nIndex")+
    coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
    labs(x='',y='',fill='Traps per\nsq. km',title='Average Distribution of Traps\nall years and months')+
    theme(axis.text.x.bottom = element_text(angle=45),
          legend.position = "right",
          title=element_text(size=12),
          legend.text = element_blank(),
          legend.title = element_text(size=10))
wa_trap_dist_map
ggsave(here('vertical line model','plots','wa_trap_density_map.png'),wa_trap_dist_map,w=4,h=6)
```


# Trap Density Histogram

```{r distribution of trap densities}
trap_density_hist <- all_obs_xy_unweighted %>% 
  filter(!is.na(value),value>0) %>% 
  ggplot(aes(value/25))+
  geom_density(fill='red',alpha=0.98,col=NA)+
  scale_x_continuous(limits=c(0,50))+
  labs(x="Density (Traps per sq. Km)",y="Kernel Density",title="Distribution of Positive Trap Densities in WA \nall years and months")
ggsave(here('vertical line model','plots','wa_trap_density_histogram.png'),trap_density_hist,w=6,h=6)
```

# Time Series of Total Pots
```{r}
wa_mean_density_ts <- all_obs_xy_unweighted %>%
  mutate(date=parse_date(month,format="%B%Y")) %>% 
  group_by(date) %>% 
  summarise(mn=mean(value/25,na.rm=T)) %>% 
  ggplot(aes(date,mn))+
  geom_col(stat='identity',col='black',fill='black')+
  labs(y="pots/sq km",x="month")+
  theme(axis.text.x = element_text(angle=90))
wa_total_pots_ts <- all_obs_xy_unweighted %>%
  mutate(date=parse_date(month,format="%B%Y")) %>% 
  group_by(date) %>% 
  summarise(totpots=sum(value)) %>% 
  ggplot(aes(date,totpots/1000))+
  geom_point()+geom_line()+
  scale_x_date(date_breaks="6 months",date_minor_breaks = "2 months")+
  scale_y_continuous(limits=c(0,140),breaks = seq(0,140,by=20))+
  labs(y="Thousands of Pots",x="month",title="Total Simulated Pots in Washington")+
  theme(axis.text.x = element_text(angle=90,vjust=0.5))
wa_mean_density_ts
wa_total_pots_ts
```


# Depth Distribution of Pots

```{r}
trap_depths <- simulated_traps %>% 
  mutate(depths=purrr::map(traps,pluck('depth'))) %>% 
  filter(Rec_ID %in% wa_traps$Rec_ID) %>% 
  select(Rec_ID,depths) %>% 
  unnest(cols=c(depths)) %>% 
  left_join(fishtix_matched_all %>% select(Rec_ID,drvid,year,month))

trap_depth_p <- trap_depths %>% 
  sample_n(1e6) %>% 
  ggplot(aes(depths))+
  geom_density(fill='darkred',alpha=0.8)+
  labs(x="Depth (m)",y="Kernel Density")
trap_depth_p
ggsave(here('vertical line model','plots','wa_traps_depth_distribution.png'),trap_depth_p,w=6,h=5)
```