---
title: "Step by Step VMS Processing"
author: "Owen Liu"
date: Last Run "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup}
library(knitr)
library(here)
```

# Master VMS and Fish Ticket Processing Workflow

This document collates all individual steps of VMS/Fish Ticket processing for a chosen year.

```{r timer, include=FALSE}
alltime <- proc.time()
```

## Choose Year to Process

This parameter will get passed through all processing steps

```{r choose year}
alltime <- proc.time()
process_year <- 2016
```



# Step 4: Match VMS and Fish Ticket Data

```{r step5, message=F,warning=F,child=here::here('process steps','04_match_vms_fishtix.Rmd')}
```
```{r}
x<-proc.time()-alltime
cat('This step for',process_year,'VMS data took',round(x[3]/60,2),'minutes to run.')
```

**So far, this pipeline for `r process_year` VMS data has taken `r round(x[3]/60,2)` minutes to run.**

# Step 5: Filter Matched Data by Speed and Proximity to Ports

```{r step5, message=F,warning=F,child=here::here('process steps','05_filter_matched_data.Rmd')}
```
```{r}
x<-proc.time()-alltime
```

**So far, this pipeline for `r process_year` VMS data has taken `r round(x[3]/60,2)` minutes to run.**

# Step 6: Interpolate and Regularize Data

```{r step5, message=F,warning=F,child=here::here('process steps','06_interpolate.Rmd')}
```
```{r}
x<-proc.time()-alltime
```

**So far, this pipeline for `r process_year` VMS data has taken `r round(x[3]/60,2)` minutes to run.**