---
title: "Calculating Fishing Behavioral Metrics"
author: "Owen Liu"
date: "2/4/2020"
output: html_document
---

## Purpose

Following O'Farrell et al. (2019), we want to use VMS and fish ticket data to define behavioral metrics for the west coast Dungeness crab fishery. These can then be used for classifying behavior through clustering and perhaps looking at behavior change surrounding whale entanglement and the 2014-16 marine heat wave.

## Import Data
```{r}
library(tidyverse)
library(magrittr)
library(here)
library(lubridate)
library(sf)
library(rnaturalearth)

# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=12,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)

# DCRB fish tickets only, with day of season included
fishtix <- read_rds(here::here('data','processed','fish tickets','fish_tickets_w_dayofseason.rds'))

# All fish tickets, including non-DCRB trips
fishtix_all <- purrr::map_df(2009:2019, function(yr){
  read_rds(paste0(here::here('data','processed','fish tickets'),"/",yr,"fishtix_vlengths.rds"))
}) %>% left_join(select(fishtix,Rec_ID,crab_season))

# crab season indicator
recid_crabseason <- fishtix %>% select(Rec_ID,crab_season) %>% distinct()

# Non-interpolated VMS data, filtered for DCRB trips
vms <- purrr::map_df(2009:2019, function(yr){
  read_rds(paste0(here::here('data','processed','matched','filtering',yr),"matched_filtered.rds"))
}) %>% 
  filter(TARGET_rev=='DCRB') %>% 
  # join crab season indicator
  left_join(recid_crabseason,by='Rec_ID')

```

## Metrics to Calculate

Here is a running list of the metrics to calculate. Some metrics have multiple types or options of calculations to perform

1.  Port use: Number of ports visited per month, average ports visited per trip, Shannon diversity index of port use
2.  Revenue: Mean Dungeness revenue per trip, SD revenue per trip
3.  Trip Length: Days per trip, distance per trip, SD days per trip, SD distance per trip
4.  Exploration: Mean cumulative choice entropy
5.  Mobility: Home range
6.  Risk-taking: Propensity to fish in high winds
7.  Derby fishing: Mean day-of-season of 90% annual catch reached
8.  Participation in other fisheries: percent of revenue from non-DCRB fisheries
9.  Vessel size (not included as a variable yet)

```{r}
# Start with a version of the fish tickets that has only the information that we need to calculate metrics
fishtix %<>%
  select(Rec_ID:agency_code,date,crab_season,DCRB_revenue,FINAL_LENGTH,ticket_day_of_season)
```

We only want to investigate boats that participate in the Dungeness crab fishery. We can make a list of those, with some cutoff for revenue from the DCRB fishery.

```{r}
dcrb_revenue <- vms %>% 
  filter(drvid!="MISSING") %>% 
  mutate(year=year(UTCDATETIM)) %>% 
  mutate(nov1_season=case_when(
    month(date)>10 ~ paste0(year,"-",year+1),
    month(date)<=10 ~ paste0(year-1,"-",year),
    TRUE ~ NA_character_)) %>% 
  group_by(drvid,nov1_season) %>% 
  summarise(dcrb_revenue=sum(DCRB_revenue[which(TARGET_rev=="DCRB")]))

# what is the distribution of annual revenue, removing vessels with zero DCRB revenue
dcrb_revenue %>% 
  filter(dcrb_revenue!=0) %>% 
  ggplot(aes(dcrb_revenue))+
  geom_density(fill='#33638DFF')+
  labs(x='annual total DCRB revenue per vessel')

# what's the 5 percent quantile?
dcrb_revenue %>% filter(dcrb_revenue != 0) %>% pluck("dcrb_revenue") %>% quantile(0.05)

```

Five percent of the vessels made less than $101813 in at least one year of the data. We retain only those vessels that made more than this

```{r}
dcrb_vessels <- dcrb_revenue %>%
  filter(dcrb_revenue != 0) %>% 
  group_by(nov1_season) %>% 
  filter(dcrb_revenue>quantile(dcrb_revenue,0.05)) %>% 
  ungroup() %>% 
  distinct()
vessel_list <- unique(dcrb_vessels$drvid)
```

### Vessel Size

Use processed fish tickets to extract vessel sizes

```{r}
vessel_size_metric <-  fishtix %>%
  filter(drvid %in% vessel_list) %>% 
  distinct(drvid,crab_season,FINAL_LENGTH) %>% 
  rename(vessel_size=FINAL_LENGTH)

# plot vessel size metric
vessel_size_plot <- vessel_size_metric %>% 
  ggplot(aes(vessel_size))+
  geom_density(fill='#33638DFF')+
  labs(x='Vessel Length (m)')
vessel_size_plot
ggsave(here::here('fishing behavior','plots','vessel_size_metric.png'),vessel_size_plot,w=6,h=4)
write_rds(vessel_size_metric,here::here('fishing behavior','vessel_sizes.rds'))
```


### Port Use Metrics

Calculate number of ports visited per month, average ports visited per trip, and Shannon diversity index of port use. We calculate these from the VMS data

```{r}
# Average number of ports per fishing trip
ports_by_trip <- vms %>% 
  group_by(Rec_ID,drvid,crab_season) %>% 
  summarise(n_ports=n_distinct(in_port,na.rm = T)) %>% 
  mutate(n_ports=max(n_ports,1)) %>% 
  ungroup() %>% 
  group_by(drvid,crab_season) %>% 
  summarise(n_ports_trip=mean(n_ports,na.rm=T)) %>% 
  ungroup()

# Average number of ports per month
ports_by_month <- vms %>% 
  mutate(mth=month(date)) %>% 
  group_by(drvid,mth,crab_season) %>% 
  summarise(n_ports_mth=n_distinct(in_port,na.rm = T)) %>% 
  mutate(n_ports_mth=max(n_ports_mth,1)) %>% 
  ungroup() %>% 
  group_by(drvid,crab_season) %>% 
  summarise(ports_mean_mth=mean(n_ports_mth,na.rm=T)) %>% 
  ungroup()

# Shannon diversity index of port use by season
# For this metric we count each port visit as equal, and calculate the index on
# total diversity of port use in a given crab season
ports_diversity <- vms %>% 
  filter(!is.na(in_port)) %>% 
  select(drvid,crab_season,in_port) %>% 
  group_by(drvid,crab_season,in_port) %>% 
  tally(name="n_visits") %>% 
  ungroup() %>% 
  group_by(drvid,crab_season) %>% 
  mutate(total_visits=sum(n_visits,na.rm=T),prop_visits=n_visits/total_visits) %>% 
  mutate(plnp=prop_visits*log(prop_visits)) %>% 
  summarise(port_shannon= -1*sum(plnp)) %>% 
  ungroup()

# Collected port use metrics
port_use_metrics <- ports_by_trip %>% 
  left_join(ports_by_month) %>% 
  left_join(ports_diversity)

# plot diversity metric
port_diversity_plot <- port_use_metrics %>% 
  ggplot(aes(port_shannon))+
  geom_density(fill='#33638DFF')+
  labs(x='Diversity of Port Use')
port_diversity_plot
ggsave(here::here('fishing behavior','plots','port_diversity_metric.png'),port_diversity_plot,w=6,h=4)
```

### Revenue Metrics

Calculate mean and SD Dungeness revenue per trip. We can calculate this straight from the fish tickets

```{r}
revenue_metrics <- fishtix %>%
  filter(drvid %in% vessel_list) %>% 
  distinct(drvid,crab_season,Rec_ID,DCRB_revenue) %>% 
  group_by(drvid,crab_season) %>% 
  summarise(revenue_mean=mean(DCRB_revenue,na.rm=T),revenue_sd=sd(DCRB_revenue,na.rm=T)) %>% 
  ungroup()
revenue_mean_plot <- revenue_metrics %>% 
  ggplot(aes(revenue_mean))+
  geom_density(fill='#33638DFF')+
  labs(x='mean revenue')
revenue_sd_plot <- revenue_metrics %>% 
  ggplot(aes(revenue_sd))+
  geom_density(fill='#33638DFF')+
  labs(x='SD revenue')
revenue_mean_plot
ggsave(here::here('fishing behavior','plots','revenue_mean_metric.png'),revenue_mean_plot,w=6,h=4)
revenue_sd_plot
ggsave(here::here('fishing behavior','plots','revenue_sd_metric.png'),revenue_sd_plot,w=6,h=4)
```


### Trip Length Metrics

Using the VMS data, we can calculate days per trip, distance traveled per trip, and the standard deviations of these.

```{r}
trip_metrics <- vms %>% 
  filter(drvid %in% vessel_list) %>% 
  mutate(segment_dist=segment_dur*avg_speed_recalc) %>% 
  group_by(Rec_ID,drvid,crab_season,trip_dur) %>%
  # trip distance IN KILOMETERS
  summarise(trip_dist=sum(segment_dist,na.rm=T)/1000) %>% 
  ungroup() %>% 
  group_by(drvid,crab_season) %>% 
  summarise(trip_dist_mean=mean(trip_dist,na.rm = T),trip_dur_mean=mean(trip_dur,na.rm = T),trip_dist_sd=sd(trip_dist,na.rm=T),trip_dur_sd=sd(trip_dur,na.rm=T)) %>%
  ungroup()
trip_plot <- trip_metrics %>% 
  ggplot(aes(trip_dist_mean))+
  geom_density(fill='#33638DFF')+
  labs(x='trip distance (km)')
trip_plot
ggsave(here::here('fishing behavior','plots','trips_metric.png'),trip_plot,w=6,h=4)
```

### Derby Metric

For each fishing vessel in each year, calculate the day-of-season at which the boat finished catching 90% of its cumulative catch for the year. This metric should give a sense for how much a given vessel participates in the early-season derby. We can calculate this using the fish ticket data.

```{r}
derby_metric <- fishtix %>% 
  filter(drvid %in% vessel_list) %>% 
  group_by(drvid,crab_season) %>% 
  # find total and 90% revenue for each vessel in each crab season
  mutate(total_dcrb_revenue=sum(DCRB_revenue,na.rm=T),quant90_dcrb_revenue=0.9*total_dcrb_revenue) %>% 
  arrange(date) %>% 
  # cumulative sum of revenue for each vessel across the year
  mutate(dcrb_cum_revenue=cumsum(DCRB_revenue)) %>% 
  mutate(is_quant90=dcrb_cum_revenue>=quant90_dcrb_revenue) %>% 
  # find the ticket that crossed the 90% threshold and mark its date
  filter(cumany(is_quant90)) %>% 
  slice(1) %>% 
  select(drvid,crab_season,ticket_day_of_season) %>% 
  rename(quant90_day=ticket_day_of_season)

derby_plot <- derby_metric %>% 
  ggplot(aes(quant90_day))+
  geom_density(fill='#33638DFF')+
  labs(x='day of season reaching 90% cumulative catch')
derby_plot
ggsave(here::here('fishing behavior','plots','derby_metric.png'),derby_plot,w=6,h=4)
```

### Participation Metrics

Calculate what percent of a vessel's revenue comes from non-DCRB fisheries. We do this with the unfiltered fish ticket data, counting the non-DCRB revenue across all trips.

Choice made here: Because other fisheries do not have seasons in the same way that crab do, we define November to the end of June as our window to look for other revenue. For example, for a crab season 2008-2009, we search for other revenue from Nov 1, 2008 to June 30, 2009.

```{r}
participation_metric <- fishtix_all %>% 
  filter(drvid %in% vessel_list) %>% 
  # add general indicator of revenue season
  mutate(revenue_season=case_when(
    month(date)>10 ~ paste0(year,"-",year+1),
    month(date)<=6 ~ paste0(year-1,"-",year),
    TRUE ~ NA_character_)) %>% 
  filter(!is.na(revenue_season)) %>% 
  group_by(drvid,revenue_season) %>% 
  summarise(total_dcrb_revenue=sum(DCRB_revenue,na.rm=T),
            total_other_revenue=sum(other_rev+SABL_revenue+LOBS_revenue+SPRW_revenue),
            total_other_tix=n_distinct(Rec_ID[which(TARGET_rev != "DCRB")]),
            total_dcrb_tix=n_distinct(Rec_ID[which(TARGET_rev == "DCRB")])) %>% 
  mutate(total_other_revenue=coalesce(total_other_revenue,0)) %>% 
  mutate(prop_other_revenue=total_other_revenue/(total_other_revenue+total_dcrb_revenue),
         prop_other_tix=total_other_tix/(total_other_tix+total_dcrb_tix)) %>% 
  ungroup() %>% 
  rename(crab_season=revenue_season) %>% 
  select(drvid,crab_season,prop_other_revenue,prop_other_tix)

participation_rev_plot <- participation_metric %>% 
  ggplot(aes(prop_other_revenue))+
  geom_density(fill='#33638DFF')+
  labs(x='Proportion of Revenue from Other Fisheries')
participation_tix_plot <- participation_metric %>% 
  ggplot(aes(prop_other_tix))+
  geom_density(fill='#33638DFF')+
  labs(x='Proportion of Tickets Targeting Other Fisheries')
participation_rev_plot
participation_tix_plot
ggsave(here::here('fishing behavior','plots','participation_rev_metric.png'),participation_rev_plot,w=6,h=4)
ggsave(here::here('fishing behavior','plots','participation_tix_metric.png'),participation_tix_plot,w=6,h=4)
```

### Risk Metric

Following O'Farrell et al., we extract wind speeds from NOAA's ERDAPP server and use them to measure a fisher's propensity to go out in stormy weather. Following Pfeiffer and Gratz (2016), who also studied west coast fishing, we calculate a max (95th percentile) wind speed per trip, and then calculate what percent of a vessel's trips in a given crab season were pursued in winds >7.5m/s.

Download all winds for the study area.
DEFAULT: not run because downloading takes a long time

02/18/2020: https://www.ncei.noaa.gov/thredds/ncss/uv/clm_agg/Latest_Climatology_of_Ocean_Winds_best.ncd/dataset.html
https://rdrr.io/github/bocinsky/thredds/f/README.md


```{r, eval=FALSE}
library(rerddap)
# Range of Lat/Lon for entire dataset
range(vms$LATITUDE)
windslat <- c(34,49)
360+range(vms$LONGITUDE) # Longitudes in the winds dataset are in degrees E
windslon <- c(233,245)

# Have to use some if statements here because of weird breaks in the two datasets
winds_all <- purrr::map_df(2009:2018, function(yr) {
  if(yr <= 2010){
    times <- c(paste0(yr,"-01-01"),paste0(yr,"-12-31"))
    winds <- griddap("ncdcOwDly", # daily blended winds dataset
                      time=times,
                      latitude=windslat,
                      longitude=windslon,
                      fields=c("w")) %>%  # scalar wind speed
      pluck("data")
  } 
  if(yr > 2010){
    times <- c(paste0(yr,"-01-01"),paste0(yr,"-12-31"))
    winds <- griddap("erdQAdivmod1day", # daily blended winds dataset
                      time=times,
                      latitude=windslat,
                      longitude=windslon,
                      fields=c("mod")) %>%  # scalar wind speed
      pluck("data") %>% rename(w=mod)
  } 
  
  # if(yr == 2011){
  #   times1 <- c(paste0(yr,"-01-01"),paste0(yr,"-09-30"))
  #   winds1 <- griddap("ncdcOwDly", # daily blended winds dataset
  #                     time=times1,
  #                     latitude=windslat,
  #                     longitude=windslon,
  #                     fields=c("w")) %>% # scalar wind speed
  #     pluck("data")
  #   
  #   times2 <- c(paste0(yr,"-10-01"),paste0(yr,"-12-31"))
  #   winds2 <- griddap("ncdcOwDlyP", # daily blended winds dataset
  #                     time=times2,
  #                     latitude=windslat,
  #                     longitude=windslon,
  #                     fields=c("w") )%>%    # scalar wind speed
  #     pluck("data")
  #   winds <- bind_rows(winds1,winds2)
  # } 
  # if(yr>2011&yr<2018){
  #   times <- c(paste0(yr,"-01-01"),paste0(yr,"-12-31"))
  #   winds <- griddap("ncdcOwDlyP", # daily blended winds dataset
  #                     time=times,
  #                     latitude=windslat,
  #                     longitude=windslon,
  #                     fields=c("w")) %>% # scalar wind speed
  #     pluck("data") 
  # }
  # if(yr==2018){
  #   times <- c(paste0(yr,"-01-01"),paste0(yr,"-10-09"))
  #   winds <- griddap("ncdcOwDlyP", # daily blended winds dataset
  #                     time=times,
  #                     latitude=windslat,
  #                     longitude=windslon,
  #                     fields=c("w")) %>% # scalar wind speed
  #     pluck("data") 
  # }  
  out <- winds %>% as_tibble() %>% mutate(utcdate=as_date(time)) %>% select(-altitude)
  out
})
write_rds(winds_all,here::here('data','raw','blended winds','erddap_raw_daily_winds.rds'))
```

Match winds to VMS pings

```{r}
# if not loaded
winds_all <- read_rds(here::here('data','raw','blended winds','erddap_raw_daily_winds.rds'))

library(RANN)
vms_latlons <- vms %>% select(LATITUDE,LONGITUDE) %>%
  # longitude has to be in degrees E
  mutate(LONGITUDE=360+LONGITUDE)

winds_latlons <- winds_all %>% select(lat,lon) %>% distinct()

# Match VMS lat/lon to ERDDAP lat/lon
erddap_vms_match <- RANN::nn2(winds_latlons,vms_latlons,k=1)$nn.idx

# add appropriate erddap lat/lons to vms lat/lons data
vms_latlons %<>% mutate(matchlat=winds_all$lat[erddap_vms_match],matchlon=winds_all$lon[erddap_vms_match]) %>% mutate(LONGITUDE=LONGITUDE-360)

# Finally, join winds data to VMS data
vms_winds <- vms %>% 
  # join the lat/lon matching key
  bind_cols(select(vms_latlons,matchlat,matchlon)) %>% 
  select(drvid,crab_season,Rec_ID,UTCDATETIM,LATITUDE,LONGITUDE,matchlat,matchlon) %>% 
  #convert UTC times to dates
  mutate(utcdate=as_date(UTCDATETIM)) %>% 
  # join winds data by date and matching latitude/longitude
  left_join(select(winds_all,-time),by=c('utcdate'='utcdate','matchlat'='lat','matchlon'='lon'))
```

Calculate an extreme wind value for each trip (95th percentile of wind speed per trip).

```{r}
winds_per_trip <- vms_winds %>% 
  filter(drvid %in% vessel_list) %>% 
  group_by(drvid,crab_season,Rec_ID) %>% 
  summarise(wind_quant95=quantile(w,0.95,na.rm = T)) %>% 
  ungroup()
winds_per_trip %>% 
  ggplot(aes(wind_quant95))+
  geom_density(fill='#33638DFF')+
  labs(x='95th percentile of wind speed (m/s)')

risk_metric <- winds_per_trip %>% 
  # define whether a trip is 'high wind'
  mutate(is_high_winds=wind_quant95>=7.5) %>% 
  # count proportion of high wind trips by vessel/season
  group_by(drvid,crab_season) %>% 
  summarise(prop_highwinds=sum(is_high_winds,na.rm=T)/n_distinct(Rec_ID)) %>% 
  ungroup()

risk_plot <- risk_metric %>% 
  ggplot(aes(prop_highwinds))+
  geom_density(fill='#33638DFF')+
  labs(x='Propensity to Fish in High Winds (>7.5m/s)')
risk_plot
ggsave(here::here('fishing behavior','plots','risk_metric.png'),risk_plot,w=6,h=4)
```

### Cumulative Choice Entropy

For our 'exploration' metric, we follow O'Farrell et al. (2019) and calculate choice entropy. We use the 5km grid cells to define different 'locations'

```{r}
vms_all_grd <- read_rds(here::here('data','processed','vms','vms_all_w_grd.rds'))
vms_all_grd %<>%
  filter(TARGET_rev=='DCRB') %>% 
  # join crab season indicator
  left_join(recid_crabseason,by='Rec_ID')
vms_grd_only <- vms_all_grd %>% 
  as_tibble() %>% 
  select(VMS_RECNO,Rec_ID,UTCDATETIM,port_group_code,drvid,crab_season,GRID5KM_ID)

# To calculate entropy, we need a cumulative record of location choices in time, that is, a record of the grid cells visited by the vessel that accumulates over time

vms_cum_locs <- vms_grd_only %>% 
  group_by(drvid,crab_season) %>% 
  arrange(UTCDATETIM) %>% 
  mutate(locations=accumulate(GRID5KM_ID,c))

# test vessel/season to see if it worked
test <- vms_cum_locs %>% filter(crab_season=="2008-2009",drvid=="264147")

# this is the sequence of grid cells visited by this vessel up until VMS point 62 of the 2008-2009 season
test$locations[[62]]

# to calculate entropy, we use the equation from O'Farrell et al.
calc_entropy <- function(vec){
  locs <- unique(na.exclude(vec))
  freqs <- map_dbl(locs,~sum(vec==.x,na.rm=T)/length(vec))
  sum(freqs*log(freqs,base=2))*-1
}

test %<>% ungroup() %>% mutate(obs=row_number(),entropy=map_dbl(locations,calc_entropy))

test %>% ggplot(aes(obs,entropy))+geom_line()+labs(x="VMS observation number",y="Entropy",title="Example Choice Entropy Time Series")

# Apply to all VMS observations
vms_entropy <- vms_cum_locs %>% 
  ungroup() %>% 
  mutate(entropy=map_dbl(locations,calc_entropy))

# test another arbitrary vessel/season
vms_entropy %>% 
  filter(drvid=="580135",crab_season=='2009-2010') %>% 
  ggplot(aes(UTCDATETIM,entropy))+
  geom_line()+
  labs(x="VMS observation number",y="Entropy",title="Example Choice Entropy Time Series")

# as a final entropy 'score' for each year, take 90th quantile of max entropy
entropy_metric <- vms_entropy %>% 
  filter(drvid %in% vessel_list) %>% 
  group_by(drvid,crab_season) %>% 
  summarise(entropy_quant90=quantile(entropy,0.9))

entropy_plot <- entropy_metric %>% 
  ggplot(aes(entropy_quant90))+
  geom_density(fill='#33638DFF')+
  labs(x='Location Choice Entropy (90th percentile by Vessel/Season)')
entropy_plot
ggsave(here::here('fishing behavior','plots','entropy_metric.png'),entropy_plot,w=6,h=4)
```

### Home range

Calculate 95% convex polygons for each vessel/year

```{r}
## Coastline
coaststates <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('California','Oregon','Washington','Nevada')) %>% 
  st_transform(st_crs(vms_all_grd)) %>% 
  summarise()

test_hull <-  vms_all_grd %>% 
  filter(drvid=="580135",crab_season=='2009-2010') %>%
  select(VMS_RECNO,X_COORD,Y_COORD)
dsts <- nn2(test_hull[,c('X_COORD','Y_COORD')],k=100)$nn.dists %>% rowMeans()
test_hull %<>%
  mutate(meandist=dsts) %>% 
  # remove top 5% spatial outliers based on distance from other points
  mutate(rnk=percent_rank(dsts)) %>% 
  filter(rnk<0.95) %>% 
  ungroup() %>% 
  summarise() %>% 
  st_convex_hull()

test_hull %>% 
  ggplot()+
  geom_sf()

st_area(test_hull)

# apply to all vessels/seasons
find_hull95 <- function(df){
  dsts <- nn2(df %>% as_tibble() %>% select(X_COORD,Y_COORD),k=min(100,nrow(df)))$nn.dists %>% rowMeans()
  hull<-df %>% 
    mutate(meandist=dsts) %>% 
    # remove top 5% spatial outliers based on distance from other points
    mutate(rnk=percent_rank(dsts)) %>% 
    filter(rnk<0.95) %>% 
    ungroup() %>% 
    summarise() %>% 
    st_convex_hull()
  st_area(hull)*1e-6
}

homerange_calc <- vms_all_grd %>% 
  filter(drvid %in% vessel_list) %>% 
  select(drvid,crab_season,VMS_RECNO,X_COORD,Y_COORD) %>% 
  group_by(drvid,crab_season) %>% 
  nest() %>% 
  ungroup() %>% 
  mutate(nobs=map_dbl(data,nrow)) %>% 
  filter(nobs>10) %>% 
  mutate(homerange=purrr::map_dbl(data,find_hull95))

homerange_metric <- homerange_calc %>% 
  select(drvid,crab_season,homerange)

homerange_plot <- homerange_metric %>% 
  ggplot(aes(homerange))+
  geom_density(fill='#33638DFF')+
  labs(x='Home Range (95% Convex Hull Area, km^2)')
homerange_plot
ggsave(here::here('fishing behavior','plots','homerange_metric.png'),homerange_plot,w=6,h=4)
```

## Collect all Metrics

```{r}
all_metrics <- port_use_metrics %>% 
  full_join(revenue_metrics) %>% 
  full_join(trip_metrics) %>% 
  full_join(derby_metric) %>% 
  full_join(participation_metric) %>% 
  full_join(risk_metric) %>% 
  full_join(entropy_metric) %>% 
  full_join(homerange_metric) %>% 
  filter(drvid %in% vessel_list)
```
### Look for Missing Data

### Save
```{r}
write_rds(all_metrics,here::here('fishing behavior','fishing_behavior_metrics_vessel_season.rds'))
```



