---
title: "Calculating Fishing Behavioral Metrics"
author: "Owen Liu"
date: "2/9/2022"
output: html_document
---

## Purpose

We want to use VMS and fish ticket data to define behavioral metrics for the west coast Dungeness crab fishery. These can then be used for classifying behavior through clustering and perhaps looking at behavior change surrounding whale entanglement and the 2014-16 marine heat wave.

## Import Data

```{r}
options(dplyr.summarise.inform = FALSE)
options(tidyverse.quiet = TRUE)
library(tidyverse)
library(magrittr)
library(here)
library(lubridate)
library(sf)
library(rnaturalearth)

# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=12,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)

# DCRB fish tickets only, with day of season included
fishtix <- read_rds(here::here('data','processed','fish tickets','fish_tickets_w_dayofseason.rds'))

# Long form fish tickets, including entries for all species (used in participation metric calculations)
fishtix_long <- read_rds(here::here('data','processed','fish tickets','fish_tickets_long_allspecies.rds'))

# for both sets of fish tickets, we do not make calculations on pre-season tickets
fishtix %<>% filter(ticket_day_of_season>=0)
fishtix_long %<>% filter(ticket_day_of_season>=0)

# pacfin species identifiers
pacfin_spid <- read_csv(here('data','raw','pacfin_spid.csv'),col_types='icccc')

# crab season indicator
recid_crabseason <- fishtix %>% select(Rec_ID,crab_season) %>% distinct()

# Non-interpolated VMS data, filtered for DCRB trips
vms <- purrr::map_df(2009:2019, function(yr){
  read_rds(paste0(here::here('data','processed','matched','filtering',yr),"matched_filtered.rds"))
}) %>% 
  filter(TARGET_rev=='DCRB') %>% 
  # join crab season indicator
  left_join(recid_crabseason,by='Rec_ID')
# filter vms to only within-season tickets
vms %<>% filter(!is.na(crab_season))

# vessels in both VMS and fish ticket data
vms_fishtix_vessels <- intersect(unique(fishtix$drvid),unique(vms$drvid))

# consumer price index (for inflation adjustments)
#https://www.minneapolisfed.org/about-us/monetary-policy/inflation-calculator/consumer-price-index-1913-
cpi <- read_csv(here('data','raw','cpi_2021.csv'),col_types='idc')
# baseline is 1982-1984 $$
# add a conversion factor to 2010 $$
cpi %<>% 
  mutate(convert2010=1/(annual_average/218.1)) %>% 
  filter(year>2007) %>% 
  dplyr::select(year,convert2010)

```

## Metrics to Calculate

Here is a running list of the metrics to calculate. Some metrics have multiple types or options of calculations to perform

1.  Vessel size
2.  Port use: Number of ports visited per month, average ports visited per trip, Shannon diversity index of port use, total number of distinct ports visited per season.
3.  Revenue: Mean Dungeness revenue per trip, SD revenue per trip
4.  Landings: Mean Dungeness landings per trip, SD landings per trip
5.  Trip Length: Days per trip, distance per trip, SD days per trip, SD distance per trip
6.  Exploration: Mean cumulative choice entropy
7.  Mobility: Home range
8.  Risk-taking: Propensity to fish in high winds
9.  Derby fishing: Mean day-of-season of 90% annual catch reached
10.  Participation in other fisheries: percent of revenue from non-DCRB fisheries, Simpson diversity of index of landings by fishery

```{r}
# convert revenue in all revenue observations to 2010 dollars
fishtix %<>%
  # do the $$ conversion factor to 2010 dollars
  left_join(cpi,by='year') %>%
  mutate(across(contains('revenue'), ~.x*convert2010)) %>%
  select(Rec_ID:agency_code,date,crab_season,DCRB_revenue,DCRB_lbs,FINAL_LENGTH,ticket_day_of_season)


fishtix_long %<>%
  # do the $$ conversion factor to 2010 dollars
  left_join(cpi,by='year') %>% 
  mutate(across(c(spp_revenue,tot_revenue_spp), ~.x*convert2010)) %>% 
  dplyr::select(-convert2010)

vms %<>%
  mutate(year=year(date)) %>% 
  # do the $$ conversion factor to 2010 dollars
  left_join(cpi,by='year') %>% 
  mutate(across(c(other_rev,SABL_revenue,DCRB_revenue,LOBS_revenue,SPRW_revenue), ~.x*convert2010)) %>% 
  dplyr::select(-convert2010,-year)
```

We only want to investigate boats that participate in the Dungeness crab fishery. We can make a list of those, with some cutoff for revenue from the DCRB fishery.

```{r}
dcrb_revenue <- vms %>% 
  filter(drvid!="MISSING",drvid %in% vms_fishtix_vessels) %>%
  mutate(year=year(UTCDATETIM)) %>%
  left_join(recid_crabseason) %>% 
  filter(!is.na(crab_season)) %>% 
  distinct(drvid,Rec_ID,crab_season,DCRB_revenue) %>% 
  group_by(drvid,crab_season) %>% 
  summarise(dcrb_revenue=sum(DCRB_revenue,na.rm=T))

# what is the distribution of annual revenue, removing vessels with zero DCRB revenue
dcrb_revenue %>% 
  filter(dcrb_revenue!=0) %>% 
  ggplot(aes(dcrb_revenue))+
  geom_density(fill='#33638DFF')+
  labs(x='annual total DCRB revenue per vessel')

# what's the 5 percent quantile?
dcrb_revenue %>% filter(dcrb_revenue != 0) %>% pluck("dcrb_revenue") %>% quantile(0.05)

```

Five percent of the vessels made less than $5227 (in 2010 inflation-adjusted dollars) in at least one year of the data. We retain only those vessels that made more than this

```{r}
dcrb_vessels <- dcrb_revenue %>%
  # filter(dcrb_revenue != 0) %>% 
  group_by(crab_season) %>% 
  filter(dcrb_revenue>quantile(dcrb_revenue,0.05)) %>% 
  ungroup() %>% 
  distinct()
vessel_list <- unique(dcrb_vessels$drvid)
```

### Vessel Size

Use processed fish tickets to extract vessel sizes

```{r}
vessel_size_metric <-  fishtix %>%
  filter(drvid %in% vessel_list) %>% 
  distinct(drvid,crab_season,FINAL_LENGTH) %>% 
  rename(vessel_size=FINAL_LENGTH)

# plot vessel size metric
vessel_size_plot <- vessel_size_metric %>% 
  ggplot(aes(vessel_size))+
  geom_density(fill='#33638DFF')+
  labs(x='Vessel Length (m)')
vessel_size_plot
ggsave(here::here('fishing behavior','plots','vessel_size_metric.png'),vessel_size_plot,w=6,h=4)
write_rds(vessel_size_metric,here::here('fishing behavior','vessel_sizes.rds'))
```


### Port Use Metrics

Calculate number of ports visited per month, average ports visited per trip, and Shannon diversity index of port use. We calculate these from the VMS data

```{r}
# Average number of ports per fishing trip
ports_by_trip <- vms %>% 
  filter(drvid %in% vessel_list) %>% 
  group_by(Rec_ID,drvid,crab_season) %>% 
  summarise(n_ports=n_distinct(in_port,na.rm = T)) %>% 
  mutate(n_ports=max(n_ports,1)) %>% 
  ungroup() %>% 
  group_by(drvid,crab_season) %>% 
  summarise(n_ports_trip=mean(n_ports,na.rm=T)) %>% 
  ungroup()

# Average number of ports per month
ports_by_month <- vms %>% 
  filter(drvid %in% vessel_list) %>% 
  mutate(mth=month(date)) %>% 
  group_by(drvid,mth,crab_season) %>% 
  summarise(n_ports_mth=n_distinct(in_port,na.rm = T)) %>% 
  mutate(n_ports_mth=max(n_ports_mth,1)) %>% 
  ungroup() %>% 
  group_by(drvid,crab_season) %>% 
  summarise(ports_mean_mth=mean(n_ports_mth,na.rm=T)) %>% 
  ungroup()

# Shannon diversity index of port use by season
# For this metric we count each port visit as equal, and calculate the index on
# total diversity of port use in a given crab season
ports_diversity_shannon <- vms %>% 
  filter(!is.na(in_port)) %>% 
  filter(drvid %in% vessel_list) %>% 
  select(drvid,crab_season,in_port) %>% 
  group_by(drvid,crab_season,in_port) %>% 
  tally(name="n_visits") %>% 
  ungroup() %>% 
  group_by(drvid,crab_season) %>% 
  mutate(total_visits=sum(n_visits,na.rm=T),prop_visits=n_visits/total_visits) %>% 
  mutate(plnp=prop_visits*log(prop_visits)) %>% 
  summarise(port_shannon= -1*sum(plnp)) %>% 
  ungroup()

ports_diversity_simpson <- vms %>% 
  filter(!is.na(in_port)) %>% 
  filter(drvid %in% vessel_list) %>% 
  select(drvid,crab_season,in_port) %>% 
  group_by(drvid,crab_season,in_port) %>% 
  tally(name="n_visits") %>% 
  ungroup() %>%
  group_by(drvid,crab_season) %>% 
  mutate(total_visits=sum(n_visits,na.rm=T),prop_visits=n_visits/total_visits) %>%
  summarise(port_simpson=1/sum(prop_visits^2)) %>% 
  ungroup()

total_ports <- vms %>% 
  filter(!is.na(in_port)) %>% 
  filter(drvid %in% vessel_list) %>% 
  select(drvid,crab_season,in_port) %>% 
  group_by(drvid,crab_season) %>% 
  summarise(total_ports=n_distinct(in_port)) %>% 
  ungroup()

# Collected port use metrics
port_use_metrics <- ports_by_trip %>% 
  left_join(ports_by_month,by = c("drvid", "crab_season")) %>%
  # using inverse simpson index
  left_join(ports_diversity_shannon,by = c("drvid", "crab_season")) %>%
  left_join(ports_diversity_simpson,by = c("drvid", "crab_season")) %>% 
  left_join(total_ports,by = c("drvid", "crab_season"))

# plot diversity metric (Shannon and Simpson)
port_diversity_plot_shannon <- ports_diversity_shannon %>% 
  ggplot(aes(port_shannon))+
  geom_density(fill='#33638DFF')+
  labs(x='Diversity of Port Use (Shannon)')
port_diversity_plot_shannon
ggsave(here::here('fishing behavior','plots','port_diversity_metric_shannon.png'),port_diversity_plot_shannon,w=6,h=4)

port_diversity_plot_simp <- ports_diversity_simpson %>% 
  ggplot(aes(port_simpson))+
  geom_density(fill='#33638DFF')+
  labs(x='Diversity of Port Use (Simpson)')
port_diversity_plot_simp
ggsave(here::here('fishing behavior','plots','port_diversity_metric_simpson.png'),port_diversity_plot_simp,w=6,h=4)

port_diversity_comparison <- port_use_metrics %>% 
  ggplot(aes(port_shannon,port_simpson))+
  geom_point()+
  labs(x="Shannon",y="Simpson")
port_diversity_comparison
ggsave(here::here('fishing behavior','plots','port_diversity_metric_comparison.png'),port_diversity_comparison,w=6,h=4)

# for now use Shannon
port_use_metrics %<>% dplyr::select(-port_simpson)
```

### Revenue Metrics

Calculate mean and SD Dungeness revenue per trip. We can calculate this straight from the fish tickets

```{r}
revenue_metrics <- fishtix %>%
  filter(drvid %in% vessel_list) %>% 
  distinct(drvid,crab_season,Rec_ID,DCRB_revenue) %>% 
  group_by(drvid,crab_season) %>% 
  summarise(revenue_mean=mean(DCRB_revenue,na.rm=T),revenue_sd=sd(DCRB_revenue,na.rm=T)) %>% 
  ungroup()
revenue_mean_plot <- revenue_metrics %>% 
  ggplot(aes(revenue_mean))+
  geom_density(fill='#33638DFF')+
  labs(x='mean revenue')
revenue_sd_plot <- revenue_metrics %>% 
  ggplot(aes(revenue_sd))+
  geom_density(fill='#33638DFF')+
  labs(x='SD revenue')
revenue_mean_plot
ggsave(here::here('fishing behavior','plots','revenue_mean_metric.png'),revenue_mean_plot,w=6,h=4)
revenue_sd_plot
ggsave(here::here('fishing behavior','plots','revenue_sd_metric.png'),revenue_sd_plot,w=6,h=4)
```

### Landings Metrics

Calculate mean and SD Dungeness landings per trip. We can calculate this straight from the fish tickets

```{r}
landings_metrics <- fishtix %>%
  filter(drvid %in% vessel_list) %>% 
  distinct(drvid,crab_season,Rec_ID,DCRB_lbs) %>% 
  group_by(drvid,crab_season) %>% 
  summarise(landings_mean=mean(DCRB_lbs,na.rm=T),landings_sd=sd(DCRB_lbs,na.rm=T)) %>% 
  ungroup()
landings_mean_plot <- landings_metrics %>% 
  ggplot(aes(landings_mean))+
  geom_density(fill='#33638DFF')+
  labs(x='mean landings')
landings_sd_plot <- landings_metrics %>% 
  ggplot(aes(landings_sd))+
  geom_density(fill='#33638DFF')+
  labs(x='SD landings')
landings_mean_plot
ggsave(here::here('fishing behavior','plots','landings_mean_metric.png'),revenue_mean_plot,w=6,h=4)
landings_sd_plot
ggsave(here::here('fishing behavior','plots','landings_sd_metric.png'),revenue_sd_plot,w=6,h=4)
```

### Trip Length Metrics

Using the VMS data, we can calculate days per trip, distance traveled per trip, and the standard deviations of these.

```{r}
trip_metrics <- vms %>% 
  filter(drvid %in% vessel_list) %>% 
  mutate(segment_dist=segment_dur*avg_speed_recalc) %>% 
  group_by(Rec_ID,drvid,crab_season,trip_dur) %>%
  # trip distance IN KILOMETERS
  summarise(trip_dist=sum(segment_dist,na.rm=T)/1000) %>% 
  ungroup() %>% 
  group_by(drvid,crab_season) %>% 
  summarise(trip_dist_mean=mean(trip_dist,na.rm = T),trip_dur_mean=mean(trip_dur,na.rm = T),trip_dist_sd=sd(trip_dist,na.rm=T),trip_dur_sd=sd(trip_dur,na.rm=T)) %>%
  ungroup()
trip_plot <- trip_metrics %>% 
  ggplot(aes(trip_dist_mean))+
  geom_density(fill='#33638DFF')+
  labs(x='trip distance (km)')
trip_plot
ggsave(here::here('fishing behavior','plots','trips_metric.png'),trip_plot,w=6,h=4)

# relationship between trip distance and trip duration
# trip_dist_duration_plot <- trip_metrics %>% 
#   ggplot(aes(trip_dist_mean,trip_dur_mean))+
#   geom_point(size=0.5)+
#   labs(x="trip distance (km)",y="trip duration (days)")
# trip_dist_duration_plot
```

### Derby Metric

For each fishing vessel in each year, calculate the day-of-season at which the boat finished catching 90% of its cumulative catch for the year. This metric should give a sense for how much a given vessel participates in the early-season derby. We can calculate this using the fish ticket data.

```{r}
derby_metric <- fishtix %>% 
  filter(drvid %in% vessel_list) %>% 
  group_by(drvid,crab_season) %>% 
  # find total and 90% revenue for each vessel in each crab season
  mutate(total_dcrb_revenue=sum(DCRB_revenue,na.rm=T),quant90_dcrb_revenue=0.9*total_dcrb_revenue) %>% 
  arrange(date) %>% 
  # cumulative sum of revenue for each vessel across the year
  mutate(dcrb_cum_revenue=cumsum(DCRB_revenue)) %>% 
  mutate(is_quant90=dcrb_cum_revenue>=quant90_dcrb_revenue) %>% 
  # find the ticket that crossed the 90% threshold and mark its date
  filter(cumany(is_quant90)) %>% 
  slice(1) %>% 
  select(drvid,crab_season,ticket_day_of_season) %>% 
  rename(quant90_day=ticket_day_of_season)

derby_plot <- derby_metric %>% 
  ggplot(aes(quant90_day))+
  geom_density(fill='#33638DFF')+
  labs(x='day of season reaching 90% cumulative catch')
derby_plot
ggsave(here::here('fishing behavior','plots','derby_metric.png'),derby_plot,w=6,h=4)
```

### Participation Metrics

Calculate what percent of a vessel's revenue comes from non-DCRB commercial fisheries. We do this with the unfiltered fish ticket data, counting the non-DCRB revenue across all trips.

Additionally, calculate revenue diversity of landings (inverse Simpson diversity on revenue by species).

Seasons over which to calculate other revenue are defined in the same manner as other metrics- that is, revenue is measured over the course of the crab seasons (as opposed to, e.g., calendar years).

```{r}
# Grouping species before revenue diversity and participation metrics

# check species prevalence
all_fishtix_spp <- fishtix_long %>%
  filter(drvid %in% vessel_list) %>%
  distinct(PACFIN_SPECIES_CODE) %>% 
  # join the PACFIN-recognized species groups, and use this to organize
  left_join(pacfin_spid,by=c('PACFIN_SPECIES_CODE'='SPID')) %>% 
  dplyr::select(PACFIN_SPECIES_CODE,`Common Name`, Grp) %>% 
  rename(commonname=`Common Name`,spp_grp=Grp)

other_fisheries_revenues_spp <- fishtix_long %>%
  filter(drvid %in% vessel_list) %>%
  distinct(Rec_ID,drvid,crab_season,PACFIN_SPECIES_CODE,tot_revenue_spp) %>%
  left_join(all_fishtix_spp) %>% 
  group_by(drvid,crab_season,spp_grp,PACFIN_SPECIES_CODE) %>%  
  summarise(tot_rev=sum(tot_revenue_spp,na.rm=T)) %>% 
  ungroup() %>% 
  complete(nesting(drvid,crab_season),PACFIN_SPECIES_CODE,spp_grp,fill=list(tot_rev=0)) %>% 
  ungroup() %>% 
  group_by(drvid,crab_season) %>% 
  mutate(prop_rev=tot_rev/sum(tot_rev)) %>%
  ungroup() %>% 
  group_by(PACFIN_SPECIES_CODE) %>% 
  summarise(mean_prop_rev=mean(prop_rev,na.rm=T))

other_fisheries_revenues_grp <- fishtix_long %>%
  filter(drvid %in% vessel_list) %>%
  distinct(Rec_ID,drvid,crab_season,PACFIN_SPECIES_CODE,tot_revenue_spp) %>%
  left_join(all_fishtix_spp) %>% 
  group_by(drvid,crab_season,spp_grp) %>%  
  summarise(tot_rev=sum(tot_revenue_spp,na.rm=T)) %>% 
  ungroup() %>% 
  complete(nesting(drvid,crab_season),spp_grp,fill=list(tot_rev=0)) %>% 
  ungroup() %>% 
  group_by(drvid,crab_season) %>% 
  mutate(prop_rev=tot_rev/sum(tot_rev)) %>%
  ungroup() %>% 
  group_by(spp_grp) %>% 
  summarise(mean_prop_rev=mean(prop_rev,na.rm=T))

# Across all crab seasons and vessels, this is what the breakdown of mean percent revenue by species looks like
other_fisheries_revenues_spp %>% 
  arrange(desc(mean_prop_rev)) %>% 
  slice(1:20) %>% 
  ggplot(aes(x=fct_reorder(PACFIN_SPECIES_CODE,-mean_prop_rev),y=mean_prop_rev*100))+
  geom_col()+
  labs(x="Species Code",y="Mean Percent of Revenue",title="Top 20 Species by Proportion of Revenue")

other_fisheries_revenues_grp %>% 
  arrange(desc(mean_prop_rev)) %>% 
  ggplot(aes(x=fct_reorder(spp_grp,-mean_prop_rev),y=mean_prop_rev*100))+
  geom_col()+
  labs(x="Species Group Code",y="Mean Percent of Revenue",title="Species Groups by Proportion of Revenue")
```

```{r}
# organize long-form fish tickets by Rec_ID, and DCRB vs. all other revenue
fishtix_dcrb_other <- fishtix_long %>%
  filter(drvid %in% vessel_list) %>%
  left_join(all_fishtix_spp) %>% 
  mutate(is_DCRB=PACFIN_SPECIES_CODE=="DCRB") %>% 
  group_by(Rec_ID,TARGET_rev,drvid,crab_season) %>% 
  summarise(DCRB_revenue=sum(spp_revenue[is_DCRB]),other_revenue=sum(spp_revenue[!is_DCRB])) %>% 
  ungroup() %>% 
  mutate(DCRB_revenue=coalesce(DCRB_revenue,0),other_revenue=coalesce(other_revenue,0))
  
participation_metric <- fishtix_dcrb_other %>% 
  filter(drvid %in% vessel_list) %>% 
  group_by(drvid,crab_season) %>% 
  summarise(total_dcrb_revenue=sum(DCRB_revenue,na.rm=T),
            total_other_revenue=sum(other_revenue),
            total_other_tix=n_distinct(Rec_ID[which(TARGET_rev != "DCRB")]),
            total_dcrb_tix=n_distinct(Rec_ID[which(TARGET_rev == "DCRB")])) %>% 
  mutate(total_other_revenue=coalesce(total_other_revenue,0)) %>% 
  mutate(prop_other_revenue=total_other_revenue/(total_other_revenue+total_dcrb_revenue),
         prop_other_tix=total_other_tix/(total_other_tix+total_dcrb_tix)) %>% 
  ungroup() %>% 
  rename(crab_season=crab_season) %>% 
  distinct(drvid,crab_season,prop_other_revenue,prop_other_tix)

# Update 02/2022- calculate diversity on species groups, not just species
fishtix_spp_grps <- fishtix_long %>%
  filter(drvid %in% vessel_list) %>%
  left_join(all_fishtix_spp) %>%
  group_by(drvid,crab_season,spp_grp) %>% 
  summarise(tot_rev_grp=sum(tot_revenue_spp)) %>% 
  ungroup()

participation_diversity <- fishtix_spp_grps %>%
  group_by(drvid,crab_season) %>% 
  mutate(prop_rev=tot_rev_grp/sum(tot_rev_grp)) %>% 
  summarise(revenue_diversity=1/sum(prop_rev^2)) %>% 
  ungroup()

participation_rev_plot <- participation_metric %>% 
  ggplot(aes(prop_other_revenue))+
  geom_density(fill='#33638DFF')+
  labs(x='Proportion of Revenue from Other Fisheries')
participation_tix_plot <- participation_metric %>% 
  ggplot(aes(prop_other_tix))+
  geom_density(fill='#33638DFF')+
  labs(x='Proportion of Tickets Targeting Other Fisheries')
participation_div_plot <- participation_diversity %>% 
  ggplot(aes(revenue_diversity))+
  geom_density(fill='#33638DFF')+
  labs(x='Inverse Simpson Diversity Index of Landings by Fishery')

participation_rev_plot
participation_tix_plot
participation_div_plot
ggsave(here::here('fishing behavior','plots','participation_rev_metric.png'),participation_rev_plot,w=6,h=4)
ggsave(here::here('fishing behavior','plots','participation_tix_metric.png'),participation_tix_plot,w=6,h=4)
ggsave(here::here('fishing behavior','plots','participation_diversity_metric.png'),participation_div_plot,w=6,h=4)

participation_metrics <- participation_metric %>% 
  full_join(participation_diversity)

# comparing participation metrics
participation_metrics %>%
  ggplot(aes(prop_other_tix,revenue_diversity))+
  geom_point()+
  geom_smooth()+
  labs(x='Proportion of Tickets Targeting Other Fisheries',y='Inverse Simpson Diversity Index of Landings')
# dome-shaped relationship makes sense
```

### Risk Metric

Following O'Farrell et al., we extract wind speeds from NOAA's ERDAPP server and use them to measure a fisher's propensity to go out in stormy weather. Following Pfeiffer and Gratz (2016), who also studied west coast fishing, we calculate a max (95th percentile) wind speed per trip, and then calculate what percent of a vessel's trips in a given crab season were pursued in winds >7.5m/s.

Download all winds for the study area.
DEFAULT: not run because downloading takes a long time

02/18/2020: https://www.ncei.noaa.gov/thredds/ncss/uv/clm_agg/Latest_Climatology_of_Ocean_Winds_best.ncd/dataset.html
https://rdrr.io/github/bocinsky/thredds/f/README.md


```{r, eval=T}
library(rerddap)
# Range of Lat/Lon for entire dataset
range(vms$LATITUDE)
windslat <- c(34,49)
# 360+range(vms$LONGITUDE) # Longitudes in the winds dataset are in degrees E
windslon <- c(233,245)-360

# sample day of winds data, to obtain lat/lons
wind_sample <- griddap("nceiPH53sstd1day", # daily Pathfinder winds dataset
                      time=c("2009-01-01","2009-01-02"),
                      latitude=windslat,
                      longitude=windslon,
                      fields=c("wind_speed")) %>%  # scalar wind speed
      pluck("data")

library(RANN)
vms_latlons <- vms %>% select(LATITUDE,LONGITUDE)

winds_latlons <- wind_sample %>% select(lat,lon) %>% distinct()

# Match VMS lat/lon to ERDDAP lat/lon
erddap_vms_match <- RANN::nn2(winds_latlons,vms_latlons,k=1)$nn.idx

# add appropriate erddap lat/lons to vms lat/lons data
vms_latlons %<>% mutate(matchlat=wind_sample$lat[erddap_vms_match],matchlon=wind_sample$lon[erddap_vms_match])

# only pull the winds data for the lat/lons that we need
needed_latlons <- vms_latlons %>% select(matchlat,matchlon) %>% distinct()
# another to try
# nceiPH53sstd1day
```

```{r,eval=T}
# Have to use some if statements here because of weird breaks in the two datasets
winds_all <- purrr::map_df(2009:2019, function(yr) {
  times <- c(paste0(yr,"-01-01"),paste0(yr,"-12-31"))
  winds <- griddap("nceiPH53sstn1day", # daily blended winds dataset
                    time=times,
                    latitude=windslat,
                    longitude=windslon,
                    fields=c("wind_speed")) %>%  # scalar wind speed
    pluck("data") %>% 
    filter(lat %in% unique(needed_latlons$matchlat)&lon %in% unique(needed_latlons$matchlon))
  # if(yr < 2010){
  #   times <- c(paste0(yr,"-01-01"),paste0(yr,"-12-31"))
  #   winds <- griddap("ncdcOwDly", # daily blended winds dataset
  #                     time=times,
  #                     latitude=windslat,
  #                     longitude=windslon,
  #                     fields=c("w")) %>%  # scalar wind speed
  #     pluck("data")
  # }
  # if(yr > 2008){
  #   times <- c(paste0(yr,"-01-01"),paste0(yr,"-12-31"))
  #   winds <- griddap("nceiPH53sstn1day", # daily blended winds dataset
  #                     time=times,
  #                     latitude=windslat,
  #                     longitude=windslon,
  #                     fields=c("wind_speed")) %>%  # scalar wind speed
  #     pluck("data")
  #     
  # } 
  
  # if(yr == 2011){
  #   times1 <- c(paste0(yr,"-01-01"),paste0(yr,"-09-30"))
  #   winds1 <- griddap("ncdcOwDly", # daily blended winds dataset
  #                     time=times1,
  #                     latitude=windslat,
  #                     longitude=windslon,
  #                     fields=c("w")) %>% # scalar wind speed
  #     pluck("data")
  #   
  #   times2 <- c(paste0(yr,"-10-01"),paste0(yr,"-12-31"))
  #   winds2 <- griddap("ncdcOwDlyP", # daily blended winds dataset
  #                     time=times2,
  #                     latitude=windslat,
  #                     longitude=windslon,
  #                     fields=c("w") )%>%    # scalar wind speed
  #     pluck("data")
  #   winds <- bind_rows(winds1,winds2)
  # } 
  # if(yr>2011&yr<2018){
  #   times <- c(paste0(yr,"-01-01"),paste0(yr,"-12-31"))
  #   winds <- griddap("ncdcOwDlyP", # daily blended winds dataset
  #                     time=times,
  #                     latitude=windslat,
  #                     longitude=windslon,
  #                     fields=c("w")) %>% # scalar wind speed
  #     pluck("data") 
  # }
  # if(yr==2018){
  #   times <- c(paste0(yr,"-01-01"),paste0(yr,"-10-09"))
  #   winds <- griddap("ncdcOwDlyP", # daily blended winds dataset
  #                     time=times,
  #                     latitude=windslat,
  #                     longitude=windslon,
  #                     fields=c("w")) %>% # scalar wind speed
  #     pluck("data") 
  # }  
  out <- winds %>% as_tibble() %>% mutate(utcdate=as_date(time))
  out
})
# write_rds(winds_all,here::here('data','raw','blended winds','erddap_raw_daily_winds.rds'))
```

Match winds to VMS pings

```{r}
# if not loaded
# winds_all <- read_rds(here::here('data','raw','blended winds','erddap_raw_daily_winds.rds'))

# library(RANN)
# vms_latlons <- vms %>% select(LATITUDE,LONGITUDE) %>%
#   # longitude has to be in degrees E
#   mutate(LONGITUDE=360+LONGITUDE)
# 
# winds_latlons <- winds_all %>% select(lat,lon) %>% distinct()
# 
# # Match VMS lat/lon to ERDDAP lat/lon
# erddap_vms_match <- RANN::nn2(winds_latlons,vms_latlons,k=1)$nn.idx
# 
# # add appropriate erddap lat/lons to vms lat/lons data
# vms_latlons %<>% mutate(matchlat=winds_all$lat[erddap_vms_match],matchlon=winds_all$lon[erddap_vms_match]) %>% mutate(LONGITUDE=LONGITUDE-360)
# 

# Finally, join winds data to VMS data
vms_winds <- vms %>%
  select(drvid,crab_season,Rec_ID,UTCDATETIM) %>% 
  # join the lat/lon matching key
  bind_cols(select(vms_latlons,matchlat,matchlon)) %>% 
  #convert UTC times to dates
  mutate(utcdate=as_date(UTCDATETIM)) %>% 
  # join winds data by date and matching latitude/longitude
  left_join(select(winds_all,-time),by=c('utcdate'='utcdate','matchlat'='lat','matchlon'='lon'))
```

Calculate an extreme wind value for each trip (95th percentile of wind speed per trip).

```{r}
winds_per_trip <- vms_winds %>% 
  filter(drvid %in% vessel_list) %>% 
  group_by(drvid,crab_season,Rec_ID) %>% 
  summarise(wind_quant95=quantile(wind_speed,0.95,na.rm = T)) %>% 
  ungroup()
winds_per_trip %>% 
  ggplot(aes(wind_quant95))+
  geom_density(fill='#33638DFF')+
  labs(x='95th percentile of wind speed (m/s)')

risk_metric <- winds_per_trip %>% 
  # define whether a trip is 'high wind'
  mutate(is_high_winds=wind_quant95>=7.5) %>% 
  # count proportion of high wind trips by vessel/season
  group_by(drvid,crab_season) %>% 
  summarise(prop_highwinds=sum(is_high_winds,na.rm=T)/n_distinct(Rec_ID)) %>% 
  ungroup()

risk_plot <- risk_metric %>% 
  ggplot(aes(prop_highwinds))+
  geom_density(fill='#33638DFF')+
  labs(x='Propensity to Fish in High Winds (>7.5m/s)')
risk_plot

write_rds(risk_metric,here::here('fishing behavior','risk_metric.rds'))

ggsave(here::here('fishing behavior','plots','risk_metric.png'),risk_plot,w=6,h=4)
```

### Cumulative Choice Entropy

For our 'exploration' metric, we follow O'Farrell et al. (2019) and calculate choice entropy. We use the 5km grid cells to define different 'locations'

```{r}
vms_all_grd <- read_rds(here::here('data','processed','vms','vms_all_w_grd.rds'))
vms_all_grd %<>%
  filter(TARGET_rev=='DCRB') %>% 
  # join crab season indicator
  left_join(recid_crabseason,by='Rec_ID')
vms_grd_only <- vms_all_grd %>% 
  as_tibble() %>% 
  select(VMS_RECNO,Rec_ID,UTCDATETIM,port_group_code,drvid,crab_season,GRID5KM_ID) %>% 
  filter(!is.na(GRID5KM_ID),!is.na(crab_season))

# To calculate entropy, we need a cumulative record of location choices in time, that is, a record of the grid cells visited by the vessel that accumulates over time

vms_cum_locs <- vms_grd_only %>% 
  group_by(drvid,crab_season) %>% 
  mutate(mult_locs = n_distinct(GRID5KM_ID)>1) %>% 
  filter(mult_locs) %>% 
  arrange(UTCDATETIM) %>%
  mutate(locations=purrr::accumulate(GRID5KM_ID,c))

# test vessel/season to see if it worked
test <- vms_cum_locs %>% filter(crab_season=="2008-2009",drvid=="264147")

# this is the sequence of grid cells visited by this vessel up until VMS point 62 of the 2008-2009 season
test$locations[[62]]

# to calculate entropy, we use the equation from O'Farrell et al.
calc_entropy <- function(vec){
  locs <- unique(na.exclude(vec))
  freqs <- map_dbl(locs,~sum(vec==.x,na.rm=T)/length(vec))
  sum(freqs*log(freqs,base=2))*-1
}

test %<>% ungroup() %>% mutate(obs=row_number(),entropy=map_dbl(locations,calc_entropy))

test %>% ggplot(aes(obs,entropy))+geom_line()+labs(x="VMS observation number",y="Entropy",title="Example Choice Entropy Time Series")

# Apply to all VMS observations
vms_entropy <- vms_cum_locs %>% 
  ungroup() %>% 
  mutate(entropy=map_dbl(locations,calc_entropy))

# test another arbitrary vessel/season
vms_entropy %>% 
  filter(drvid=="580135",crab_season=='2009-2010') %>% 
  ggplot(aes(UTCDATETIM,entropy))+
  geom_line()+
  labs(x="VMS observation number",y="Entropy",title="Example Choice Entropy Time Series")

# as a final entropy 'score' for each year, take 90th quantile of max entropy
entropy_metric <- vms_entropy %>% 
  filter(drvid %in% vessel_list) %>% 
  group_by(drvid,crab_season) %>% 
  summarise(entropy_quant90=quantile(entropy,0.9))

entropy_plot <- entropy_metric %>% 
  ggplot(aes(entropy_quant90))+
  geom_density(fill='#33638DFF')+
  labs(x='Location Choice Entropy (90th percentile by Vessel/Season)')
entropy_plot
write_rds(entropy_metric,here::here('fishing behavior','entropy_metric.rds'))
ggsave(here::here('fishing behavior','plots','entropy_metric.png'),entropy_plot,w=6,h=4)
```

### Home range

Calculate 95% convex polygons for each vessel/year

```{r}
## Coastline
coaststates <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('California','Oregon','Washington','Nevada')) %>% 
  st_transform(st_crs(vms_all_grd)) %>% 
  summarise()

test_hull <-  vms_all_grd %>% 
  filter(drvid=="580135",crab_season=='2009-2010') %>%
  select(VMS_RECNO,X_COORD,Y_COORD)
dsts <- nn2(test_hull[,c('X_COORD','Y_COORD')],k=100)$nn.dists %>% rowMeans()
test_hull %<>%
  mutate(meandist=dsts) %>% 
  # remove top 5% spatial outliers based on distance from other points
  mutate(rnk=percent_rank(dsts)) %>% 
  filter(rnk<0.95) %>% 
  ungroup() %>% 
  summarise() %>% 
  st_convex_hull %>% 
  #cut out land
  st_difference(coaststates)

bbox <- st_bbox(test_hull)
test_hull %>% 
  ggplot()+
  geom_sf()+
  # geom_sf(data=coaststates,fill='tan')+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))

st_area(test_hull)

# apply to all vessels/seasons
find_hull95 <- function(df){
  dsts <- nn2(df %>% as_tibble() %>% select(X_COORD,Y_COORD),k=min(100,nrow(df)))$nn.dists %>% rowMeans()
  hull<-df %>% 
    mutate(meandist=dsts) %>% 
    # remove top 5% spatial outliers based on distance from other points
    mutate(rnk=percent_rank(dsts)) %>% 
    filter(rnk<0.95) %>% 
    ungroup() %>% 
    summarise() %>% 
    st_convex_hull()%>% 
    st_difference(coaststates)
  out <- st_area(hull)*1e-6
  if(length(out)==0) {out <- NA}
  as.numeric(out)
}

homerange_calc <- vms_all_grd %>% 
  filter(drvid %in% vessel_list) %>% 
  select(drvid,crab_season,VMS_RECNO,X_COORD,Y_COORD) %>% 
  group_by(drvid,crab_season) %>% 
  nest() %>% 
  ungroup() %>% 
  mutate(nobs=map_dbl(data,nrow)) %>% 
  filter(nobs>10) %>% 
  mutate(homerange=purrr::map_dbl(data,find_hull95))

homerange_metric <- homerange_calc %>% 
  select(drvid,crab_season,homerange)

homerange_plot <- homerange_metric %>% 
  ggplot(aes(homerange))+
  geom_density(fill='#33638DFF')+
  labs(x='Home Range (95% Convex Hull Area, km^2)')

homerange_plot

write_rds(homerange_metric,here::here('fishing behavior','homerange_metric.rds'))
ggsave(here::here('fishing behavior','plots','homerange_metric.png'),homerange_plot,w=6,h=4)
```

## Collect all Metrics

```{r}
all_metrics <- port_use_metrics %>% 
  full_join(revenue_metrics, by = c("drvid", "crab_season")) %>% 
  full_join(landings_metrics, by = c("drvid", "crab_season")) %>% 
  full_join(trip_metrics, by = c("drvid", "crab_season")) %>% 
  full_join(derby_metric, by = c("drvid", "crab_season")) %>% 
  full_join(participation_metrics, by = c("drvid", "crab_season")) %>% 
  full_join(risk_metric, by = c("drvid", "crab_season")) %>% 
  full_join(entropy_metric, by = c("drvid", "crab_season")) %>% 
  full_join(homerange_metric, by = c("drvid", "crab_season")) %>% 
  full_join(vessel_size_metric, by = c("drvid", "crab_season")) %>% 
  filter(drvid %in% vessel_list) %>% 
  filter(!is.na(crab_season)) %>% 
  distinct()
```
### Look for Missing Data

### Save
```{r}
write_rds(all_metrics,here::here('fishing behavior','fishing_behavior_metrics_vessel_season_final.rds'))
```
