---
title: "Fisher Behavioral Strategies"
author: "Owen Liu"
date: Last Run "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F,warning=F,message=F)
library(tidyverse)
library(magrittr)
library(here)
library(NbClust)
library(viridis)

# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=10,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
```

## Purpose

The goal of this analysis is to perform a clustering analysis on a suite of behavioral metrics describing the 2009-2019 west coast Dungeness crab fishing fleet. The metrics have been compiled based on fish ticket and VMS data, and include information on port use, revenue, and trip lengths, as well as more advanced metrics from analysis of VMS data, like distance traveled, exploratory behavior and propensity to fish in high winds.

All metrics (to the extent data allow) were calculated for each vessel and crab season. The unit of observation in clustering is one vessel's crab season, e.g. Vessel A in crab season 2012-2013.

Our guiding questions are:

1. **Can the behavior of the west coast Dungeness crab fishing fleet be classified into distinct behavioral types?**
2. **Which behavioral variables are particularly important in defining these behavioral types?**
3. **Have individual vessels moved between behavioral types across years?**
4. **How do fishing behavioral strategies relate to revenue and landings from fishing?**

## Import Data

```{r}
dat <- read_rds(here::here('fishing behavior','fishing_behavior_metrics_vessel_season.rds'))
glimpse(dat)
```

More specifically, the metrics are as follows for each vessel (`drvid`) in each `crab_season`:

1.  Port use: 
  *  `n_ports_trip` Average ports visited per trip
  *  `ports_mean_mth` Number of ports visited per month 
  *  `port_shannon` Shannon diversity index of port use
  *  `total_ports` Total number of ports visited across the entire season
2.  Revenue: 
  *  `revenue_mean` Mean Dungeness revenue per trip
  *  `revenue_sd` SD revenue per trip
3.  Landings:
  *  `landings_mean` Mean Dungeness landings (lbs) per trip
  *  `landings_sd` SD landings per trip
4.  Trip Length: 
  *  `trip_dist_mean` Mean distance per fishing trip, 
  *  `trip_dur_mean` Mean number of days per fishing trip
  *  `trip_dist_sd` SD distance per trip
  *  `trip_dur_sd` SD days per trip
5.  Derby fishing:
  *  `quant90_day` Day-of-season on which fisher reach 90% of eventual annual catch
6.  Participation in other fisheries:
  *  `prop_other_revenue` Proportion of revenue from non-DCRB fisheries
  *  `prop_other_tix` Proportion of all fish tickets form non-DCRB fisheries
  *  `landings_diversity` Inverse Simpson diversity index of landings by fishery
7.  Risk-taking:
  *  `prop_highwinds` Propensity to fish in high winds. Proportion of trips pursued where the 95% quantile of wind speed was greater than 7.5 m/s
8.  Exploration: 
  *  `entropy_quant90` Cumulative choice entropy, measuring how likely a vessel is to fish in new versus past locations
9.  Mobility:
  * `homerange` Home range defined as the area of the convex hull surrounding all of a vessel's VMS pings during the season, excluding the top 5% spatial outliers


Because we are interested here in a post-hoc analysis of how fishing behaviors correlate with revenue and landings, we remove revenue and landings metrics from the clustering analysis and instead investigate their relationships with the derived behavioral groups later.

## Check Collinearity

Before clustering, we take a look at all the variables to search for extreme collinearity, which would suggest that we should trim our variables down.

```{r}
vars.corrs <- dat %>% select(-(1:2),-revenue_mean,-revenue_sd,-landings_mean,-landings_sd) %>% 
  cor(use="complete.obs") %>% 
  as_tibble(rownames="var1") %>% 
  pivot_longer(-var1,names_to = "var2",values_to = "corr")
vars.corrs %>% filter(corr !=1) %>% distinct(corr,.keep_all = T) %>% arrange(desc(corr)) %>% slice(1:10) %>% knitr::kable()

vars.corrs %>% 
  ggplot(aes(var1,var2,fill=corr))+
  geom_tile()+
  coord_equal()+
  scale_fill_gradient2()+
  labs(x="",y="",title="Variable Pearson Correlation Matrix",fill="")+
  theme(axis.text.x = element_text(angle=90,vjust=0.5,hjust=1))
```

There are some correlated variables, but few are greater than 0.7. Fishing trip distance and duration are correlated, as well as the different measures of port use (diversity, total ports, and ports per trip and per month). Additionally, the trip distance metrics are correlated with our measure of location choice entropy.

We keep all variables for now.

## Re-scale variables

For clustering analyses, we need to re-scale variables to a 0 to 1 scale to avoid artifacts caused simply by scale differences between variables.

Re-scaling is done by dividing each metric by its maximum value.

```{r}
dat_scaled <- dat %>% 
  select(-revenue_mean,-revenue_sd,-landings_mean,-landings_sd) %>% 
  mutate_at(-(1:2),~./max(.,na.rm=T)) %>% 
  # remove any observations with missing variables
  drop_na()
```

## Do Clustering

We then perform a clustering of the data, using Ward aggregation that attempts to minimize total within-cluster variance. However, instead of choosing the number of resulting classes rather arbitrarily from the results, we utilize the package `NbClust` that calculates 22 clustering indices defined in the literature, and then recommends the optimal number of clusters via majority vote.

```{r,results=FALSE,include=FALSE}
dat_clust <- dat_scaled %>% select(-(1:2)) %>% NbClust(method="ward.D2",min.nc = 3,max.nc=10,index='all')
```

After clustering, there was a range of number of classes suggested by the clustering indices, but the **majority vote was 3 classes**. For now, we utilize those 3 classes to explore the clusters.

## Explore Clustering

```{r,include=FALSE,eval=FALSE}
dat_clust$All.index %>% as.tibble(rownames="n_clust")
dat_clust$All.CriticalValues
dat_clust$Best.partition %>% length()
```

We apply the majority-rule partitions from clustering to the dataset.

```{r}
library(factoextra)
dat_clustered <- dat_scaled %>% na.omit() %>% 
  mutate(partition=dat_clust$Best.partition) %>% 
  as_tibble()
# write_rds(dat_clustered,here::here('fishing behavior','data_scaled_4clusters.rds'))
```

We can visualize the results of clustering through a clustering dendrogram. Below we can see all of the data (unique season-vessel combinations), colored by the cluster into which they were grouped.

```{r}
library(ggdendro)
dat_nolabs <- dat_scaled %>% na.omit() %>% select(-(1:2))

# calc distances
dat.d <- dat_scaled %>% na.omit() %>% select(-(1:2)) %>% get_dist()

# calc clustering
dat.ward <- dat.d %>% hclust(method="ward.D2")

# ggdendro::ggdendrogram(dat.ward,leaf_labels = F,labels = FALSE)

# Rectangular lines dendrogram
dend <- as.dendrogram(dat.ward) %>% dendro_data(type = "rectangle")

# figure out colors for segments
# partitions <- tibble(obs=1:3261,partition=dat_clust$Best.partition)
# head(dend$segments)
# test <- dend$segment %>% as_tibble() %>%
#   mutate(obs=paste(label) %>% as.numeric()) %>% 
#   left_join(partitions,by=c('obs'))
test <- dat_clustered %>% group_by(partition) %>% summarise(num_obs=n()) %>% ungroup() %>% mutate(cumobs=cumsum(num_obs))

# segs <- dend$segments %>% as_tibble() %>% 
#   mutate(partition=case_when(
#     round(xend)<= 272 ~ 1,
#     round(xend)>272 & round(xend)<=1877 ~ 2,
#     round(xend)>1877 & round(xend)<=2565 ~ 3,
#     round(xend)>2565 ~ 4
#   ))
segs <- dend$segments %>% as_tibble() %>% 
  mutate(partition=case_when(
    round(xend)<= 2211 ~ 1,
    round(xend)>2211 & round(xend)<=2670 ~ 2,
    round(xend)>2670 ~ 3
  ))
p <- ggplot(segs) + 
  geom_segment(aes(x = x, y = y, xend = xend, yend = yend,col=factor(partition)))+
  scale_color_manual(values=viridis_pal(begin=0.2,end=0.8)(4),name="Behavioral Group")+
  labs(x="",y="",title="Dendrogram of all Data with 3 Behavioral Classes")+
  theme(axis.text=element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor=element_blank())
p
```

We can also use the same data to perform a principal component analysis to help visualize our clusters.

```{r}
dat.pca <- FactoMineR::PCA(dat_nolabs,graph=FALSE)

# factoextra::fviz_dend(dat.hclust,k=4,palette=viridis_pal()(4),show_labels=FALSE,rect=T,rect_lty=2)
# summary(dat.pca)
```

```{r}
# pca.inds <- dat.pca$ind$coord %>% as_tibble() %>% mutate(partition=dat_clust$Best.partition) %>% mutate(partition=factor(partition))
# glimpse(pca.inds)
# pca.vars <- dat.pca$var$coord %>% as_tibble() %>% mutate(lab=colnames(dat_nolabs))
# glimpse(pca.vars)

fviz_pca(dat.pca,geom.ind="point",
         col.ind=dat_clust$Best.partition %>% as.factor,
         addEllipses=TRUE,col.var="black",
         palette=viridis_pal(begin=0.2,end=0.8)(4),alpha.ind=0.4)+
  labs(color="",shape="",fill="",shape="")+
  scale_x_continuous(expand = c(.1,0))

# scree plot
fviz_eig(dat.pca,choice = 'variance')
```

Here, with both observations and variables plotted along the first two principal components, we can start to see the variables that are defining the classes. In general, the first class seems to be associated with smaller vessels and lower mean trip distance, risk-taking and exploration, along with smaller home ranges and larger variance in trip duration.

The second group is associated with greater port number and diversity, larger home ranges, and greater choice entropy.

The third group seems to be larger vessels that are associated with lower variance in trip duration and a shorter season (derby), with a greater proportion of tickets from other fisheries and landings diversity.

As another visualization of differences between clusters, we can look at distribution of specific variable values across groups

```{r,fig.height=8,fig.width=8}
library(multcompView)
find_multcompletters <- function(df){
  aovtemp <- aov(df,formula = value~partition)
  tukeyhsd <- TukeyHSD(aovtemp)
  multcompLetters(tukeyhsd$partition[,4])$Letters %>% 
    enframe(name='partition',value='letter') %>% 
    mutate(partition=as.factor(partition))
}
group_means <- dat %>% 
  select(-drvid,-crab_season,-revenue_mean,-revenue_sd,-landings_mean,-landings_sd) %>% 
  na.omit() %>% 
  mutate(partition=dat_clust$Best.partition %>% as.factor()) %>% 
  pivot_longer(-partition,names_to='variable',values_to = 'value') %>% 
  group_by(variable) %>% 
  nest() %>% 
  mutate(plotvars=purrr::map2(variable,data,function(j,k){
    labels=find_multcompletters(k)
    k %>%
      left_join(labels,by='partition') %>% 
      ggplot(aes(factor(partition),value,fill=factor(partition)))+
      geom_boxplot(col='gray50')+
      geom_text(aes(partition,y=(max(value)-min(value))*0.9+min(value),label=letter),check_overlap = T,nudge_x=-0.15,size=4)+
      scale_fill_manual(values=viridis_pal(begin=0.2,end=0.8)(4))+
      labs(y="",x="",title=j)+
      guides(fill='none')
  })) %>% 
  ungroup()

library(cowplot)

plot_grid(plotlist=group_means$plotvars,labels=LETTERS[1:nrow(group_means)])
```

Letters in the plot above show the results of Tukey's honest significant difference test between the values of each variable amongst classes. We can use these to further define each class, and give each a nickname.

1. Group 1 have a pattern that seems to indicate more localized fishing. We can call them the **locals**. They are the smallest vessels, with low diversity in port use, smallest home range, and generally shorter trips (distance and duration), but with the highest variance in trip duration relative to other groups. The least derby-like fishers (i.e., latest day reaching 90% cumulative landings). Reliant on Dungeness crab, with low participation in other fisheries (but more than Group 2). Least likely to fish in high wind conditions and have the lowest choice entropy amongst groups.

2. Group 2 are the **rovers**. They are medium-sized vessels with the highest port use by number and diversity. They are between the two other groups in mean trip distance and duration, but have the highest location choice entropy (indicating exploratory behavior) as well as the largest home ranges. They have the lowest participation in other fisheries out of all the groups.

3. Group 3 is composed of the **derby fishers**, and is defined by its participation in other fisheries and its derby behavior. Vessels in Group 3 are the largest of any of the groups. Highest propensity for derby fishing (i.e., earliest day reaching 90% cumulative landings), and highest participation in other fisheries, by number of tickets, landings diversity and overall revenue. Fishers in this group are in the middle of the other two groups in their overall port use, but they take the longest trips. Their location choice entropy and home ranges are greater than Group 1 but less than Group 2. Finally, they have the highest propensity to fish in high wind conditions out of all of the groups.

## Variable Importance

We can investigate the contribution of individual variables to the eventual classification. We use random forests in the package `randomForest` to perform supervised classification and see which variables contribute most to classification.

Random forests are built from a number of individual classification trees (we use 1000). Each tree is built by recursive binary partitioning of the data, using a predictor variable (e.g., `trip_dur_mean`) to make a split in the data to create the most homogenous subgroups, where homogeneity is measured by the Gini coefficient on the relative proportions of class (Group 1, 2, or 3) membership. The tree keeps dividing until the Gini index no longer declines with further splitting, at which point the tree is "fully grown".

Random forests build many of these trees, with two differences from single classification trees. First, for each tree, only a bootstrapped subsample of the original data are chosen for classification. Second, only a subset of the predictor variables are "available" to the tree for binary partitioning, chosen randomly. Therefore, each tree is twice randomized---first through bootstrapping, and then through the random selection of partitioning variables; hence creating the "random forest". Once many trees are grown from the bootstrapped sample, the classes for the observations from the original data set that were not chosen for fitting (known as out-of-bag observations) are predicted by each tree in the forest. We compile the rate at which the forest mis-classifies observations into the incorrect class. The measure of variable importance is the *change* in this mis-classification rate when a variable is randomly permuted in the out-of-bag sample. In other words, we measure how much worse our classification performs without the correctly-ordered variable.

We apply this procedure to our data, using the 3 classes from above and growing 1000 trees.

```{r}
# variable contribution from PCA
# fviz_contrib(dat.pca,choice = 'var')

dat_nolabs <- dat_clustered %>% 
  na.omit() %>% 
  select(-(1:2)) %>% 
  mutate(partition=as.factor(partition))

library(randomForest)
rf <- randomForest(partition~.,data=dat_nolabs,importance = TRUE,ntree=1000)
var_imp <- rf$importance %>% as_tibble(rownames="variable") %>% 
  select(variable,MeanDecreaseAccuracy) %>% 
  mutate(rel_imp=MeanDecreaseAccuracy*100)
var_imp %>% 
  arrange(desc(rel_imp)) %>% 
  mutate(varname=factor(variable,levels=variable)) %>% 
  ggplot(aes(varname,rel_imp))+
  geom_col()+
  # scale_y_continuous(breaks=seq(0,10,by=2))+
  labs(x="Variable",y="Relative Importance")+
  theme(axis.text.x = element_text(angle=90))
```

The two variables that, when permuted, had the greatest effect on classification errors were mean trip distance and the proportion of tickets from other fisheries. vessel size was also important in classification.

## Class Transition Across Years

We have described the behavior of some broad groups of Dungeness crab fishers, and explored the importance of individual behavioral variables in distinguishing between groups. Lastly, we want to investigate how the total number of vessels in each class may have shifted over the course of years on our data, as well as explore interannual movements of individual vessels between the three classes.

```{r}
class_by_yr <- dat_clustered %>% 
  group_by(crab_season,partition) %>% 
  summarise(n_vessels=n_distinct(drvid)) %>% 
  ungroup()
class_by_yr %>% 
  ggplot(aes(crab_season,n_vessels,fill=factor(partition)))+
  geom_col()+
  labs(x="Crab Season",y="Number of Vessels",title="Number of Vessels by Behavior Class and Year",fill="Group")+
  scale_fill_manual(values=viridis_pal(begin=0.2,end=0.8)(4))+
  theme(axis.text.x = element_text(angle=90,vjust=0.5),
        legend.position = 'bottom')
class_by_yr %>% 
  ggplot(aes(crab_season,n_vessels,group=factor(partition),color=factor(partition)))+
  geom_line(size=2)+
  labs(x="Crab Season",y="Number of Vessels",title="Number of Vessels by Behavior Class and Year",color="Group")+
  scale_color_manual(values=viridis_pal(begin=0.2,end=0.8)(4))+
  theme(axis.text.x = element_text(angle=90,vjust=0.5),
        legend.position = 'bottom')
```

The number of vessels in Group 1 declined in 2014-15, but has recovered since then.

We can also look at the flows of vessels between classes and years.

```{r,warning=F,message=F}
library(ggalluvial)

dat_alluvium <- dat_clustered %>% 
  select(drvid,crab_season,partition) %>% 
  distinct() %>% 
  mutate(class=case_when(
    partition==1 ~ "Group 1",
    partition==2 ~ "Group 2",
    partition==3 ~ "Group 3",
    partition==4 ~ "Group 4"
  )) %>% 
  complete(drvid,crab_season,fill=list('class'="Not Fishing"))

# with boats not fishing in a given season
# dat_alluvium %>%
#   ggplot(aes(crab_season,stratum=class,alluvium=drvid,fill=class,label=class))+
#   scale_fill_manual(values=viridis_pal()(4),name="Class")+
#   geom_flow()+
#   geom_stratum()+
#   theme(axis.text.x = element_text(angle=90,vjust=0.5))

# non-fishing vessels removed
dat_alluvium %>%
  filter(class!="Not Fishing") %>% 
  ggplot(aes(crab_season,stratum=class,alluvium=drvid,fill=class,label=class))+
  scale_fill_manual(values=viridis_pal(begin=0.2,end=0.8)(4),name="")+
  geom_flow()+
  geom_stratum()+
  labs(x="Crab Season",y="Number of Vessels")+
  theme(axis.text.x = element_text(angle=90,vjust=0.5),
        legend.position = 'bottom')
```

This is the same plot as above but with the flows from each class included between years.

### Fleetwide Changes in Individual Metrics

We can look at changes in the important variables fleetwide over time

```{r}
annual_var_means <- dat_clustered %>% 
  select(-drvid,-partition) %>% 
  pivot_longer(-crab_season,names_to = 'variable',values_to='value') %>% 
  group_by(crab_season,variable) %>% 
  summarise(mean_val=mean(value),sd_val=sd(value)) %>% 
  ungroup()

annual_var_means %>% 
  ggplot(aes(crab_season,mean_val,group=1))+
  geom_point()+
  geom_line()+
  facet_wrap(~variable)+
  geom_hline(yintercept=0)+
  labs(x="Crab Year",y="Fleet-wide Mean")+
  theme(axis.text.x = element_text(angle=90,vjust=0.5),
        legend.position = 'bottom')

# annual_var_means %>% 
#   filter(variable=='port_shannon') %>% 
#   ggplot(aes(crab_season,mean_val,group=1))+
#   geom_point()+
#   geom_line()+
#   geom_hline(yintercept=0)+
#   labs(x="Crab Year",y="Fleet-wide Mean")+
#   theme(axis.text.x = element_text(angle=90,vjust=0.5),
#         legend.position = 'bottom')
```

Fleetwide means have remained relatively consistent across years.

## Vessel Size Distribution by Class

This is how vessel length maps on to behavioral groups in the fishery

```{r}
vessel_sizes <- read_rds(here::here('fishing behavior','vessel_sizes.rds'))
dat_clustered %>% left_join(vessel_sizes) %>% 
  ggplot(aes(vessel_size,fill=factor(partition)))+
  geom_density(alpha=0.5)+
  labs(x="Vessel Length",y='density',fill='Group',title="Vessel Length Distribution by Behavioral Group")+
  scale_fill_manual(values=viridis_pal(begin=0.2,end=0.8)(4))+
  facet_wrap(~crab_season)

```

### Vessel Size Correlations

Correlation of vessel size with other variables in the cluster analysis. As expected from the above, larger vessels are associated with longer trips, more exploratory behavior, risk-taking behavior (i.e. propensity to fish in high winds), and shorter seasons.

```{r}
vars.corrs %>% 
  filter(var1=='vessel_size',var2 !='vessel_size') %>% 
  arrange(desc(corr)) %>% 
  mutate(varname=factor(var2,levels=var2)) %>% 
  mutate(is_neg=ifelse(corr<0,"neg","pos")) %>% 
  ggplot(aes(varname,corr,fill=is_neg))+
  geom_col()+
  labs(x="Variable",y="Correlation with Vessel Length")+
  scale_fill_manual(values=c("red","blue"),name="")+
  scale_y_continuous(breaks=seq(-1,1,by=0.2))+
  guides(fill='none')+
  theme(axis.text.x = element_text(angle=90,vjust=0.5,hjust=1))
```

## Behavioral Class and Revenue

Our last research question is, "*How do fishing behavioral strategies relate to revenue and landings from fishing?*". Now that we have the three groups of vessels, we can use these groups to explore how these behavioral strategies link to revenues gained.

Here is how revenue and revenue variance have changed over time for each behavioral group

### Per-trip Revenue and Landings

```{r}
dat_rev <- dat_clustered %>% select(drvid,crab_season,partition) %>%
  left_join(select(dat,drvid,crab_season,contains('revenue_'),contains('landings_'))) %>% 
  select(-landings_diversity) %>% 
  ungroup()
glimpse(dat_rev)

dat_rev %>% 
  group_by(crab_season,partition) %>% 
  summarise(rev_per_trip=mean(revenue_mean,na.rm=T)) %>% 
  ungroup() %>% 
  ggplot(aes(crab_season,rev_per_trip,group=factor(partition),color=factor(partition)))+
  geom_line(size=2)+
  labs(x="Crab Season",y="Revenue per Trip",title="Per-trip Revenue by Behavior Class and Year",color="Group")+
  scale_color_manual(values=viridis_pal(begin=0.2,end=0.8)(4))+
  theme(axis.text.x = element_text(angle=90,vjust=0.5),
        legend.position = 'bottom')
```

What about variance in per-trip revenue?

```{r}
dat_rev %>% 
  group_by(crab_season,partition) %>% 
  summarise(rev_sd=mean(revenue_sd,na.rm=T)) %>% 
  ungroup() %>% 
  ggplot(aes(crab_season,rev_sd,group=factor(partition),color=factor(partition)))+
  geom_line(size=2)+
  labs(x="Crab Season",y="Mean SD of Revenue per Trip",title="Per-trip Standard Deviation of Revenue by Behavior Class and Year",color="Group")+
  scale_color_manual(values=viridis_pal(begin=0.2,end=0.8)(4))+
  theme(axis.text.x = element_text(angle=90,vjust=0.5),
        legend.position = 'bottom')
```

And the above two plots, but for landings instead of revenue

```{r}
dat_rev %>% 
  group_by(crab_season,partition) %>% 
  summarise(land_per_trip=mean(landings_mean,na.rm=T)) %>% 
  ungroup() %>% 
  ggplot(aes(crab_season,land_per_trip,group=factor(partition),color=factor(partition)))+
  geom_line(size=2)+
  labs(x="Crab Season",y="Landings (lbs) per Trip",title="Per-trip Landings by Behavior Class and Year",color="Group")+
  scale_color_manual(values=viridis_pal(begin=0.2,end=0.8)(4))+
  theme(axis.text.x = element_text(angle=90,vjust=0.5),
        legend.position = 'bottom')
dat_rev %>% 
  group_by(crab_season,partition) %>% 
  summarise(rev_sd=mean(revenue_sd,na.rm=T)) %>% 
  ungroup() %>% 
  ggplot(aes(crab_season,rev_sd,group=factor(partition),color=factor(partition)))+
  geom_line(size=2)+
  labs(x="Crab Season",y="Mean SD of Landings (lbs) per Trip",title="Per-trip Standard Deviation of Landings by Behavior Class and Year",color="Group")+
  scale_color_manual(values=viridis_pal(begin=0.2,end=0.8)(4))+
  theme(axis.text.x = element_text(angle=90,vjust=0.5),
        legend.position = 'bottom')
```

### Season-long Revenue and Landings
  
Finally, what about total revenue gained from the Dungeness fishery over the course of the entire season? Remember that especially Group 3 vessels have short, derby seasons.

```{r}
fishtix <- read_rds(here::here('data','processed','fish tickets','fish_tickets_w_dayofseason.rds'))

total_rev_landings_season <- fishtix %>%
  filter(drvid %in% dat_clustered$drvid) %>% 
  distinct(drvid,crab_season,Rec_ID,DCRB_revenue,DCRB_lbs) %>% 
  group_by(drvid,crab_season) %>% 
  summarise(total_rev=sum(DCRB_revenue,na.rm=T),total_land=sum(DCRB_lbs,na.rm=T)) %>% 
  ungroup() %>% 
  right_join(dat_rev)

total_rev_landings_season %>%
  group_by(crab_season,partition) %>% 
  summarise(rev_tot=mean(total_rev,na.rm=T)) %>% 
  ungroup() %>% 
  ggplot(aes(crab_season,rev_tot,group=factor(partition),color=factor(partition)))+
  geom_line(size=2)+
  labs(x="Crab Season",y="Mean Total Season Revenue",title="Total Revenue by Behavior Class and Year",color="Group")+
  scale_color_manual(values=viridis_pal(begin=0.2,end=0.8)(4))+
  theme(axis.text.x = element_text(angle=90,vjust=0.5),
        legend.position = 'bottom')

total_rev_landings_season %>%
  group_by(crab_season,partition) %>% 
  summarise(land_tot=mean(total_land,na.rm=T)) %>% 
  ungroup() %>% 
  ggplot(aes(crab_season,land_tot,group=factor(partition),color=factor(partition)))+
  geom_line(size=2)+
  labs(x="Crab Season",y="Mean Total Season Landings (lbs)",title="Total Landings by Behavior Class and Year",color="Group")+
  scale_color_manual(values=viridis_pal(begin=0.2,end=0.8)(4))+
  theme(axis.text.x = element_text(angle=90,vjust=0.5),
        legend.position = 'bottom')
```


