---
title: "Prep Fishing Ground Data"
output: html_document
---

Produce a .csv file with filtered VMS data that (to the best of our ability) represents Dungeness crab fishing grounds.


```{r setup, include=FALSE}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE,
                      fig.width=12, fig.height=8)

# r packages
library(tidyverse)
library(magrittr)
library(sf)
library(lubridate)
library(ggplot2)
library(rgdal)
library(rgeos)
library(here)
library(rnaturalearth)
library(cowplot)
# ggplot theme

plot_theme <- theme_classic() +
  theme(text=element_text(family="sans",size=12,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
theme_set(plot_theme)

```
<br>

User Inputs:
```{r}
crab_years <- c(2015)
states <- c("C")
outdir <- "fishing-ground-change/project-dat/vms/interpolation_60min/NaN_speed_filter"
```
<br>

## Data

```{r read_vms}
for(y in crab_years){
  tmpvms <-  read_rds(here::here('data','processed','matched','interpolation_60min',paste0(y,'interpolated.rds')))
  if(y==crab_years[1]){
    vms <- read_rds(here::here("data","processed","matched","interpolation_60min",paste0(y-1,"interpolated.rds"))) %>%
      bind_rows(tmpvms)
  } else{
    vms %<>% bind_rows(tmpvms)
  }
}
```


filter for California Dungeness crab trips, and grab columns of interest. add in some date columns specific to crab fishing, and the management area (northern / central). filter for the specified crab year.
```{r thin}
vms.thin <- vms %>%
  # according to user inputs
  filter(agency_code == states) %>%
  filter(TARGET_lbs == "DCRB") %>%
  filter(removal_type_code=="C" | removal_type_code=="D" | removal_type_code=="U") %>%
  # grab subset of columns
  dplyr::select(Rec_ID,date, pacfin_port_code, port_group_code, agency_code, 
                removal_type_code, removal_type_name, TARGET_lbs,
                VMS_RECNO,VESSEL_NAM,drvid,UTCDATETIM,westcoastdate,westcoastdate_notime,
                X_COORD,Y_COORD, LATITUDE,LONGITUDE,NGDC_M,
                AVG_SPEED,avg_speed_recalc,AVG_COURSE,DECLARATIO, in_port, is_interpolated,
                DCRB_lbs,DCRB_revenue) %>%
  # add on crab season based on month of year, and filter
  # mutate(date=as_date(date, c("ymd","mdy")),
  #        year=year(date)) %>%
  mutate(year=year(date)) %>%
  mutate(crab_season=case_when(
    month(date)>10 ~ paste0(year,"-",year+1),
    month(date)<=10 ~ paste0(year-1,"-",year)),
    crab_year=case_when(
      month(date)>10 ~ year+1,
      month(date)<=10 ~ year)) %>%
  filter(crab_year %in% crab_years) %>%
  # add on management district based on port group
  mutate(port_area=ifelse(port_group_code %in% c("CCA","ERA","BGA"),"northern","central"))
```
<br>

season start dates, by management district (these are the earliest *codified* season start dates for any port in that management district, validated with CDFW)
```{r}
startdat <- read_csv(here('fishing-ground-change','project-dat','dcrb_season_starts.csv')) %>%
  mutate(season_start_date=mdy(season_start_date))
```


## Filter for Fishing Activity

Apply filters for best guess at VMS points that represent fishing activity. filter out data points 

(1) with average speed is > 4.2 m/s (in transit), 

(2) with depths greater than 150m, (depth) and 

(3) with low speeds inside the buffer zones **OR** if speed is `NA` (point has an NGDC_M > 100000, the ArcGIS indication of an in-port point.) (in port)

Mark "remove" if the record is in port *and* the avg speed is < 0.75 m/s.
```{r flag}
vms.thin %<>%  mutate(to_remove = ifelse(
  (!is.na(in_port) & avg_speed_recalc < 0.75) | is.na(avg_speed_recalc), "in_port",ifelse(
    avg_speed_recalc > 4.2, "transit",ifelse(
      NGDC_M > 150, "depth","retain"))))
```
<br>

Filter records
```{r filter}
vms.filter <- vms.thin %>%
  filter(to_remove =="retain" | is.na(to_remove)) %>% dplyr::select(-to_remove)
```
<br>

Summarize filtering from this section
```{r track_filter1}
track_filtering <- vms.thin %>%
  group_by(crab_season) %>%
  mutate(seasonVMS=length(unique(VMS_RECNO))) %>%
  ungroup() %>%
  group_by(crab_season,seasonVMS,to_remove) %>%
  summarise(nVMS=length(unique(VMS_RECNO))) %>%
  mutate(pVMS=nVMS/seasonVMS)
```
<br>


## Filter low activity

Remove vessel with fewer than 5 relocations per season.
```{r plot_relocations, echo=FALSE}
vms.filter %>%
  group_by(crab_season,drvid) %>%
  summarise(nlocs=length(unique(LATITUDE)),
            ntrips=length(unique(Rec_ID))) %>%
  pivot_longer(c(nlocs,ntrips),names_to="activity") %>%
  ggplot(aes(x=value, fill=crab_season)) + geom_histogram() +
  facet_grid(cols=vars(activity), scales="free") + xlab("") + theme(legend.title=element_blank())
```
```{r relocations}
low.activity.vessels <- vms.filter %>%
  group_by(crab_season,drvid) %>%
  summarise(nlocs=length(unique(LATITUDE)),
            ntrips=length(unique(Rec_ID))) %>%
  filter(nlocs < 5) %>%
  dplyr::select(crab_season,drvid) %>% distinct()
```
```{r echo=FALSE}
message("\n\nwill remove ", length(unique(low.activity.vessels$drvid)), " low activity vessels (",round(length(unique(low.activity.vessels$drvid))/length(unique(vms.filter$drvid))*100,3),"% all vessels) from data set.")
```
```{r filter2}
vms.filter %<>% anti_join(low.activity.vessels,by=c("drvid","crab_season"))
```
<br>

## Track lost data

```{r echo=FALSE, track_filter2}
if(dim(low.activity.vessels)[1] > 0){
track_filtering %<>% bind_rows(low.activity.vessels %>%
                                 left_join(vms.thin %>% dplyr::select(crab_season,drvid,VMS_RECNO) %>% distinct()) %>%
                                 group_by(crab_season) %>%
                                 summarize(nVMS=length(unique(VMS_RECNO))) %>%
                                 left_join(track_filtering %>% 
                                             dplyr::select(crab_season,seasonVMS) %>%
                                             distinct(),by="crab_season") %>%
                                 mutate(to_remove="low_activity",
                                        pVMS=nVMS/seasonVMS) %>%
                                 dplyr::select(all_of(colnames(track_filtering))))
}
track_filtering %>%
  rename("percent VMS records"=pVMS,"number of VMS records"=nVMS) %>%
  pivot_longer(cols=c(`percent VMS records`,`number of VMS records`), names_to="name") %>%
  ggplot(aes(x=crab_season,y=value, fill=to_remove)) +
  geom_col() +
  facet_wrap(~name,scales="free_y") + labs(x="crab season",y="") + theme(legend.title=element_blank())
```
<br>

Let's also plot a few random trips, pre- and post-filtering.

```{r map, echo=FALSE}
# subset trips
vms.filter.sub <- vms.filter %>% filter(port_group_code=="CCA") %>% 
  group_by(Rec_ID) %>% summarise(nrec=length(unique(VMS_RECNO))) %>% filter(nrec > 10) %>% sample_n(5) %>%
  left_join(vms.filter,by="Rec_ID")
  
vms.thin.sub <- vms.thin %>% filter(Rec_ID %in% vms.filter.sub$Rec_ID)

# convert to spatial obj
test_filter_sf <- as.data.frame(vms.filter.sub) %>% 
  dplyr::select(Rec_ID,VMS_RECNO,X_COORD,Y_COORD,date) %>%
  st_as_sf(coords=c("X_COORD","Y_COORD"),crs = "+proj=utm +north +zone=10 +ellps=WGS84")
test_thin_sf <- as.data.frame(vms.thin.sub) %>% 
  dplyr::select(Rec_ID,VMS_RECNO,X_COORD,Y_COORD,date) %>%
  st_as_sf(coords=c("X_COORD","Y_COORD"),crs = "+proj=utm +north +zone=10 +ellps=WGS84")


# import a background/land layer from rnaturalearth package
coaststates <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('California','Oregon','Washington')) %>%
  # make sure CRS is the same as the port layer
  st_transform(st_crs(test_thin_sf))

testbbox <- st_bbox(test_thin_sf)
ggplot()+
  geom_sf(data=coaststates,fill='gray50')+
  # geom_sf(data=ports_buffer,fill='seagreen',alpha=0.5)+
  geom_sf(data=test_thin_sf,size=1,color="red")+
  geom_sf(data=test_filter_sf,size=1,color="black")+
  facet_wrap(~Rec_ID) +
  xlim(testbbox[1]-5000,testbbox[3]+5000)+ylim(testbbox[2]-5000,testbbox[4]+5000)+
  labs(x='',y='',title="Points for Sample CCA Trips")+
  theme(axis.text.x = element_text(angle=90))

```
```{r map2, echo=FALSE}
# subset trips
vms.filter.sub <- vms.filter %>% filter(port_group_code=="CCA") %>% 
  group_by(Rec_ID) %>% summarise(nrec=length(unique(VMS_RECNO))) %>% filter(nrec > 10) %>% sample_n(5) %>%
  left_join(vms.filter,by="Rec_ID")
  
vms.thin.sub <- vms.thin %>% filter(Rec_ID %in% vms.filter.sub$Rec_ID)

# convert to spatial obj
test_filter_sf <- as.data.frame(vms.filter.sub) %>% 
  dplyr::select(Rec_ID,VMS_RECNO,X_COORD,Y_COORD,date) %>%
  st_as_sf(coords=c("X_COORD","Y_COORD"),crs = "+proj=utm +north +zone=10 +ellps=WGS84")
test_thin_sf <- as.data.frame(vms.thin.sub) %>% 
  dplyr::select(Rec_ID,VMS_RECNO,X_COORD,Y_COORD,date) %>%
  st_as_sf(coords=c("X_COORD","Y_COORD"),crs = "+proj=utm +north +zone=10 +ellps=WGS84")

testbbox <- st_bbox(test_thin_sf)
ggplot()+
  geom_sf(data=coaststates,fill='gray50')+
  # geom_sf(data=ports_buffer,fill='seagreen',alpha=0.5)+
  geom_sf(data=test_thin_sf,size=1,color="red")+
  geom_sf(data=test_filter_sf,size=1,color="black")+
  facet_wrap(~Rec_ID) +
  xlim(testbbox[1]-5000,testbbox[3]+5000)+ylim(testbbox[2]-5000,testbbox[4]+5000)+
  labs(x='',y='',title="Points for Sample CCA Trips")+
  theme(axis.text.x = element_text(angle=90))

```


```{r map3, echo=FALSE}
# subset trips
vms.filter.sub <- vms.filter %>% filter(port_group_code=="BDA") %>% 
  group_by(Rec_ID) %>% summarise(nrec=length(unique(VMS_RECNO))) %>% filter(nrec > 10) %>% sample_n(5) %>%
  left_join(vms.filter,by="Rec_ID")
  
vms.thin.sub <- vms.thin %>% filter(Rec_ID %in% vms.filter.sub$Rec_ID)

# convert to spatial obj
test_filter_sf <- as.data.frame(vms.filter.sub) %>% 
  dplyr::select(Rec_ID,VMS_RECNO,X_COORD,Y_COORD,date) %>%
  st_as_sf(coords=c("X_COORD","Y_COORD"),crs = "+proj=utm +north +zone=10 +ellps=WGS84")
test_thin_sf <- as.data.frame(vms.thin.sub) %>% 
  dplyr::select(Rec_ID,VMS_RECNO,X_COORD,Y_COORD,date) %>%
  st_as_sf(coords=c("X_COORD","Y_COORD"),crs = "+proj=utm +north +zone=10 +ellps=WGS84")

testbbox <- st_bbox(test_thin_sf)
ggplot()+
  geom_sf(data=coaststates,fill='gray50')+
  # geom_sf(data=ports_buffer,fill='seagreen',alpha=0.5)+
  geom_sf(data=test_thin_sf,size=1,color="red")+
  geom_sf(data=test_filter_sf,size=1,color="black")+
  facet_wrap(~Rec_ID) +
  xlim(testbbox[1]-5000,testbbox[3]+5000)+ylim(testbbox[2]-5000,testbbox[4]+5000)+
  labs(x='',y='',title="Points for Sample BDA Trips")+
  theme(axis.text.x = element_text(angle=90))
```

```{r map4, echo=FALSE}
# subset trips
vms.filter.sub <- vms.filter %>% filter(port_group_code=="BDA") %>% 
  group_by(Rec_ID) %>% summarise(nrec=length(unique(VMS_RECNO))) %>% filter(nrec > 10) %>% sample_n(5) %>%
  left_join(vms.filter,by="Rec_ID")
  
vms.thin.sub <- vms.thin %>% filter(Rec_ID %in% vms.filter.sub$Rec_ID)

# convert to spatial obj
test_filter_sf <- as.data.frame(vms.filter.sub) %>% 
  dplyr::select(Rec_ID,VMS_RECNO,X_COORD,Y_COORD,date) %>%
  st_as_sf(coords=c("X_COORD","Y_COORD"),crs = "+proj=utm +north +zone=10 +ellps=WGS84")
test_thin_sf <- as.data.frame(vms.thin.sub) %>% 
  dplyr::select(Rec_ID,VMS_RECNO,X_COORD,Y_COORD,date) %>%
  st_as_sf(coords=c("X_COORD","Y_COORD"),crs = "+proj=utm +north +zone=10 +ellps=WGS84")

testbbox <- st_bbox(test_thin_sf)
ggplot()+
  geom_sf(data=coaststates,fill='gray50')+
  # geom_sf(data=ports_buffer,fill='seagreen',alpha=0.5)+
  geom_sf(data=test_thin_sf,size=1,color="red")+
  geom_sf(data=test_filter_sf,size=1,color="black")+
  facet_wrap(~Rec_ID) +
  xlim(testbbox[1]-5000,testbbox[3]+5000)+ylim(testbbox[2]-5000,testbbox[4]+5000)+
  labs(x='',y='',title="Points for Sample BDA Trips")+
  theme(axis.text.x = element_text(angle=90))
```
<br>

## Add day of season
```{r dos}
vms.filter %<>% 
  left_join(startdat,by=c("crab_year","port_area"="pcdistrict")) %>%
  mutate(ticket_day_of_season=difftime(date,season_start_date,units='days')) %>% 
  mutate(ticket_day_of_season=as.numeric(ticket_day_of_season))
```

```{r echo=FALSE, plot_dos}
vms.filter %>% dplyr::select(Rec_ID, date, ticket_day_of_season, crab_year, port_area) %>%
  group_by(port_area,crab_year,date,ticket_day_of_season) %>%
  summarise(ntrips=length(unique(Rec_ID))) %>%
  ggplot(aes(x=date,y=ticket_day_of_season, col=as.factor(crab_year))) +
  geom_point() + facet_grid(cols=vars(port_area), rows=vars(as.factor(crab_year))) + theme(legend.title=element_blank())
```

<br>

## Save fishing activity records
```{r write}
for(y in crab_years){
  vms.out <- vms.filter %>% filter(crab_year==y)
  write_rds(vms.out,here::here(outdir,paste0(y,'season_crabfishing.rds')))
}
```
<br>



## QC

**Are we removing too many trips / records by filtering for only those which TARGET Dungeness crab?**

How many VMS records with DCRB landed lbs > 0 do / do not have the TARGET as DCRB?
```{r echo=FALSE, warning=FALSE}
total_vms <- vms %>%
  filter(agency_code %in% states) %>%
  filter(DCRB_lbs > 0) %>%
  mutate(year=year(ymd(date))) %>%
  group_by(year) %>%
  summarise(totalvms = length(unique(VMS_RECNO)))
plotn <- vms %>%
  filter(agency_code %in% states) %>%
  filter(DCRB_lbs > 0) %>%
  mutate(year=year(ymd(date))) %>%
  dplyr::select(Rec_ID,year,drvid,VMS_RECNO, TARGET_lbs,DCRB_lbs,UTCDATETIM,LATITUDE,LONGITUDE) %>%
    mutate(dcrb_targeted = ifelse(TARGET_lbs == "DCRB","dcrb target","other target")) %>%
  group_by(year, dcrb_targeted) %>%
  summarise(nvms = length(unique(VMS_RECNO))) %>%
  ggplot(aes(x=as.factor(year),y=nvms)) + 
  geom_col() + facet_wrap(~dcrb_targeted,scales="free_y") + labs(x="year",y="N VMS records")
plotp <- vms %>%
  filter(agency_code %in% states) %>%
  filter(DCRB_lbs > 0) %>%
  mutate(year=year(ymd(date))) %>%
  dplyr::select(Rec_ID,year,drvid,VMS_RECNO, TARGET_lbs,DCRB_lbs,UTCDATETIM,LATITUDE,LONGITUDE) %>%
    mutate(dcrb_targeted = ifelse(TARGET_lbs == "DCRB","dcrb target","other target")) %>%
  group_by(year, dcrb_targeted) %>%
  summarise(nvms = length(unique(VMS_RECNO))) %>%
  left_join(total_vms, by="year") %>% mutate(pvms=(nvms/totalvms)*100) %>%
  filter(dcrb_targeted=="other target") %>%
  ggplot(aes(x=as.factor(year),y=pvms)) + 
  geom_col() + labs(x="year",y="% VMS records", title="other target")
plot_grid(plotn,plotp, ncol=2)
```


Fishing Trips?
```{r echo=FALSE, warning=FALSE}
total_trips <- vms %>%
  filter(agency_code %in% states) %>%
  filter(DCRB_lbs > 0) %>%
  mutate(year=year(ymd(date))) %>%
  group_by(year) %>%
  summarise(totaltrips = length(unique(Rec_ID)))
plotn <- vms %>%
  filter(agency_code %in% states) %>%
  filter(DCRB_lbs > 0) %>%
  mutate(year=year(ymd(date))) %>%
  dplyr::select(Rec_ID,year,drvid,TARGET_lbs,DCRB_lbs) %>%
    mutate(dcrb_targeted = ifelse(TARGET_lbs == "DCRB","dcrb target","other target")) %>%
  group_by(year, dcrb_targeted) %>%
  summarise(ntrips = length(unique(Rec_ID))) %>%
  ggplot(aes(x=as.factor(year),y=ntrips)) + 
  geom_col() + facet_wrap(~dcrb_targeted,scales="free_y") + labs(x="year",y="N Trips")
plotp <- vms %>%
  filter(agency_code %in% states) %>%
  filter(DCRB_lbs > 0) %>%
  mutate(year=year(ymd(date))) %>%
  dplyr::select(Rec_ID,year,drvid,TARGET_lbs,DCRB_lbs) %>%
    mutate(dcrb_targeted = ifelse(TARGET_lbs == "DCRB","dcrb target","other target")) %>%
  group_by(year, dcrb_targeted) %>%
  summarise(ntrips = length(unique(Rec_ID))) %>%
  left_join(total_trips, by="year") %>% mutate(ptrips=(ntrips/totaltrips)*100) %>%
  filter(dcrb_targeted=="other target") %>%
  ggplot(aes(x=as.factor(year),y=ptrips)) + 
  geom_col() + labs(x="year",y="% Trips", title="other target")
plot_grid(plotn,plotp, ncol=2)
```


Dungeness crab lbs landed?
```{r echo=FALSE, warning=FALSE}
total_lbs <- vms %>%
  filter(agency_code %in% states) %>%
  filter(DCRB_lbs > 0) %>%
  mutate(year=year(ymd(date))) %>%
  group_by(year) %>%
  summarise(totallbs = sum(DCRB_lbs))
plotn <- vms %>%
  filter(agency_code %in% states) %>%
  filter(DCRB_lbs > 0) %>%
  mutate(year=year(ymd(date))) %>%
  dplyr::select(Rec_ID,year,drvid,TARGET_lbs,DCRB_lbs) %>%
    mutate(dcrb_targeted = ifelse(TARGET_lbs == "DCRB","dcrb target","other target")) %>%
  group_by(year, dcrb_targeted) %>%
  summarise(nlbs = sum(DCRB_lbs)) %>%
  ggplot(aes(x=as.factor(year),y=nlbs)) + 
  geom_col() + facet_wrap(~dcrb_targeted,scales="free_y") + labs(x="year",y="DCRB Landed Lbs")
plotp <- vms %>%
  filter(agency_code %in% states) %>%
  filter(DCRB_lbs > 0) %>%
  mutate(year=year(ymd(date))) %>%
  dplyr::select(Rec_ID,year,drvid,TARGET_lbs,DCRB_lbs) %>%
    mutate(dcrb_targeted = ifelse(TARGET_lbs == "DCRB","dcrb target","other target")) %>%
  group_by(year, dcrb_targeted) %>%
  summarise(nlbs = sum(DCRB_lbs)) %>%
  left_join(total_lbs, by="year") %>% mutate(plbs=(nlbs/totallbs)*100) %>%
  filter(dcrb_targeted=="other target") %>%
  ggplot(aes(x=as.factor(year),y=plbs)) + 
  geom_col() + labs(x="year",y="% Landed Lbs", title="other target")
plot_grid(plotn,plotp, ncol=2)
```






