---
title: "EMD by vessel"
author: "M. Fisher"
date: "2/20/2023"
output: html_document
---

Calculate the year-over-year Earth Mover's Distance for Dungeness crab fishing grounds.

For year-over-year comparisons of fishing grounds based only on vessel size, we do not need to restrict the analysis to vessels with VMS in the 2013-14 season. Instead, I just need vessels with VMS in at least two successive seasons. Then I can conduct the district-based analysis on the 2013-14 subset. 

In widening the sample size for the size-based year-over-year comparisons, I want to focus on Dungeness commercial fishing vessels which took more than 5 Dungeness trips in a given year. 

This script will establish the exvessel revenue cutoff, and then generate a list of vessels to track in the size-based year-over-year comparison.


The following filters were already imposed in creating the fishing ground data: 

- landings in a given state (e.g., California)
- trips targeting Dungeness crab
- vessels that have five or more relocations per crab season


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())

library(tidyverse)
library(here)
library(raster)
library(adehabitatHR)
library(move)
library(pryr)
```


User inputs
```{r}
yrs <- seq(2016,2017)
g <- 75


indir <- 'project-dat/vms/interpolation_60min/NaN_speed_filter'
outdir <- 'project-dat/vms/interpolation_60min/NaN_speed_filter/vessel90ud'
```





