---
title: "EMD by vessel"
author: "M. Fisher"
date: "2/20/2023"
output: html_document
---

Calculate the year-over-year Earth Mover's Distance for Dungeness crab fishing grounds.

For year-over-year comparisons of fishing grounds based only on vessel size, we do not need to restrict the analysis to vessels assigned to a group from the 2013-14 crab season. Instead, we just need vessels with VMS in at least two successive seasons.

This script includes options to remove Dungeness crab vessels based on (1) the number of fishing trips they took in a given year, and/or (2) an exvessel revenue cut-off. This is in addition to the filters that were already imposed in creating the fishing ground data: 

- landings in a given state (e.g., California)
- trips targeting Dungeness crab
- vessels that have five or more relocations per crab season


However, for the analysis as it stands, I am not going to apply any additional filters beyond those already used in script 01.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list=ls())

library(tidyverse)
library(magrittr)
library(here)
library(raster)
library(adehabitatHR)
library(move)
library(pryr)
```


User inputs
```{r}
crab_years <- seq(2011,2018)
kernel_g <- 75
udsize       <- 0.90
cut_length <- 40  # all vessels below this length are "small"


indir   <- 'project-dat/vms/interpolation_60min/NaN_speed_filter'
outdir  <- 'project-dat/vms/interpolation_60min/NaN_speed_filter/vessel90ud'

filter=FALSE
filter_trips=5
filter_revenue=5000
```



## Data

VMS
```{r read_vms}
for(y in crab_years){
  tmpvms <-  read_rds(here::here(indir,paste0(y,'season_crabfishing.rds')))
  if(y==crab_years[1]){
    vms <- tmpvms
  } else{
    vms %<>% bind_rows(tmpvms)
  }
}

str(vms)
```


Fish tickets
```{r read_fishtix}
for(y in crab_years){
  tmpy <- y-1
  tmptix <-  read_rds(here::here('../data','processed','fish tickets',paste0(tmpy,'fishtix_vlengths.rds'))) %>%
  dplyr::select(Rec_ID,drvid,year,FINAL_LENGTH) %>% distinct()
  if(y==crab_years[1]){
    tix <- tmptix
  } else{
    tix %<>% bind_rows(tmptix)
  }
}

str(tix)
```



make just a vessel length key. since vessel lengths can vary by year, vessel lengths are assigned according to the year at the *beginning* each crab season (in other words, crab_year - 1). 
```{r}
lkey <- tix %>% left_join(vms %>% dplyr::select(Rec_ID,drvid,crab_year) %>% distinct()) %>%
  filter(year!=crab_year) %>%
  dplyr::select(drvid,FINAL_LENGTH,crab_year) %>% distinct() %>%
  mutate(vessel_size=ifelse(FINAL_LENGTH < cut_length,"small","large"))

lkey.filtered <- lkey %>% filter(!is.na(FINAL_LENGTH))
```


## Filtering

### trips per vessel

### exvessel revenue

## Earth Mover's Distance per vessel

The Earth Mover's Distance calculation takes a utilization distribution as: A Raster, RasterStack, RasterBrick, SpatialPoints, SpatialPointsDataFrame. This code first constructs the utilization distribution, and then saves it uses a Spatial Points DataFrame. That object can be called back in to calculate EMD.

### create utilization distribution


```{r}
for(d in unique(vms$drvid)){
  # filter VMS data
  tmpvms <- filter(vms, crab_year %in% crab_years & drvid == d) %>%
    dplyr::select(drvid,crab_year,X_COORD, Y_COORD)
  # create the Spatial Points Data frame
  tmpvms.sp <- SpatialPointsDataFrame(coords=tmpvms %>% dplyr::select(X_COORD,Y_COORD), data=tmpvms %>% dplyr::select(crab_year), proj4string = CRS("+init=epsg:32610"))
  tmpyrs <- unique(tmpvms.sp$crab_year)
  tmpout <- list()
  # create a list of raster layers
  for(j in seq(1,length(tmpyrs))){
    y <- tmpyrs[j]
    # create utilization distribution, with grid size of 'kernel_g'
    tmpud <- kernelUD(tmpvms.sp[tmpvms.sp$crab_year==y,], grid=kernel_g)
    # convert estUD object to a raster
    tmpras <- raster(as(tmpud[[1]],"SpatialPixelsDataFrame"))
    # set any points outside the 90% line to NA, then recalculate the probability surface
    tmpras[tmpras > udsize] <- NA; newras <- tmpras/cellStats(tmpras,sum)
    tmpout[[j]] <- newras
  }
  names(tmpout) <- tmpyrs
  # save the list(by year) of 90% UD rasters as an RDS
  write_rds(tmpout, here::here(outdir,paste0(d,"_",udsize*100,"ud_RasterLayer_",kernel_g,"grid_", crab_years[1],"-",substr(tail(crab_years,1),3,4), "crabYrs.rds")))
}
rm(tmpout, tmpvms,tmpvms.sp)
```
<br>

Get filenames from above.
```{r}
file_vec <- list.files(path=here(outdir),pattern="RasterLayer")
```

Read back in the 90% UD lists, and find the year-over-year EMD for each vessel.
```{r}
# for(i in seq(1,length(file_vec))){
for(i in seq(1,10)){
  # get drvid & rds name
  f <- file_vec[[i]]
  d <- strsplit(f,split="_")[[1]][1]
  # read in rds
  ras_list <- readRDS(here::here(outdir,f))
  tmpyrs <- names(ras_list)
  
  # empty data frame for EMD
  tmpdat <- matrix(nrow=length(tmpyrs),ncol=length(tmpyrs),dimnames=list(tmpyrs,tmpyrs),data=0)
 
  # get emd for consecutive years
  sstart <- Sys.time()
  message("calculating emd...")
  for(j in seq(1,length(ras_list)-1)){
    k=j+1
    while(k < (length(ras_list)+1)){
      tmpdat[j,k] <- emd(ras_list[[j]],ras_list[[k]])
    }
  }
  
  # get unique values from EMD matrix. coerce to dataframe.
  ind <- which(upper.tri(tmpdat, diag = FALSE), arr.ind = TRUE)
  tmpdat.df <- data.frame(drvid=d,
                          crab_year_1 = tmpyrs[ind[,1]],
                          crab_year_2 = tmpyrs[ind[,2]],
                          emd=tmpdat[ind])
  
  if(i==1){
    emd_dat <- tmpdat.df
  } else{
    emd_dat %<>% rbind(tmpdat.df)
  }
  rm(ras_list,ras_list,tmpdat,ind,tmpdat.df)
  message(d, " done.\n")
  Sys.time()-sstart
  
}
```

```{r}
write.csv(emd_dat,here::here('project-dat','statistics',paste0('EMD_annual_per_vessel_',udsize*100,'ud_',kernel_g,'grid_', crab_years[1],'-',substr(tail(crab_years,1),3,4), 'crabYrs.csv')))

```
<br>