---
title: "VMS Main Text Figures: Map Gridded Fishing Activity"
author: "M. Fisher"
date: "7/3/2023"
output: html_document
---

## Description

Create Figure 3 for main text. Figure 3 will show fishing activity for (a) small local vessels, (b) large local vessels, and (c) the large vessel cross-district pool. Fishing activity will be represented by a heat map on 5x5 grids to maintain confidentiality. 

This script is based on Owen Liu's `match_to_5x5_grid.Rmd` in github.com/owenrliu/VMS-NWFSC-2. 


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(magrittr)
library(sf)
library(here)
library(knitr)
library(cowplot)
library(ggrepel)

source(here('R','map_fishing_grid.R'))
source(here('R','map_fishing_grid10.R'))
source(here('R','map_fishing_grid15.R'))
```

User inputs
```{r}
keep_years <- seq(2013,2018)
vmsdir      <- here('project-dat','vms','interpolation_60min','NaN_speed_filter')
metadir    <- here('project-dat')


## change the resolution of the map figure
png_res = 300 # suggested:200
## offset of each coastline from the others; this should change with res
offset = 2.5  # suggested:2.5
## size of port group labels on maps; this should change with res
pg_label_size = 4  # suggested:4
```



## Data

VMS
```{r read_vms}
for(y in keep_years){
  tmpvms <-  read_rds(here::here(vmsdir,paste0(y,'season_crabfishing.rds')))
  if(y==keep_years[1]){
    vms <- tmpvms
  } else{
    vms %<>% bind_rows(tmpvms)
  }
}

str(vms)
```


5x5 grid shapefile
```{r read_gridshp}
grd <- read_sf(here::here('../data','raw','grid','regions_master_final_lamb.shp'))
names(grd)
```


10x10 grid shapefile
```{r read_gridshp10}
grd10 <- read_sf(here::here('../data','raw','grid','grid_10k','tenkm_grid_polys_shore_lamb_ClipLayer.shp'))
names(grd10)
```

port group points
```{r}
pg <- read.csv(here::here(metadir,'pcgroup_mean_coordinates.csv'))
pg <- read.csv(here::here('project-dat','pcgroup_mean_coordinates.csv'))
pg$port_group <- c("Bodega Bay","Fort Bragg","Crescent City","Eureka","Monterey","Morro Bay","San Francisco")
pg_as_sf <- pg %>% st_as_sf(coords=c("Lon", "Lat"),crs=4326)
# proj4string(pg) <- CRS("+init=epsg:4326") # WGS 84
# pg_geo <- spTransform(pg, CRS("+init=epsg:32610"))
pg_df <- st_drop_geometry(pg_as_sf) %>%
  as.data.frame() %>%
  mutate(Lat=st_coordinates(pg_as_sf)[,2],
         Lon=st_coordinates(pg_as_sf)[,1])


pg_df$port_group_label <- pg_df$port_group_name
pg_df$port_group_label[which(pg_df$port_group_label=="Morro Bay")] <- "Morro\nBay"
```

season start dates
```{r}
season_dates_df <- read_csv(here::here(metadir,"dcrb_season_starts.csv")) %>%
  mutate(season_start_date=mdy(season_start_date),
         season_end_date=ifelse(pcdistrict=="northern",paste0("7/15/",crab_year),
                                paste0("6/30/",crab_year))) %>%
  mutate(season_end_date=mdy(season_end_date))
```


Clustering groups
```{r}
vgroups <- read_csv(here('project-dat','vessel_groups','k4_NaNspeedfilter_2014clusterYR_byArea_noncon-group-key.csv'))

vgroup_summary <- vgroups %>% dplyr::select(group,subgroup,group_area, vessel_size) %>% distinct()
vgroup_annual_summary <- vgroups %>% dplyr::select(subgroup,group_area,vessel_size,crab_year) %>% 
  distinct() %>% arrange(subgroup)
```

Basemap
```{r create_basemap, fig.height=3, fig.width=2}
data(stateMapEnv)
states_df <- map_data("state") %>%
  filter(region %in% c("california"))
states_df_sp <- st_as_sf(states_df, coords=c("long","lat"), crs=4326)
# coordinates(states_df_sp) <- c("long", "lat")
# proj4string(states_df_sp) <- CRS("+init=epsg:4326") # WGS 84
# states_geo <- spTransform(states_df_sp, CRS("+init=epsg:32610"))

states_df_coast <- filter(states_df,long < -116) %>%
  filter(!(long > -120.5 & lat > 32.5)) %>%
  filter(!(long > -122.5 & lat > 38.1))

# check the outline
ggplot(data=states_df_coast, aes(x=long,y=lat,group=group)) + geom_path()
```

## Prep VMS Spatial Data

get only VMS records for clustered vessels

```{r}
dim(vms)
vms_sub <- vms %>% left_join(vgroups %>% filter(group %in% c(1,2,5)),by=c("drvid","crab_year")) %>%
  filter(!is.na(subgroup))
dim(vms_sub)
```

convert VMS to spatial object

```{r}
pt <- proc.time()
vms_all_sf <- vms_sub %>%
  st_as_sf(coords=c('LONGITUDE','LATITUDE'),crs=4326) %>%
  # then, convert to planar projection to match the grid
  st_transform(st_crs(grd))
x<-proc.time()-pt
```


Took `r round(x[3]/60,2)` minutes to do the conversion. Now for the join...



## Match Points to Grid


### 5x5km

```{r join to grid}
pt <- proc.time()
vms_all_grd_match <- vms_all_sf %>%
  st_join(grd)
x<-proc.time()-pt
```

The join took `r round(x[3]/60,2)` minutes.

```{r}
glimpse(vms_all_grd_match)
```

convert to wsg
```{r}
vms_all_grd_match_wgs84 <- st_transform(vms_all_grd_match, crs=st_crs(pg_as_sf))
```

how many records didn't match to a grid cell?
```{r}
sum(is.na(vms_all_grd_match$GRID5KM_ID))/length(vms_all_grd_match$GRID5KM_ID)
```

### 10x10km

```{r join to grid}
pt <- proc.time()
vms_all_grd10_match <- vms_all_sf %>%
  st_join(grd10)
x<-proc.time()-pt
```

The join took `r round(x[3]/60,2)` minutes.

```{r}
glimpse(vms_all_grd10_match)
```

convert to wsg
```{r}
vms_all_grd10_match_wgs84 <- st_transform(vms_all_grd10_match, crs=st_crs(pg_as_sf))
```

how many records didn't match to a grid cell?
```{r}
sum(is.na(vms_all_grd10_match$FID_))/length(vms_all_grd10_match$FID_)
```

## Clear confidential grids

### 5x5
Calculate # vessels represented per group, crab year, and grid cell. Save any grids with fewer than three vessels.
```{r}
vms_grd_confidential <- as.data.frame(vms_all_grd_match) %>%
  group_by(crab_year, subgroup, GRID5KM_ID) %>%
  summarise(nvessels=length(unique(drvid))) %>%
  filter(nvessels < 3)
```

Remove confidential grids from sf file
```{r}
vms_noncon_grd_match <- st_drop_geometry(vms_all_grd_match) %>%
  as.data.frame() %>%
  mutate(LATITUDE=st_coordinates(vms_all_grd_match_wgs84)[,2],
         LONGITUDE=st_coordinates(vms_all_grd_match_wgs84)[,1]) %>%
  anti_join(vms_grd_confidential)

dim(vms_all_grd_match)
dim(vms_noncon_grd_match)
```


### 10x10
Calculate # vessels represented per group, crab year, and grid cell. Save any grids with fewer than three vessels.
```{r}
vms_grd10_confidential <- as.data.frame(vms_all_grd10_match) %>%
  group_by(crab_year, subgroup, FID_) %>%
  summarise(nvessels=length(unique(drvid))) %>%
  filter(nvessels < 3)
```

Remove confidential grids from sf file
```{r}
vms_noncon_grd10_match <- st_drop_geometry(vms_all_grd10_match) %>%
  as.data.frame() %>%
  mutate(LATITUDE=st_coordinates(vms_all_grd10_match_wgs84)[,2],
         LONGITUDE=st_coordinates(vms_all_grd10_match_wgs84)[,1]) %>%
  anti_join(vms_grd10_confidential)

dim(vms_all_grd10_match)
dim(vms_noncon_grd10_match)
```



### Save and Write

```{r}
# non-spatial version
vms_noncon_grd_match %>% 
  write_rds(here::here('project-dat','vessel_groups','vms_noncon_interpolated_k4groups_2014clusterYR_5x5grd.rds'))
```

```{r}
# non-spatial version
vms_noncon_grd10_match %>% 
  write_rds(here::here('project-dat','vessel_groups','vms_noncon_interpolated_k4groups_2014clusterYR_10x10grd.rds'))
```


## Map

### 5x5km grid

#### grid values

```{r}
vms_by_grd <- vms_noncon_grd_match %>%
  group_by(crab_year,subgroup,vessel_size, GRID5KM_ID) %>%
  summarise(n.vms=n(),
            n.vessels=length(unique(drvid)))
```

```{r}
grd_wgs84 <- st_transform(grd,crs=4326)
grd_wgs84 <- grd_wgs84[grd_wgs84$GRID5KM_ID %in% vms_by_grd$GRID5KM_ID,]
```

```{r}
vms_by_grd <- left_join(vms_by_grd, grd_wgs84,by="GRID5KM_ID")
vms_by_grd_sf <- st_as_sf(vms_by_grd)
```

#### plot n vms

1 Small
```{r}
checkvms <- vms_by_grd_sf[vms_by_grd_sf$subgroup=="1-small" & !is.na(vms_by_grd_sf$GRID5KM_ID),]
keep_years.a <- keep_years[which(keep_years %in% checkvms$crab_year)]

myplot_list.a <- map_fishing_grid(vms_sf=vms_by_grd_sf, keep_group="1-small",
                                keep_years=keep_years.a,
                                states_df_coast=states_df_coast,
                                max_vms=NA)
```
```{r}
png(here::here('project-dat','figures',"change_5x5grid_annual_1Small.png"),res=300, height=1500, width=2000)
plot_grid(plotlist=myplot_list.a,nrow=1,rel_widths=c(2,0.9,0.9,3.4), labels=c("a"))
dev.off()
```


2 Small 
```{r}
checkvms <- vms_by_grd_sf[vms_by_grd_sf$subgroup=="2-small",]
myplot_list.b <- map_fishing_grid(vms_sf=vms_by_grd_sf, keep_group="2-small",
                                keep_years=keep_years,
                                states_df_coast=states_df_coast,
                                max_vms=2000)
myplot_list.a <- myplot_list.a[which(!is.null(myplot_list.a))]
```
```{r}
png(here::here('project-dat','figures',"change_5x5grid_annual_2Small.png"),res=300, height=1500, width=2500)
plot_grid(plotlist=myplot_list.b,nrow=1,rel_widths=c(2,0.9,0.9,0.9,0.9,3.3), labels=c("b"))
dev.off()
```




1 Large 
```{r}
checkvms <- vms_by_grd_sf[vms_by_grd_sf$subgroup=="1-large",]
myplot_list.c <- map_fishing_grid(vms_sf=vms_by_grd_sf, keep_group="1-large",
                                keep_years=keep_years,
                                states_df_coast=states_df_coast,
                                max_vms=1000)
```

```{r}
png(here::here('project-dat','figures',"change_5x5grid_annual_1Large.png"),res=300, height=1500, width=2500)
plot_grid(plotlist=myplot_list.c,nrow=1,rel_widths=c(2,0.9,0.9,0.9,0.9,3.3), labels=c("c"))
dev.off()
```

2 Large
```{r}
myplot_list.d <- map_fishing_grid(vms_sf=vms_by_grd_sf, keep_group="2-large",
                                keep_years=keep_years,
                                states_df_coast=states_df_coast,
                                max_vms=1000)
```

```{r}
png(here::here('project-dat','figures',"change_5x5grid_annual_2Large.png"),res=300, height=1500, width=2300)
plot_grid(plotlist=myplot_list.d,nrow=1,rel_widths=c(1.6,0.8,0.8,0.8,0.8,3.3), labels=c("d"))
dev.off()
```


5 Large
```{r}
myplot_list.e <- map_fishing_grid(vms_sf=vms_by_grd_sf, keep_group="5-large",
                                keep_years=keep_years,
                                states_df_coast=states_df_coast,
                                max_vms=NA)
```

```{r}
png(here::here('project-dat','figures',"change_5x5grid_annual_5Large.png"),res=300, height=1500, width=2300)
plot_grid(plotlist=myplot_list.e,nrow=1,rel_widths=c(1.6,0.8,0.8,0.8,0.8,3.3), labels=c("e"))
dev.off()
```

### 10x10km grid

```{r}
vms_noncon_grd10_match <- read_rds(here::here('project-dat','vessel_groups','vms_noncon_interpolated_k4groups_2014clusterYR_10x10grd.rds'))
```


Let's test out a map for one year, one group, to make sure everything was calculated and/or saved properly
```{r}
g  <- "2-large"
yr <- "2015"

vms_noncon_grd10_match_sf <- st_as_sf(vms_noncon_grd10_match,coords=c('LONGITUDE','LATITUDE'),crs=4326)
map1_dat <- vms_noncon_grd10_match_sf %>% filter(subgroup==g & crab_year==yr)

ggplot() +
  geom_path(data=states_df_coast, aes(x=long,y=lat)) +
  geom_point(data=pg_df, aes(x=Lon, y=Lat), color="red") +
  geom_sf(data=map1_dat)
```


#### grid values

```{r}
vms_by_grd10 <- vms_noncon_grd10_match %>%
  group_by(crab_year,subgroup,vessel_size, FID_) %>%
  summarise(n.vms=n(),
            n.vessels=length(unique(drvid)))
```

```{r}
# save
vms_by_grd10 %>% 
  write_rds(here::here('project-dat','vessel_groups','vms_noncon_interpolated_k4groups_2014clusterYR_10x10grd_summary.rds'))
```

```{r}
grd10_wgs84 <- st_transform(grd10,crs=4326)
grd10_wgs84 <- grd10_wgs84[grd10_wgs84$FID_ %in% vms_by_grd10$FID_,]
```

```{r}
vms_by_grd10 <- left_join(vms_by_grd10, grd10_wgs84,by="FID_")
vms_by_grd10_sf <- st_as_sf(vms_by_grd10)
```

#### plot n vms

2 Small 
```{r}
checkvms <- vms_by_grd10_sf[vms_by_grd10_sf$subgroup=="2-small",]; unique(checkvms$crab_year)
myplot10_list.b <- map_fishing_grid10(vms_sf=vms_by_grd10_sf, keep_group="2-small",
                                keep_years=keep_years,
                                states_df_coast=states_df_coast,
                                max_vms=2000)
```
```{r}
png(here::here('project-dat','figures',"change_10x10grid_annual_2Small.png"),res=300, height=1500, width=2550)
plot_grid(plotlist=myplot10_list.b,nrow=1,rel_widths=c(1.4,0.8,0.8,0.8,0.8,3), labels=c("b"))
dev.off()
```



1 Large 
```{r}
checkvms <- vms_by_grd10_sf[vms_by_grd10_sf$subgroup=="1-large",]
myplot_list.c <- map_fishing_grid10(vms_sf=vms_by_grd10_sf, keep_group="1-large",
                                keep_years=keep_years,
                                states_df_coast=states_df_coast,
                                max_vms=1000)
```

```{r}
png(here::here('project-dat','figures',"change_10x10grid_annual_1Large.png"),res=300, height=1500, width=2500)
plot_grid(plotlist=myplot_list.c,nrow=1,rel_widths=c(1.4,0.8,0.8,0.8,0.8,3), labels=c("c"))
dev.off()
```

2 Large
```{r}
myplot_list.d <- map_fishing_grid10(vms_sf=vms_by_grd10_sf, keep_group="2-large",
                                keep_years=keep_years,
                                states_df_coast=states_df_coast,
                                max_vms=1000)
```

```{r}
png(here::here('project-dat','figures',"change_10x10grid_annual_2Large.png"),res=300, height=1500, width=2500)
plot_grid(plotlist=myplot_list.d,nrow=1,rel_widths=c(1.4,0.8,0.8,0.8,0.8,3), labels=c("d"))
dev.off()
```


5 Large
```{r}
myplot_list.e <- map_fishing_grid10(vms_sf=vms_by_grd10_sf, keep_group="5-large",
                                keep_years=keep_years,
                                states_df_coast=states_df_coast,
                                max_vms=NA)
```

```{r}
png(here::here('project-dat','figures',"change_10x10grid_annual_5Large.png"),res=300, height=1500, width=2500)
plot_grid(plotlist=myplot_list.e,nrow=1,rel_widths=c(1.4,0.8,0.8,0.8,0.8,3), labels=c("e"))
dev.off()
```


## Map vessels confidential
```{r}
vms_all_by_grd <- st_drop_geometry(vms_all_grd_match) %>%
  as.data.frame() %>%
  mutate(LATITUDE=st_coordinates(vms_all_grd_match_wgs84)[,2],
         LONGITUDE=st_coordinates(vms_all_grd_match_wgs84)[,1]) %>%
  group_by(crab_year,subgroup,vessel_size, GRID5KM_ID) %>%
  summarise(n.vms=n(),
            n.vessels=length(unique(drvid)))
```

```{r}
grd_wgs84 <- st_transform(grd,crs=4326)
grd_all_wgs84 <- grd_wgs84[grd_wgs84$GRID5KM_ID %in% vms_all_by_grd$GRID5KM_ID,]
```

```{r}
vms_all_by_grd <- left_join(vms_all_by_grd, grd_all_wgs84,by="GRID5KM_ID")
vms_all_by_grd_sf <- st_as_sf(vms_all_by_grd)
```

1 Small
```{r}
checkvms <- vms_all_by_grd_sf[vms_all_by_grd_sf$subgroup=="1-small" & !is.na(vms_all_by_grd_sf$GRID5KM_ID),]
keep_years.a <- keep_years[which(keep_years %in% checkvms$crab_year)]

myplot_list.a2 <- map_fishing_grid(vms_sf=vms_all_by_grd_sf, metric="vessels", keep_group="1-small",
                                keep_years=keep_years.a,
                                states_df_coast=states_df_coast,
                                max_vms=NA)

png(here::here('project-dat','figures',"change_5x5grid_annual_1Small_nvessels_CONFIDENTIAL.png"),res=300, height=1500, width=2800)
plot_grid(plotlist=myplot_list.a2,nrow=1,rel_widths=c(2,0.9,0.9,0.9,0.9,3.4), labels=c("a"))
dev.off()
```

2 Small
```{r}
checkvms <- vms_all_by_grd_sf[vms_all_by_grd_sf$subgroup=="2-small" & !is.na(vms_all_by_grd_sf$GRID5KM_ID),]
keep_years.a <- keep_years[which(keep_years %in% checkvms$crab_year)]

myplot_list.b2 <- map_fishing_grid(vms_sf=vms_all_by_grd_sf, metric="vessels", keep_group="2-small",
                                keep_years=keep_years.a,
                                states_df_coast=states_df_coast,
                                max_vms=NA)

png(here::here('project-dat','figures',"change_5x5grid_annual_2Small_nvessels_CONFIDENTIAL.png"),res=300, height=1500, width=2800)
plot_grid(plotlist=myplot_list.b2,nrow=1,rel_widths=c(2,0.9,0.9,0.9,0.9,3.4), labels=c("a"))
dev.off()
```
5 Large
```{r}
checkvms <- vms_all_by_grd_sf[vms_all_by_grd_sf$subgroup=="5-large" & !is.na(vms_all_by_grd_sf$GRID5KM_ID),]
keep_years.a <- keep_years[which(keep_years %in% checkvms$crab_year)]

myplot_list.e2 <- map_fishing_grid(vms_sf=vms_all_by_grd_sf, metric="vessels", keep_group="5-large",
                                keep_years=keep_years.a,
                                states_df_coast=states_df_coast,
                                max_vms=NA)

png(here::here('project-dat','figures',"change_5x5grid_annual_5Large_nvessels_CONFIDENTIAL.png"),res=300, height=1500, width=2800)
plot_grid(plotlist=myplot_list.e2,nrow=1,rel_widths=c(2,0.9,0.9,0.9,0.9,3.4), labels=c("e"))
dev.off()
```

#### troubleshooting 1-small

Where are the data points that weren't assigned a grid cell??
```{r}
g  <- "1-small"
yr <- "2016"

vms_noncon_grd10_match_sf <- st_as_sf(vms_noncon_grd10_match,coords=c('LONGITUDE','LATITUDE'),crs=4326)
map1_dat <- vms_noncon_grd10_match_sf %>% filter(subgroup==g & crab_year==yr)

ggplot() +
  geom_path(data=states_df_coast, aes(x=long,y=lat)) +
  geom_point(data=pg_df, aes(x=Lon, y=Lat), color="red") +
  geom_sf(data=map1_dat, aes(color=FID_)) + facet_wrap(~drvid) +
  coord_sf(ylim=c(37,39))
```

In 2016, there should be this many vessels in the 2-small group:
```{r}
vgroups %>% filter(subgroup=="1-small") %>%
  group_by(crab_year) %>% summarize(nv=length(unique(drvid)))
  
```

1 Small
```{r}
checkvms <- vms_all_by_grd_sf[vms_all_by_grd_sf$subgroup=="1-small" & !is.na(vms_all_by_grd_sf$GRID5KM_ID),]
keep_years.a <- keep_years[which(keep_years %in% checkvms$crab_year)]

myplot_list.a2 <- map_fishing_grid(vms_sf=vms_all_by_grd_sf, keep_group="1-small",
                                keep_years=keep_years.a,
                                states_df_coast=states_df_coast,
                                max_vms=NA)

png(here::here('project-dat','figures',"change_5x5grid_annual_1Small_CONFIDENTIAL.png"),res=300, height=1500, width=2800)
plot_grid(plotlist=myplot_list.a2,nrow=1,rel_widths=c(2,0.9,0.9,0.9,0.9,3.4), labels=c("a"))
dev.off()
```



## 1-small

### Final Map

2016 & 2018 use 100km cells, or 15km if they work. All others use 10km grid

```{r}
vms_by_grd10 <- read_rds(here::here('project-dat','vessel_groups','vms_noncon_interpolated_k4groups_2014clusterYR_10x10grd_summary.rds'))

vms_by_grd100 <- read_rds(here::here('project-dat','vessel_groups','vms_noncon_interpolated_k4groups_2014clusterYR_100x100grd_summary.rds'))
```


```{r}
grd10_wgs84 <- st_transform(grd10,crs=4326)
grd10_wgs84 <- grd10_wgs84[grd10_wgs84$FID_ %in% vms_by_grd10$FID_,]
```

```{r}
vms_by_grd10 <- left_join(vms_by_grd10, grd10_wgs84,by="FID_")
vms_by_grd10_sf <- st_as_sf(vms_by_grd10)
```


Subset each data set and get max/min overall
```{r}
subvms10 <- vms_by_grd10_sf[vms_by_grd10_sf$crab_year %in% c(2013,2014,2015,2017),]
subvms10 <- subvms10[subvms10$subgroup=="1-small",]
subvms10 <- subvms10[!is.na(subvms10$FID_),]

subvms100 <- vms_by_grd100_sf[vms_by_grd100_sf$crab_year %in% c(2016,2018),]
subvms100 <- subvms100[subvms100$subgroup=="1-small",]
subvms100 <- subvms100[!is.na(subvms100$FID_),]

tmpmax <- max(subvms100$n.vms,subvms10$n.vms); tmpmin <- min(subvms100$n.vms,subvms10$n.vms)
```



Map 10x10km
```{r}
myplot10_list.a <- map_fishing_grid10(vms_sf=subvms10, keep_group="1-small",
                                keep_years=c(2013,2014,2015,2017),
                                states_df_coast=states_df_coast,
                                vms_scale_max=tmpmax,vms_scale_min=tmpmin,
                                max_vms=20000,print.legend=FALSE)

names(myplot10_list.a) <- c(2013,2014,2015,2017)
```


Map 100x100km
```{r}
myplot100_list.a <- map_fishing_grid10(vms_sf=subvms100, keep_group="1-small",
                                keep_years=c(2016,2018),
                                states_df_coast=states_df_coast,
                                vms_scale_max=tmpmax,vms_scale_min=tmpmin,
                                max_vms=20000,subset=TRUE)

names(myplot100_list.a) <- c(2016,2018)
```


combine and print
```{r}
myplotOUT_list.a <- list(myplot10_list.a[["2013"]],myplot10_list.a[["2014"]],myplot10_list.a[["2015"]],
                         myplot100_list.a[["2016"]],myplot10_list.a[["2017"]],myplot100_list.a[["2018"]])
```

```{r}
png(here::here('project-dat','figures',"change_CUSTOMgrid_annual_1Small.png"),res=300, height=1500, width=2550)
plot_grid(plotlist=myplotOUT_list.a,nrow=1,rel_widths=c(1.4,0.8,0.8,0.8,0.8,3), labels=c("a"))
dev.off()
```


## 5-large

### Final Map
2014,2015,2018 10x10km

2013,2016,2017 15x15km for now, but worth testing 25x25km to better capture northern district fishing in 2013,2017 and central district fishing in 2016. 

Subset each data set and get max/min overall
```{r}
subvms10 <- vms_by_grd10_sf[vms_by_grd10_sf$crab_year %in% c(2014,2015,2018),]
subvms10 <- subvms10[subvms10$subgroup=="5-large",]
subvms10 <- subvms10[!is.na(subvms10$FID_),]

subvms15 <- vms_by_grd15_sf[vms_by_grd15_sf$crab_year %in% c(2013,2016,2017),]
subvms15 <- subvms15[subvms15$subgroup=="5-large",]
subvms15 <- subvms15[!is.na(subvms15$BLOCK15_ID),]

tmpmax <- max(subvms10$n.vms,subvms15$n.vms); tmpmin <- min(subvms15$n.vms,subvms10$n.vms)
```



Map 10x10km
```{r}
myplot10_list.e <- map_fishing_grid10(vms_sf=subvms10, keep_group="5-large",
                                keep_years=c(2014,2015,2018),
                                states_df_coast=states_df_coast,
                                vms_scale_max=tmpmax,vms_scale_min=tmpmin,
                                max_vms=20000,subset=TRUE)

names(myplot10_list.e) <- c(2014,2015,2018)
```


Map 15x15km
```{r}
myplot15_list.e <- map_fishing_grid15(vms_sf=subvms15, keep_group="5-large",
                                keep_years=c(2013,2016,2017),
                                states_df_coast=states_df_coast,
                                vms_scale_max=tmpmax,vms_scale_min=tmpmin,
                                max_vms=20000,print.legend=FALSE)

names(myplot15_list.e) <- c(2013,2016,2017)
```


combine and print
```{r}
myplotOUT_list.e <- list(myplot15_list.e[["2013"]],myplot10_list.e[["2014"]],myplot10_list.e[["2015"]],
                         myplot15_list.e[["2016"]],myplot15_list.e[["2017"]],myplot10_list.e[["2018"]])
```

```{r}
png(here::here('project-dat','figures',"change_CUSTOMgrid_annual_5Large.png"),res=300, height=1500, width=2550)
plot_grid(plotlist=myplotOUT_list.e,nrow=1,rel_widths=c(1.6,0.8,0.8,0.8,0.8,3), labels=c("e"))
dev.off()
```





### Test: 15x15km

Read in grid 
```{r read_gridshp15}
grd15 <- read_sf(here::here('../data','raw','grid','grid_15k','fifteenkm_grid_polys_shore_lamb_ClipLayer.shp'))
names(grd15)
```


#### Match Points to Grid
```{r join to grid}
pt <- proc.time()
vms_all_grd15_match <- vms_all_sf %>%
  st_join(grd15)
x<-proc.time()-pt
```

The join took `r round(x[3]/60,2)` minutes.

```{r}
glimpse(vms_all_grd15_match)
```

convert to wsg
```{r}
vms_all_grd15_match_wgs84 <- st_transform(vms_all_grd15_match, crs=st_crs(pg_as_sf))
```

how many records didn't match to a grid cell?
```{r}
sum(is.na(vms_all_grd15_match$BLOCK15_ID))/length(vms_all_grd15_match$BLOCK15_ID)
```

#### Clear Confidential grids
Calculate # vessels represented per group, crab year, and grid cell. Save any grids with fewer than three vessels.
```{r}
vms_grd15_confidential <- as.data.frame(vms_all_grd15_match) %>%
  group_by(crab_year, subgroup, BLOCK15_ID) %>%
  summarise(nvessels=length(unique(drvid))) %>%
  filter(nvessels < 3)
```

Remove confidential grids from sf file
```{r}
vms_noncon_grd15_match <- st_drop_geometry(vms_all_grd15_match) %>%
  as.data.frame() %>%
  mutate(LATITUDE=st_coordinates(vms_all_grd15_match_wgs84)[,2],
         LONGITUDE=st_coordinates(vms_all_grd15_match_wgs84)[,1]) %>%
  anti_join(vms_grd15_confidential)

dim(vms_all_grd15_match)
dim(vms_noncon_grd15_match)
```

```{r}
# non-spatial version
vms_noncon_grd15_match %>% 
  write_rds(here::here('project-dat','vessel_groups','vms_noncon_interpolated_k4groups_2014clusterYR_15x15grd.rds'))
```


#### Map

```{r}
# non-spatial version
vms_noncon_grd15_match <- read_rds(here::here('project-dat','vessel_groups','vms_noncon_interpolated_k4groups_2014clusterYR_15x15grd.rds'))
```

Calculate grid values
```{r}
vms_by_grd15 <- vms_noncon_grd15_match %>%
  group_by(crab_year,subgroup,vessel_size, BLOCK15_ID) %>%
  summarise(n.vms=n(),
            n.vessels=length(unique(drvid)))
## check - no grids with fewer than 3 vessels, right?
filter(vms_by_grd15, n.vessels < 3)
```

Save non-confidential, summarized data for mapping
```{r eval=FALSE}
vms_by_grd15 %>%
  write_rds(here::here('project-dat','vessel_groups','vms_noncon_interpolated_k4groups_2014clusterYR_15x15grd_summary.rds'))
```

Add shape info
```{r}
grd15_wgs84 <- st_transform(grd15,crs=4326)
  grd15_wgs84 <- grd15_wgs84[grd15_wgs84$BLOCK15_ID %in% vms_by_grd15$BLOCK15_ID,]
```

```{r}
vms_by_grd15 <- left_join(vms_by_grd15, grd15_wgs84,by="BLOCK15_ID")
vms_by_grd15_sf <- st_as_sf(vms_by_grd15)
```

1 - small

```{r}
checkvms <- vms_by_grd15_sf[vms_by_grd15_sf$subgroup=="1-small" & !is.na(vms_by_grd15_sf$BLOCK15_ID),]
keep_years.a <- keep_years[which(keep_years %in% checkvms$crab_year)]

myplot15_list.a <- map_fishing_grid15(vms_sf=vms_by_grd15_sf, keep_group="1-small",
                                keep_years=keep_years.a,
                                states_df_coast=states_df_coast,
                                max_vms=NA)
```
```{r}
png(here::here('project-dat','figures',"change_15x15grid_annual_1Small.png"),res=300, height=1500, width=2200)
plot_grid(plotlist=myplot15_list.a,nrow=1,rel_widths=c(2,0.9,0.9,0.9,3.4), labels=c("a"))
dev.off()
```


5 - large
```{r}
checkvms <- vms_by_grd15_sf[vms_by_grd15_sf$subgroup=="5-large" & !is.na(vms_by_grd15_sf$BLOCK15_ID),]
keep_years.a <- keep_years[which(keep_years %in% checkvms$crab_year)]

myplot15_list.e <- map_fishing_grid15(vms_sf=vms_by_grd15_sf, keep_group="5-large",
                                keep_years=keep_years.a,
                                states_df_coast=states_df_coast,
                                max_vms=NA)
```
```{r}
png(here::here('project-dat','figures',"change_15x15grid_annual_5Large.png"),res=300, height=1500, width=2600)
plot_grid(plotlist=myplot15_list.e,nrow=1,rel_widths=c(2,0.9,0.9,0.9,0.9,3.4), labels=c("e"))
dev.off()
```





### Test: 100x100km

Read in grid 
```{r read_gridshp15}
grd100 <- read_sf(here::here('../data','raw','grid','grid_100k','hunkm_grid_polys_shore_lamb_ClipLayer.shp'))
names(grd100)
```


#### Match Points to Grid
```{r join to grid}
pt <- proc.time()
vms_all_grd100_match <- vms_all_sf %>%
  st_join(grd100)
x<-proc.time()-pt
```

The join took `r round(x[3]/60,2)` minutes.

```{r}
glimpse(vms_all_grd100_match)
```

convert to wsg
```{r}
vms_all_grd100_match_wgs84 <- st_transform(vms_all_grd100_match, crs=st_crs(pg_as_sf))
```

how many records didn't match to a grid cell?
```{r}
sum(is.na(vms_all_grd100_match$FID_))/length(vms_all_grd100_match$FID_)
```

#### Clear Confidential grids
Calculate # vessels represented per group, crab year, and grid cell. Save any grids with fewer than three vessels.
```{r}
vms_grd100_confidential <- as.data.frame(vms_all_grd100_match) %>%
  group_by(crab_year, subgroup, FID_) %>%
  summarise(nvessels=length(unique(drvid))) %>%
  filter(nvessels < 3)
```

Remove confidential grids from sf file
```{r}
vms_noncon_grd100_match <- st_drop_geometry(vms_all_grd100_match) %>%
  as.data.frame() %>%
  mutate(LATITUDE=st_coordinates(vms_all_grd100_match_wgs84)[,2],
         LONGITUDE=st_coordinates(vms_all_grd100_match_wgs84)[,1]) %>%
  anti_join(vms_grd100_confidential)

dim(vms_all_grd100_match)
dim(vms_noncon_grd100_match)
```

```{r}
# non-spatial version
vms_noncon_grd100_match %>% 
  write_rds(here::here('project-dat','vessel_groups','vms_noncon_interpolated_k4groups_2014clusterYR_100x100grd.rds'))
```


#### Map

```{r}
vms_noncon_grd100_match <-read_rds(here::here('project-dat','vessel_groups','vms_noncon_interpolated_k4groups_2014clusterYR_100x100grd.rds'))
```

Calculate grid values
```{r}
vms_by_grd100 <- vms_noncon_grd100_match %>%
  group_by(crab_year,subgroup,vessel_size, FID_) %>%
  summarise(n.vms=n(),
            n.vessels=length(unique(drvid)))

## check - no grids with fewer than 3 vessels, right?
filter(vms_by_grd100, n.vessels < 3)
```

Save non-confidential, summarized data for mapping
```{r eval=FALSE}
vms_by_grd100 %>%
  write_rds(here::here('project-dat','vessel_groups','vms_noncon_interpolated_k4groups_2014clusterYR_100x100grd_summary.rds'))
```


add shape info
```{r}
grd100_wgs84 <- st_transform(grd100,crs=4326)
  grd100_wgs84 <- grd100_wgs84[grd100_wgs84$FID_ %in% vms_by_grd100$FID_,]
```

```{r}
vms_by_grd100 <- left_join(vms_by_grd100, grd100_wgs84,by="FID_")
vms_by_grd100_sf <- st_as_sf(vms_by_grd100)
```


1 Small
```{r}
checkvms <- vms_by_grd100_sf[vms_by_grd100_sf$subgroup=="1-small" & !is.na(vms_by_grd100_sf$FID_),]
keep_years.a <- keep_years[which(keep_years %in% checkvms$crab_year)]

myplot100_list.a <- map_fishing_grid10(vms_sf=vms_by_grd100_sf, keep_group="1-small",
                                keep_years=keep_years.a,
                                states_df_coast=states_df_coast,
                                max_vms=NA)
```
```{r}
png(here::here('project-dat','figures',"change_100x100grid_annual_1Small.png"),res=300, height=1500, width=2700)
plot_grid(plotlist=myplot100_list.a,nrow=1,rel_widths=c(2,0.9,0.9,0.9,0.9,3.5), labels=c("a"))
dev.off()
```

5 Large
```{r}
checkvms <- vms_by_grd100_sf[vms_by_grd100_sf$subgroup=="5-large" & !is.na(vms_by_grd100_sf$FID_),]
keep_years.a <- keep_years[which(keep_years %in% checkvms$crab_year)]

myplot100_list.e <- map_fishing_grid10(vms_sf=vms_by_grd100_sf, keep_group="5-large",
                                keep_years=keep_years.a,
                                states_df_coast=states_df_coast,
                                max_vms=NA)
```
```{r}
png(here::here('project-dat','figures',"change_100x100grid_annual_5Large.png"),res=300, height=1500, width=2700)
plot_grid(plotlist=myplot100_list.e,nrow=1,rel_widths=c(2,0.9,0.9,0.9,0.9,3.5), labels=c("a"))
dev.off()
```

