---
title: 'QC: vessel representation'
author: "M. Fisher"
date: "6/23/2023"
output: html_document
---

## Description

Get some summary statistics of what vessels and fishing trips are / are not represented in the VMS data. Some of this script borrows from `vms_representation_index.Rmd` by Owen Liu.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(magrittr)
library(here)
library(grid); library(PNWColors)
```


User inputs
```{r}
keep_years <- seq(2013,2018)

tixdir <- '../data/processed/fish tickets'
vmsdir <- '../data/processed/matched/interpolation_60min'
datdir <- 'project-dat/vms/interpolation_60min/NaN_speed_filter'

```


## Data

The fish tickets
```{r read_fishtix}
for(y in c(keep_years, (last(keep_years)+1))){
  tmpy <- y-1
  tmptix <-  read_rds(here::here(tixdir,paste0(tmpy,'fishtix_vlengths.rds')))
  if(y==(keep_years[1]-1)){
    tix <- tmptix
  } else{
    tix %<>% bind_rows(tmptix)
  }
}

str(tix)

## add crab year, day of season, filter for ca commercial Dungeness crab landings
tix %<>% filter(agency_code == "C") %>%
  filter(TARGET_lbs == "DCRB") %>%
  filter(removal_type_code=="C" | removal_type_code=="D" | removal_type_code=="U") %>%
  # grab subset of columns
  dplyr::select(Rec_ID,date, month, year, drvid, pacfin_port_code, port_group_code, agency_code, 
                removal_type_code, removal_type_name, TARGET_lbs,
                DCRB_lbs,DCRB_revenue,
                FINAL_LENGTH) %>%
  # add on crab season based on month of year, and filter
  # mutate(date=as_date(date, c("ymd","mdy")),
  #        year=year(date)) %>%
  mutate(year=year(date)) %>%
  mutate(crab_season=case_when(
    month(date)>10 ~ paste0(year,"-",year+1),
    month(date)<=10 ~ paste0(year-1,"-",year)),
    crab_year=case_when(
      month(date)>10 ~ year+1,
      month(date)<=10 ~ year)) %>%
  filter(crab_year %in% keep_years)
```


The interpolated VMS data
```{r}
for(y in c(keep_years, (last(keep_years)+1))){
  tmpy <- y-1
  tmpvms <-  read_rds(here::here(vmsdir,paste0(y,'interpolated.rds')))
  if(y==(keep_years[1]-1)){
    vms <- read_rds(here::here(vmsdir,paste0(y-1,"interpolated.rds"))) %>%
      bind_rows(tmpvms)
  } else{
    vms %<>% bind_rows(tmpvms)
  }
}

## add crab year, day of season; filter for california commercial Dungeness crab landings
vms %<>% filter(agency_code == "C") %>%
  filter(TARGET_lbs == "DCRB") %>%
  filter(removal_type_code=="C" | removal_type_code=="D" | removal_type_code=="U") %>%
  # grab subset of columns
  dplyr::select(Rec_ID,date, pacfin_port_code, port_group_code, agency_code, 
                removal_type_code, removal_type_name, TARGET_lbs,
                VMS_RECNO,VESSEL_NAM,drvid,UTCDATETIM,westcoastdate,westcoastdate_notime,
                X_COORD,Y_COORD, LATITUDE,LONGITUDE,NGDC_M,
                AVG_SPEED,avg_speed_recalc,AVG_COURSE,DECLARATIO, in_port, is_interpolated,
                DCRB_lbs,DCRB_revenue) %>%
  # add on crab season based on month of year, and filter
  # mutate(date=as_date(date, c("ymd","mdy")),
  #        year=year(date)) %>%
  mutate(year=year(date)) %>%
  mutate(crab_season=case_when(
    month(date)>10 ~ paste0(year,"-",year+1),
    month(date)<=10 ~ paste0(year-1,"-",year)),
    crab_year=case_when(
      month(date)>10 ~ year+1,
      month(date)<=10 ~ year)) %>%
  filter(crab_year %in% keep_years)
```


The interpolated VMS data that was filtered for fishing locations
```{r}
for(y in keep_years){
  tmpdat <- read_rds(here::here(datdir,paste0(y,'season_crabfishing.rds')))
  if(y==keep_years[1]){dat <- tmpdat} else{dat %<>% bind_rows(tmpdat)}
}
```

## Ex-vessel revenue

What was the average ex-vessel revenue from commercial fishing trips targeting Dungeness crab, across the whole VMS data set?

```{r}
mean(vms$DCRB_revenue)
sd(vms$DCRB_revenue)
min(vms$DCRB_revenue)
min(filter(vms,DCRB_revenue>0)$DCRB_revenue)
```


## Proportion of vessels / trips represented

What proportion of California commercial Dungeness crab *fishing trips* are present in the **VMS** data, by year and overall?
```{r} 
vms_tickets <- dplyr::select(vms, Rec_ID, date, agency_code, drvid) %>% mutate(has_vms=1)
tix2vms <- tix %>% left_join(vms_tickets) %>% 
  mutate(has_vms=replace_na(has_vms,0))
```
```{r count trips}
ntrips_by_year <- tix2vms %>% 
  mutate(has_vms=ifelse(has_vms==0,"no_vms","has_vms")) %>%
  group_by(crab_year,crab_season,agency_code,has_vms) %>% 
  summarise(ntrips=n_distinct(Rec_ID)) %>%
  pivot_wider(names_from=has_vms,values_from=ntrips) %>%
  mutate(total_trips=has_vms+no_vms,
         percent_vms=has_vms/total_trips*100)
ntrips_by_year
```


What proportion of California commercial Dungeness crab *vessels* are present in the **VMS** data, by year and overall?
```{r count vessels}
nvessels_by_year <- tix2vms %>% 
  mutate(has_vms=ifelse(has_vms==0,"no_vms","has_vms")) %>%
  group_by(crab_year,crab_season,agency_code,has_vms) %>% 
  summarise(nvessels=n_distinct(drvid)) %>%
  pivot_wider(names_from=has_vms,values_from=nvessels) %>%
  mutate(total_vessels=has_vms+no_vms,
         percent_vms=has_vms/total_vessels*100)
nvessels_by_year
```



What proportion of California commercial Dungeness crab *fishing trips* are present in the **fishing locations** data, by year and overall?
```{r}
```
```{r count trips2}

```





