---
title: "EMD by group"
author: "M. Fisher"
date: "2/20/2023"
output: html_document
---

Calculate the year-over-year Earth Mover's Distance for Dungeness crab fishing grounds **using only VMS data from shared dates between seasons** (difference from script 04a). This means that EMD for the season will only be calculated using data collected on the dates of the shortest season (2016 -- spring / summer crabbing locations). 

Then, summarize EMD for vessel groups.



This script includes options to remove Dungeness crab vessels based on (1) the number of fishing trips they took in a given year, and/or (2) an exvessel revenue cut-off. This is in addition to the filters that were already imposed in creating the fishing ground data: 

- landings in a given state (e.g., California)
- trips targeting Dungeness crab
- vessels that have five or more relocations per crab season


However, for the analysis as it stands, I am not going to apply any additional filters beyond those already used in script 01.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)


library(tidyverse)
library(magrittr)
library(here)
library(raster)
library(adehabitatHR)
library(move)
library(pryr)
library(plotrix)
```


User inputs
```{r}
crab_years <- seq(2011,2018)
keep_years <- seq(2013,2018)
kernel_g <- 75
udsize       <- 0.90
cut_length <- 40  # all vessels below this length are "small"

save_SPDF <- TRUE # save an SPDF for each vessel, in addition to the 90% UD? Want this if calculating BA overlap later.
write_data <- FALSE

indir   <- 'project-dat/vms/interpolation_60min/NaN_speed_filter'
outdir  <- 'project-dat/vms/interpolation_60min/NaN_speed_filter/vessel90ud/spring-summer'
statdir <- 'project-dat/statistics'

filter=FALSE
filter_trips=5
filter_revenue=5000
```

## Data

VMS
```{r read_vms}
for(y in crab_years){
  tmpvms <-  read_rds(here::here(indir,paste0(y,'season_crabfishing.rds')))
  if(y==crab_years[1]){
    vms <- tmpvms
  } else{
    vms %<>% bind_rows(tmpvms)
  }
}

str(vms)
```


Fish tickets
```{r read_fishtix}
for(y in crab_years){
  tmpy <- y-1
  tmptix <-  read_rds(here::here('../data','processed','fish tickets',paste0(tmpy,'fishtix_vlengths.rds'))) %>%
  dplyr::select(Rec_ID,drvid,year,FINAL_LENGTH) %>% distinct()
  if(y==crab_years[1]){
    tix <- tmptix
  } else{
    tix %<>% bind_rows(tmptix)
  }
}

str(tix)
```

Clustering groups
```{r}
vgroups <- read_csv(here('project-dat','vessel_groups','k4_NaNspeedfilter_2014clusterYR_byArea_noncon-group-key.csv'))
vgroups_summary <- vgroups %>% 
              dplyr::select(group,subgroup,vessel_size,group_area,crab_year) %>% 
              distinct()
unique(vgroups$subgroup)
```

Vessel length keys
```{r}
vlengths <- read_csv(here('project-dat','vessel_groups','length_key_dcrb_vms_vessels.csv'))
```


Season dates
```{r}
season_dates_df <- read_csv(here::here('project-dat','dcrb_season_starts.csv'))

season_dates_df %<>%
  mutate(port_area_start_date = mdy(season_start_date)) %>%
  dplyr::select(-season_start_date)
str(season_dates_df)
```
<br>

For now, just grab the first (central) start date from the season with the longest delay
```{r}
start_date <- season_dates_df %>% filter(pcdistrict=="central") %>%
  slice_min(n=1,order_by=month(port_area_start_date)) %>%
  pull(port_area_start_date)
```


## Earth Mover's Distance per vessel

The Earth Mover's Distance calculation takes a utilization distribution as: A Raster, RasterStack, RasterBrick, SpatialPoints, SpatialPointsDataFrame. This code first constructs the utilization distribution, and then saves it uses a Spatial Points DataFrame. That object can be called back in to calculate EMD.

### create utilization distribution

First, filter VMS to only include spring / summer records after the 2016 start date

```{r}
sub.vms <- vms %>% mutate(late_start_month=paste0(month(start_date),"-",day(start_date))) %>%
  mutate(late_start=mdy(paste0(late_start_month,"-",crab_year))) %>%
  filter(date > late_start)
```

Make sure that worked
```{r}
hist(month(sub.vms$date), main="Tickets per month", xlab="Month")
```


```{r}
for(d in unique(sub.vms$drvid)){
  # filter VMS data
  tmpvms <- filter(sub.vms, crab_year %in% crab_years & drvid == d) %>%
    dplyr::select(drvid,crab_year,X_COORD, Y_COORD)
  # create the Spatial Points Data frame
  tmpvms.sp <- SpatialPointsDataFrame(coords=tmpvms %>% dplyr::select(X_COORD,Y_COORD), data=tmpvms %>% dplyr::select(crab_year), proj4string = CRS("+init=epsg:32610"))
  
  # save the list(by year) of spatial points dataframes as an RDS
  if(save_SPDF){
  write_rds(tmpvms.sp, here::here(outdir,paste0(d,"_SPDF_", crab_years[1],"-",substr(tail(crab_years,1),3,4), "crabYrs_AFTER-",paste0(month(start_date),"-",day(start_date)),".rds")))
  }
  
  tmpyrs <- unique(tmpvms.sp$crab_year)
  tmpout <- list()
  # create a list of raster layers
  for(j in seq(1,length(tmpyrs))){
    y <- tmpyrs[j]
    # create utilization distribution, with grid size of 'kernel_g'
    tmpud <- kernelUD(tmpvms.sp[tmpvms.sp$crab_year==y,], grid=kernel_g)
    # convert estUD object to a raster
    tmpras <- raster(as(tmpud[[1]],"SpatialPixelsDataFrame"))
    # set any points outside the 90% line to NA, then recalculate the probability surface
    tmpras[tmpras > udsize] <- NA; newras <- tmpras/cellStats(tmpras,sum)
    tmpout[[j]] <- newras
  }
  names(tmpout) <- tmpyrs
  
  # save the list(by year) of 90% UD rasters as an RDS
  if(write_data){
  write_rds(tmpout, here::here(outdir,paste0(d,"_",udsize*100,"ud_RasterLayer_",kernel_g,"grid_", crab_years[1],"-",substr(tail(crab_years,1),3,4), "crabYrs_AFTER-",
                                             paste0(month(start_date),"-",day(start_date)),".rds")))
  }
  
}
rm(tmpout, tmpvms,tmpvms.sp)
```
<br>

### calculate EMD 

Get filenames from above.
```{r}
file_vec <- list.files(path=here(outdir),pattern="RasterLayer")
```

Read back in the 90% UD lists, and find the year-over-year EMD for each vessel.
```{r eval=FALSE}
# for(i in seq(1,length(file_vec))){
for(i in seq(202,length(file_vec))){
  # get drvid & rds name
  f <- file_vec[[i]]
  d <- strsplit(f,split="_")[[1]][1]
  message(d)
  # read in rds
  ras_list <- read_rds(here::here(outdir,f))
  tmpyrs <- names(ras_list)
  if(length(tmpyrs) > 1){
    # empty data frame for EMD
    tmpdat <- matrix(nrow=length(tmpyrs),ncol=length(tmpyrs),dimnames=list(tmpyrs,tmpyrs),data=0)
    
    # get emd for consecutive years
    sstart <- Sys.time()
    message("calculating emd...")
    for(j in seq(1,length(ras_list)-1)){
      k=j+1
      message("working on ",names(ras_list)[j],"...")
        tmpdat[j,k] <- emd(ras_list[[j]],ras_list[[k]])
    }
    
    # get unique values from EMD matrix. coerce to dataframe.
    ind <- which(upper.tri(tmpdat, diag = FALSE), arr.ind = TRUE)
    tmpdat.df <- data.frame(drvid=d,
                            crab_year_1 = tmpyrs[ind[,1]],
                            crab_year_2 = tmpyrs[ind[,2]],
                            emd=tmpdat[ind]) %>%
      mutate(tmpdiff=as.numeric(crab_year_2)-as.numeric(crab_year_1)) %>%
      filter(tmpdiff == 1) %>% dplyr::select(-tmpdiff)
    message(d, " done.\n")
    print(Sys.time()-sstart)
    # 
    
    if(!(exists("emd_dat"))){
      emd_dat <- tmpdat.df
    } else{
      emd_dat %<>% rbind(tmpdat.df)
    }
    rm(ras_list,tmpdat,ind)
    
  }
}

if(write_data){
write.csv(emd_dat,here::here('project-dat','statistics',paste0('EMD_annual_per_vessel_',udsize*100,'ud_',kernel_g,'grid_', crab_years[1],'-',substr(tail(crab_years,1),3,4), 'crabYrs_AFTER-',
                                                               paste0(month(start_date),'-',day(start_date)),'.csv')))
}
```
<br>

######################################################
## EMD Group Means

### add vessel group info

add the vessel groups dataframe
```{r}
emd.dat.groups <- emd.dat %>% 
  left_join(vgroups %>% 
              dplyr::select(drvid,group,subgroup,vessel_size,group_area) %>% 
              distinct(),by=c("drvid"),multiple="all") %>%
  filter(!is.na(group))
```

combine crab years
```{r}
emd.dat.groups %<>% unite(col="crab_years",crab_year_1, crab_year_2, sep="_",remove=FALSE)
```

sample sizes for year=over=year comparisons 
```{r}
emd_groups_sample_sizes <-  emd.dat.groups %>%
  group_by(subgroup) %>% summarise(n.vessels=length(unique(drvid)))
filter(emd_groups_sample_sizes,n.vessels < 3) # should probably be 0

emd_groups_yearly_sample_sizes <-  emd.dat.groups %>%
  group_by(crab_years,subgroup) %>% summarise(n.vessels=length(unique(drvid)))
filter(emd_groups_yearly_sample_sizes,n.vessels < 3)
```


grab only the EMD values with non-confidential years for each vessel subgroup and year=over=year comparison (may not have been filtered out with non-confidential vessel key)
```{r}
emd.dat.groups.noncon <- emd.dat.groups %>% 
  left_join(emd_groups_yearly_sample_sizes,by=c("crab_years","subgroup")) %>%
  filter(n.vessels > 2)
```



### group means
```{r}
## vessel clusters
emd_group_means_overall <- emd.dat.groups %>%
  filter(crab_year_1 %in% keep_years & crab_year_2 %in% keep_years) %>%
  group_by(vessel_size,group,subgroup,group_area) %>%
  summarize(mean.emd=mean(emd),sd.emd=std.error(emd),n.vessels=length(unique(drvid)),.groups='drop') %>%
  ungroup() %>%
  mutate(`Vessel Group`=ifelse(group_area=="central","Central",
                         ifelse(group_area=="northern","Northern",
                                ifelse(group_area=="central/northern" & group==3 ,"Central/Northern",
                                       ifelse(group_area=="central/northern" & group==4, "Diversified Northern",
                                              ifelse(group_area=="central/northern" & group==5,"cross-district",NA))))),
         crab_years="Overall") %>%
  rename(`Vessel Size`=vessel_size)

emd_group_means <- emd.dat.groups.noncon %>%
  filter(crab_year_1 %in% keep_years & crab_year_2 %in% keep_years) %>%
  group_by(vessel_size,group,subgroup,group_area,crab_years) %>%
  summarize(mean.emd=mean(emd),sd.emd=std.error(emd),n.vessels=length(unique(drvid)),.groups='drop') %>%
  ungroup() %>%
  mutate(`Vessel Group`=ifelse(group_area=="central","Central",
                         ifelse(group_area=="northern","Northern",
                                ifelse(group_area=="central/northern" & group==3 ,"Central/Northern",
                                       ifelse(group_area=="central/northern" & group==4, "Diversified Northern",
                                              ifelse(group_area=="central/northern" & group==5,"cross-district",NA)))))) %>%
  rename(`Vessel Size`=vessel_size)
emd_group_means %<>%
  rbind(dplyr::select(emd_group_means_overall, colnames(emd_group_means)))
emd_group_means$`Vessel Group` <- factor(emd_group_means$`Vessel Group`,levels=c("Central","Northern","Central/Northern","Diversified Northern","cross-district"))
emd_group_means$crab_years <- factor(emd_group_means$crab_years, levels=c("Overall","2013_2014","2014_2015","2015_2016","2016_2017","2017_2018"))
```


```{r}
write_csv(emd_group_means, here('project-dat','statistics','meanEMD_annual_bygroup_90ud_75grid_k4_2014clusterYR_noncon.csv'))
```


### t test

Is there a significant difference in Earth Mover's Distance between large / small vessels in a given year?

#### local vessels

Compare EMD of large v. small Local Northern; and large v. small Local Central

```{r}
wilcox.local.lengths.out <- list()
sig.local.lengths.out <- data.frame(crab_years=as.character(),
                              pval=as.numeric())
emd.dat.local.groups <- filter(emd.dat.groups.noncon, group %in% c(1,2))
i=1
for(a in unique(emd.dat.local.groups$group_area)){
for(y in unique(emd.dat.local.groups$crab_years)){
  testdat <- emd.dat.groups %>% filter(crab_years==y & group_area==a)
  wilcox.local.lengths.out[[i]] <- wilcox.test(emd~vessel_size,testdat)
  sig.local.lengths.out %<>% bind_rows(data.frame(area=a,
                                                  crab_years=y,
                                                  pval=round(wilcox.local.lengths.out[[i]]$p.value,5)))
  names(wilcox.local.lengths.out)[[i]] <- paste0(a,"-",y)
  i = i +1
}
}
wilcox.local.lengths.out
```
```{r}
wilcox.local.lengths <- wilcox.local.lengths.out[c(seq(1,5),seq(8,12))]
sig.local.lengths.out %<>% mutate(correct.pval=pval*(length(wilcox.local.lengths)/2)) %>%
  mutate(significance=ifelse(correct.pval<0.05,"*","")) %>%
  mutate(y=(max(emd_group_means$mean.emd)+max(emd_group_means$sd.emd))/1000) %>% filter(crab_years %in% emd_group_means$crab_years)
```

again, on the overall data
```{r}
i=length(wilcox.local.lengths) + 1
for(a in unique(emd.dat.local.groups$group_area)){
  testdat <- emd.dat.groups %>% filter(crab_years %in% emd_group_means$crab_years & group_area==a)
  wilcox.local.lengths[[i]] <- wilcox.test(emd~vessel_size,testdat)
  names(wilcox.local.lengths)[[i]] <- paste0(a,"- Overall")
  
  sig.local.lengths.out %<>% bind_rows(data.frame(area=a,
                                                  crab_years="Overall",
                                                  pval=round(wilcox.local.lengths[[i]] $p.value,5)))
  i = i +1
}
```


Save output from all t-tests
```{r}
wilcox_tests_bylength <- list(wilcox.local.lengths.out,wilcox.local.lengths,sig.local.lengths.out)
save(wilcox_tests_bylength, file=here('project-dat','statistics','wilcox_tests_groups_localBYlength_noncon.rds'))
```



#### coastwide vessels

Compare EMD of groups 3-large versus 4-large.
```{r}
wilcox.coast.lengths.out <- list()
sig.coast.lengths.out <- data.frame(crab_years=as.character(),
                              pval=as.numeric())
emd.dat.coast.groups <- filter(emd.dat.groups.noncon, group %in% c(3,4))
i=1
for(y in unique(emd.dat.coast.groups$crab_years)){
  testdat <- emd.dat.coast.groups %>% filter(crab_years==y)
  if(length(unique(testdat$subgroup)) > 1){
  wilcox.coast.lengths.out[[i]] <- wilcox.test(emd~subgroup,testdat)
  sig.coast.lengths.out %<>% bind_rows(data.frame(crab_years=y,
                                                  pval=round(wilcox.coast.lengths.out[[i]]$p.value,5)))
  names(wilcox.coast.lengths.out)[[i]] <- paste0(y)
  i = i +1
  }
}
wilcox.coast.lengths.out
```
$`2011_2012`

	Wilcoxon rank sum exact test

data:  emd by subgroup
W = 1, p-value = 0.1143
alternative hypothesis: true location shift is not equal to 0


$`2014_2015`

	Wilcoxon rank sum exact test

data:  emd by subgroup
W = 10, p-value = 0.5714
alternative hypothesis: true location shift is not equal to 0


$`2017_2018`

	Wilcoxon rank sum exact test

data:  emd by subgroup
W = 2, p-value = 0.4
alternative hypothesis: true location shift is not equal to 0


Compare overall:
```{r}
testdat <- emd.dat.coast.groups %>% filter(crab_years %in% emd_group_means$crab_years)
wilcox.test(emd~subgroup,testdat)
```





### anova

Is there significant variability between years, within a single subgroup?

#### between years
```{r}
emd.dat.groups.keepYRS <- filter(emd.dat.groups.noncon, (crab_year_1 %in% keep_years & crab_year_2 %in% keep_years))
aov.years.out <- list()
tukey.years.out <- list()
i=1

for(s in unique(emd.dat.groups$subgroup)){
  testdat <- filter(emd.dat.groups.keepYRS, subgroup==s)
  aov.years.out[[i]] <- aov(emd~crab_years,data=testdat)
  tukey.years.out[[i]] <- TukeyHSD(aov.years.out[[i]],conf.level=0.90)
  names(aov.years.out)[[i]] <- paste0(s)
  names(tukey.years.out)[[i]] <- paste0(s)
  i=i+1
}
```


```{r}
anovas_list <- list(aov.years.out,tukey.years.out)
save(anovas_list, file=here('project-dat','statistics','anova_tukey_tests_crossYrs_byGroup.rds'))
```


#### between groups: original subgroups
are there significant differences between vessel groups, for EMD mean over all years?
```{r}
emd.dat.groups.keepYRS.og <- filter(emd.dat.groups.keepYRS, subgroup != "5-large")
aov.groups.out <- aov(emd~subgroup,data=emd.dat.groups.keepYRS.og)
summary(aov.groups.out)
tukey.groups.out <- TukeyHSD(aov.groups.out)
tukey.groups.out
```

```{r}
anovas_list <- list(aov.groups.out,tukey.groups.out)
save(anovas_list, file=here('project-dat','statistics','anova_tukey_tests_groups_byYR.rds'))
```

#### between groups: cross-district pool

are there significant differences between vessel groups, for EMD mean over all years?

```{r}
emd.dat.groups.keepYRS.pool <- filter(emd.dat.groups.keepYRS, !(subgroup %in% c("3-large","4-large")))
aov.groups.out <- aov(emd~subgroup,data=emd.dat.groups.keepYRS.pool)
summary(aov.groups.out)
tukey.groups.out <- TukeyHSD(aov.groups.out)
tukey.groups.out
```







