---
title: "Spatiotemporal Patterns in Vertical Lines"
output:
  html_document:
    toc: true
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r pkgs, include=FALSE,message=F,warning=FALSE}
library(raster)
library(tidyverse)
library(knitr)
library(extrafont)
library(ggsci)
library(lubridate)
library(sf)
library(geosphere)
library(magrittr)
library(rnaturalearth)
library(gstat)
library(data.table)
library(fasterize)
library(gganimate)
library(here)

# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=12,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
```

# Introduction

The purpose of this document is to use the outputs from a vertical line/crab pot simulation estimation to investigate preliminary spatial and temporal patterns in the estimated number of traps in the water. We will do this by using a spatial grid to aggregate total crab traps across space and time.

```{r import data}
# fp <- "C:/Users/Owen.Liu/Documents/NWFSC Research/Dungeness Fleet Dynamics/VMS DATA CONFIDENTIAL/Processed Data/VMS/"
# fp_tix <- "C:/Users/Owen.Liu/Documents/NWFSC Research/Dungeness Fleet Dynamics/VMS DATA CONFIDENTIAL/Processed Data/Fish tickets/"

# fish tickets
# load(file=paste0(fp_tix,"FishTix_all_2009_2018.RData"))
fishtix_matched_all <- read_rds("C:/Users/Owen.Liu/Documents/github/VMS-NWFSC-2/data/processed/fish tickets/fish_tickets_w_dayofseason.rds")
# filter for out of season?
fishtix_matched_all %<>% filter(ticket_day_of_season>=0)

# a coastline, for plotting
# coastline for plotting
coaststates <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('California','Oregon','Washington','Nevada'))
coastline <- ne_coastline(scale='medium',returnclass = 'sf') %>% 
  st_crop(st_bbox(coaststates))

# Load if not already loaded
simulated_traps <- read_rds(here::here('data','processed',"sim_traps_speed_depth_filters.rds")) %>% ungroup()

# traps locations only (discard VMS data for now), attached to fish ticket data
traplocs <- simulated_traps %>% dplyr::select(Rec_ID,traps) %>% 
  left_join(fishtix_matched_all,by="Rec_ID") %>% 
  dplyr::select(Rec_ID,date,year,month,ticket_day_of_season,pacfin_port_code,port_group_code,agency_code,drvid,FINAL_LENGTH,traps) %>% 
  distinct(Rec_ID,.keep_all = T) %>% 
  # count number of traps associated with each trip
  mutate(ntraps=purrr::map_int(traps,nrow))

# Vessel monthly weighting factors
vessel_month_allweights<-read_rds(here::here('data','processed',"vessel_month_all_weights.rds"))

# 5x5 grid shapefile
# grd <- read_sf(here::here("Input_Data","5x5 Grid","five_km_grid_polys_geo.shp"))
grd <- read_sf(here::here("data","grid","regions_master_final_lamb.shp"))
names(grd)

meta <- readxl::read_xlsx(here::here("data","grid","regions_master_final_lamb_metadata.xlsx")) 

meta %>% kable()
```

# Vessel Sizes by Tier

For the CA data, we have tier information- that is, we have data on how many traps each vessel is permitted to fish. Let's use these data to look at the breakdown of vessel sizes within these tiers.

```{r}
ca_tier <- read_rds(here::here('data','CAtier_drvid_matching_key.rds'))
glimpse(ca_tier)
ca_tier %<>% rename(ntraps_permitted=ntraps)

# assign crab season to tier info
CA_traplocs <- traplocs %>% filter(agency_code=="C") %>% left_join(ca_tier,by=c('year','drvid'))

library(viridis)
# vessel size histograms (vms boats)
CA_vessel_size_tier_vms <- CA_traplocs %>% 
  distinct(drvid,tier,ntraps_permitted,FINAL_LENGTH) %>% 
  drop_na() %>% 
  ggplot(aes(FINAL_LENGTH,fill=factor(ntraps_permitted)))+
  geom_density(alpha=0.5)+
  scale_fill_manual(values=viridis_pal()(7))+
  labs(x='Vessel Length',y='Kernel Density',fill="Tier",title="Vessel Size by Tier, VMS vessels")
CA_vessel_size_tier_vms
ggsave(here('plots','vessel_size_histogram_by_tier_vms.png'),CA_vessel_size_tier_vms,w=6,h=5)

# vessel size histogram (all CA vessels)
fishtix_tier <- fishtix_matched_all %>% filter(agency_code=="C") %>% left_join(ca_tier,by=c('year','drvid'))
CA_vessel_size_tier_all <- fishtix_tier %>% 
  distinct(drvid,tier,ntraps_permitted,FINAL_LENGTH) %>% 
  drop_na() %>% 
  ggplot(aes(FINAL_LENGTH,fill=factor(ntraps_permitted)))+
  geom_density(alpha=0.5)+
  scale_fill_manual(values=viridis_pal()(7))+
  labs(x='Vessel Length',y='Kernel Density',fill="Tier",title="Vessel Size by Tier, all Vessels")
CA_vessel_size_tier_all
ggsave(here('plots','vessel_size_histogram_by_tier_all.png'),CA_vessel_size_tier_all,w=6,h=5)
```

## Create Tier Representation by Month

We want to use the tier information to make a scaling key by measuring the relative representation of VMS vessels relative to overall DCRB vessels, by tier, year, month, and CA region.

```{r}
# include regional indicator for each port group
cenCA_ports <- c("MRA","SBA","SFA","MNA")
norCA_ports <- c("ERA","CCA","BGA","BDA","CA2")
CA_traplocs %<>%
  mutate(CA_region=case_when(
    port_group_code %in% cenCA_ports ~ "cenCA",
    port_group_code %in% norCA_ports ~ "norCA"
  ))
CA_vms_vessels <- CA_traplocs %>% 
  distinct(year,month,drvid,CA_region,tier,ntraps_permitted) %>% 
  drop_na() %>% 
  mutate(has_vms=1)
CA_vms_prop_tier <- fishtix_tier %>% 
  left_join(CA_vms_vessels) %>% 
  mutate(CA_region=case_when(
    port_group_code %in% cenCA_ports ~ "cenCA",
    port_group_code %in% norCA_ports ~ "norCA"
  )) %>% 
  filter(!is.na(CA_region)) %>% 
  mutate(has_vms=replace_na(has_vms,0)) %>%
  distinct(crab_season,year,month,CA_region,drvid,.keep_all = T) %>% 
  group_by(crab_season,year,month,CA_region,tier,ntraps_permitted) %>% 
  summarise(tot_vessels=n_distinct(drvid),tot_vms=sum(has_vms)) %>% 
  ungroup() %>% 
  mutate(prop_vms=tot_vms/tot_vessels)
```

The last thing we need is a key to determine total traps permitted in each port group and month. We will use the fish tickets and CA tier info to create this key.

```{r}
traps_region_month <- fishtix_tier %>% 
  mutate(CA_region=case_when(
    port_group_code %in% cenCA_ports ~ "cenCA",
    port_group_code %in% norCA_ports ~ "norCA"
  )) %>% 
  filter(!is.na(CA_region),!is.na(ntraps_permitted),!is.na(tier)) %>% 
  distinct(crab_season,year,month,CA_region,drvid,ntraps_permitted,tier) %>% 
  group_by(crab_season,year,month,CA_region) %>% 
  summarise(max_traps_region=sum(ntraps_permitted)) %>% 
  ungroup()
```
# Join Grid

```{r grid fxn,warning=FALSE,fig.height=8,fig.width=4}
# Function to build a grid from a given year/month's simulated traps data

# raster representation of grid.
# transform to a projection that uses meters instead of lat/long
grd_rast <- fasterize(grd,raster = raster(grd,res=5000,crs=crs(grd)),field="GRID5KM_ID")
# set -99999 to NA
grd_rast[grd_rast==-99999] <- NA

# add trap weights to traps list-column
vessel_month_allweights %>% ungroup() %>%  mutate(month=month.name[month]) %>% right_join(traplocs)->traplocs

traplocs %<>% mutate(traps=purrr::map2(traps,combined_weight,function(df,wt) mutate(df,wt=wt)))

count_traps <- function(yr,mth){
  mth_name <- month.name[mth]
  samp <- traplocs %>%
    ungroup() %>%
    filter(year==yr,month==mth_name)
    # mutate(traps_xy=purrr::map(traps,function(sfdf){st_coordinates(sfdf) %>% as_tibble(.name_repair = 'minimal') %>% set_names(c('lon','lat'))})) %>% 
    # dplyr::select(traps_xy) %>% 
    # unnest()
  if(nrow(samp)==0){
    print(paste(mth_name,yr,"had zero observations."))
    zero_rast <- setValues(grd_rast,NA)
    names(zero_rast) <- paste0(mth_name,yr)
    return(zero_rast)
  } else{
    samp_sf<-setDT( samp )[, rbindlist( traps ) ] %>%
    sf::st_as_sf(crs=4326) %>% 
  # samp_sf <- samp %>%
  #   st_as_sf(coords=c('lon','lat'),crs=4326) %>% 
    # then, convert to planar projection to match the grid
    st_transform(st_crs(grd)) %>% 
    dplyr::select(wt) %>% 
    as_Spatial()
  out_rast <- raster::rasterize(samp_sf,grd_rast,field='wt',fun='sum')/25
  names(out_rast) <- paste0(mth_name,yr)
  # out <- samp_sf %>%
  #   mutate(ind=row_number()) %>% 
  #   st_join(grd) %>% 
  #   ungroup() %>% 
  #   group_by(GRID5KM_ID) %>% 
  #   summarise(ntraps=n_distinct(ind),traps_per_sq_km=ntraps/25)
  # if(plot){
  #   st_geometry(out) <- NULL
  #   out_plot <- grd %>% 
  #     dplyr::dplyr::select(GRID5KM_ID) %>% 
  #     left_join(out) %>% 
  #     filter(!is.na(traps_per_sq_km)) %>% 
  #     ggplot()+
  #     geom_sf(aes(fill=traps_per_sq_km),col=NA)+
  #     geom_sf(data=coaststates,col=NA)+
  #     scale_fill_viridis_c(name="Number of Traps\nper sq. km")+
  #     coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  #     theme(axis.text.x.bottom = element_text(angle=45,vjust=+0.5,hjust=0))
    # return(out_plot)
  # } else return(out)
  print(paste(mth_name,yr,"done."))
  return(out_rast)
  }

}

# sample map
p <- proc.time()
dec2013 <- count_traps(2013,12)
proc.time()-p
dec2013
plot(dec2013)
# data.table version (dunno why i can't get something like this to work in tidy) 
#   traps_sf<-setDT( samp )[, rbindlist( traps ) ] %>% 
#     sf::st_as_sf()
```

Apply to all years/months' data.

```{r apply fxn}
p <- proc.time()
years_months <- crossing(yr=2009:2018,mth=1:12)
trap_density_all_years <- purrr::map2(years_months$yr,years_months$mth,count_traps)
proc.time()-p
trap_density_stack <- raster::stack(trap_density_all_years)
write_rds(trap_density_stack,here::here('data','processed',"trap_density_all_weights.rds"))
```

# Animate output

We can then represent the output in an animation

```{r interpolate output,warning=FALSE,fig.height=8,fig.width=4}
# RasterStack to long form xy
all_obs_xy <- purrr::map(as.list(trap_density_stack),function(x){
  mth <- names(x)
  y<-rasterToPoints(x) %>% as_tibble() %>% set_names(c('x','y','density'))
  z <-y %>% mutate(month=mth)
  z
}) %>% bind_rows()

#bounding box for animation
bbox=c(825064,444947,1186239,2016996)

# test plot
library(viridis)
coaststates %<>% st_transform(st_crs(grd))
test<-all_obs_xy %>% 
  filter(month=="December2009")

test %>% ggplot()+
    geom_tile(aes(x,y,fill=density),na.rm=T,alpha=0.8,interpolate=FALSE)+
    geom_sf(data=coaststates,col=NA,fill='gray50')+
    scale_fill_viridis(na.value=NA,option="C",name="Traps per\nsq. km",breaks=seq(0,200,by=50),limits=c(0,200))+
    coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
    labs(x='',y='',fill='Traps per\nsq. km',title='{current_frame}')+
    theme(axis.text.x.bottom = element_text(angle=45),
          legend.position = c(0.7,0.6),
          title=element_text(size=16))

# ordering factor for frames of the animation
all_obs_xy %<>% mutate(labeler=factor(month,levels=unique(month))) 
# produce the animation
library(viridis)
library(gifski)
traps.gif <- all_obs_xy %>%
  filter(density>0) %>% 
  ggplot()+
    geom_tile(aes(x,y,fill=density),na.rm=T,alpha=0.8,interpolate=TRUE)+
    geom_sf(data=coaststates,col=NA,fill='gray50')+
    scale_fill_viridis(na.value=NA,option="C",name="Traps per\nsq. km",breaks=seq(0,200,by=50),limits=c(0,200))+
    coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
    labs(x='',y='',fill='Traps per\nsq. km',title='{current_frame}')+
    theme(axis.text.x.bottom = element_text(angle=45),
          legend.position = c(0.7,0.6),
          title=element_text(size=16))+
    transition_manual(labeler)
animate(traps.gif,fps=1,width=500,height=1000,renderer = gifski_renderer())

#save the animation
# fp_plots <- "C:/Users/Owen.Liu/Documents/NWFSC Research/Dungeness Fleet Dynamics/VMS DATA CONFIDENTIAL/Processed Data/owen processed/plots/"
anim_save(here::here('plots',"trap_density_all_years_combined_weights.gif"))

# same map by log traps
# log.traps.gif <- all_obs_xy %>%
#   filter(density>0) %>% 
#   ggplot()+
#     geom_tile(aes(x,y,fill=log(density)),na.rm=T,alpha=0.8,interpolate=TRUE)+
#     geom_sf(data=coaststates,col=NA,fill='gray50')+
#     scale_fill_viridis(na.value=NA,option="C",name="Traps per\nsq. km")+
#     coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
#     labs(x='',y='',fill='Traps per\nsq. km',title='{current_frame}')+
#     theme(axis.text.x.bottom = element_text(angle=45),
#           legend.position = c(0.7,0.6),
#           title=element_text(size=16))+
#     transition_manual(labeler)
# animate(traps.gif,fps=1,width=500,height=1000)
# anim_save(filename=paste0(fp_plots,"log_trap_density_all_years.gif"))
```

```{r distribution of trap densities}
all_obs_xy %>% 
  filter(!is.na(density),density>0) %>% 
  ggplot(aes(density))+
  geom_density(fill='red',alpha=0.98,col=NA)+
  scale_x_continuous(limits=c(0,150))+
  labs(x="Density (Traps per sq. Km)",y="Kernel Density",title="Distribution of Positive Trap Densities\nall years and months")
```

```{r comparison to Blake}
# read in Blake's estimated data
blake<-readxl::read_xlsx(here::here('data',"Monthly DCRB VMS pings per 5km grid cell 4Owen 101019.xlsx"))
grd_match <- tibble(GRID5KM_ID=getValues(grd_rast)) %>% left_join(blake,by="GRID5KM_ID")

# an example raster: for December, 2013
blake_rast <- raster(grd_rast)
values(blake_rast) <- grd_match$DUN2013_12
# reclassify zeroes to NA to match vertical line model
blake_rast <- reclassify(blake_rast,rcl=c(0,0,NA),right=NA)

test <- stack(blake_rast,trap_density_stack$December2013)
plot(test)
# do a correlation
cor(getValues(blake_rast),getValues(trap_density_stack$December2013),use="na.or.complete")

```

Correlation is not that high. Let's try with all the months in the data

```{r corr all months}
# Tibble to hold outputs of correlation
cor_table <- crossing(year=2009:2016,month=1:12,correlation=NA)
# Blake's analysis only goes through 06/2016
cor_table %<>% slice(-((nrow(cor_table)-5):nrow(cor_table)))

get_cor <- function(yr,mth){
  
  #parse correct blake year/month
  blake_var <- paste0("DUN",yr,"_",ifelse(mth>9,mth,paste0(0,mth)))
  # make raster and assign values
  blake_rast <- raster(grd_rast)
  values(blake_rast) <- grd_match[[blake_var]]
  # reclassify zeroes to NA to match vertical line model
  blake_rast <- reclassify(blake_rast,rcl=c(0,0,NA),right=NA)
  # grab correct owen year/month
  owen_var <- paste0(month.name[[mth]],yr)
  owen_rast <- trap_density_stack[[owen_var]]
  # do the correlation
  out <- cor(getValues(blake_rast),getValues(owen_rast),use="na.or.complete")
  # print(paste(month.name[mth],yr,"done."))
  out
}

cor_table %<>% mutate(correlation=map2_dbl(year,month,get_cor))

cor_table %>% 
  mutate(day=1) %>% 
  unite(date,year,month,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%m-%d",tz="America/Los_Angeles")) %>% 
  ggplot(aes(date,correlation))+
  geom_point()+
  geom_line()+
  labs(x="Date",y="Correlation",title="Correlation between vertical line model\nand Feist et al.")

head(arrange(cor_table,desc(correlation)))

## NOTE: ANOTHER OPTION?
#https://statnmap.com/2018-01-27-spatial-correlation-between-rasters/
```


Look at all values

```{r}
grd_xy <- rasterToPoints(grd_rast) %>% as_tibble() %>% set_names(c('x','y','GRID5KM_ID'))

vlm_blake <- all_obs_xy %>% left_join(grd_xy)

vlm_blake <- vlm_blake %>%
  mutate(year=str_sub(month,-4,-1) %>% as.numeric(),month=str_remove_all(month,"[:digit:]"),
         month=match(month,month.name)) %>% 
  select(-labeler)

blake_long <- blake %>%
  pivot_longer(contains("DUN"),names_to="year_month",values_to = "pings") %>% 
  mutate(year= str_sub(year_month,4,7) %>% as.numeric(),month=str_sub(year_month,9,10) %>% as.numeric()) %>% 
  select(-year_month,-contains("BAND"))

vlm_blake %<>%
  left_join(blake_long)

trap_density_comparison <- vlm_blake %>% 
  ggplot(aes(pings,density))+
  geom_point(size=0.25)+geom_smooth(method='lm',se=F)+
  labs(x="Number of VMS pings",y="Trap Density per sq. km (VLM)")+
  coord_cartesian(xlim=c(0,500),ylim=c(0,1000))
trap_density_comparison
ggsave(here('plots','trap_density_vs_pings.png'),trap_density_comparison,w=8,h=6)

# what about for just the highest-density areas/times
vlm_blake_top5 <- vlm_blake %>% 
  filter(density>=quantile(density,0.95))

vlm_blake_top5 %>% 
  ggplot(aes(pings,density))+
  geom_point(size=0.25)+geom_smooth(method="lm",se=F)+
  labs(x="Number of VMS pings",y="Density per sq. km (VLM)")
  
```


# Aggregation at Larger Spatial Scales

Now that we have traps on the 5km grid scale, we can aggregate over larger spatial scales (counties, marine regions).

```{r}
subregion_grd <- grd %>% dplyr::select(SUBREGION) %>% 
  group_by(SUBREGION) %>% 
  summarise()

ggplot(subregion_grd)+
  geom_sf(aes(fill=SUBREGION))
# for each month, convert raster to points and then aggregate
```

```{r}
all_obs_xy %>%
  mutate(date=parse_date(month,format="%B%Y")) %>% 
  group_by(date) %>% 
  summarise(mn=mean(density,na.rm=T)) %>% 
  ggplot(aes(date,mn))+
  geom_col(stat='identity',col='black',fill='black')+
  labs(y="pots/sq km",x="month")+
  theme(axis.text.x = element_text(angle=90))
```

Let's make a plot for average March number of traps by these blocks.

IMPORTANT: The measurements in our `trap_density_stack` RasterStack are in traps per square kilometer. However, to sum the total number across aggregated blocks, we have to multiply these densities by 25 (1 trap per square kilometer equals 25 traps per 5kmx5km grid square). Then we can sum across 5x5 blocks.

```{r}
marchtraps <- subset(trap_density_stack,which(grepl("March",names(trap_density_stack)))) %>% 
  mean(na.rm=T)
# convert to counts from densities
marchtraps <- marchtraps*25

plot(marchtraps)

march_traps_points <- marchtraps %>% 
  rasterToPoints(spatial = TRUE) %>% 
  st_as_sf()

march_traps_regions <- subregion_grd %>% 
  st_join(march_traps_points) %>% 
  group_by(SUBREGION) %>% 
  summarise(tot_traps=sum(layer))

ggplot(march_traps_regions)+
  geom_sf(aes(fill=tot_traps))+
  labs(fill="Total Traps")

```

What about California macro blocks?

```{r}
grd_macro_match <- grd %>% 
  st_drop_geometry() %>% 
  select(GRID5KM_ID,CA_OFFSHOR) %>% 
  distinct() %>% 
  rename(mb=CA_OFFSHOR)
  
ca_macro <- grd %>% dplyr::select(CA_OFFSHOR) %>% 
  group_by(CA_OFFSHOR) %>% 
  summarise() %>% 
  filter(CA_OFFSHOR!=-999) %>% 
  mutate(mb=as.character(CA_OFFSHOR))

ggplot(ca_macro)+
  geom_sf(aes(fill=mb))

march_traps_mb <- ca_macro %>% 
  st_join(march_traps_points) %>% 
  group_by(mb) %>% 
  summarise(tot_traps=sum(layer)) %>% 
  filter(!is.na(tot_traps))

bbox_ca=st_bbox(march_traps_mb)

march_total_traps_plot <-ggplot(march_traps_mb)+
  geom_sf(aes(fill=tot_traps))+
  geom_sf_label(aes(label=round(tot_traps)))+
  geom_sf(data=coaststates,fill='gray50')+
  scale_fill_viridis(breaks=seq(0,70000,by=20000),labels=seq(0,70000,by=20000),limits=c(0,70000))+
  coord_sf(xlim=c(bbox_ca[1],bbox_ca[3]),ylim=c(bbox_ca[2],bbox_ca[4]))+
  labs(fill="Total Traps",title="Average Total Traps by Macroblock\nMarch, all years",x="",y="")+
  theme(legend.position = c(0.8,0.7),
        legend.background = element_rect(fill='white'))

march_total_traps_plot

ggsave(filename=here::here('plots',"march_traps_CA_macroblocks.png"),march_total_traps_plot,h=6,w=6)
```

All years/months traps

```{r}
all_obs_sf <- vlm_blake %>% st_as_sf(coords=c('x','y'),crs=crs(grd))

all_traps_mb <- ca_macro %>% 
  st_join(all_obs_sf) %>% 
  mutate(ntraps=density*25) %>% 
  group_by(year,month,mb) %>% 
  summarise(tot_traps=sum(ntraps,na.rm=T),tot_pings=sum(pings,na.rm=T))

all_traps_mb %<>%
  mutate(day=1,date=paste(year,month,day,sep="-"),date=ymd(date))
# time series of trap density
mb1038_ts_traps<-all_traps_mb %>% 
  ungroup() %>% 
  filter(mb=="1038") %>% 
  st_as_sf() %>% st_drop_geometry() %>% 
  complete(mb,year,month=1:12,fill=list(tot_traps=NA,tot_pings=NA)) %>% 
  mutate(tot_traps=ifelse(month %in% c(7:10),NA,tot_traps)) %>%
  ggplot(aes(date,tot_traps/1000))+
  geom_point()+geom_path()+
  labs(x="Date",y="Thousands of Traps\nHalf Moon Bay to Point Arena")
mb1038_ts_traps
ggsave(here('plots','macro_block_1038_timeseries.png'),mb1038_ts_traps,w=8,h=3)

# time series of pings
mb1038_ts_pings <- all_traps_mb %>%
  ungroup() %>% 
  filter(mb=="1038") %>% 
  st_as_sf() %>% st_drop_geometry() %>% 
  complete(mb,year,month=1:12,fill=list(tot_traps=NA,tot_pings=NA)) %>% 
  mutate(tot_traps=ifelse(month %in% c(7:10),NA,tot_traps)) %>%
  ggplot(aes(date,tot_pings/1000))+
  geom_point()+geom_path()+
  labs(x="Date",y="Thousands of Pings\nHalf Moon Bay to Point Arena")
mb1038_ts_pings

mb1038_2011_2012 <- vlm_blake %>%
  ungroup() %>% 
  mutate(day=1,date=paste(year,month,day,sep="-"),date=ymd(date)) %>% 
  left_join(grd_macro_match) %>% 
  mutate(keep=case_when(
    month%in%c(11,12)&year==2011~TRUE,
    month %in% c(1:4)&year==2012~TRUE,
    TRUE ~ FALSE)) %>% 
  filter(keep,mb=='1038') %>% 
  mutate(monthname=month.name[month]) %>% 
  mutate(monthname=factor(monthname,levels=c("November","December","January","February","March","April")))

bbox_mb1038 <- mb1038_2011_2012 %>% st_as_sf(coords=c('x','y')) %>% st_bbox()
bbox_mb1038 <-bbox_mb1038+30000*c(-1,0,1,0)

mb1038_2011_2012_plot <- mb1038_2011_2012 %>% 
  ggplot()+
  geom_tile(aes(x,y,fill=density),na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col='black',fill='gray50',size=0.25)+
  scale_fill_viridis(na.value=NA,option="C",name="Traps per\nsq. km",breaks=seq(0,200,by=50),limits=c(0,200))+
  coord_sf(xlim=c(bbox_mb1038[1],bbox_mb1038[3]),ylim=c(bbox_mb1038[2],bbox_mb1038[4]),datum=NA)+
  labs(x='',y='',fill='Traps per\nsq. km',title='Trap Density, 2011-12 Season')+
  facet_wrap(~monthname)+
  theme(axis.text.x.bottom = element_text(angle=45),
        title=element_text(size=14),
        legend.title = element_text(size=10),
        panel.background = element_rect(fill='lightblue'))
mb1038_2011_2012_plot
ggsave(here('plots','trap_density_mb1038.png'),mb1038_2011_2012_plot,w=10,h=6)
```
Focal spatial correlation

```{r}
# Focal (moving-window) spatial correlation with rasters, for MB 1038
mb1038_r <- grd %>% filter(CA_OFFSHOR==1038) %>% summarise() %>% fasterize(grd_rast)
traps_dec2011 <- trap_density_stack[["December2011"]]*mb1038_r %>% trim()

# make raster and assign values
pings_dec2011 <- raster(grd_rast)
values(pings_dec2011) <- grd_match[["DUN2011_12"]]
# reclassify zeroes to NA to match vertical line model
pings_dec2011 <- reclassify(pings_dec2011,rcl=c(0,0,NA),right=NA)
pings_dec2011 <- pings_dec2011*mb1038_r %>% trim()
plot(traps_dec2011)
plot(pings_dec2011)
traps_pings_s <- stack(traps_dec2011,pings_dec2011) %>% values()

# do focal correlation in 3x3 window
cellnum_r <- raster(traps_dec2011)
values(cellnum_r) <- 1:ncell(traps_dec2011)

focal_cor <- focal(
  x = cellnum_r,
  w = matrix(1, 3, 3),
  fun = function(x, y = traps_pings_s){ # [MW]
    cor(y[x, 1], y[x, 2], # [MW]
        use = "pairwise.complete.obs")
  }
)
corr_plot <-focal_cor %>% 
  rasterToPoints() %>% as_tibble() %>% set_names(c('x','y','corr')) %>%
  ggplot()+
  geom_tile(aes(x,y,fill=corr),na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col='black',fill='gray50',size=0.25)+
  scale_fill_gradient2(midpoint = 0,low="blue",high="red")+
  coord_sf(xlim=c(bbox_mb1038[1]+50000,bbox_mb1038[3]),ylim=c(bbox_mb1038[2],bbox_mb1038[4]),datum=NA)+
  labs(x="",y="",fill="Correlation")+
  theme(panel.background = element_rect(fill='lightblue'))

# 3 panel plot
traps_dec2011_plot <-traps_dec2011 %>% 
  rasterToPoints() %>% as_tibble() %>% set_names(c('x','y','density')) %>%
  ggplot()+
  geom_tile(aes(x,y,fill=density),na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col='black',fill='gray50',size=0.25)+
  scale_fill_viridis(na.value=NA,option="C",name="Traps per\nsq. km",breaks=seq(0,150,by=50),limits=c(0,150))+
  coord_sf(xlim=c(bbox_mb1038[1]+50000,bbox_mb1038[3]),ylim=c(bbox_mb1038[2],bbox_mb1038[4]),datum=NA)+
  labs(x="",y="",title="Dec 2011 Trap Density")+
  theme(panel.background = element_rect(fill='lightblue'))
# traps_dec2011_plot

pings_dec2011_plot <-pings_dec2011 %>% 
  rasterToPoints() %>% as_tibble() %>% set_names(c('x','y','pings')) %>%
  ggplot()+
  geom_tile(aes(x,y,fill=pings),na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col='black',fill='gray50',size=0.25)+
  scale_fill_viridis(na.value=NA,option="C",name="Pings",limits=c(0,200),breaks=seq(0,200,by=50))+
  coord_sf(xlim=c(bbox_mb1038[1]+50000,bbox_mb1038[3]),ylim=c(bbox_mb1038[2],bbox_mb1038[4]),datum=NA)+
  labs(x="",y="",title="Dec 2011 VMS Pings")+
  theme(panel.background = element_rect(fill='lightblue'))
# pings_dec2011_plot
library(cowplot)
pings_traps_comp <- plot_grid(pings_dec2011_plot,traps_dec2011_plot,corr_plot,nrow=1)
ggsave(here('plots','pings_traps_comparison.png'),pings_traps_comp,w=10,h=5)
```

# Depth Distribution of Pots

```{r}
trap_depths <- simulated_traps %>% 
  mutate(depths=purrr::map(traps,pluck('depth'))) %>% 
  select(Rec_ID,depths) %>% 
  unnest(cols=c(depths)) %>% 
  left_join(fishtix_matched_all %>% select(Rec_ID,drvid,year,month))

trap_depth_p <- trap_depths %>% 
  sample_n(1e6) %>% 
  ggplot(aes(depths))+
  geom_density(fill='darkred',alpha=0.8)+
  labs(x="Depth (m)",y="Kernel Density")
```


# Comparison to Sam Macks' thesis

We can compare our estimates of trap densities to those calculated by Sam Macks in her thesis.
For example, for 2013-2014 to 2016-17

```{r}
# Pull out data for November-February, 2013 to 2017
traps_macks_range <- all_obs_xy %>% 
  mutate(date=parse_date(month,format="%B%Y")) %>% 
  filter(date>=as.Date("2013-11-01"),date<=as.Date("2017-02-01")) %>% 
  mutate(month=month(date),year=year(date)) %>%
  filter(month %in% c(11,12,1,2)) %>% 
  # average by month/year
  group_by(x,y,month) %>% 
  summarise(mean_density=mean(density,na.rm=T)) %>% 
  ungroup() %>% 
  mutate(monthname=factor(month.name[month],levels=c("November","December","January","February"))) %>% 
  filter(y<bbox_ca[4],y>bbox_ca[2])

ca_macks_comp_plot <- traps_macks_range %>% 
  ggplot()+
  geom_tile(aes(x,y,fill=mean_density),na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col=NA,fill='gray50')+
  facet_wrap(~monthname,nrow=1)+
  scale_fill_viridis(na.value=NA,option="C",name="Traps per\nsq. km")+
  coord_sf(xlim=c(bbox_ca[1]+350000,bbox_ca[3]-100000),ylim=c(bbox_ca[2],bbox_ca[4]))+
  labs(x='',y='',fill='Traps per\nsq. km',title='Average Traps Density\nNovember to February, 2013-2017')+
  theme(axis.text.x.bottom = element_text(angle=45),
        title=element_text(size=16))
ca_macks_comp_plot

# ggsave(filename = paste0(fp_plots,"CA_macks_comparison_plot.png"),ca_macks_comp_plot,h=6,w=10)
```