---
title: "Summarize VLM Outputs"
author: "Owen Liu"
date: "10/21/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster)
library(data.table)
library(tidyverse)
library(lubridate)
library(magrittr)
library(sf)
library(rnaturalearth)
library(fasterize)
library(tictoc)

library(furrr)
library(future)
```

```{r}
# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=12,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
```


# Purpose

The goal of this script is to take the output of the vertical line model simulation and apply functions that can summarise the total number and/or density of simulated crab pots. Update on a previous analysis that used weighting factors.

# Import Data

Simulated traps

```{r}
simtraps <- read_rds(here::here('vertical line model',"sim_traps_speed_depth_filters.rds"))
```

5x5km grid
```{r}
grd <- read_sf(here::here("data","raw","grid","fivekm_grid_polys_shore_lamb.shp"))
names(grd)
# raster representation of grid.
# transform to a projection that uses meters instead of lat/long
grd_rast <- fasterize(grd,raster = raster(grd,res=5000,crs=crs(grd)),field="GRID5KM_ID")
# set -99999 to NA
grd_rast[grd_rast==-99999] <- NA
```

Fish tickets
```{r}
fishtix_matched_all <- read_rds(here::here('data','processed','fish tickets','fish_tickets_w_dayofseason.rds'))
# filter for out of season
fishtix_matched_all %<>% filter(ticket_day_of_season>=0)
```

Coastline (just for plotting)
```{r}
# a coastline, for plotting
# coastline for plotting
coaststates <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('California','Oregon','Washington','Nevada'))
coastline <- ne_coastline(scale='medium',returnclass = 'sf') %>% 
  st_crop(st_bbox(coaststates))
```

Trap locations, discarding VMS data
```{r}
# traps locations only (discard VMS data for now), attached to fish ticket data
traplocs <- simtraps %>% dplyr::select(Rec_ID,traps) %>% 
  left_join(fishtix_matched_all,by="Rec_ID") %>% 
  dplyr::select(Rec_ID,date,year,month,ticket_day_of_season,pacfin_port_code,port_group_code,agency_code,drvid,FINAL_LENGTH,traps) %>% 
  distinct(Rec_ID,.keep_all = T) %>% 
  # count number of traps associated with each trip
  mutate(ntraps=purrr::map_int(traps,nrow))
```

Vessel tier info

```{r}
tier_key <- read_rds(here::here('vertical line model','tier_information_year_month_allstates.rds'))
tier_key %<>% mutate(month=month.name[month]) %>% dplyr::select(-crabyr)
```

# Summarising Traps on the 5km Grid

We want to write a function that takes the full simulated traps dataset, filters it by 1 week intervals, then gets counts of total traps per 5km grid cell.

## Create Periods

First, we need to create identifiers for 1 week periods.

```{r}
traplocs <- traplocs %>% 
  filter(!is.na(date)) %>% 
  # week of year
  mutate(period_1wk=week(date))
```

Add tier info

```{r}
traplocs <- traplocs %>% 
  left_join(tier_key,by=c('year','month','agency_code','drvid')) %>% 
  mutate(vessel_size_category=ifelse(FINAL_LENGTH<40,'small','large'))
```

## Calculate Total Number of Permitted Traps

```{r}
max_traps_by_year_week_size <- fishtix_matched_all %>% 
  left_join(tier_key) %>% 
  mutate(period_1wk=week(date)) %>% 
  mutate(vessel_size_category=ifelse(FINAL_LENGTH<40,'small','large')) %>% 
  mutate(ntraps_permitted=coalesce(ntraps_permitted,500)) %>%
  distinct(year,period_1wk,drvid,agency_code,vessel_size_category,ntraps_permitted) %>% 
  group_by(year,period_1wk,agency_code,drvid,vessel_size_category) %>% 
  slice_max(ntraps_permitted,n=1) %>% 
  group_by(agency_code,year,period_1wk,vessel_size_category) %>% 
  summarise(max_potential_traps=sum(ntraps_permitted)) %>% 
  ungroup()
```

## Rasterize Traps

Write the function to count traps, per grid cell, by period.

```{r}
count_traps <- function(yr,per,state,vesselsize){
  
  # use max_traps_year_by_year_month_size to get a maximum traps number
  maxtraps <- max_traps_by_year_week_size %>% 
    filter(year==yr,period_1wk==per,agency_code==state,vessel_size_category==vesselsize) %>% 
    pull('max_potential_traps') %>% unique()
  
  samp <- traplocs %>%
    ungroup() %>%
    filter(year==yr,period_1wk==per,agency_code==state,vessel_size_category==vesselsize)
  
  # add a relative weight for density calcs
  samp <- samp %>% 
    mutate(traps=purrr::map(traps,function(df,wt) mutate(df,wt=1)))

  if(nrow(samp)==0){
    print(paste('Week',per,'in',yr,"had zero observations."))
    zero_rast <- setValues(grd_rast,NA)
    names(zero_rast) <- paste0('week'," ",per," ",yr)
    return(zero_rast)
  } else{
    samp_sf<-setDT( samp )[, rbindlist( traps ) ] %>%
    sf::st_as_sf(crs=4326) %>% 
    # transform to match the grid
    st_transform(st_crs(grd)) %>% 
    dplyr::select(wt) %>% 
    as_Spatial()
    
    # use raster::rasterize() to count the number of traps in each cell
  out_rast <- raster::rasterize(samp_sf,grd_rast,field='wt',fun='sum')
  
  # normalize the raster, then multiply by max traps
  out_rast <- out_rast/sum(getValues(out_rast),na.rm=T)*maxtraps
  names(out_rast) <- paste0('week'," ",per," ",yr)

  print(paste0('state ',state,' week'," ",per," ",yr," done."))
  return(out_rast)
  }

}
```

```{r}
week48_2013_ca <- count_traps(2013,48,state="C",vesselsize='small')
plot(week48_2013_ca)
```


```{r apply fxn}
tic("1 Week Interval Pot Counts, Small Vessels")
years_states_periods <- traplocs %>% 
  ungroup() %>% 
  filter(ntraps > 0) %>% 
  distinct(year,period_1wk,agency_code) %>% 
  rename(yr=year,per=period_1wk,state=agency_code) %>% 
  unite(lab,state,yr,per,sep="_",remove=F)
# future::plan(multisession)
# trap_density_all_periods_small <- furrr::future_pmap(list(years_states_periods$yr,years_states_periods$per,years_states_periods$state),count_traps,vesselsize='small')
trap_density_all_periods_small <- purrr::pmap(list(years_states_periods$yr,years_states_periods$per,years_states_periods$state),count_traps,vesselsize='small')
toc()
trap_density_stack_small_vessels <- raster::stack(trap_density_all_periods_small)
names(trap_density_stack_small_vessels) <- years_states_periods$lab
write_rds(trap_density_stack_small_vessels,here::here('vertical line model',"trap_density_all_1week_periods_states_small_vessels.rds"))

```

```{r apply fxn}
tic("1 Week Interval Pot Counts, Large Vessels")
trap_density_all_periods_large <- purrr::pmap(list(years_states_periods$yr,years_states_periods$per,years_states_periods$state),count_traps,vesselsize='large')
toc()
trap_density_stack_large_vessels <- raster::stack(trap_density_all_periods_large)
names(trap_density_stack_large_vessels) <- years_states_periods$lab
write_rds(trap_density_stack_large_vessels,here::here('vertical line model',"trap_density_all_1week_periods_states_large_vessels.rds"))

# future::plan(sequential)
```