---
title: "Visualizing VLM"
author: "Owen Liu"
date: "10/22/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(raster)
library(tidyverse)
library(magrittr)
library(lubridate)
library(sf)
library(rnaturalearth)
library(fasterize)
library(here)
library(ggsci)
library(viridis)
library(stars)
options(dplyr.summarise.inform=FALSE)
```

# Purpose

Take the outputs from the vertical line model, and produce visualizations/manuscript figures.

```{r}
# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=12,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=12,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
```

# Import Data

Simulated traps

```{r}
# simulated traps with VMS data
simtraps <- read_rds(here::here('vertical line model',"sim_traps_speed_depth_filters.rds"))

# trap locations only, including fish ticket identifiers and weighting factors
traplocs <- read_rds(here::here('vertical line model','trap_locations_with_weights.rds'))
```

5x5km grid
```{r}
grd <- read_sf(here::here("data","raw","grid","fivekm_grid_polys_shore_lamb.shp"))
names(grd)
# raster representation of grid.
# transform to a projection that uses meters instead of lat/long
grd_rast <- fasterize(grd,raster = raster(grd,res=5000,crs=crs(grd)),field="GRID5KM_ID")
# set -99999 to NA
grd_rast[grd_rast==-99999] <- NA

# compare to old grid
# grd2 <- read_sf(here::here("data","raw","grid","regions_master_final_lamb.shp"))
# names(grd2)
# # raster representation of grid.
# # transform to a projection that uses meters instead of lat/long
# grd_rast2 <- fasterize(grd2,raster = raster(grd2,res=5000,crs=crs(grd)),field="GRID5KM_ID")

# x <- grd2 %>% 
#   filter(STATE=="CA") %>% 
#   fasterize(raster = raster(grd2,res=5000,crs=crs(grd)),field="GRID5KM_ID")
# 
# plot(x)

# ca <- grd2 %>% filter(STATE=="CA")
# range(ca$GRID5KM_ID)
# oreg <- grd2 %>% filter(STATE=="OR")
# range(oreg$GRID5KM_ID)
# wa <- grd2 %>% filter(STATE=="WA")
# range(wa$GRID5KM_ID)
# add a state designation
states <- grd %>% st_centroid() %>% st_transform(4326) %>% st_coordinates() %>% as_tibble() %>% 
  mutate(state=case_when(
    Y<42 ~ "California",
    Y>42&Y<=46.25 ~ "Oregon",
    Y>46.25 ~ "Washington"
  ))
grd <- grd %>% mutate(STATE=states$state)

```

Whale Biologically Important Areas (BIAs)

```{r}
bia_mn <- read_sf(here::here('data','raw','BIAs','CetMap_BIA_WGS84.shp')) %>% 
  # select whales of interest
  filter(sci_name=="Megaptera novaeangliae") %>% 
  dplyr::select(BIA_ID,BIA_name) %>% 
  st_transform(st_crs(grd))

# join to grid
grd <- grd %>% st_join(bia_mn) %>% mutate(is_BIA=!is.na(BIA_ID))
glimpse(grd)
```

Raster grid xy attributes

```{r}
grdxy <- rasterToPoints(grd_rast) %>% as_tibble() %>% set_names(c('x','y','GRID5KM_ID')) %>% 
  left_join(grd %>% st_set_geometry(NULL),by="GRID5KM_ID")
```

Fish tickets
```{r}
fishtix_matched_all <- read_rds(here::here('data','processed','fish tickets','fish_tickets_w_dayofseason.rds'))
# filter for out of season
fishtix_matched_all %<>% filter(ticket_day_of_season>=0)
```

Coastline (just for plotting)
```{r}
# a coastline, for plotting
# coastline for plotting
coaststates <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('California','Oregon','Washington','Nevada')) %>% 
  st_transform(st_crs(grd))
coastline <- ne_coastline(scale='medium',returnclass = 'sf') %>% 
  st_transform(st_crs(grd)) %>% 
  st_crop(st_bbox(coaststates))
```

Raster stacks of trap counts
```{r}
# raster stacks of trap counts (expanded and non-expanded)
traps_nonexpanded <- read_rds(here::here('vertical line model','trap_density_all_1week_periods_non_expanded.rds'))
traps_expanded <- read_rds(here::here('vertical line model',"trap_density_all_1week_periods_expanded.rds"))
```

Port groups in Northern vs. Central CA
```{r}
cenCA_ports <- c("MRA","SBA","SFA","MNA")
norCA_ports <- c("ERA","CCA","BGA","BDA","CA2")
```

VMS representation index that went into simulation and rasterization calculations
```{r}
vms_representation_index <- read_rds(here::here('vertical line model','vms_representation_index.rds'))
```

Combined tier info
```{r}
tier_key <- read_rds(here::here('vertical line model','tier_information_year_month_allstates.rds'))
tier_key %<>% mutate(month=month.name[month]) %>% dplyr::select(-crabyr)
```


Whale entanglements

```{r}
entanglements <- read_csv(here('vertical line model','entanglements','all_entanglements_102021.csv'))
```

# Fig 1 Landings Time Series

Figure 1 could be a timeseries of crab landings by state.

```{r}
# approx timing of whale presence in the Cal Current
rects <- tibble(xmin=as_date("2008-04-01")+years(1:11),
                xmax=as_date("2008-11-01")+years(1:11),
                ymin=-Inf,ymax=Inf)
```

```{r,fig.height=4,fig.width=8}
landings_by_state <- fishtix_matched_all %>% 
  group_by(year,month,agency_code) %>% 
  summarise(tot_land =sum(DCRB_lbs)/2204.62)

landings_by_state_plot <- landings_by_state %>% 
  mutate(day=1,month=match(month,month.name)) %>% 
  unite(date,year,month,day,sep="-") %>% 
  mutate(date=as_date(date)) %>% 
  ggplot()+
  geom_line(aes(date,tot_land,color=agency_code),size=1)+
  geom_rect(data=rects,aes(xmin=xmin,xmax=xmax,ymin=ymin,ymax=ymax),fill="red",alpha=0.3)+
  labs(y="Monthly Landings (MT)",x="Date",color="State")+
  scale_color_uchicago()+
  scale_x_date(breaks="4 months",date_labels ="%b-%y",expand=c(0,0))+
  theme(axis.text.x=element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

landings_by_state_plot
```
# Fig 2 Simulated Pots Timeseries

A. example vessel track with pot.
B. timeseries of total trap counts.
C. confirmed entanglements.

## Example Vessel Track

```{r}
create_linesegs <- function(sf_df){
  linesegs <- sf_df %>% 
    mutate(pt=row_number()) %>% 
    # create start and end points for each segment by duplicating rows (so that, e.g., the first segment will connect point 1 to 2, the second segment will connect point 2 to 3, etc.)
    slice(1,rep(2:(n()-1),each=2),n()) %>% 
    # line segment identifier, such that the segment for each VMS point is composed of its own point and the next point
    mutate(seg=lag(pt,1)) %>% replace_na(list(seg=1)) %>% 
    # build the lines
    group_by(seg) %>% 
    summarise() %>% 
    filter(st_geometry_type(.) == "MULTIPOINT") %>% 
    st_cast("LINESTRING") %>% 
    ungroup()
  return(linesegs)
}
```

```{r}
set.seed(2003)
sample_trip <- simtraps %>%
  ungroup() %>% 
  mutate(ntraps=purrr::map(traps,nrow)) %>% 
  filter(ntraps>=400) %>%  
  slice_sample(n=1)
sample_lines <- create_linesegs(sample_trip %>% pluck("data",1)) %>% st_transform(st_crs(grd))
sample_traps <- sample_trip %>% pluck("traps",1) %>% st_set_crs(4326) %>% st_transform(st_crs(grd))
bbox <- st_bbox(sample_traps)

sample_trip_plot <- ggplot()+
  geom_sf(data=sample_traps,size=0.75,color='blue')+
  geom_sf(data=sample_lines)+
  geom_sf(data=coaststates)+
  coord_sf(datum=NA)+
  lims(x=c(bbox[1],bbox[3]),y=c(bbox[2],bbox[4]))

sample_trip_plot
```


## Total Trap Counts

For total trap counts, we need to use the rasterized version of the VLM, because our corrections (weightings etc.) matter for accurate counts of total traps.

Gather x-y trap data

```{r}
years_periods <- traplocs %>% 
  ungroup() %>% 
  filter(ntraps > 0) %>% 
  distinct(year,period_1wk) %>% 
  rename(yr=year,per=period_1wk) %>% 
  mutate(per=per-1)
years_periods %<>% mutate(nm = paste(yr,per,sep = "_"))
names(traps_nonexpanded) <- years_periods$nm
names(traps_expanded) <- years_periods$nm
```


```{r}
# Rasterstack to long form xy data
all_obs_xy_no_expansion <- purrr::map(as.list(traps_nonexpanded),function(x){
  wk <- names(x)
  y<-rasterToPoints(x) %>% as_tibble() %>% set_names(c('x','y','count'))
  z <-y %>% mutate(wk=wk)
  z
}) %>% bind_rows() %>% 
  # join grid data
  left_join(grdxy,by=c('x','y')) %>% 
  # add a density calc
  mutate(traps_sqkm=count/(AREA/1e6)) %>% 
  # add a better date
  mutate(year=str_sub(wk,2,5)) %>% 
  mutate(week=str_extract(wk,"\\d+$")) %>% 
  mutate(day="1") %>% 
  mutate(jan1=paste0(year,"-01-01") %>% as_date(format="%F")) %>% 
  unite(date,year,week,day,sep="-") %>% 
  mutate(date=as_date(date,format="%Y-%U-%u")) %>%
  mutate(date=coalesce(date,jan1)) %>% 
  dplyr::select(-jan1)

all_obs_xy_expanded <- purrr::map(as.list(traps_expanded),function(x){
  wk <- names(x)
  y<-rasterToPoints(x) %>% as_tibble() %>% set_names(c('x','y','count'))
  z <-y %>% mutate(wk=wk)
  z
}) %>% bind_rows() %>% 
  # join grid data
  left_join(grdxy,by=c('x','y')) %>% 
  # add a density calc
  mutate(traps_sqkm=count/(AREA/1e6)) %>% 
  # add a better date
  mutate(year=str_sub(wk,2,5)) %>% 
  mutate(week=str_extract(wk,"\\d+$")) %>% 
  mutate(day="1") %>% 
  mutate(jan1=paste0(year,"-01-01") %>% as_date(format="%F")) %>% 
  unite(date,year,week,day,sep="-") %>% 
  mutate(date=as_date(date,format="%Y-%U-%u")) %>%
  mutate(date=coalesce(date,jan1)) %>% 
  dplyr::select(-jan1)

glimpse(all_obs_xy_expanded)
```

Now we can take these non-expanded and expanded datasets and make timeseries of total traps.

```{r}
trap_counts_state_month_nonexpanded <- all_obs_xy_no_expansion %>% 
  filter(!is.na(GRID5KM_ID)) %>% 
  # mean traps by cell by month
  mutate(mth=month(date),yr=year(date)) %>% 
  group_by(STATE,date,yr,mth) %>% 
  summarise(trapcount=sum(count,na.rm=T)) %>% 
  ungroup() %>% 
  group_by(yr,mth,STATE) %>% 
  summarise(tottraps=mean(trapcount,na.rm=T))

trap_counts_state_month_expanded <- all_obs_xy_expanded %>% 
  filter(!is.na(GRID5KM_ID)) %>% 
  # mean traps by cell by month
  mutate(mth=month(date),yr=year(date)) %>% 
  group_by(STATE,date,yr,mth) %>% 
  summarise(trapcount=sum(count,na.rm=T)) %>% 
  ungroup() %>% 
  group_by(yr,mth,STATE) %>% 
  summarise(tottraps=mean(trapcount,na.rm=T))
```

```{r,fig.width=8,fig.height=8}
trap_counts_nonexpanded_ts_plot <- trap_counts_state_month_nonexpanded %>% 
  mutate(day=1) %>% 
  unite(date,yr,mth,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%m-%d")) %>%  
  ggplot()+
  geom_line(aes(date,tottraps/1000))+
  geom_rect(data=rects,aes(xmin=xmin,xmax=xmax,ymin=ymin,ymax=ymax),fill="red",alpha=0.3)+
  scale_x_date(date_breaks= "4 months",date_labels="%b-%y",expand=c(0,0))+
  labs(x="Date",y="Thousand Traps (VMS Vessels Only)")+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())+
  facet_wrap(~STATE,nrow=3,scales='free_y')
trap_counts_nonexpanded_ts_plot

trap_counts_expanded_ts_plot <- trap_counts_state_month_expanded %>% 
  mutate(day=1) %>% 
  unite(date,yr,mth,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%m-%d")) %>%  
  ggplot()+
  geom_line(aes(date,tottraps/1000))+
  geom_rect(data=rects,aes(xmin=xmin,xmax=xmax,ymin=ymin,ymax=ymax),fill="red",alpha=0.3)+
  scale_x_date(date_breaks= "4 months",date_labels="%b-%y",expand=c(0,0))+
  labs(x="Date",y="Thousand Traps (Whole Fleet Estimate)")+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())+
  facet_wrap(~STATE,nrow=3,scales='free_y')
trap_counts_expanded_ts_plot
```
```{r}
trap_counts_3states_expanded <- trap_counts_state_month_expanded %>% 
  mutate(day=1) %>% 
  unite(date,yr,mth,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%m-%d")) %>%  
  ggplot()+
  geom_line(aes(date,tottraps/1000,color=STATE),size=1)+
  geom_rect(data=rects,aes(xmin=xmin,xmax=xmax,ymin=ymin,ymax=ymax),fill="red",alpha=0.3)+
  labs(y="Simulated Traps (Thousands)",x="Date",color="State")+
  scale_color_uchicago()+
  scale_x_date(breaks="4 months",date_labels ="%b-%y",expand=c(0,0))+
  theme(axis.text.x=element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
trap_counts_3states_expanded
```


Zoom in on 2014-2017

```{r}
trap_counts_nonexpanded_ts_zoom_plot <- trap_counts_state_month_nonexpanded %>%
  filter(yr>2013,yr<2018) %>% 
  mutate(day=1) %>% 
  unite(date,yr,mth,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%m-%d")) %>% 
  ggplot(aes(date,tottraps/1000))+
  geom_line()+
  scale_x_date(date_breaks= "3 months",date_labels="%b-%Y",expand=c(0,0))+
  labs(x="Date",y="Thousand Traps (VMS Vessels Only)")+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())+
  facet_wrap(~STATE,nrow=3,scales='free_y')
trap_counts_nonexpanded_ts_zoom_plot

trap_counts_expanded_ts_zoom_plot <- trap_counts_state_month_expanded %>%
  filter(yr>2013,yr<2018) %>% 
  mutate(day=1) %>% 
  unite(date,yr,mth,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%m-%d")) %>%  
  ggplot(aes(date,tottraps/1000))+
  geom_line()+
  scale_x_date(date_breaks= "3 months",date_labels="%b-%Y",expand=c(0,0))+
  labs(x="Date",y="Thousand Traps (Whole Fleet Estimate)")+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())+
  facet_wrap(~STATE,nrow=3,scales='free_y')
trap_counts_expanded_ts_zoom_plot
```

## Confirmed Entanglements Timeseries

```{r}
# See Jameal's workup of this for comparison

entanglements_ts <- entanglements %>% 
  mutate(spp=tolower(`Common Name`)) %>% 
  filter(grepl("Dungeness",Entanglement_Fishery),
         Year>2007,Year<=max(traplocs$year),
         spp %in% c("humpback whale","blue whale"),
         State %in% c("WA","CA","OR")) %>% 
  dplyr::select(Year,Month,spp,State) %>% 
  add_row(Year=2009,Month=1,spp="humpback whale") %>% 
  filter(Year>2008) %>% 
  # organize by quarter of year
  mutate(day=1) %>% 
  unite(date,Year,Month,day,sep="-",remove=F) %>% 
  mutate(date=as_date(date,format="%Y-%m-%d")) %>% 
  mutate(qtr=quarter(date)) %>% 
  # summarize
  group_by(spp,Year,qtr) %>% 
  summarise(ent=n()) %>% 
  ungroup() %>% 
  complete(Year,spp,qtr,fill=list(ent=0)) %>% 
  unite(year_qtr,Year,qtr,sep="-")
  # re-do dates
  # mutate(month=c('January','April','July','October')[qtr]) %>% 
  # mutate(day=1) %>% 
  # unite(date,Year,month,day,sep="-",remove=F) %>% 
  # mutate(date=as_date(date,format="%Y-%B-%d"))

entanglements_ts_plot <- entanglements_ts %>% 
  ggplot(aes(year_qtr,ent,fill=spp))+
  geom_bar(stat='identity',position="dodge")+
  # geom_point()+
  # geom_line()+
  labs(x="Quarter",y="Confirmed Entanglements",fill="Species")+
  # scale_x_date(date_breaks="6 months",date_labels = "%b-%y")+
  scale_y_continuous(breaks=seq(0,10,by=2))+
  scale_fill_aaas()+
  theme(axis.text.x=element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        legend.position = c(0.2,0.8))
entanglements_ts_plot
```
# Fig 3 Traps Within BIAs

## Trap Numbers

Trap numbers within Biologically Important Areas for humpback whales

```{r}
trap_counts_BIAs_month_nonexpanded <- all_obs_xy_no_expansion %>% 
  filter(!is.na(GRID5KM_ID)) %>% 
  # mean traps by cell by month
  mutate(mth=month(date),yr=year(date)) %>% 
  group_by(GRID5KM_ID,STATE,is_BIA,yr,mth) %>% 
  summarise(trapcount=mean(count,na.rm=T)) %>% 
  ungroup() %>% 
  group_by(yr,mth,STATE,is_BIA) %>% 
  summarise(tottraps=sum(trapcount,na.rm=T))

trap_counts_BIAs_month_expanded <- all_obs_xy_expanded %>% 
  filter(!is.na(GRID5KM_ID)) %>% 
  # mean traps by cell by month
  mutate(mth=month(date),yr=year(date)) %>% 
  group_by(GRID5KM_ID,STATE,is_BIA,yr,mth) %>% 
  summarise(trapcount=mean(count,na.rm=T)) %>% 
  ungroup() %>% 
  group_by(yr,mth,STATE,is_BIA) %>% 
  summarise(tottraps=sum(trapcount,na.rm=T))
```

```{r,fig.width=8,fig.height=8}
library(ggsci)
trap_counts_nonexpanded_BIAs_ts_plot <- trap_counts_BIAs_month_nonexpanded %>% 
  ungroup() %>% 
  complete(yr,mth,is_BIA,fill=list(tottraps=NA)) %>% 
  mutate(day=1) %>% 
  unite(date,yr,mth,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%m-%d")) %>% 
  mutate(BIA_label=ifelse(is_BIA,"Inside","Outside")) %>% 
  filter(!is.na(STATE)) %>% 
  ggplot(aes(date,tottraps/1000,col=BIA_label))+
  geom_point()+
  geom_line()+
  scale_color_locuszoom(name="")+
  scale_x_date(date_breaks= "6 months",date_labels="%b-%Y",expand=c(0,0))+
  facet_wrap(~STATE,nrow=3,scales="free_y")+
  labs(x="Date",y="Thousand Traps (VMS Vessels Only)",title="VMS Traps inside/outside BIAs")+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())
trap_counts_nonexpanded_BIAs_ts_plot

trap_counts_expanded_BIAs_ts_plot <- trap_counts_BIAs_month_expanded %>% 
  ungroup() %>% 
  complete(yr,mth,is_BIA,fill=list(tottraps=NA)) %>% 
  mutate(day=1) %>% 
  unite(date,yr,mth,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%m-%d")) %>% 
  mutate(BIA_label=ifelse(is_BIA,"Inside","Outside")) %>%  
  filter(!is.na(STATE)) %>% 
  ggplot(aes(date,tottraps/1000,col=BIA_label))+
  # geom_bar(stat='identity',position='stack')+
  geom_line()+
  geom_point()+
  # scale_fill_locuszoom(name="")+
  scale_color_locuszoom(name="")+
  scale_x_date(date_breaks= "6 months",date_labels="%b-%Y",expand=c(0,0))+
  facet_wrap(~STATE,nrow=3,scales="free_y")+
  labs(x="Date",y="Thousand Traps (Whole Fleet Estimate)",title="VMS Traps inside/outside BIAs")+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())
trap_counts_expanded_BIAs_ts_plot

# STACKED BARS with quarters (to match entanglements)
trap_counts_expanded_BIAs_ts_barplot <- trap_counts_BIAs_month_expanded %>%
  ungroup() %>%
  complete(yr,mth,is_BIA,fill=list(tottraps=NA)) %>%
  mutate(day=1) %>%
  filter(!is.na(STATE)) %>% 
  unite(date,yr,mth,day,sep="-",remove = FALSE) %>%
  mutate(date=as_date(date,format="%Y-%m-%d")) %>%
  mutate(qtr=quarter(date)) %>% 
  group_by(yr,qtr,is_BIA,STATE) %>% 
  # Mean traps over the quarter
  summarise(tottraps=mean(tottraps,na.rm=T)) %>% 
  ungroup() %>% 
  unite(year_qtr,yr,qtr,sep="-") %>% 
  mutate(BIA_label=ifelse(is_BIA,"Inside","Outside")) %>%
  ggplot(aes(year_qtr,tottraps/1000,fill=BIA_label,col=BIA_label))+
  geom_bar(stat='identity',position='stack')+
  # geom_line()+
  scale_fill_locuszoom(name="")+
  scale_color_locuszoom(name="")+
  facet_wrap(~STATE,nrow=3,scales="free_y")+
  # scale_x_date(date_breaks= "6 months",date_labels="%b-%Y",expand=c(0,0))+
  labs(x="Date",y="Thousand Traps (Whole Fleet Estimate)",title="VMS Traps inside/outside BIAs")+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),)
trap_counts_expanded_BIAs_ts_barplot
```

What about a monthly mean/anomaly time series for California?

```{r}
# have to expand with zeroes
trap_counts_expanded_BIAs_anomaly <- trap_counts_BIAs_month_expanded %>% 
  filter(STATE=='California',is_BIA) %>% 
  complete(yr,mth=full_seq(1:12,1),STATE,fill=list(tottraps=0)) %>% 
  group_by(mth) %>% 
  mutate(mean_traps=mean(tottraps)) %>% 
  ungroup() %>% 
  mutate(anom=tottraps-mean_traps)

# mean by month
trap_counts_CA_BIAS_mean <- trap_counts_expanded_BIAs_anomaly %>% 
  ggplot(aes(mth,mean_traps/1000))+
  geom_line(size=1.5)+
  scale_x_continuous(limits=c(1,12),breaks=seq(1,12,by=1))+
  labs(x="Month",y="Mean Traps in CA BIAs (x1000)")
trap_counts_CA_BIAS_mean

# monthly anomalies
trap_counts_CA_BIAs_anom_ts <- trap_counts_expanded_BIAs_anomaly %>% 
  mutate(day=1) %>% 
  unite(date,yr,mth,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%m-%d")) %>% 
  ggplot(aes(date,anom/1000))+
  geom_line()+
  geom_hline(yintercept=0,linetype=2)+
  scale_x_date(breaks="4 months",date_labels ="%b-%y",expand=c(0,0))+
  labs(x="Month",y="BIA Trap Count Anomaly (1000s traps)")+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())
trap_counts_CA_BIAs_anom_ts
```


Zoom

```{r,fig.width=8,fig.height=8}
trap_counts_nonexpanded_BIAs_ts_zoom_plot <- trap_counts_BIAs_month_nonexpanded %>% 
  ungroup() %>% 
  filter(yr %in% 2014:2017) %>% 
  complete(yr,mth,is_BIA,fill=list(tottraps=NA)) %>% 
  mutate(day=1) %>% 
  unite(date,yr,mth,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%m-%d")) %>% 
  mutate(BIA_label=ifelse(is_BIA,"Inside","Outside")) %>% 
  filter(!is.na(STATE)) %>% 
  ggplot(aes(date,tottraps/1000,col=BIA_label))+
  geom_point()+
  geom_line()+
  scale_color_locuszoom(name="")+
  scale_x_date(date_breaks= "6 months",date_labels="%b-%Y",expand=c(0,0))+
  facet_wrap(~STATE,nrow=3,scales="free_y")+
  labs(x="Date",y="Thousand Traps (VMS Vessels Only)",title="VMS Traps inside/outside BIAs")+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())
trap_counts_nonexpanded_BIAs_ts_zoom_plot

trap_counts_expanded_BIAs_ts_zoom_plot <- trap_counts_BIAs_month_expanded %>% 
  ungroup() %>% 
  filter(yr %in% 2014:2017) %>% 
  complete(yr,mth,is_BIA,fill=list(tottraps=NA)) %>% 
  mutate(day=1) %>% 
  unite(date,yr,mth,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%m-%d")) %>% 
  mutate(BIA_label=ifelse(is_BIA,"Inside","Outside")) %>% 
  filter(!is.na(STATE)) %>% 
  ggplot(aes(date,tottraps/1000,col=BIA_label))+
  # geom_bar(stat='identity',position='stack')+
  geom_line()+
  geom_point()+
  # scale_fill_locuszoom(name="")+
  scale_color_locuszoom(name="")+
  scale_x_date(date_breaks= "6 months",date_labels="%b-%Y",expand=c(0,0))+
  facet_wrap(~STATE,nrow=3,scales="free_y")+
  labs(x="Date",y="Thousand Traps (Whole Fleet Estimate)",title="VMS Traps inside/outside BIAs")+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())
trap_counts_expanded_BIAs_ts_zoom_plot

## bar plot?
trap_counts_expanded_BIAs_ts_zoom_barplot <- trap_counts_BIAs_month_expanded %>% 
  ungroup() %>% 
  filter(yr %in% 2014:2017) %>% 
  complete(yr,mth,is_BIA,fill=list(tottraps=NA)) %>% 
  mutate(day=1) %>% 
  unite(date,yr,mth,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%m-%d")) %>% 
  mutate(BIA_label=ifelse(is_BIA,"Inside","Outside")) %>% 
  filter(!is.na(STATE)) %>% 
  ggplot(aes(date,tottraps/1000,fill=BIA_label))+
  geom_bar(stat='identity',position='stack')+
  # geom_line()+
  # geom_point()+
  # scale_fill_locuszoom(name="")+
  scale_fill_locuszoom(name="")+
  scale_x_date(date_breaks= "4 months",date_labels="%b-%Y",expand=c(0,0))+
  facet_wrap(~STATE,nrow=3,scales="free_y")+
  labs(x="Date",y="Thousand Traps (Whole Fleet Estimate)",title="VMS Traps inside/outside BIAs")+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())
trap_counts_expanded_BIAs_ts_zoom_barplot
```

Proportion of traps inside/outside BIAs for expanded data.

```{r,fig.width=8,fig.height=8}
rects <- tibble(xmin=as_date("2008-04-01")+years(1:11),
                xmax=as_date("2008-11-01")+years(1:11),
                ymin=-Inf,ymax=Inf)
    
prop_traps_BIAs <- trap_counts_BIAs_month_expanded %>% 
  ungroup() %>%
  # filter(yr %in% 2014:2017) %>% 
  complete(yr,mth,STATE,is_BIA,fill=list(tottraps=0)) %>% 
  mutate(day=1) %>% 
  unite(date,yr,mth,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%m-%d")) %>% 
  filter(!is.na(STATE)) %>% 
  group_by(date,STATE) %>% 
  mutate(propinside=tottraps[is_BIA]/sum(tottraps)) %>% 
  ungroup() %>% 
  ggplot()+
  geom_rect(data=rects,aes(xmin=xmin,xmax=xmax,ymin=ymin,ymax=ymax),fill="red",alpha=0.3)+
  # geom_bar(stat='identity',position='stack')+
  geom_line(aes(date,propinside))+
  geom_point(aes(date,propinside))+
  scale_x_date(date_breaks= "4 months",date_labels="%b-%y",expand=c(0,0))+
  labs(x="Date",y="Proportion of Traps",title="VMS Traps Inside BIAs")+
  facet_wrap(~STATE,nrow=3) + 
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())
prop_traps_BIAs
```

During whale season (April-November), summarise traps inside BIAs

```{r,fig.width=8,fig.height=8}
traps_inside_BIAs_seasons <- trap_counts_BIAs_month_expanded %>% 
  ungroup() %>%
  # filter(yr %in% 2014:2017) %>% 
  complete(yr,mth,STATE,is_BIA,fill=list(tottraps=0)) %>% 
  mutate(day=1) %>% 
  unite(date,yr,mth,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%m-%d")) %>% 
  filter(!is.na(STATE)) %>% 
  group_by(date,STATE) %>% 
  mutate(propinside=tottraps[is_BIA]/sum(tottraps)) %>% 
  ungroup() %>% 
  mutate(crabyr=ifelse(mth<=10, paste0(yr-1,"-",yr),paste0(yr,'-',yr+1))) %>% 
  # select only April-November of each year
  filter(mth %in% 4:11) %>% 
  group_by(crabyr,STATE) %>% 
  summarise(mean_prop_inside=mean(propinside,na.rm=T)) %>% 
  ungroup()

traps_inside_BIAs_seasons %>% 
  ggplot(aes(crabyr,mean_prop_inside))+
  geom_bar(stat='identity')+
  facet_wrap(~STATE,nrow=3)+
  labs(x="Crab Season",y="Mean Proportion of Traps Inside BIAs")+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())
```

## Maps

A couple of representative maps for the large BIA around San Fran Bay.

```{r,fig.height=6,fig.width=10}
years_periods %<>% mutate(lyr=row_number()) %>% 
  mutate(day=1) %>% 
  unite(date,yr,per,day,sep="-",remove=F) %>% 
  mutate(date=as_date(date,format="%Y-%U-%u")) %>% 
  mutate(mth=month(date)) %>%
  mutate(mth=coalesce(mth,1)) %>% 
  dplyr::select(-day)
bia_fara <- bia_mn %>% filter(BIA_name=="Gulf of the Farallones - Monterey Bay")
bbox_bias <- st_bbox(bia_fara)

# VLM estimates
traps_expanded_stars <- traps_expanded %>% st_as_stars() %>% setNames('traps')

traps_fara <- traps_expanded_stars %>% st_crop(bbox_bias)
  # filter for the first 3 months of 2016 (right before huge entanglement spike)
spr2016 <- years_periods %>% filter(yr==2016,mth %in% c(3,4,5,6)) %>% pull(lyr)
traps_fara_2016 <- traps_fara["traps",,,spr2016] %>% 
  # and get a mean
  st_apply(1:2,function(x)mean(x,na.rm=T))

spr <- years_periods %>% filter(mth %in% c(3,4,5,6)) %>% pull(lyr)
traps_fara_allyr <- traps_fara["traps",,,spr] %>% 
  # and get a mean
  st_apply(1:2,function(x)mean(x,na.rm=T))

fara_2016 <- ggplot()+
  geom_stars(data=traps_fara_2016,na.rm=T)+
  geom_sf(data=bia_mn,col="black",fill=NA,size=1.5)+
  geom_sf(data=coaststates)+
  scale_fill_viridis(na.value=NA,limits=c(0,200),name="# traps")+
  coord_sf(datum=NA)+
  labs(x="",y="",title="March to June 2016")+
  lims(x=c(bbox_bias[1],bbox_bias[3]),y=c(bbox_bias[2],bbox_bias[4]))+
  theme(legend.position=c(0.2,0.2))

fara_mean <- ggplot()+
  geom_stars(data=traps_fara_allyr,na.rm=T)+
  geom_sf(data=bia_mn,col="black",fill=NA,size=1.5)+
  geom_sf(data=coaststates)+
  scale_fill_viridis(na.value=NA,limits=c(0,200),name="# traps")+
  coord_sf(datum=NA)+
  labs(x="",y="",title="March to June Mean")+
  lims(x=c(bbox_bias[1],bbox_bias[3]),y=c(bbox_bias[2],bbox_bias[4]))+
  theme(legend.position=c(0.2,0.2))

fara_diff <- (traps_fara_2016-traps_fara_allyr)/traps_fara_allyr*100
fara_diff_plot <- ggplot()+
  geom_stars(data=fara_diff,na.rm=T)+
  geom_sf(data=bia_mn,col="black",fill=NA,size=1.5)+
  geom_sf(data=coaststates)+
  scale_fill_viridis(option="C",breaks=seq(-50,150,by=50), na.value=NA,name="Percent\nDifference")+
  coord_sf(datum=NA)+
  labs(x="",y="",title="2016 Relative to Mean")+
  lims(x=c(bbox_bias[1],bbox_bias[3]),y=c(bbox_bias[2],bbox_bias[4]))+
  theme(legend.position=c(0.2,0.2))

# fara_2016
# fara_mean
# fara_diff_plot

library(cowplot)
plot_grid(fara_mean,fara_2016,fara_diff_plot,nrow=1,labels = 'auto')
```

## Density vs. Entanglements

Doesn't look super promising

```{r}
trap_density_CA_BIAs_month_expanded <- all_obs_xy_expanded %>% 
  filter(!is.na(GRID5KM_ID),STATE=="California",is_BIA) %>% 
  # mean traps by cell by month
  mutate(mth=month(date),yr=year(date)) %>% 
  group_by(GRID5KM_ID,yr,mth) %>% 
  summarise(trapdens=mean(traps_sqkm,na.rm=T)) %>% 
  ungroup() %>% 
  group_by(yr,mth) %>% 
  summarise(trapdens=mean(trapdens,na.rm=T)) %>% 
  ungroup() %>% 
  arrange(yr,mth)

# organize entanglement data to add this to the trap density data
entanglements_CA<-entanglements %>% 
  mutate(spp=tolower(`Common Name`)) %>% 
  filter(grepl("Dungeness",Entanglement_Fishery),
         Year<=max(traplocs$year),
         spp %in% c("humpback whale","blue whale"),
         State =="CA") %>% 
  dplyr::select(Year,Month,spp) %>% 
  # organize by quarter of year
  mutate(day=1) %>% 
  unite(date,Year,Month,day,sep="-",remove=F) %>% 
  mutate(date=as_date(date,format="%Y-%m-%d")) %>% 
  group_by(Year,Month) %>% 
  summarise(ent=n()) %>% 
  ungroup() %>% 
  complete(Month=full_seq(1:12,1),Year=full_seq(2008:2019,1),fill=list(ent=0))

# join
trapdens_CA_BIAs_entanglements <- trap_density_CA_BIAs_month_expanded %>% 
  left_join(entanglements_CA,by=c("yr"="Year","mth"="Month")) %>% 
  # leads
  mutate(lead1=lead(ent,1,default=0),lead2=lead(ent,2,default=0),lead3=lead(ent,3,default=0),lead4=lead(ent,4,default=0)) %>% 
  # entanglements in the last 4 months in CA
  mutate(ent3mon = lead1+lead2+lead3,
         ent4mon = lead1+lead2+lead3+lead4) %>% 
  mutate(ent_bin3=ifelse(ent3mon>0,1,0),
         ent_bin4=ifelse(ent4mon>0,1,0))

# plot
binomial_smooth <- function(...) {
  geom_smooth(method = "glm", method.args = list(family = "binomial"), ...)
}
trapdens_CA_BIAs_entanglements %>% 
  filter(mth %in% 1:7,yr %in% 2014:2018) %>% 
  ggplot(aes(trapdens,ent_bin3))+
  geom_point()+
  binomial_smooth()+
  scale_y_continuous(limits = c(0,1),breaks=c(0,1),labels=c(0,1))+
  labs(title="Mean Trap Density in CA BIAs and Entanglements",x="Mean Trap Density (#/sq. km)",y="Confirmed Entanglement(s) in CA\nin the next 4 months")
```


# Fig 4 Trap Depth Distributions

Look at the depth distributions of simulated traps. For these first plots, we use the non-rasterized version of the simulated traps.

Pull out the mean trap depth for all individual fish tickets/trips

```{r}
trap_depths <- traplocs %>% 
  mutate(coast_region=case_when(
    port_group_code %in% cenCA_ports ~ "Central CA",
    port_group_code %in% norCA_ports ~ "Northern CA",
    agency_code=="O" ~ "Oregon",
    agency_code=="W" ~ "Washington"
  )) %>% 
  mutate(mean_trap_depth=purrr::map_dbl(traps,function(df) mean(df$depth,na.rm=T))) %>% 
  dplyr::select(-traps) %>% 
  filter(month !="October") %>% 
  # add an identifier for early, mid, or late season
  mutate(season_segment=case_when(
    month %in% c('November','December','January') ~ 'early',
    month %in% c('February','March','April') ~ 'mid',
    month %in% c('May','June','July','August','September') ~ 'late'
  )) %>% 
  mutate(season_segment=factor(season_segment,levels=c('early','mid','late')))
```

Depth distributions by state and season segment

```{r,fig.width=8,fig.height=8}
trap_depth_dist <- trap_depths %>% 
  ggplot(aes(mean_trap_depth,fill=season_segment,col=season_segment))+
  geom_density(alpha=0.5)+
  labs(x="Mean Trap Depth per Trip (m)",y="Kernel Density",fill="Season Portion",col="Season Portion")+
  facet_wrap(~coast_region,nrow=4,scales = 'free_y')+
  geom_vline(xintercept=-40*1.829,col='red',linetype=2)+
  theme(panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        axis.text.y=element_blank())
trap_depth_dist
```

Now, a timeseries of mean trap depth per trip across months

```{r,fig.width=8,fig.height=8}
trap_depths_months <- trap_depths %>% 
  group_by(year,month,coast_region) %>% 
  summarise(mean_depth=mean(mean_trap_depth,na.rm=T),se_depth=sd(mean_trap_depth,na.rm=T)/sqrt(n()))

trap_depths_ts <- trap_depths_months %>% 
  mutate(day=1) %>% 
  unite(date,year,month,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%B-%d")) %>% 
  mutate(depthpos=-mean_depth)

trap_depths_ts_plot <- trap_depths_ts %>% 
  ggplot(aes(date,depthpos,ymax=depthpos+se_depth,ymin=depthpos-se_depth))+
  geom_line()+
  scale_x_date(date_breaks= "6 months",date_labels="%b-%Y",expand=c(0,0))+
  geom_ribbon(fill='red',alpha=0.7)+
  labs(x="Date",y="Mean Trap Depth (m)")+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())+
  facet_wrap(~coast_region,nrow=4)
trap_depths_ts_plot
```


```{r,fig.width=8,fig.height=8}
# trap_density_dist_nonexpanded <- all_obs_xy_no_expansion %>%  
#   # add an identifier for early, mid, or late season
#   mutate(season_segment=case_when(
#     month(date) %in% c(11,12,1) ~ 'early',
#     month(date) %in% c(2,3,4) ~ 'mid',
#     month(date) %in% c(5,6,7,8,9) ~ 'late'
#   )) %>% 
#   mutate(statename=case_when(
#     STATE=="WA" ~ "Washington",
#     STATE=="CA" ~ "California",
#     STATE=="OR" ~ "Oregon"
#   )) %>% 
#   filter(month(date)!=10,!is.na(count),count>0,!is.na(statename)) %>% 
#   ggplot(aes(count/25,fill=season_segment,col=season_segment))+
#   geom_density(alpha=0.5,col=NA)+
#   scale_x_continuous(limits=c(0,10))+
#   labs(x="Density (Traps per sq. Km)",y="Kernel Density",title="Distribution of Positive Trap Densities\nall years and weeks (non-expanded)")+
#   facet_wrap(~statename)
# 
# trap_density_dist_expanded <- all_obs_xy_expanded %>%  
#   # add an identifier for early, mid, or late season
#   mutate(season_segment=case_when(
#     month(date) %in% c(11,12,1) ~ 'early',
#     month(date) %in% c(2,3,4) ~ 'mid',
#     month(date) %in% c(5,6,7,8,9) ~ 'late'
#   )) %>% 
#   mutate(statename=case_when(
#     STATE=="WA" ~ "Washington",
#     STATE=="CA" ~ "California",
#     STATE=="OR" ~ "Oregon"
#   )) %>% 
#   filter(month(date)!=10,!is.na(count),count>0,!is.na(statename)) %>% 
#   ggplot(aes(count/25,fill=season_segment,col=season_segment))+
#   geom_density(fill='red',alpha=0.98,col=NA)+
#   scale_x_continuous(limits=c(0,150))+
#   labs(x="Density (Traps per sq. Km)",y="Kernel Density",title="Distribution of Positive Trap Densities\nall years and weeks (expanded)")+
#   facet_wrap(~statename)
# trap_density_dist_nonexpanded
# trap_density_dist_expanded
```


# Simulated vs. Permitted Traps

Try to do some validation by looking at unweighted and weighted traps per trip, including the relationship between simulated and permitted traps per trip in the data.

```{r}
sim_vs_permit <- traplocs %>% 
  dplyr::select(Rec_ID,date,year,month,agency_code,FINAL_LENGTH,ntraps,ntraps_permitted,vms_representation_wt)
```

What is the relationship between number of permitted and number of simulated traps, per trip, by state?

```{r}
sim_vs_permit_p <- sim_vs_permit %>% 
  ggplot(aes(ntraps,ntraps_permitted))+
  geom_point()+
  geom_smooth(method='lm')+
  geom_abline(slope=1,intercept=0)+
  labs(x="Simulated Traps",y="Permitted Traps")+
  facet_wrap(~agency_code,nrow=2)
sim_vs_permit_p
```

Histogram of simulated pots per trip

```{r}
sim_vs_permit_hist <- sim_vs_permit %>% 
  ggplot(aes(ntraps))+
  geom_histogram(fill='darkblue',alpha=0.5,col=NA,binwidth = 50)+
  labs(x="Simulated Traps",y="")+
  facet_wrap(~agency_code,nrow=2)
sim_vs_permit_hist
```

Percent of trap limit fished, by month and state

```{r}
sim_vs_permit_months <- sim_vs_permit %>% 
  group_by(agency_code,year,month) %>% 
  summarise(perc_utilization=sum(ntraps,na.rm=T)/sum(ntraps_permitted,na.rm=T)) %>% 
  mutate(day=1) %>% 
  unite(date,year,month,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%B-%d")) %>% 
  ggplot(aes(date,perc_utilization))+
  geom_point()+
  geom_line()+
  facet_wrap(~agency_code,nrow=3)+
  scale_x_date(date_breaks= "6 months",date_labels="%b-%Y",expand=c(0,0))+
  labs(x="Month",y="Percent Utilization of Trap Limit")+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())
sim_vs_permit_months
```

Utilization is pretty high. How does this compare to TOTAL possible traps, based on all registered vessels?

Look at the sum of all traps for all registered vessels per season, and compare to the number of vessels/traps actually fished.

```{r}
max_traps_season <- fishtix_matched_all %>% 
  distinct(year,month,drvid) %>% 
  left_join(tier_key) %>% 
  distinct(year,month,drvid,agency_code,ntraps_permitted) %>% 
  group_by(year,month,agency_code,drvid) %>% 
  slice_max(ntraps_permitted,n=1) %>% 
  group_by(agency_code,year,month) %>% 
  summarise(max_potential_traps=sum(ntraps_permitted)) %>% 
  ungroup() %>% 
  mutate(day=1) %>% 
  unite(date,year,month,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%B-%d"))

# add the total, expanded, simualted traps from the VLM
trap_counts_state_month_expanded %>% 
  rename(year=yr,month=mth) %>% 
  mutate(month=month.name[month]) %>% 
  mutate(agency_code=case_when(
    STATE=="California" ~ "C",
    STATE=="Washington" ~ "W",
    STATE=="Oregon" ~ "O"
  )) %>% 
  right_join(max_traps_season)->max_traps_season

max_traps_season %>% 
  ungroup() %>% 
  pivot_longer(cols=c(tottraps,max_potential_traps),names_to="type",values_to="traps") %>% 
  ggplot(aes(date,traps/1000,col=type))+
  geom_point()+
  geom_line()+
  facet_wrap(~agency_code,nrow=3)+
  scale_x_date(date_breaks= "6 months",date_labels="%b-%Y",expand=c(0,0))+
  labs(x="Month",y="Number of Traps (1000s)")+
  theme(axis.text.x.bottom = element_text(angle=90,vjust=+0.5,hjust=0),
        panel.grid.minor = element_blank())

```
The gap is larger for Washington. I'm not sure why this is. Maybe look at the weighting schemes?

```{r}
wts_ts <- traplocs %>% 
  dplyr::select(Rec_ID,date,year,month,agency_code,FINAL_LENGTH,ntraps,ntraps_permitted,ntraps_wt,vms_representation_wt) %>% 
  mutate(comb_wt=ntraps_wt*vms_representation_wt) %>% 
  group_by(year,month,agency_code) %>% 
  summarise(mean_ntraps_wt=mean(ntraps_wt,na.rm=T),
            mean_vms_rep_wt=mean(vms_representation_wt,na.rm=T),
            mean_comb_wt=mean(comb_wt,na.rm=T)) %>% 
  ungroup() %>%  
  mutate(day=1) %>% 
  unite(date,year,month,day,sep="-",remove = FALSE) %>% 
  mutate(date=as_date(date,format="%Y-%B-%d"))

wts_ts %>% 
  ggplot(aes(date,mean_ntraps_wt,col=agency_code))+
  geom_point()+
  geom_line()+
  labs(x="Month",y="Traps per week weighting factor")

wts_ts %>% 
  ggplot(aes(date,mean_vms_rep_wt,col=agency_code))+
  geom_point()+
  geom_line()+
  scale_y_log10()+
  labs(x="Month",y="VMS representation weighting factor")
wts_ts %>% 
  ggplot(aes(date,mean_comb_wt,col=agency_code))+
  geom_point()+
  geom_line()+
  labs(x="Month",y="Combined weighting factor")+
  scale_y_log10()
```

Scatter sum(pots per trip) by week, then compare to permits (i.e. investigate calculation of ntraps_wt)

```{r}
test <- traplocs %>% 
  filter(drvid=="264147",period_1wk==1)
```

