---
title: "Pre-Process 2021"
author: "Leena Riekkola"
date: "7/12/2022"
output: html_document
---
# Purpose

This code is based on the previous pre-processing codes that were in this folder

Import and join together all months' VMS data for 2021 -->  data already provided in a single file. 
Then, we rename column headers to match other years.

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(magrittr)
```

## Import and Bind Rows

We read date-times as character for now

```{r import2019}
#2021 vms data is already in one file, but only goes upto October

all2021_raw <- read_csv(here::here('data','raw','vms', 'VMS 2021 data',paste0("vms_01012021_10042021.csv")),col_types='ccccddddddccc') 
glimpse(all2021_raw)


```


## Fix Column Names and Reorder

We choose only some columns for now, to match the VMs data we have for other years.

```{r}
# column names from previous years
correct_colnames <- c('UTCDATETIM', 'LATITUDE', 'LONGITUDE', 'VESSEL_NAM', 'AVG_SPEED', 
                      'AVG_COURSE', 'DOCNUM', 'DECLARATIO')
correct_colnames


all2021 <- all2021_raw %>% 
  select(UTC_TIME, LAT, LON, VESSEL_NAME, AVG_SPEED, AVG_COURSE, DOCNUM, DECLARATION_CODE) %>% 
  set_colnames(correct_colnames)

glimpse(all2021)
```

## Save Output

```{r}
write_csv(all2021,here::here('data','raw','vms','VMS_all_data_2021.csv'))
```