---
title: "Interpolate VMS Data"
output: html_document
---

## Purpose

In this step, we produce a version of the VMS data that is interpolated such that records are evenly distributed every hour.

First, the script filters for time gaps over which it is unlikely that linear interpolation will be representative. According to the following rules:

*If the maximum time gap is greater than 4 hours OR composes more than 25 percent of the trip, Remove the trip. UNLESS...
*The maximum time gap occurs when the vessel is in port. In these cases, keep the trip
*The maximum time gap is the last record, retain it. This may be due to overland transport to a different port.

We then interpolate the data using the `move` package. This regularizes the data such that we have a VMS ping every hour. After interpolation is done, we re-calculate average speeds for each trip segment, as well as re-join the bathymetry layer.

## Setup and Data Import

Clear workspace
```{r, echo=FALSE,message=FALSE}
library(tidyverse)
library(magrittr)
library(here)
library(lubridate)
library(sf)
library(move)
library(rgdal)

# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=12,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())
theme_set(plot_theme)

rm(list=setdiff(ls(),c('process_year','alltime')))

#if needed, process_year
# process_year=2012
```

Read in the filtered VMS data set from the previous step

```{r rawdat}
#Edited to read in data with FTID and length
vms_filtered <- read_rds(here::here('data','processed','matched','filtering',paste0(process_year,'matched_filtered_withFTID_length.rds')))
```

## Remove Trips with Large Time Gaps

Find trips whose maximum time gap is greater than 4 hours or encompasses more than 40 percent of the trip

```{r}
# random_trip <- vms_filtered %>% filter(TARGET_rev=='DCRB') %>% distinct(Rec_ID) %>% sample_n(3) %>% pull(Rec_ID)

vms_gapfiltered <- vms_filtered %>%
  # filter(Rec_ID %in% random_trip) %>% 
  group_by(Rec_ID) %>% 
  # overall trip duration
  mutate(trip_dur=sum(segment_dur)) %>% 
  # add a sequential record number
  arrange(westcoastdate) %>% 
  mutate(recnum=row_number()) %>% 
  #indicate whether to keep or remove trip based on segment durations
  mutate(keep_remove_timelag=ifelse(max(segment_dur,na.rm=T) > (4*3600) | (segment_dur/trip_dur) > 0.40, "remove", "keep")) %>% 
  #indicate whether the reason for removal was because of the last segment
  mutate(longest_seg=first(recnum,order_by=desc(segment_dur))) %>% 
  mutate(last_seg_longest=longest_seg==max(recnum)) %>% 
  mutate(is_last_seg=ifelse(last_seg_longest&keep_remove_timelag=='remove',TRUE,FALSE)) %>% 
  ungroup()

trips_removed_table <- vms_gapfiltered %>%
  filter(keep_remove_timelag=='remove') %>% 
  dplyr::select(Rec_ID,last_seg_longest,is_last_seg) %>% 
  distinct()

ntottrips <- length(unique(vms_filtered$Rec_ID))
nremovedtrips <- length(unique(trips_removed_table$Rec_ID))
nremovedlastseg <- trips_removed_table %>% filter(is_last_seg) %>% nrow()
```

With these criteria, out of `r ntottrips` unique trips in the filtered dataset, we flagged `r nremovedtrips` for removal, with `r nremovedlastseg` of these because of the last segment in the trip. For now, we instead retain these `r nremovedlastseg` trips for investigation later.

Now we can filter the dataset.

```{r}
vms_gapfiltered %<>%
  # change trips flagged for removal because of last segment to "keep" instead
  mutate(keep_remove_timelag=ifelse(is_last_seg,'keep',keep_remove_timelag)) %>% 
  filter(keep_remove_timelag=='keep')
```

## Interpolate

Interpolate, using package `move`.

Arrange by trip and ascending in time.

```{r}
vms_move <- vms_gapfiltered %>% 
  ungroup() %>% 
  arrange(Rec_ID, UTCDATETIM) %>% 
  distinct(VMS_RECNO,UTCDATETIM,.keep_all = T) %>% 
  mutate(DECLARATIO=na_if(DECLARATIO,'N/A')) %>% 
  as.data.frame()
```

### Create `move` Object

```{r move}
#This does not work
#start.time <- Sys.time()
#movedat <- move(x=vms_move$X_COORD,
#                y=vms_move$Y_COORD,
#                time=vms_move$UTCDATETIM,
#                data=vms_move,
#                proj=sp::CRS("+proj=utm +north +zone=10 +ellps=WGS84"),
#                animal=vms_move$Rec_ID)
#Sys.time() - start.time


#The version of code that works for Leena - chunk starting from line 110
#start.time <- Sys.time()
#movedat <- move(x=vms_move$LONGITUDE,
#                y=vms_move$LATITUDE,
#                time=vms_move$UTCDATETIM,
#                data=vms_move,
#                proj=sp::CRS("+proj=longlat +ellps=WGS84 +datum=WGS84"),
#                animal=vms_move$Rec_ID)
#Sys.time() - start.time

#movedatProj <- spTransform(movedat, CRSobj="+proj=utm +north +zone=10 +ellps=WGS84")
# Seems to work although at the end still says Discarded datum Unknown based on WGS84 ellipsoid in Proj4 definition


#Mary's solution
#start.time <- Sys.time()
#movedat2 <- move(x=vms_move$X_COORD,
#                y=vms_move$Y_COORD,
#                time=vms_move$UTCDATETIM,
#                data=vms_move,
#                proj=sp::CRS("+proj=utm +zone=10 +datum=WGS84 +units=m +no_defs +type=crs"),
#                animal=vms_move$Rec_ID)
#Sys.time() - start.time
# Doesn't work - Error in SpatialPointsDataFrame(coords = cbind(x, y, deparse.level = 0),  :   row.names of data and coords do not match

##THIS IS TE ONE COPIED INTO THE OFFICIAL PROCESS STEP
movedat <- move(x=vms_move$LONGITUDE,
                y=vms_move$LATITUDE,
                time=vms_move$UTCDATETIM,
                data=vms_move,
                proj=sp::CRS("+proj=longlat +datum=WGS84 +no_defs +type=crs"),
                animal=vms_move$Rec_ID)

movedatProj <- spTransform(movedat, CRSobj="+proj=utm +zone=10 +datum=WGS84 +units=m +no_defs +type=crs")
# This works best, no messages at all


#Jameals version
#sp::coordinates(vms_move) <- c(38,39)
#start.time <- Sys.time()
#movedat4 <- move(x=coordinates(vms_move)[,1], #[1:100,1]
#                y=coordinates(vms_move)[,2], #[1:100,2]
#                time=as.data.frame(vms_move)$UTCDATETIM,#[1:100]
#                data=as.data.frame(vms_move), #[1:100,]
#                proj=sp::CRS("+proj=utm +north +zone=10 +datum=WGS84 +units=m +no_defs +type=crs"),
#                animal=as.data.frame(vms_move)$Rec_ID) #[1:100]
#Sys.time() - start.time
#This also work, no messages, BUT need to make vms_move back into a data.frame for later code to work
#vms_move <- as.data.frame(vms_move)
```

### Run Interpolation

Starting from a MoveStack object

```{r}
ptm <- proc.time()
ntrips <- n.indiv(movedatProj)
#ntrips <- n.indiv(movedat4)
interpolated_data <- purrr::map_df(1:ntrips,function(i){

  tmp_movedat <- movedatProj[[i]]
  #tmp_movedat <- movedat4[[i]]
  
  inter_movedat <- interpolateTime(x=tmp_movedat, time=as.difftime(60, units="mins"), spaceMethod='euclidean')
  inter_df <- as.data.frame(inter_movedat)
  #clean up the data frame
  inter_df <- inter_df %>% 
    dplyr::select(Rec_ID, drvid, UTCDATETIM, X_COORD, Y_COORD, timestamps, coords.x1, coords.x2, sensor) %>% 
    mutate_if(is.factor,as.character)
  if(i%%500==0){
    message("completed interpolation for ", i, " out of ", ntrips, " trips.")
    }
  return(inter_df)
})
tm <- proc.time()-ptm

#interpolated_data created based on Mary's move object. Interpolation took 16.25minutes
#interpolated_data2 created based on Jameal's move object. Interpolation took 16.79minutes
#identical(interpolated_data,interpolated_data2)
#[1] FALSE
#all.equal(interpolated_data,interpolated_data2)
#[1] "Component “X_COORD”: names for target but not for current"
#[2] "Component “Y_COORD”: names for target but not for current"
```

Interpolation took taken `r round(tm[3]/60,2)` minutes to run.

### Integrate Interpolated Data

Rename columns from interpolation so we can re-join the rest of the VMS trip information. For original (non-interpolated) data, we can re-join with times and coordinates. For newly-interpolated data, we define the new time and coordinates and then join the trip-level information.

```{r}
#orig_vms2 <- interpolated_data2 %>%
#  filter(sensor=="unknown") %>% 
#  dplyr::select(-UTCDATETIM,-X_COORD,-Y_COORD) %>% 
#  rename(UTCDATETIM=timestamps,X_COORD=coords.x1,Y_COORD=coords.x2) %>% 
#  left_join(vms_move,by = c("Rec_ID", "drvid", "UTCDATETIM", "X_COORD", "Y_COORD"))
#This version gives NAs for Mary's way, but 'works' for Jameal's way

#Other version -- USE THIS!
orig_vms <- interpolated_data %>%
  filter(sensor=="unknown") %>% 
  left_join(vms_move,by = c("Rec_ID", "drvid", "UTCDATETIM", "X_COORD", "Y_COORD")) %>% 
  dplyr::select(-UTCDATETIM,-X_COORD,-Y_COORD) %>% 
  rename(UTCDATETIM=timestamps,X_COORD=coords.x1,Y_COORD=coords.x2)
#This works on Mary's way
# identical(orig_vms, orig_vms2)
#[1] FALSE
# all.equal(orig_vms, orig_vms2)
#[1] TRUE


vms_trip_info <- vms_move %>% 
  dplyr::select(-VMS_RECNO,-UTCDATETIM,-LATITUDE,-LONGITUDE,-NGDC_M,-X_COORD,-Y_COORD,-AVG_SPEED,-AVG_COURSE,-westcoastdate,-westcoastdate_notime,-segment_dur,-avg_speed_recalc,-in_port,-recnum,-keep_remove_timelag,-longest_seg,-last_seg_longest,-is_last_seg) %>%
  distinct()


new_vms <- interpolated_data %>% 
  filter(sensor=="interpolateTime")%>% 
  dplyr::select(-UTCDATETIM,-X_COORD,-Y_COORD) %>% 
  rename(UTCDATETIM=timestamps,X_COORD=coords.x1,Y_COORD=coords.x2) %>% 
  left_join(vms_trip_info,by = c("Rec_ID", "drvid"))
#Mary's way

#new_vms2 <- interpolated_data2 %>% 
#  filter(sensor=="interpolateTime")%>% 
#  dplyr::select(-UTCDATETIM,-X_COORD,-Y_COORD) %>% 
#  rename(UTCDATETIM=timestamps,X_COORD=coords.x1,Y_COORD=coords.x2) %>% 
#  left_join(vms_trip_info,by = c("Rec_ID", "drvid"))
#Jameal's way
# identical(new_vms, new_vms2)
#[1] FALSE
# all.equal(new_vms, new_vms2)
#[1] TRUE
```

Create calculable missing columns: west coast dates, latitude, longitude.

```{r}
# west coast dates
new_vms %<>%
  mutate(westcoastdate = with_tz(UTCDATETIM, tzone = "America/Los_Angeles"),
         westcoastdate_notime = as_date(westcoastdate))

# convert UTM x/y coordinates to lat/lon
new_vms_coords <- new_vms %>% 
  st_as_sf(coords=c("X_COORD","Y_COORD"),crs="+proj=utm +north +zone=10 +ellps=WGS84") %>% 
  st_transform(4326) %>% 
  st_coordinates() %>% 
  as_tibble() %>% 
  set_colnames(c('LONGITUDE','LATITUDE'))

# add the new lat/lon into the dataset
new_vms %<>%
  bind_cols(new_vms_coords)


#For Jameal's way
# west coast dates
#new_vms2 %<>%
#  mutate(westcoastdate = with_tz(UTCDATETIM, tzone = "America/Los_Angeles"),
#         westcoastdate_notime = as_date(westcoastdate))

# convert UTM x/y coordinates to lat/lon
#new_vms_coords2 <- new_vms2 %>% 
#  st_as_sf(coords=c("X_COORD","Y_COORD"),crs="+proj=utm +north +zone=10 +ellps=WGS84") %>% 
#  st_transform(4326) %>% 
#  st_coordinates() %>% 
#  as_tibble() %>% 
#  set_colnames(c('LONGITUDE','LATITUDE'))

# add the new lat/lon into the dataset
#new_vms2 %<>%
#  bind_cols(new_vms_coords2)
```

Add in VMS Record Numbers for new VMS points

```{r}
new_recnos <- seq(from=(max(vms_move$VMS_RECNO)+10), length.out=nrow(new_vms))
new_vms %<>% mutate(VMS_RECNO = new_recnos)


#For Jameal's way
#new_recnos2 <- seq(from=(max(vms_move$VMS_RECNO)+10), length.out=nrow(new_vms2))
#new_vms2 %<>% mutate(VMS_RECNO = new_recnos2)
```

Recombine
```{r}
vms_regular <- bind_rows(orig_vms, new_vms)

# check record numbers to make sure they're unique
length(unique(vms_regular$VMS_RECNO))/length(vms_regular$VMS_RECNO)

#returns 1 for Mary's way


#For Jameal's way
#vms_regular2 <- bind_rows(orig_vms2, new_vms2)

# check record numbers to make sure they're unique
#length(unique(vms_regular2$VMS_RECNO))/length(vms_regular2$VMS_RECNO)

#returns 1 for Jameal's way
```

## Re-calculate Speeds

Finally, just as in the previous step for the non-regularized data, we calculate speeds for each interpolated segment.

```{r}
vms_regular <- vms_regular %>% 
  ungroup() %>% 
  group_by(Rec_ID) %>% 
  # lag latitude and longitude by 1 time step
  mutate(laglon=lag(LONGITUDE,1,order_by=westcoastdate),laglat=lag(LATITUDE,1,order_by=westcoastdate)) %>% 
  # lag time by 1 time step
  mutate(lagdate=lag(westcoastdate,1,order_by=westcoastdate)) %>% 
  # calculate duration since last ping, in seconds
  mutate(segment_dur=as.duration(lagdate %--% westcoastdate)/dseconds()) %>% 
  ungroup()

# Calculate distance (Note: geosphere seems much faster than doing this with sf())
segment_dists <- geosphere::distHaversine(p1=cbind(vms_regular$LONGITUDE, vms_regular$LATITUDE),
                                    p2=cbind(vms_regular$laglon, vms_regular$laglat))

vms_regular %<>% 
  mutate(segment_dist=segment_dists)

# Speed is just segment distance (default in meters) divided by segment duration (in seconds)
vms_regular %<>%
  mutate(avg_speed_recalc=segment_dist/segment_dur) %>% 
  # some calculations will be NaN or Inf because of 0 distance or time. Fix these as zeroes
  mutate(avg_speed_recalc=ifelse(segment_dist==0|segment_dur==0,0,avg_speed_recalc)) %>% 
  # select out columns we don't want (columns that were just used for intermediate calculations)
  dplyr::select(-(recnum:lagdate))



#For Jameal's way
#vms_regular2 <- vms_regular2 %>% 
#  ungroup() %>% 
#  group_by(Rec_ID) %>% 
  # lag latitude and longitude by 1 time step
#  mutate(laglon=lag(LONGITUDE,1,order_by=westcoastdate),laglat=lag(LATITUDE,1,order_by=westcoastdate)) %>% 
  # lag time by 1 time step
#  mutate(lagdate=lag(westcoastdate,1,order_by=westcoastdate)) %>% 
  # calculate duration since last ping, in seconds
#  mutate(segment_dur=as.duration(lagdate %--% westcoastdate)/dseconds()) %>% 
#  ungroup()

# Calculate distance (Note: geosphere seems much faster than doing this with sf())
#segment_dists2 <- geosphere::distHaversine(p1=cbind(vms_regular2$LONGITUDE, vms_regular2$LATITUDE),
#                                    p2=cbind(vms_regular2$laglon, vms_regular2$laglat))

#vms_regular2 %<>% 
#  mutate(segment_dist=segment_dists)

# Speed is just segment distance (default in meters) divided by segment duration (in seconds)
#vms_regular2 %<>%
#  mutate(avg_speed_recalc=segment_dist/segment_dur) %>% 
  # some calculations will be NaN or Inf because of 0 distance or time. Fix these as zeroes
#  mutate(avg_speed_recalc=ifelse(segment_dist==0|segment_dur==0,0,avg_speed_recalc)) %>% 
  # select out columns we don't want (columns that were just used for intermediate calculations)
#  dplyr::select(-(recnum:lagdate))
```

## Add Bathymetry Again

We re-match to the bathymetry layer but do NO FILTERING for now.

Read in the bathymetry SpatialGridDataFrame object

```{r}
bathy.grid <- read_rds(here::here('data','raw','vms_composite_bath_spGrid.RDS'))
```

Get bathymetry at VMS data points

```{r}
# tm <- proc.time()
vms_sp <- vms_regular
coordinates(vms_sp) <- c("LONGITUDE", "LATITUDE") 
proj4string(vms_sp)<-proj4string(bathy.grid) <- CRS("+init=epsg:4326") # WGS 84
bathy.points <- over(vms_sp, bathy.grid)$vms_composite_bath
# proc.time()-tm

vms_regular %<>% mutate(NGDC_M = bathy.points)




#For Jameal's way
# tm <- proc.time()
#vms_sp2 <- vms_regular2
#coordinates(vms_sp2) <- c("LONGITUDE", "LATITUDE") 
#proj4string(vms_sp2)<-proj4string(bathy.grid) <- CRS("+init=epsg:4326") # WGS 84
#bathy.points2 <- over(vms_sp2, bathy.grid)$vms_composite_bath
# proc.time()-tm

#vms_regular2 %<>% mutate(NGDC_M = bathy.points)


#The bathy step gives the following message for both solution options (during line CRS("+init=espg:4326")) -- proj4string thing - can this be altered for PROJ6?
#A new CRS was assigned to an object with an existing CRS:
#+proj=longlat +datum=WGS84 +no_defs +ellps=WGS84 +towgs84=0,0,0
#without reprojecting.
#For reprojection, use function spTransform
```

## Save Results

Add a final indicator for whether a point was interpolated or not, then save

```{r}
vms_regular %<>% mutate(is_interpolated=ifelse(sensor=="interpolateTime",1,0)) %>% dplyr::select(-sensor)
#vms_regular2 %<>% mutate(is_interpolated=ifelse(sensor=="interpolateTime",1,0)) %>% dplyr::select(-sensor)


write_rds(vms_regular,here::here('data','processed','matched','interpolation',paste0(process_year,'interpolated_withFTID_length.rds')))
#write_rds(vms_regular,here::here('data','processed','matched','interpolation',paste0(process_year,'interpolated_MaryV2.rds')))
#write_rds(vms_regular2,here::here('data','processed','matched','interpolation',paste0(process_year,'interpolated_Jameal.rds')))

```

