---
title: "Step6_comparisons"
author: "Leena"
date: "04/03/2021"
output: html_document
---

```{r}
library(tidyverse)
library(magrittr)
library(here)
library(janitor)
library(sf)
library(rnaturalearth)
library(lubridate)
library(move)


```
Package and Library versions on Leena's computer

sessionInfo()
R version 3.6.3 (2020-02-29)
Platform: x86_64-w64-mingw32/x64 (64-bit)
Running under: Windows 10 x64 (build 19041)

Matrix products: default

locale:
[1] LC_COLLATE=English_New Zealand.1252  LC_CTYPE=English_New Zealand.1252   
[3] LC_MONETARY=English_New Zealand.1252 LC_NUMERIC=C                        
[5] LC_TIME=English_New Zealand.1252    

attached base packages:
[1] stats     graphics  grDevices utils     datasets  methods   base     

other attached packages:
 [1] move_4.0.6          rgdal_1.5-23        raster_3.4-5        sp_1.4-1           
 [5] geosphere_1.5-10    lubridate_1.7.9.2   rnaturalearth_0.1.0 sf_0.9-7           
 [9] janitor_2.1.0       here_1.0.1          magrittr_2.0.1      forcats_0.5.1      
[13] stringr_1.4.0       dplyr_1.0.4         purrr_0.3.4         readr_1.4.0        
[17] tidyr_1.1.2         tibble_3.0.6        ggplot2_3.3.3       tidyverse_1.3.0    

loaded via a namespace (and not attached):
 [1] httr_1.4.2               jsonlite_1.7.2           modelr_0.1.8            
 [4] rnaturalearthhires_0.2.0 assertthat_0.2.1         cellranger_1.1.0        
 [7] yaml_2.2.1               pillar_1.4.7             backports_1.2.1         
[10] lattice_0.20-38          glue_1.4.2               digest_0.6.27           
[13] rvest_0.3.6              snakecase_0.11.0         colorspace_2.0-0        
[16] htmltools_0.5.1.1        pkgconfig_2.0.3          broom_0.7.4             
[19] haven_2.3.1              scales_1.1.1             generics_0.1.0          
[22] farver_2.0.3             ellipsis_0.3.1           cachem_1.0.4            
[25] withr_2.4.1              cli_2.3.0                crayon_1.4.1            
[28] readxl_1.3.1             memoise_2.0.0            evaluate_0.14           
[31] fs_1.5.0                 fansi_0.4.2              xml2_1.3.2              
[34] class_7.3-15             tools_3.6.3              hms_1.0.0               
[37] lifecycle_1.0.0          munsell_0.5.0            reprex_1.0.0            
[40] compiler_3.6.3           e1071_1.7-4              tinytex_0.29            
[43] rlang_0.4.10             classInt_0.4-3           units_0.6-7             
[46] grid_3.6.3               rstudioapi_0.13          rmarkdown_2.6           
[49] gtable_0.3.0             codetools_0.2-16         DBI_1.1.1               
[52] R6_2.5.0                 knitr_1.31               fastmap_1.1.0           
[55] rgeos_0.5-5              utf8_1.1.4               rprojroot_2.0.2         
[58] KernSmooth_2.23-16       stringi_1.5.3            parallel_3.6.3          
[61] Rcpp_1.0.6               vctrs_0.3.6              dbplyr_2.1.0            
[64] tidyselect_1.1.0         xfun_0.21     


lib_versions <- sf::sf_extSoftVersion() 
lib_versions 
          GEOS           GDAL         proj.4 GDAL_with_GEOS     USE_PROJ_H 
       "3.8.0"        "3.0.4"        "6.3.1"         "true"         "true" 
```{r}
#2014 interpolated vms from Owens pipeline run
owendat <- read_rds(here::here('data','processed','matched','interpolation','2014interpolated_Owen.rds'))

#2014 interpolated vms using Mary's way of creating move object
marydat <- read_rds(here::here('data','processed','matched','interpolation','2014interpolated_MaryV2.rds'))

#2014 interpolated vms using Jameals's way of creating move object
jamealdat <- read_rds(here::here('data','processed','matched','interpolation','2014interpolated_Jameal.rds'))


```


```{r}
identical(owendat, marydat)
identical(owendat, jamealdat)
identical(marydat, jamealdat)
#False for all

all.equal(owendat, marydat) #Gives a list of 'discrepancies'
all.equal(owendat, jamealdat) #Gives a list of 'discrepancies'
all.equal(marydat, jamealdat) #Gives a list of 'discrepancies'

```

```{r}

#no. and names of columns are same
identical(colnames(owendat), colnames(marydat)) #TRUE
identical(colnames(owendat), colnames(jamealdat)) #TRUE

compare_df_cols(owendat, marydat, jamealdat, return = "mismatch")
#only difference is that some columns are character columns in owendat and marydat, but factor columns in jamealdat

```

```{r}
nrow(owendat)
nrow(marydat)
nrow(jamealdat)

#Owen version has 145 more rows than mary or jameal dataset (which have same number of rows)
nrow(owendat)-nrow(marydat)
```

```{r}
summary(owendat)
summary(marydat)
summary(jamealdat)

#Based on visual comparison of summaries, there really doesn't seem to be any differences between marydat and jamealdat...

#owendat min utcdatetim 2013-12-28 15:42:00
#marydat min utcdatetim 2014-01-01 08:04:00
#this might relate to Step 4, section 'Load VMS data, including the last month of the previous year (for handling trips spanning January 1st of the current year)'
#It is quite likely I have done something wrong along the way and didn't accurately bring in data for trips that started in late December 2013
#Which would explain the discrepancy between owendat and marydat/jamealdat at least to some extent
#Even if looking at only 2014 data, the first observation in owendat is 2014-01-01 00:10:00 and in marydat 2014-01-01 089:04:00 -- so my version of things is missing some of the early data


owendat %<>% mutate(year = year(UTCDATETIM)) 
owendat$year <- as.factor(owendat$year)
summary(owendat$year)
#292 records with year 2013, which is more than the 145 row difference between the data frames...


```
```{r}
owendat_tail <- owendat %>% 
  arrange(UTCDATETIM) %>% 
  slice_tail(n = 100000)

marydat_tail <- marydat %>% 
  arrange(UTCDATETIM) %>% 
  slice_tail(n = 100000)

summary(owendat_tail)
summary(marydat_tail)
#If just compare sub-sample from the end of (ordered) datasets, they seem the same

```


Sum of the number of VMS point per fish ticket, if map Mary and Owen data in x and y in a plot, it should run as a 1-to-1 line.
```{r}

VMSpoints_owen <- owendat %>% group_by(Rec_ID) %>% summarise(n_owen = n())
VMSpoints_mary <- marydat %>% group_by(Rec_ID) %>% summarise(n_mary = n())

joined <- left_join(VMSpoints_owen,VMSpoints_mary, by=c("Rec_ID"))

joined %>% 
  ggplot(aes(n_owen,n_mary))+
  geom_point()+
  geom_smooth(method='lm')+
  labs(x='VMS by tix_owen',y="VMS by tix_mary")
#Almost all points fall on a 1-to-1 line, 10 points that don't - quite likely related to mary data not having point in Dec 2013
```

```{r}
#QGIS mapping seems better

owendat_sf <- owendat %>% st_as_sf(coords=c("X_COORD","Y_COORD"),crs = "+proj=utm +north +zone=10 +ellps=WGS84")
marydat_sf <- marydat %>% st_as_sf(coords=c("X_COORD","Y_COORD"),crs = "+proj=utm +north +zone=10 +ellps=WGS84")
jamealdat_sf <- jamealdat %>% st_as_sf(coords=c("X_COORD","Y_COORD"),crs = "+proj=utm +north +zone=10 +ellps=WGS84")

# Try with a subset of 100k points from the end of data frame
owendat_sf_sample <- owendat_sf %>% 
  slice_tail(n = 100000)

marydat_sf_sample <- marydat_sf %>% 
  slice_tail(n = 100000) 

jamealdat_sf_sample <- jamealdat_sf %>% 
  slice_tail(n = 100000) 
  

# coastline
coaststates <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('California','Oregon','Washington')) %>% 
  st_transform(st_crs(owendat_sf))

map1 <- ggplot()+
  geom_sf(data=coaststates,fill='gray50',col=NA)+
  geom_sf(data=owendat_sf_sample,size=0.5,col='blue')+
  geom_sf(data=marydat_sf_sample,size=0.5,col='red')+
  labs(x='Longitude',y='Latitude',title=paste0(process_year," VMS records"))

map2 <- ggplot()+
  geom_sf(data=coaststates,fill='gray50',col=NA)+
  geom_sf(data=marydat_sf_sample,size=0.5,col='red')+
  geom_sf(data=owendat_sf_sample,size=0.5,col='blue')+
  labs(x='Longitude',y='Latitude',title=paste0(process_year," VMS records"))

map3 <- ggplot()+
  geom_sf(data=coaststates,fill='gray50',col=NA)+
  geom_sf(data=marydat_sf_sample,size=0.5,col='red')+
  geom_sf(data=jamealdat_sf_sample,size=0.5,col='green')+
  labs(x='Longitude',y='Latitude',title=paste0(process_year," VMS records"))



map1
map2
map3

#maps between marydat and jamealdat seem the same
#owendat is missing some points that exists in marydat & jamealdat
```

```{r}
#Tried one vessel in one month
#Not familiar with sf and mapping in R so these weren't great PLUS mapping was better/more informative on QGIS

#Add extra column for month
#owendat_sf %<>% mutate(Month = month(UTCDATETIM))
#owendat_may <- owendat_sf %>% 
 # filter(Month == 5)
#owendat_may_580792 <- owendat_may %>% 
 # filter(drvid == 580792)

#marydat_sf %<>% mutate(Month = month(UTCDATETIM))
#marydat_may <- marydat_sf %>% 
 # filter(Month == 5)
#marydat_may_580792 <- marydat_may %>% 
 # filter(drvid == 580792)

#jamealdat_sf %<>% mutate(Month = month(UTCDATETIM))
#jamealdat_may <- jamealdat_sf %>% 
 # filter(Month == 5)
#jamealdat_may_580792 <- jamealdat_may %>% 
 # filter(drvid == 580792)

#coaststates2 <- ne_states(country='United States of America',returnclass = 'sf') %>% 
#  filter(name %in% c('Oregon','Washington')) %>% 
#  st_transform(st_crs(owendat_sf))
#coaststatescropped <- st_crop(coaststates2, xmin = -126, xmax = 122, ymin = 40, ymax = 45)

#may_580792_map <- ggplot()+
# geom_sf(data=coaststatescropped,fill='gray50',col=NA)+
#  geom_sf(data=owendat_may_580792,size=0.8,col='blue')+
#  geom_sf(data=marydat_may_580792,size=0.7,color=alpha("red",0.3))+
  #coord_sf(datum = 9001)+
  #coord_sf(ylim = c(4500000, 5000000), xlim = c(400000, 600000)) +
#  labs(x='Longitude',y='Latitude',title=paste0(process_year,"Vessel 580792 in May")) 

#Trying to figure out how to zoom - coaststates and the samples have crs
#issues with ylim and xlim part
#geom_sf by default changes things to lat longs from UTM

#may_580792_map <- ggplot()+
#  geom_sf(data=coaststatescropped,fill='gray50',col=NA)+
#  geom_sf(data=jamealdat_may_580792,size=0.8,col='green')+
#  geom_sf(data=marydat_may_580792,size=0.7,color=alpha("red",0.3))+
  #coord_sf(datum = 9001)+
  #coord_sf(ylim = c(4500000, 5000000), xlim = c(400000, 600000)) +
#  labs(x='Longitude',y='Latitude',title=paste0(process_year,"Vessel 580792 in May")) 


```




```{r}
#Sample of one vessel in one month
#write_csv(owendat_may_580792,here::here('data','processed','matched','interpolation','owendat_may_580792.csv'))
#write_csv(marydat_may_580792,here::here('data','processed','matched','interpolation','marydat_may_580792.csv'))
#write_csv(jamealdat_may_580792,here::here('data','processed','matched','interpolation','jamealdat_may_580792.csv'))
```


```{r}
#Add column for month, save entire datasets as csv for QGIS
owendat %<>% mutate(month = month(UTCDATETIM))
marydat %<>% mutate(month = month(UTCDATETIM))
jamealdat %<>% mutate(month = month(UTCDATETIM))

write_csv(owendat,here::here('data','processed','matched','interpolation','owendat.csv'))
write_csv(marydat,here::here('data','processed','matched','interpolation','marydat.csv'))
write_csv(jamealdat,here::here('data','processed','matched','interpolation','jamealdat.csv'))
```

Comparing 2018 VMS pipeline output at each step

Step 1


```{r}
#2018 step 1 from Owens pipeline run
owendat2018_step1 <- read_rds(here::here('data','processed','fish tickets','2018fishtix_Owen.rds'))

#2018 step 1 from my pipeline run
Leenadat2018_step1 <- read_rds(here::here('data','processed','fish tickets','2018fishtix.rds'))

```

```{r}
identical(owendat2018_step1, Leenadat2018_step1) #True

all.equal(owendat2018_step1, Leenadat2018_step1) #True

identical(colnames(owendat2018_step1), colnames(Leenadat2018_step1)) #TRUE

```
```{r}
summary(owendat2018_step1)
summary(Leenadat2018_step1)
```

```{r}

```
Step 2

```{r}
#2018 step 2 from Owens pipeline run
owendat2018_step2 <- read_rds(here::here('data','processed','fish tickets','2018fishtix_vlengths_Owen.rds'))

#2018 step 2 from my pipeline run
Leenadat2018_step2 <- read_rds(here::here('data','processed','fish tickets','2018fishtix_vlengths.rds'))
```

```{r}
identical(owendat2018_step2, Leenadat2018_step2) #True

all.equal(owendat2018_step2, Leenadat2018_step2) #True

identical(colnames(owendat2018_step2), colnames(Leenadat2018_step2)) #TRUE
```
```{r}
summary(owendat2018_step2)
summary(Leenadat2018_step2)
```
```{r}

```
Step 3

```{r}
#2018 step 3 from Owens pipeline run
owendat2018_step3 <- read_rds(here::here('data','processed','vms','2018_vms_clean_Owen.rds'))

#2018 step 3 from my pipeline run
Leenadat2018_step3 <- read_rds(here::here('data','processed','vms','2018_vms_clean.rds'))
```

```{r}
identical(owendat2018_step3, Leenadat2018_step3) #False

all.equal(owendat2018_step3, Leenadat2018_step3) #gives list of discrepancies

identical(colnames(owendat2018_step3), colnames(Leenadat2018_step3)) #TRUE

nrow(owendat2018_step3)
nrow(Leenadat2018_step3) #same number of rows
```
```{r}
summary(owendat2018_step3)
summary(Leenadat2018_step3)
```
```{r}
#points_owen2018 <- owendat2018_step3 %>% group_by(Rec_ID) %>% summarise(n_owen = n())
#points_Leena2018 <- Leenadat2018_step3 %>% group_by(Rec_ID) %>% summarise(n_leena = n())
points_owen2018 <- owendat2018_step3 %>% group_by(VESSEL_NAM) %>% summarise(n_owen = n())
points_Leena2018 <- Leenadat2018_step3 %>% group_by(VESSEL_NAM) %>% summarise(n_leena = n())

#joined <- left_join(points_owen2018,points_Leena2018, by=c("Rec_ID"))
joined <- left_join(points_owen2018,points_Leena2018, by=c("VESSEL_NAM"))

joined %>% 
  ggplot(aes(n_owen,n_leena))+
  geom_point()+
  geom_smooth(method='lm')+
  labs(x='points by tix_owen',y="points by tix_leena")
```
Step 3

```{r}
#2018 step 4 from Owens pipeline run
owendat2018_step4_vmstixonly <- read_rds(here::here('data','processed','matched','matching','2018matched_vmstix_only_Owen.rds'))

#2018 step 4 from my pipeline run
Leenadat2018_step4_vmstixonly <- read_rds(here::here('data','processed','matched','matching','2018matched_vmstix_only.rds'))
```

```{r}
identical(owendat2018_step4_vmstixonly, Leenadat2018_step4_vmstixonly) #False

all.equal(owendat2018_step4_vmstixonly, Leenadat2018_step4_vmstixonly) #"Component “X_COORD”: names for current but not for target"

identical(colnames(owendat2018_step4_vmstixonly), colnames(Leenadat2018_step4_vmstixonly)) #TRUE

nrow(owendat2018_step4_vmstixonly)
nrow(Leenadat2018_step4_vmstixonly) #same number of rows
```
```{r}
summary(owendat2018_step4_vmstixonly)
summary(Leenadat2018_step4_vmstixonly)
```
```{r}
points_owen2018 <- owendat2018_step4_vmstixonly %>% group_by(Rec_ID) %>% summarise(n_owen = n())
points_Leena2018 <- Leenadat2018_step4_vmstixonly %>% group_by(Rec_ID) %>% summarise(n_leena = n())
#points_owen2018 <- owendat2018_step3 %>% group_by(VESSEL_NAM) %>% summarise(n_owen = n())
#points_Leena2018 <- Leenadat2018_step3 %>% group_by(VESSEL_NAM) %>% summarise(n_leena = n())

joined <- left_join(points_owen2018,points_Leena2018, by=c("Rec_ID"))
#joined <- left_join(points_owen2018,points_Leena2018, by=c("VESSEL_NAM"))

joined %>% 
  ggplot(aes(n_owen,n_leena))+
  geom_point()+
  geom_smooth(method='lm')+
  labs(x='points by tix_owen',y="points by tix_leena")
```

