---
title: "BA Overlap using Individual Vessels"
author: "M. Fisher"
date: "3/15/2023"
output: html_document
---


The analysis to look at fishing ground overlap between groups was completed by first generating a 90% utilization distribution using fishing location data for *all vessels in the group*, and then calculating BA overlap. With this method, I would then have to bootstrap the confidence intervals.

I'm unsure why I didn't just calculate pairwise BA overlap between all vessels in group A against all vessels in group B, and then look at the median. This script will trial that method using the vessel-year 90% utilization distributions generated for script 02 (EMD). 



```{r setup, include=FALSE}
rm(list=ls())

library(raster)
library(adehabitatHR)
library(tidyverse)
library(gridExtra)
library(readr)
library(here)

knitr::opts_chunk$set(echo = TRUE)
```
<br>

Input
```{r}
yrs <- c(2013,2014,2015,2016,2017,2018)
k=5
p=90
gsize=75
cluster_year=2014
ud_years <- "2011-18"
tfs <- c("annual","Central","North")
overlap_method <- "BA"
href=0.08

indir     <- 'project-dat/vms/interpolation_60min/NaN_speed_filter/vessel90ud'
clustfile <- 'project-dat/vessel_groups/k4_NaNspeedfilter_2014clusterYR_byArea_noncon-group-key.csv'
outdir    <- 'project-dat/vessel_groups'
qcdir     <- 'project-dat/vessel_groups/kernel-overlap-QC'

```
<br>


# Set up 


Read in the cluster key
```{r}
cluster_key <- read.csv(here::here(clustfile))
if("X" %in% colnames(cluster_key)){cluster_key <- dplyr::select(cluster_key,-X)}
head(cluster_key)

ngroups <- length(unique(cluster_key$subgroup))
```

Function for calculating an *h* value for `kernelUD`
```{r}
source(here('r','pairwise_href.r'))
```



# Calculate overlap

### testing 123...

```{r}
i = 1; j = 2
groupA <- unique(cluster_key$subgroup)[i]
groupAvessels <- filter(cluster_key,subgroup %in% groupA)
groupB <- unique(cluster_key$subgroup)[j]
groupBvessels <- filter(cluster_key,subgroup %in% groupB)

d=groupAvessels$drvid[1]

e=groupBvessels$drvid[1]


fa <- paste0(d,"_SPDF_",ud_years,"crabYrs.rds")
tmpSPDFa <- read_rds(here::here(indir,fa))
tmpSPDFa@data$drvid <- d


fb <- paste0(e,"_SPDF_",ud_years,"crabYrs.rds")
tmpSPDFb <- read_rds(here::here(indir,fb))
tmpSPDFb@data$drvid <- e

y=intersect(unique(tmpSPDFa@data$crab_year),unique(tmpSPDFb@data$crab_year))[1]


tmpA <- (subset(tmpSPDFa, crab_year==y)); tmpA@data$crab_year <- NULL
tmpB <- (subset(tmpSPDFb, crab_year==y)); tmpB@data$crab_year <- NULL
tmpAB <- rbind(tmpA,tmpB)
### make sure coordinate system carries over!
if(is.null(crs(tmpAB))){
  error("coordinate systems are not the same between vessels!!")
}
```


with the default value of href, calculated based on the variation in coordinates. First, have the function calculate it for me
```{r}
tmpBAi <- kerneloverlap(tmpAB, method=overlap_method, percent=p, grid=75, h="href", conditional=F)

tmpud <- kernelUD(tmpAB, h="href",grid=75)
img <- image(tmpud)
ggplot(as.data.frame(tmpAB@coords) %>% mutate(drvid=tmpAB@data$drvid)) +
  geom_point(aes(x=X_COORD,y=Y_COORD,color=drvid)) +
  ggtitle(paste0(d,"-",e,"- Overlap:",tmpBAi[1,2]))

tmpud[[1]]@h
tmpud[[2]]@h
```

Now, do this calculation separately and then take the mean
```{r}
href_calc <- pairwise_href(tmpA,tmpB); href_calc 
tmpBAi <- kerneloverlap(tmpAB, method=overlap_method, percent=p, grid=75, h=href_calc, conditional=F)

tmpud <- kernelUD(tmpAB, h=href_calc,grid=75)
image(tmpud)
ggplot(as.data.frame(tmpAB@coords) %>% mutate(drvid=tmpAB@data$drvid)) +
  geom_point(aes(x=X_COORD,y=Y_COORD,color=drvid)) +
  ggtitle(paste0(d,"-",e,"- Overlap:",tmpBAi[1,2]))
  
```


### run as a loop

I'm going to keep the grid size set to 75 for all comparisons, and use the default href calculation. Because I'm using the default *h* value with a set grid size, all vessel comparisons will output QC plots for the first year of comparison. 
```{r}
ba.df <- data.frame(group1=as.character(),
                    group2=as.character(),
                    drvid1=as.character(),
                    drvid2=as.character(),
                    crab_year=as.character(),
                    BAindex=as.numeric(),
                    h=as.character(),
                    grid=as.numeric(),
                    percentUD=as.character())
```

#### 1 - large
```{r eval=FALSE}
# group #1 for comparison
groupA <- "1-large"
groupAvessels <- filter(cluster_key,subgroup %in% groupA)
# plot_vessels <- unique(groupAvessels$drvid)[sample(seq(1,length(unique(groupAvessels$drvid))),size=5)]
plot_vessels <- unique(groupAvessels$drvid)

for(j in seq(5,ngroups)){
  if(i != j){
    # group #2 for comparison
    groupB <- unique(cluster_key$subgroup)[j]
    groupBvessels <- filter(cluster_key,subgroup %in% groupB)
    
    # for each vessel in group A...
    for(d in unique(groupAvessels$drvid)){
      ## read in the UD list
      fa <- paste0(d,"_SPDF_",ud_years,"crabYrs.rds")
      tmpSPDFa <- read_rds(here::here(indir,fa))
      tmpSPDFa@data$drvid <- d
      # if(d %in% plot_vessels){dir.create(here(qcdir,d))}
      
      for(e in unique(groupBvessels$drvid)){
        if(d != e){
        ## read in the UD list
        fb <- paste0(e,"_SPDF_",ud_years,"crabYrs.rds")
        tmpSPDFb <- read_rds(here::here(indir,fb))
        tmpSPDFb@data$drvid <- e
        
        ## for each year in vessel 'd' also present in vessel 'e'
        for(y in intersect(unique(tmpSPDFa@data$crab_year),unique(tmpSPDFb@data$crab_year))){
          if(y %in% yrs){
            tmpA <- (subset(tmpSPDFa, crab_year==y)); tmpA@data$crab_year <- NULL
            tmpB <- (subset(tmpSPDFb, crab_year==y)); tmpB@data$crab_year <- NULL
            tmpAB <- rbind(tmpA,tmpB)
            ### make sure coordinate system carries over!
            if(is.null(crs(tmpAB))){
              error("coordinate systems are not the same between vessels!!")
            }
            
            href_calc <- href_label <- "href"
            kud_extent <- 1  # default
            kud_grid <- gsize
            #### some vessels need custom href; see below code chunk
            if(d=="266985" & e!="607016")
              {href_calc=pairwise_href(tmpA,tmpB,method="min"); kud_extent=2} else if(d=="266985" & e=="607016")
              {kud_grid=60; kud_extent=2} else if(d=="944210" | e=="944210"){kud_extent=2}
            
            ### calculate BA index
            tmpBAi <- kerneloverlap(tmpAB, method=overlap_method, percent=p, grid=kud_grid, h=href_calc,extent=kud_extent, conditional=F)
            ### save BA index to dataframe
            ba.df <- bind_rows(ba.df, data.frame(group1=groupA,
                                                 group2=groupB,
                                                 drvid1=d,
                                                 drvid2=e,
                                                 crab_year=as.character(y),
                                                 BAindex=tmpBAi[1,2],
                                                 h=as.character(href_calc),
                                                 grid=kud_grid,
                                                 percentUD=as.character(p)))
            
            ### save plots for QC, for subset of vessels
            if(d %in% plot_vessels & y==yrs[1] & e != "278493"){
              #### kernelUD for plotting
              tmpud <- kernelUD(tmpAB, h=href_calc,extent=kud_extent,grid=kud_grid)
              tmpud <- kernelUD(tmpB, h=href_calc,extent=kud_extent,grid=kud_grid)
              # tmpud <- kernelUD(tmpAB,extent=2)
              kd_names <- names(tmpud)
              p.ud <- lapply(tmpud, function(x){try(getverticeshr(x, p))})
              sapply(1:length(p.ud), function(i) {
                row.names(p.ud[[i]]) <<- kd_names[i]
              })
              sdf_poly <- Reduce(rbind, p.ud)
              #### plot 1
              tmp.plot <- ggplot() +
                geom_point(data=as.data.frame(tmpAB@coords) %>% mutate(drvid=tmpAB@data$drvid), aes(x=X_COORD,y=Y_COORD,color=drvid)) +
                geom_polygon(data=fortify(sdf_poly),aes(x = long, y = lat, fill = id),alpha=0.4) +
                ggtitle(paste0("Overlap:",round(tmpBAi[1,2],2))) + theme_bw()
              png(here(qcdir,d,paste0(d,"_",e,"_",y,"_",p,"ud_",href_label,"href_points.png")))
              print(tmp.plot)
              dev.off()
              #### plot 2
              png(here(qcdir,d,paste0(d,"_",e,"_",y,"_",p,"ud_",href_label,"href_image.png")))
              image(tmpud)
              dev.off()
              
            }
          }
          }
          
        } else{message("WARNING: fishing vessel ",d," is in two groups: ",groupA," and ",groupB )}
      }
    }
  }
}
```

Error in getverticeshr.estUD(x, p) : 
  The grid is too small to allow the estimation of home-range.
You should rerun kernelUD with a larger extent parameter

- when d = "266985" and e = "1203760" : using pairwise_href(method="mean") works
- when d = "266985" and e = "607016" : reset everything to defaults and adjust extent to 2.
- when d = "205318" and e = "944210"

Also noted: 

For animal  1203760 , the most of the UD is distributed  in less than 5 pixels.
 The results will probably be inconsistent.
 Try to increase the parameter grid.

For animal  240412 , the most of the UD is distributed  in less than 5 pixels.
 The results will probably be inconsistent.
 
 For animal  205318 , the most of the UD is distributed  in less than 5 pixels.
 The results will probably be inconsistent.
 Try to increase the parameter grid.
For animal  233048 , the most of the UD is distributed  in less than 5 pixels.
 The results will probably be inconsistent.
 
 
 drvid 278493 only had 11 re-locations in 2013; might want to remove this vessel. 

```{r}
write_csv(ba.df, here(outdir,paste0('BAoverlap_',groupA,"_",groupB,".csv")))
```
