---
title: "Blue Whale 5km Grid Matching"
author: "Owen Liu"
output: 
  html_document:
    toc: true
    toc_float: true
---

## Purpose

We want to take blue whale daily predictions from Briana Abrahms and resample them to Blake Feist's 5km grid for comparison with humpback predictions and measurements of fishing effort.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs,message=FALSE,warning=FALSE}
library(raster)
library(sf)
library(fasterize)
library(tidyverse)
library(magrittr)
library(here)
library(lubridate)

rm(list=ls())

plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=10,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
```

## Import Data

```{r}
# reference raster (the 5km grid)
# grd <- read_sf(here::here('data','raw','grid','regions_master_final_lamb.shp'))
# names(grd)
# ext <- st_bbox(grd) %>% st_as_sfc() %>% as_Spatial() %>% raster::extent()
# grd_r <- fasterize(grd,raster::raster(crs=st_crs(grd),resolution=5000,ext=ext),field="GRID5KM_ID")
# grd_r
# 
# writeRaster(grd_r,here::here('data','raw','grid','raster5km'),overwrite=T)


grd_r <- raster(here::here('data','raw','grid','fivekm_grid.tif'))

# example day of data
head(list.files(here::here('data','Blue whale predictions')))
r <- raster(here::here('data','Blue whale predictions', '2015-01-01_ensemble.grd'))

```

## Resample to 5km grid

```{r,warning=F,message=F}
r2 <- r %>% 
  # project to the new coordinate reference system
  projectRaster(grd_r) %>% 
  # resample to the 5km grid, using nearest neighbors
  resample(grd_r,method='ngb')

# see the results

r2
plot(grd_r)
plot(r)
plot(r2)
```

Okay this seems to work. Let's apply to a month's worth of data and see if it works and how long it takes

```{r}
resample_5km_grd <- function(fn,returnRaster=FALSE){
  rast <- raster(here::here('data','Blue whale predictions', fn))
  raster_name <- paste0('day_',str_trunc(fn,10,ellipsis=""))
  names(rast) <- raster_name
  out_rast <- rast %>% 
    # project to the new coordinate reference system
    projectRaster(grd_r) %>% 
    # resample to the 5km grid, using nearest neighbors
    resample(grd_r,method='ngb')
  writeRaster(out_rast,paste0(here::here('data','Blue whale predictions','5km grid rasters'),'/',raster_name),overwrite=TRUE)
  if(returnRaster){return(out_rast)}
}

all_grd_files <-list.files(here::here('data','Blue whale predictions'))
all_grd_files <- all_grd_files[str_which(all_grd_files,'.grd')]
head(all_grd_files)

# 2009 January
p <- proc.time()
jan2009 <- all_grd_files[str_which(all_grd_files,'2009-01')] %>% 
  purrr::map(resample_5km_grd,returnRaster=T)
proc.time()-p

plot(jan2009[[25]])
```

## Apply to All Years and Days

This will take awhile

```{r}
p <- proc.time()
all_grd_files %>% 
  purrr::walk(resample_5km_grd,returnRaster=F)
proc.time()-p
```

Took about 47 minutes

## Calculate Monthly 5km Grid Means

We can put the raster data into tabular form

```{r}
# grd_r <- raster(here::here('data','raw','grid','raster5km'))
nonNAs <- which(!is.na(getValues(grd_r)))
template <- tibble(GRID5KM_ID=getValues(grd_r)[nonNAs])
# assign_raster_to_dataframe <- function(fn){
#   rast <- raster(fn)
#   obs_date <- substr(fn,97,106) %>% as_date()
#   template %>% mutate(date=obs_date,blue_whale_occurrence=getValues(rast)[nonNAs]) %>% 
#     filter(!is.na(blue_whale_occurrence))
# }
fns <- list.files(here::here('data','Blue whale predictions','5km grid rasters'),full.names = T)
fns=fns[str_which(fns,'.grd')]
# # test a random day
# test <- assign_raster_to_dataframe(fns[987])
# glimpse(test)

# Function to average across a month
average_month_blwh <- function(yr,mth) {
  mth_chr <- ifelse(mth<10,paste0('0',mth),paste(mth))
  yr_mo <- paste(yr,mth_chr,sep="-")
  yr_mo_fns <- fns[str_which(fns,yr_mo)]
  rasts <- purrr::map(yr_mo_fns,raster) %>% stack()
  rasts_mean <- calc(rasts,mean)
  rasts_sd <- calc(rasts,sd)
  out <- template %>% mutate(year_mo=str_replace(yr_mo,"-","_"),Blue_occurrence_mean=getValues(rasts_mean)[nonNAs],Blue_occurrence_sd=getValues(rasts_sd)[nonNAs])%>% 
    filter(!is.na(Blue_occurrence_mean))
  paste(yr_mo,'done')
  return(out)
}

```

Correlation between old and new versions is not that high...

```{r}
# check_yr_mths <- function(yr,mth) {
#   mth_chr <- ifelse(mth<10,paste0('0',mth),paste(mth))
#   yr_mo <- paste(yr,mth_chr,sep="-")
#   yr_mo_fns <- fns[str_which(fns,yr_mo)]
#   paste(yr_mo,'has',length(yr_mo_fns),'rasters')
# }
# 
yrs_mths <- crossing(yr=2009:2019,mth=1:12) %>% filter(!(yr==2019&mth>8))
# chex <- purrr::map2(yrs_mths$yr,yrs_mths$mth,check_yr_mths)

# run
blwh_by_grd_mth <-purrr::map2(yrs_mths$yr,yrs_mths$mth,average_month_blwh)
blwh_by_grd_mth %<>% bind_rows()
blwh_by_grd_mth %<>% mutate(yr=substr(year_mo,1,4) %>% as.numeric(),mth=substr(year_mo,6,7) %>% as.numeric())

write_rds(blwh_by_grd_mth,here::here('data','Blue whale predictions','blwh_by_grd_mth.rds'))
glimpse(blwh_by_grd_mth)

# Compare to old version of blue whales
blwh_old <- read_rds(here::here('test scripts','BLWH_5km_year_mo_2009_2018.RDS')) %>% ungroup() %>% mutate(GRID5KM_ID=as.numeric(as.character(GRID5KM_ID)))

test_join <- blwh_old %>% left_join(blwh_by_grd_mth,by=c("GRID5KM_ID","year_mo"))
glimpse(test_join)

ggplot(test_join,aes(Blue_occurrence,Blue_occurrence_mean))+
  geom_point()+
  geom_smooth(method="lm")+
  labs(x="Blue Whale Occurrence (Old)",y="Blue Whale Occurrence (New)")

cor(test_join$Blue_occurrence,test_join$Blue_occurrence_mean,use='complete.obs')
```

Write to a .rds file.

```{r}
rm(list=ls())
vms <- read_rds(here::here('data','processed','matched','interpolation','vms_all_interpolated_w_grd.rds'))
blwh_by_grd_mth <- read_rds(here::here('data','Blue whale predictions','blwh_by_grd_mth.rds'))
# reduce whale data to just places with positive fishing effort
vms_grd_obs <- vms %>% dplyr::select(GRID5KM_ID,westcoastdate_notime) %>% 
  mutate(yr=year(westcoastdate_notime),mth=month(westcoastdate_notime)) %>% 
  distinct(yr,mth,GRID5KM_ID)
all_vms_cells <- vms_grd_obs$GRID5KM_ID %>% unique()
all_blwh_cells <- unique(blwh_by_grd_mth$GRID5KM_ID)
blwh_vms_cells_only <- blwh_by_grd_mth %>% filter(GRID5KM_ID %in% all_vms_cells)

write_rds(blwh_vms_cells_only,here::here('data','Blue whale predictions','blwh_vms_cells_only.rds'))
```


## Extend to Inshore Areas

One issue with the blue whale data when using it to compare to fishing activity is that the model domain does not cut very closely to the coastline. When we downscale from a 10km to a 5km grid, this leaves some inshore grid cells that are NA. To fix the issue, we assign whale occurrence values for those inshore cells based on their nearest neighbors in the existing whale data that we just converted.


```{r}
# Find the cells that have VMS data, but no values for blue whales.
vms_na_cells <- tibble(GRID5KM_ID=setdiff(all_vms_cells,all_blwh_cells)) %>% filter(!is.na(GRID5KM_ID))

#Import shapefile for the grid
grd_sf <- read_sf(here::here('data','raw','grid','regions_master_final_lamb.shp'))

# Make the VMS cells with NA whale values into a spatial object
vms_na_cells %<>% left_join(grd_sf) %>% st_as_sf(crs=st_crs(grd_sf))

# Plot to see where the NA cells are
library(rnaturalearth)
coaststates <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('California','Oregon','Washington','Nevada')) %>% 
  st_transform(st_crs(grd_sf))
bbox <- st_bbox(vms_na_cells)
ggplot()+
  geom_sf(data=coaststates,col=NA)+
  geom_sf(data=vms_na_cells,fill='red',col=NA)+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))

# match cells with NAs for blwhs with their nearest neighbor that DOES have blwh values
# Do this by matching centroids of the NA cells above to centroids of the blwh data

# find centroids of the NA cells
vms_na_cells_centroids <- vms_na_cells %>% st_centroid() %>% st_coordinates()
vms_na_cells_centroids <- vms_na_cells %>% st_set_geometry(NULL) %>% mutate(x=vms_na_cells_centroids[,1],y=vms_na_cells_centroids[,2])

# find centroids of the blwh cells
blwh_cells <- tibble(GRID5KM_ID=all_blwh_cells) %>% left_join(grd_sf) %>% filter(!st_is_empty(geometry)) %>% 
  st_as_sf(crs=st_crs(grd_sf))
blwh_centroids <- blwh_cells  %>% st_centroid() %>% st_coordinates()

# match the NA cells with their nearest neighbor from the blwh set
library(RANN)
matches_na_to_blwh <- nn2(data=blwh_centroids,query=select(vms_na_cells_centroids,x,y),k=1)[[1]]

# make the matching key
match_NA_vms_to_blwh <- vms_na_cells %>% mutate(blwh_nn_cell=blwh_cells$GRID5KM_ID[matches_na_to_blwh]) %>% 
  select(GRID5KM_ID,blwh_nn_cell) %>% st_set_geometry(NULL)
write_csv(match_NA_vms_to_blwh,here::here('data','Blue whale predictions','vms_NA_blwh_match_key.csv'))

# now, compile the new blwh data for the previously-NA cells
blwh_by_grd_mth_expanded <- match_NA_vms_to_blwh %>% 
  rename(vms_cell=GRID5KM_ID) %>% 
  # Join all the blue whale data by the MATCHED cell
  left_join(blwh_by_grd_mth,by=c('blwh_nn_cell'='GRID5KM_ID')) %>%
  # Then, reattach to all the rest of the blue whale data from before
  select(-blwh_nn_cell) %>% 
  rename(GRID5KM_ID=vms_cell) %>% 
  bind_rows(blwh_by_grd_mth) %>% 
  arrange(yr,mth)

# Save expanded blwh occurrence data
write_rds(blwh_by_grd_mth_expanded,here::here('data','Blue whale predictions','blwh_by_grd_mth_expanded.rds'))

# check to see if any vms cells remain na
all_blwh_cells_expanded<- unique(blwh_by_grd_mth_expanded$GRID5KM_ID)
vms_na_cells_new <- tibble(GRID5KM_ID=setdiff(all_vms_cells,all_blwh_cells_expanded)) %>% filter(!is.na(GRID5KM_ID))

# pick a random month and map the output to see the difference
sample_yr <- 2011
sample_mth <- 10
sample_blwh_old <- blwh_by_grd_mth %>% 
  filter(yr==sample_yr,mth==sample_mth)%>% 
  left_join(grd_sf) %>% 
  filter(!st_is_empty(geometry)) %>% 
  st_as_sf(crs=st_crs(grd_sf))
sample_blwh_mth_expanded <- blwh_by_grd_mth_expanded %>% 
  filter(yr==sample_yr,mth==sample_mth) %>% 
  left_join(grd_sf) %>% 
  filter(!st_is_empty(geometry)) %>% 
  st_as_sf(crs=st_crs(grd_sf))

# define a California bounding box
bbox <- vms_na_cells_centroids %>% 
  filter(STATE=='CA') %>% 
  st_as_sf(coords=c('x','y')) %>% 
  st_bbox()

library(viridis)
sample_map_old <- ggplot()+
  geom_sf(data=coaststates,fill='gray70',col=NA)+
  geom_sf(data=sample_blwh_old,aes(fill=Blue_occurrence_mean),col=NA)+
  scale_fill_viridis()+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  ggtitle('Old Version')
sample_map_new <- ggplot()+
  geom_sf(data=coaststates,fill='gray70',col=NA)+
  geom_sf(data=sample_blwh_mth_expanded,aes(fill=Blue_occurrence_mean),col=NA)+
  scale_fill_viridis()+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  ggtitle('New Version')
sample_map_old
sample_map_new
ggsave(here::here('test scripts','blwh_5km_match_old.png'),plot = sample_map_old,w=6,h=4)
ggsave(here::here('test scripts','blwh_5km_match_new.png'),plot = sample_map_new,w=6,h=4)
```


