---
title: "Create Port Group/Crab Season Template"
author: "Owen Liu"
date: "1/9/2020"
output: html_document
---

## Purpose

To add crab/fishing season to the fish ticket data, we need to do some manual defining of fishing start dates. In this script, we make an 'empty' template for defining these crab seasons.

We find, for all of the fish tickets in the data, all of the year/port-group combinations that have recorded commercial landings of Dungeness crab.

For comparison, we then denote the first trip after November 1 of each year that is tagged as 'commercial' crab landings, to compare later to the official start dates.

### Import Data
```{r}
library(tidyverse)
library(magrittr)
library(here)
library(lubridate)

# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=12,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)

```

```{r}
fishtix <- purrr::map_df(2009:2019, function(yr){
  read_rds(paste0(here::here('data','processed','fish tickets'),"/",yr,"fishtix.rds"))
})

# only Dungeness target
fishtix %<>% filter(TARGET_rev=="DCRB")
```

### Find Year and Port-Group Combinations

Find the unique year and port group combintions for all commercial Dungeness crab trips.

First, how many tickets of different removal types do we have?
```{r}
removal_code_counts <- fishtix %>% distinct() %>% count(removal_type_name,removal_type_code)
removal_code_counts %>% arrange(desc(n))
```

The large majority, as we would expect, are commercial removals, followed by personal use.

Now, we filter for commercial sales and find the unique combinations of port-groups and years.

```{r}
year_ports <- fishtix %>% 
  filter(removal_type_code %in% c("C","D")) %>%
  # add general indicator of crab season
  mutate(crab_season=case_when(
    month(date)>10 ~ paste0(year,"-",year+1),
    month(date)<=10 ~ paste0(year-1,"-",year))) %>% 
  distinct(agency_code,port_group_code,crab_season) %>% 
  arrange(crab_season,agency_code,port_group_code)
year_ports
```

### Find First Commercial Landing for Each Port

For each of the years and port groups in the above table, find the first commercial landing after November 1.

```{r}
first_commercial_landings <- fishtix %>% 
  filter(removal_type_code %in% c("C","D")) %>%
  # add general indicator of crab season
  mutate(crab_season=case_when(
    month(date)>10 ~ paste0(year,"-",year+1),
    month(date)<=10 ~ paste0(year-1,"-",year))) %>% 
  group_by(agency_code,port_group_code,crab_season) %>% 
  summarise(first_ticket=first(date,order_by = date)) %>% 
  ungroup()
```

What about the first *large* commercial landings? (say, in the top 50th percentile of year long catches, to try to avoid mis-labeled experimental catches)

```{r}
first_large_commercial_landings <- fishtix %>% 
  filter(removal_type_code %in% c("C","D")) %>%
  # add general indicator of crab season
  mutate(crab_season=case_when(
    month(date)>10 ~ paste0(year,"-",year+1),
    month(date)<=10 ~ paste0(year-1,"-",year))) %>% 
  group_by(agency_code,port_group_code,crab_season) %>% 
  # add 'large' designation
  mutate(large_cutoff=quantile(DCRB_lbs,0.5)) %>% 
  filter(DCRB_lbs>=large_cutoff) %>% 
  summarise(first_ticket_large=first(date,order_by = date)) %>% 
  ungroup()
```

Test with an example from each state (arbitrarily chosen): 
Crescent City, 2008-2009 season
Coos Bay, OR, 2012-2013 season
Coastal Washington ports, 2011-12 season

```{r}
cca <- fishtix %>% 
  filter(removal_type_code %in% c("C","D")) %>%
  # add general indicator of crab season
  mutate(crab_season=case_when(
    month(date)>10 ~ paste0(year,"-",year+1),
    month(date)<=10 ~ paste0(year-1,"-",year))) %>% 
  filter(crab_season =="2008-2009",port_group_code=='CCA')

# raw
cca %>% 
  ggplot(aes(date,DCRB_lbs))+
  geom_point()+
  labs(title='Crescent City all tickets')

# cumulative
cca %>%
  select(date,DCRB_lbs) %>% 
  arrange(date) %>% 
  mutate(cum_landings=cumsum(DCRB_lbs)) %>% 
  mutate(prop_tot=cum_landings/last(cum_landings)) %>% 
  ggplot(aes(date,prop_tot))+
  geom_line()+geom_point()+
  geom_hline(yintercept=0.01,linetype=2,col='red')+
  labs(title='Crescent City cumulative landings')

# for CBA
cba <- fishtix %>% 
  filter(removal_type_code %in% c("C","D")) %>%
  # add general indicator of crab season
  mutate(crab_season=case_when(
    month(date)>10 ~ paste0(year,"-",year+1),
    month(date)<=10 ~ paste0(year-1,"-",year))) %>% 
  filter(crab_season =="2012-2013",port_group_code=='CBA')
# raw
cba %>% 
  ggplot(aes(date,DCRB_lbs))+
  geom_point()+
  labs(title='Coos Bay all tickets')

# cumulative
cba %>%
  select(date,DCRB_lbs) %>% 
  arrange(date) %>% 
  mutate(cum_landings=cumsum(DCRB_lbs)) %>% 
  mutate(prop_tot=cum_landings/last(cum_landings)) %>% 
  ggplot(aes(date,prop_tot))+
  geom_line()+geom_point()+
  geom_hline(yintercept=0.01,linetype=2,col='red')+
  labs(title='Coos Bay cumulative landings')

# for CWA
cwa <- fishtix %>% 
  filter(removal_type_code %in% c("C","D")) %>%
  # add general indicator of crab season
  mutate(crab_season=case_when(
    month(date)>10 ~ paste0(year,"-",year+1),
    month(date)<=10 ~ paste0(year-1,"-",year))) %>% 
  filter(crab_season =="2011-2012",port_group_code=='CWA')
# raw
cwa %>% 
  ggplot(aes(date,DCRB_lbs))+
  geom_point()+
  labs(title='Coastal WA all tickets')

# cumulative
cwa %>%
  select(date,DCRB_lbs) %>% 
  arrange(date) %>% 
  mutate(cum_landings=cumsum(DCRB_lbs)) %>% 
  mutate(prop_tot=cum_landings/last(cum_landings)) %>% 
  ggplot(aes(date,prop_tot))+
  geom_line()+geom_point()+
  geom_hline(yintercept=0.01,linetype=2,col='red')+
  labs(title='Coastal WA cumulative landings')
```

These cumulative distribution suggest that a nice way might be to indicate the date of 1% landings achievement for each port group, since each season, no matter when it starts, begins with an intense derby that should be clear in the landings data. Using a 1 percent cutoff seems robust to the landings that 'dribble in' before the official season starts.

```{r}
one_percent_landings_achievement <- fishtix %>% 
  filter(removal_type_code %in% c("C","D")) %>%
  # add general indicator of crab season
  mutate(crab_season=case_when(
    month(date)>10 ~ paste0(year,"-",year+1),
    month(date)<=10 ~ paste0(year-1,"-",year))) %>% 
  group_by(agency_code,port_group_code,crab_season) %>% 
  arrange(date) %>% 
  mutate(cum_landings=cumsum(DCRB_lbs)) %>% 
  mutate(prop_tot=cum_landings/last(cum_landings)) %>% 
  filter(prop_tot>=0.01) %>% 
  summarise(first_ticket_oneperc=first(date,order_by = date)) %>% 
  ungroup()
```

Compare the three measures

```{r,fig.width=8,fig.height=8}
compare_first_tickets <- first_commercial_landings %>% 
  left_join(first_large_commercial_landings) %>% 
  left_join(one_percent_landings_achievement) %>% 
  pivot_longer(contains('first_ticket'),names_to='type',values_to='date') %>% 
  mutate(start_yr=substr(crab_season,1,4) %>% as.numeric()) %>% 
  mutate(start_date=as_date(paste0(start_yr,"-11-01"))) %>% 
  mutate(days_after_nov1=difftime(date,start_date,units='days'))

CA_dates <- compare_first_tickets %>% 
  filter(agency_code=='C')%>% 
  ggplot(aes(crab_season,days_after_nov1,col=type,shape=type))+
  scale_color_discrete(name="Calculation Type",labels=c('First ticket',"First large ticket","One percent\nlandings achieved"))+
  scale_shape_discrete(name="Calculation Type",labels=c('First ticket',"First large ticket","One percent\nlandings achieved"))+
  labs(y='Days After Nov. 1',x="Crab Season",title="California Season Start Date Comparison")+
  geom_point(alpha=0.5,size=3)+
  facet_wrap(~port_group_code)+
  coord_flip()

OR_dates <- compare_first_tickets %>% 
  filter(agency_code=='O')%>% 
  ggplot(aes(crab_season,days_after_nov1,col=type,shape=type))+
  scale_color_discrete(name="Calculation Type",labels=c('First ticket',"First large ticket","One percent\nlandings achieved"))+
  scale_shape_discrete(name="Calculation Type",labels=c('First ticket',"First large ticket","One percent\nlandings achieved"))+
  labs(y='Days After Nov. 1',x="Crab Season",title="Oregon Season Start Date Comparison")+
  geom_point(alpha=0.5,size=3)+
  facet_wrap(~port_group_code)+
  coord_flip()
WA_dates <- compare_first_tickets %>% 
  filter(agency_code=='W')%>% 
  ggplot(aes(crab_season,days_after_nov1,col=type,shape=type))+
  scale_color_discrete(name="Calculation Type",labels=c('First ticket',"First large ticket","One percent\nlandings achieved"))+
  scale_shape_discrete(name="Calculation Type",labels=c('First ticket',"First large ticket","One percent\nlandings achieved"))+
  labs(y='Days After Nov. 1',x="Crab Season",title="Washington Season Start Date Comparison")+
  geom_point(alpha=0.5,size=3)+
  facet_wrap(~port_group_code)+
  coord_flip()

CA_dates
OR_dates
WA_dates

ggsave(here::here('fishing behavior','plots','CA_startdate_comparison.png'),CA_dates,h=8,w=8)

ggsave(here::here('fishing behavior','plots','OR_startdate_comparison.png'),OR_dates,h=8,w=8)

ggsave(here::here('fishing behavior','plots','WA_startdate_comparison.png'),WA_dates,h=8,w=8)
```

Compare California with official start dates

```{r,fig.width=8,fig.height=8}
# data from M. Fisher on official start dates from press releases and news
ca_starts <- read_csv(here::here('test scripts','ca_port_starts.csv')) %>% 
  mutate(official_start=as_date(official_start,format="%m/%d/%Y",tz="America/Los_Angeles")) %>% 
  mutate(type='official') %>%
  rename(date=official_start) %>% 
  mutate(start_yr=substr(crab_season,1,4) %>% as.numeric()) %>% 
  mutate(start_date=as_date(paste0(start_yr,"-11-01"))) %>% 
  mutate(days_after_nov1=difftime(date,start_date,units='days'))

CA_dates2 <- compare_first_tickets %>% 
  filter(agency_code=='C')%>% 
  full_join(ca_starts) %>% 
  ggplot(aes(crab_season,days_after_nov1,col=type,fill=type,shape=type))+
  scale_color_discrete(name="Calculation Type",labels=c('First ticket',"First large ticket","One percent\nlandings achieved","Official"))+
  scale_fill_discrete(name="Calculation Type",labels=c('First ticket',"First large ticket","One percent\nlandings achieved","Official"))+
  scale_shape_manual(values = c(21:24),name="Calculation Type",labels=c('First ticket',"First large ticket","One percent\nlandings achieved","Official"))+
  labs(y='Days After Nov. 1',x="Crab Season",title="California Season Start Date Comparison")+
  geom_point(alpha=0.6,size=3)+
  facet_wrap(~port_group_code)+
  coord_flip()
CA_dates2
ggsave(here::here('fishing behavior','plots','CA_startdate_comparison2.png'),CA_dates2,h=8,w=8)
```



### Add Empty Columns to Fill In

Finally, add `season_start` and `season_end` columns that we will fill in ourselves manually for later joining to the overall data to define crab seasons.

```{r}
crab_season_template <- first_commercial_landings %>% 
  mutate(season_start=NA_character_,season_end=NA_character_)
```

### Make 1% Landings Key

We use the 1% landings achievement key to indicate the start of each crab season for each individual port group in each year.

```{r}
write_csv(one_percent_landings_achievement,here::here('data','processed','one_perc_season_start_key.csv'))
```


### Save

```{r}
write_csv(crab_season_template,here::here('test scripts','crab_season_template.csv'))
```

### Visualize

Let's look at the first commercial landing date, by port group and year
```{r,fig.height=8,fig.width=10}

# calculate days after Nov 1
test <- crab_season_template %>% 
  mutate(start_yr=substr(crab_season,1,4) %>% as.numeric()) %>% 
  mutate(start_date=as_date(paste0(start_yr,"-11-01"))) %>% 
  mutate(days_after_nov1=difftime(first_ticket,start_date,units='days'))
test %>% 
  ggplot(aes(crab_season,days_after_nov1,color=agency_code))+
  geom_point(size=2)+
  labs(x="Season",y="Days after Nov. 1",title="Dates of First Commercial Landing\nin Days After November 1st",color="State")+
  facet_wrap(~port_group_code)+
  theme(axis.text.x = element_text(angle = 90,vjust=+0.5))
```

