---
title: "Fishing Season Timeline Example"
author: "Owen Liu"
date: "10/7/2020"
output: html_document
---


## Purpose

Create a mock-up timeline of an average crab fishing season, showing the potential impacts of temporal season delays on fishery effort

```{r setup, include=FALSE}
library(tidyverse)
library(magrittr)
library(here)
library(lubridate)

rm(list=ls())

plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=10,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
```

Import data

```{r}
tix <- read_rds(here::here('data','processed','fish tickets','fish_tickets_w_dayofseason.rds'))

glimpse(tix)
```

Calculate average landings by day of season, for use in constructing a generalized time series.

```{r}
landings_dos <- tix %>% 
  filter(agency_code=="C",ticket_day_of_season>=0) %>% 
  group_by(crab_season,ticket_day_of_season) %>% 
  summarise(tot_landings=sum(DCRB_lbs)) %>% 
  group_by(ticket_day_of_season) %>% 
  summarise(mean_landings=mean(tot_landings,na.rm=T)) %>% 
  ungroup()
```

Plot the "average" year

```{r}
landings_dos %>% 
  ggplot(aes(ticket_day_of_season,mean_landings))+
  geom_line()+
  labs(x="Day of Season",y="Landings (lbs)")
```

Add in hypothetical scenarios. Early closures will be symbolized with a vertical line

```{r}
temporal_shifts <- landings_dos %>% 
  rename(baseline=mean_landings) %>% 
  # a late opening with lag method = 30 day delay
  mutate(late_opening_lag=lag(baseline,30)) %>% 
  # a late opening with pile method = add the effort for the first 30 days, no change thereafter
  rowwise() %>% 
  mutate(late_opening_pile=ifelse(ticket_day_of_season<=60,baseline+late_opening_lag,baseline)) %>% 
  # a depth restriction = some percentage difference from baseline?
  mutate(depth_restriction=baseline*0.5) %>% 
  # stylized date
  mutate(date=as_date("2015/11/15")+days(ticket_day_of_season))

```

Convert to long form and plot

```{r}
temporal_shifts_long <- temporal_shifts %>% 
  select(-ticket_day_of_season) %>% 
  pivot_longer(-date,names_to="type",values_to="value") %>% 
  mutate(type_label=case_when(
    type=="baseline" ~ "Status Quo",
    type=="depth_restriction" ~ "Depth Restriction",
    type=="late_opening_lag" ~ "Late Opening (Lag)",
    type=="late_opening_pile" ~ "Late Opening (Pile)"
  ))

scenarios_plot <- temporal_shifts_long %>% 
  ggplot(aes(date,value,col=type_label))+
  geom_line(size=1)+
  # vertical line for early closure
  geom_vline(xintercept=as_date("2016/07/15"))+
  annotate("text",x=as_date("2016/07/15"),y=5e5,label="Early Closure",angle=90,vjust=1.5)+
  labs(x="Date",y="Fishing Effort, Landings, or Revenue",col="")+
  scale_x_date(date_labels = "%b %d",date_breaks="2 months")+
  theme(axis.text.y=element_blank(),
        legend.position = c(0.85,0.8),
        legend.text = element_text(size=8))

scenarios_plot
ggsave(here::here('test scripts','tradeoff_effort_shift_scenarios_plot.png'),scenarios_plot,w=6,h=4)
```

