---
title: "Maps For Jameal"
author: "Owen Liu"
date: "1/14/2020"
output: html_document
---

## Maps for Jameal, 01/14/2020
```{r setup}
knitr::opts_chunk$set(echo=FALSE,message=F,warning=F)
```

```{r}
library(tidyverse)
library(magrittr)
library(sf)
library(here)
library(rnaturalearth)
library(viridis)
```

```{r}
load(here::here('test scripts','Scenario_Analysis_Data_2009_2018.RData'))
grd <- read_sf(here::here('data','raw','grid','regions_master_final_lamb.shp'))
# background map
coaststates <- ne_states(country='United States of America',returnclass = 'sf') %>% 
  filter(name %in% c('California','Oregon','Washington','Nevada')) %>% 
  st_transform(st_crs(grd))

# BIAs
library(smoothr)
BIA <- read_csv(here::here('data','raw','grid','Grid5km_BIA_overlap.csv'),col_types = cols())
BIA %<>% select(GRID5KM_ID,BIA_mn,BIA_bm)

grd %<>% left_join(BIA)

BIA_mn <- grd %>% 
  filter(BIA_mn==1) %>% 
  group_by(BIA_mn) %>% 
  summarise() %>% 
  ungroup()
BIA_bm <- grd %>% 
  filter(BIA_bm==1) %>% 
  group_by(BIA_bm) %>% 
  summarise()
BIA_both <- grd %>% 
  mutate(BIA=BIA_mn==1|BIA_bm==1) %>% 
  filter(BIA) %>% 
  mutate(BIA_type=case_when(
    BIA_mn==1 & BIA_bm!=1 ~ 'Humpback',
    BIA_bm==1 & BIA_mn!=1 ~ 'Blue',
    BIA_mn==1&BIA_bm==1 ~ 'Both'
  )) %>% 
  group_by(BIA_type) %>% 
  summarise()

write_rds(BIA_both,here::here('data','processed','BIA_mn_bm.rds'))
  
# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=10,color="black"),
        legend.text = element_text(size=10),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3))
theme_set(plot_theme)

ggplot()+geom_sf(data=BIA_both,aes(col=BIA_type),fill=NA)
```


## humpback whales during winter DCRB season (Nov-Mar)

```{r}
names(full.df)
# get coordinates of data in correct projection

# Average Nov-Mar humpback abundance
h_nov_mar <- full.df %>% 
  filter(month %in% c(11,12,1,2,3)) %>% 
  group_by(Region, GRID5KM_ID) %>% 
  summarise(abund=mean(H_Avg_Abund,na.rm=T)) %>% 
  ungroup() %>% 
  mutate(z=(abund-mean(abund,na.rm=T))/sd(abund,na.rm=T))

# central CA
cenCA_h_nov_mar <- h_nov_mar %>% 
  filter(Region=="CenCA") %>% 
  left_join(grd,by="GRID5KM_ID") %>% 
  st_as_sf()

bbox <- st_bbox(cenCA_h_nov_mar)
cenCA_h_nov_mar_plot <- ggplot()+
  geom_sf(data=BIA_mn,fill='gray80',col='black')+
  geom_sf(data=cenCA_h_nov_mar,aes(fill=abund),col=NA,na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col=NA,fill='gray60')+
  scale_fill_viridis(na.value=NA,option="D",name="Humpback Abundance",breaks=seq(0,1,by=0.25),limits=c(0,1))+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  labs(title="Average Humpback Abundance\nNov-Mar 2009-2018")+
  theme(axis.text.x.bottom = element_text(angle=45),
          title=element_text(size=16))
cenCA_h_nov_mar_plot

# northern CA
norCA_h_nov_mar <- h_nov_mar %>% 
  filter(Region=="NorCA") %>% 
  left_join(grd,by="GRID5KM_ID") %>% 
  st_as_sf()

bbox <- st_bbox(norCA_h_nov_mar)
norCA_h_nov_mar_plot<-ggplot()+
  geom_sf(data=BIA_mn,fill='gray80',col='black')+
  geom_sf(data=norCA_h_nov_mar,aes(fill=abund),col=NA,na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col=NA,fill='gray60')+
  scale_fill_viridis(na.value=NA,option="D",name="Humpback Abundance",breaks=seq(0,1,by=0.25),limits=c(0,1))+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  labs(title="Average Humpback Abundance\nNov-Mar 2009-2018")+
  theme(axis.text.x.bottom = element_text(angle=45),
          title=element_text(size=16))
```

## humpback whales during spring DCRB season (Apr-Jul)

```{r}
# Average April to July humpback abundance
h_apr_jul <- full.df %>% 
  filter(month %in% c(4,5,6,7)) %>% 
  group_by(Region, GRID5KM_ID, BIA_mn_noNAs) %>% 
  summarise(abund=mean(H_Avg_Abund,na.rm=T)) %>% 
  ungroup()

# central CA
cenCA_h_apr_jul <- h_apr_jul %>% 
  filter(Region=="CenCA") %>% 
  left_join(grd,by="GRID5KM_ID") %>% 
  st_as_sf()

bbox <- st_bbox(cenCA_h_apr_jul)
cenCA_h_apr_jul_plot <- ggplot()+
  geom_sf(data=BIA_mn,fill='gray80',col='black')+
  geom_sf(data=cenCA_h_apr_jul,aes(fill=abund),col=NA,na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col=NA,fill='gray60')+
  scale_fill_viridis(na.value=NA,option="D",name="Humpback Abundance",breaks=seq(0,1,by=0.25),limits=c(0,1))+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  labs(title="Average Humpback Abundance\nApr-Jul 2009-2018")+
  theme(axis.text.x.bottom = element_text(angle=45),
          title=element_text(size=16))
cenCA_h_apr_jul_plot

# northern CA
norCA_h_apr_jul <- h_apr_jul %>% 
  filter(Region=="NorCA") %>% 
  left_join(grd,by="GRID5KM_ID") %>% 
  st_as_sf()

bbox <- st_bbox(norCA_h_apr_jul)
norCA_h_apr_jul_plot <- ggplot()+
  geom_sf(data=BIA_mn,fill='gray80',col='black')+
  geom_sf(data=norCA_h_apr_jul,aes(fill=abund),col=NA,na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col=NA,fill='gray60')+
  scale_fill_viridis(na.value=NA,option="D",name="Humpback Abundance",breaks=seq(0,1,by=0.25),limits=c(0,1))+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  labs(title="Average Humpback Abundance\nApr-Jul 2009-2018")+
  theme(axis.text.x.bottom = element_text(angle=45),
          title=element_text(size=16))
norCA_h_apr_jul_plot
```

## blue whales during winter DCRB season (Nov-Mar)

```{r}
# Average Nov-Mar blue abundance
blue_nov_mar <- full.df %>% 
  filter(month %in% c(11,12,1,2,3)) %>% 
  group_by(Region, GRID5KM_ID, BIA_bm_noNAs) %>% 
  summarise(abund=mean(Blue_occurrence,na.rm=T)) %>% 
  ungroup()

# central CA
cenCA_blue_nov_mar <- blue_nov_mar %>% 
  filter(Region=="CenCA") %>% 
  left_join(grd,by="GRID5KM_ID") %>% 
  st_as_sf()

bbox <- st_bbox(cenCA_blue_nov_mar)
cenCA_blue_nov_mar_plot <- ggplot()+
  geom_sf(data=BIA_bm,fill='lightblue',col='firebrick1',alpha=0.8)+
  geom_sf(data=cenCA_blue_nov_mar,aes(fill=abund),col=NA,na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col=NA,fill='gray60')+
  scale_fill_viridis(na.value=NA,option="D",name="Blue Whale occurrence",breaks=seq(0,1,by=0.25),limits=c(0,1))+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  labs(title="Average Blue Whale occurrence\nNov-Mar 2009-2018")+
  theme(axis.text.x.bottom = element_text(angle=45),
          title=element_text(size=16))
cenCA_blue_nov_mar_plot

# northern CA
norCA_blue_nov_mar <- blue_nov_mar %>% 
  filter(Region=="NorCA") %>% 
  left_join(grd,by="GRID5KM_ID") %>% 
  st_as_sf()

bbox <- st_bbox(norCA_blue_nov_mar)
norCA_blue_nov_mar_plot <- ggplot()+
  geom_sf(data=BIA_bm,fill='lightblue',col='firebrick1',alpha=0.8)+
  geom_sf(data=norCA_blue_nov_mar,aes(fill=abund),col=NA,na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col=NA,fill='gray60')+
  scale_fill_viridis(na.value=NA,option="D",name="Blue Whale occurrence",breaks=seq(0,1,by=0.25),limits=c(0,1))+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  labs(title="Average Blue Whale occurrence\nNov-Mar 2009-2018")+
  theme(axis.text.x.bottom = element_text(angle=45),
          title=element_text(size=16))
norCA_blue_nov_mar_plot 
```

## blue whales during spring DCRB season (Apr-Jul)

```{r}
# Average spring blue abundance
blue_apr_jul <- full.df %>% 
  filter(month %in% c(4:7)) %>% 
  group_by(Region, GRID5KM_ID, BIA_bm_noNAs) %>% 
  summarise(abund=mean(Blue_occurrence,na.rm=T)) %>% 
  ungroup()

# central CA
cenCA_blue_apr_jul <- blue_apr_jul %>% 
  filter(Region=="CenCA") %>% 
  left_join(grd,by="GRID5KM_ID") %>% 
  st_as_sf()

bbox <- st_bbox(cenCA_blue_apr_jul)
cenCA_blue_apr_jul_plot <- ggplot()+
  geom_sf(data=BIA_bm,fill='lightblue',col='firebrick1',alpha=0.8)+
  geom_sf(data=cenCA_blue_apr_jul,aes(fill=abund),col=NA,na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col=NA,fill='gray60')+
  scale_fill_viridis(na.value=NA,option="D",name="Blue Whale occurrence",breaks=seq(0,1,by=0.25),limits=c(0,1))+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  labs(title="Average Blue Whale occurrence\nApr-Jul 2009-2018")+
  theme(axis.text.x.bottom = element_text(angle=45),
          title=element_text(size=16))
cenCA_blue_apr_jul_plot

# northern CA
norCA_blue_apr_jul <- blue_apr_jul %>% 
  filter(Region=="NorCA") %>% 
  left_join(grd,by="GRID5KM_ID") %>% 
  st_as_sf()

bbox <- st_bbox(norCA_blue_apr_jul)
norCA_blue_apr_jul_plot <- ggplot()+
  geom_sf(data=BIA_bm,fill='lightblue',col='firebrick1',alpha=0.8)+
  geom_sf(data=norCA_blue_apr_jul,aes(fill=abund),col=NA,na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col=NA,fill='gray60')+
  scale_fill_viridis(na.value=NA,option="D",name="Blue Whale occurrence",breaks=seq(0,1,by=0.25),limits=c(0,1))+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  labs(title="Average Blue Whale occurrence\nApr-Jul 2009-2018")+
  theme(axis.text.x.bottom = element_text(angle=45),
          title=element_text(size=16))
norCA_blue_apr_jul_plot
```

## DCRB vessel activity (VMS pings) during winter DCRB season (Nov-Mar)

```{r}
# Average Nov-Mar VMS pings abundance
dcrb_nov_mar <- full.df %>% 
  filter(month %in% c(11,12,1,2,3)) %>% 
  group_by(Region, GRID5KM_ID) %>% 
  summarise(pings=mean(Num_DCRB_VMS_pings,na.rm=T)) %>% 
  ungroup()%>% 
  mutate(z=(pings-mean(pings,na.rm=T))/sd(pings,na.rm=T))

# central CA
cenCA_dcrb_nov_mar <- dcrb_nov_mar %>% 
  filter(Region=="CenCA") %>% 
  left_join(grd,by="GRID5KM_ID") %>% 
  st_as_sf()

bbox <- st_bbox(cenCA_dcrb_nov_mar)
cenCA_dcrb_nov_mar_plot <- ggplot()+
  geom_sf(data=BIA_mn,fill='gray80',col='black',alpha=0.8)+
  geom_sf(data=BIA_bm,fill='lightblue',col='firebrick1',alpha=0.8)+
  geom_sf(data=cenCA_dcrb_nov_mar,aes(fill=pings),col=NA,na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col=NA,fill='gray60')+
  scale_fill_viridis(na.value=NA,option="D",name="DCRB pings",breaks=seq(0,100,by=20),limits=c(0,100))+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  labs(title="Average Number of VMS pings\nNov-Mar 2009-2018")+
  theme(axis.text.x.bottom = element_text(angle=45),
          title=element_text(size=16))
cenCA_dcrb_nov_mar_plot

# northern CA
norCA_dcrb_nov_mar <- dcrb_nov_mar %>% 
  filter(Region=="NorCA") %>% 
  left_join(grd,by="GRID5KM_ID") %>% 
  st_as_sf()

bbox <- st_bbox(norCA_dcrb_nov_mar)
norCA_dcrb_nov_mar_plot<- ggplot()+
  geom_sf(data=BIA_mn,fill='gray80',col='black',alpha=0.8)+
  geom_sf(data=BIA_bm,fill='lightblue',col='firebrick1',alpha=0.8)+
  geom_sf(data=norCA_dcrb_nov_mar,aes(fill=pings),col=NA,na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col=NA,fill='gray60')+
  scale_fill_viridis(na.value=NA,option="D",name="DCRB pings",breaks=seq(0,100,by=20),limits=c(0,100))+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  labs(title="Average Number of VMS pings\nNov-Mar 2009-2018")+
  theme(axis.text.x.bottom = element_text(angle=45),
          title=element_text(size=16))
norCA_dcrb_nov_mar_plot
```

## DCRB vessel activity (VMS pings) during spring DCRB season (Apr-Jul)

```{r}
# Average Nov-Mar VMS pings abundance
dcrb_apr_jul <- full.df %>% 
  filter(month %in% c(4:7)) %>% 
  group_by(Region, GRID5KM_ID) %>% 
  summarise(pings=mean(Num_DCRB_VMS_pings,na.rm=T)) %>% 
  ungroup()

# central CA
cenCA_dcrb_apr_jul <- dcrb_apr_jul %>% 
  filter(Region=="CenCA") %>% 
  left_join(grd,by="GRID5KM_ID") %>% 
  st_as_sf()

bbox <- st_bbox(cenCA_dcrb_apr_jul)
cenCA_dcrb_apr_jul_plot <- ggplot()+
  geom_sf(data=BIA_mn,fill='gray80',col='black',alpha=0.8)+
  geom_sf(data=BIA_bm,fill='lightblue',col='firebrick1',alpha=0.8)+
  geom_sf(data=cenCA_dcrb_apr_jul,aes(fill=pings),col=NA,na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col=NA,fill='gray60')+
  scale_fill_viridis(na.value=NA,option="D",name="DCRB pings",breaks=seq(0,100,by=20),limits=c(0,100))+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  labs(title="Average Number of VMS pings\nApr-Jul 2009-2018")+
  theme(axis.text.x.bottom = element_text(angle=45),
          title=element_text(size=16))
cenCA_dcrb_apr_jul_plot

# northern CA
norCA_dcrb_apr_jul <- dcrb_apr_jul %>% 
  filter(Region=="NorCA") %>% 
  left_join(grd,by="GRID5KM_ID") %>% 
  st_as_sf()

bbox <- st_bbox(norCA_dcrb_apr_jul)
norCA_dcrb_apr_jul_plot <- ggplot()+
  geom_sf(data=BIA_mn,fill='gray80',col='black',alpha=0.8)+
  geom_sf(data=BIA_bm,fill='lightblue',col='firebrick1',alpha=0.8)+
  geom_sf(data=norCA_dcrb_apr_jul,aes(fill=pings),col=NA,na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col=NA,fill='gray60')+
  scale_fill_viridis(na.value=NA,option="D",name="DCRB pings",breaks=seq(0,100,by=20),limits=c(0,100))+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  labs(title="Average Number of VMS pings\nApr-Jul 2009-2018")+
  theme(axis.text.x.bottom = element_text(angle=45),
          title=element_text(size=16))
norCA_dcrb_apr_jul_plot
```

```{r}
ggsave(here::here('test scripts','cenCA_mn_nov_mar.png'),cenCA_h_nov_mar_plot,h=6,w=8)
ggsave(here::here('test scripts','norCA_mn_nov_mar.png'),norCA_h_nov_mar_plot,h=8,w=6)
ggsave(here::here('test scripts','cenCA_mn_apr_jul.png'),cenCA_h_apr_jul_plot,h=6,w=8)
ggsave(here::here('test scripts','norCA_mn_apr_jul.png'),norCA_h_apr_jul_plot,h=8,w=6)

ggsave(here::here('test scripts','cenCA_bm_nov_mar.png'),cenCA_blue_nov_mar_plot,h=6,w=8)
ggsave(here::here('test scripts','norCA_bm_nov_mar.png'),norCA_blue_nov_mar_plot,h=8,w=6)
ggsave(here::here('test scripts','cenCA_bm_apr_jul.png'),cenCA_blue_apr_jul_plot,h=6,w=8)
ggsave(here::here('test scripts','norCA_bm_apr_jul.png'),norCA_blue_apr_jul_plot,h=8,w=6)

ggsave(here::here('test scripts','cenCA_dcrb_nov_mar.png'),cenCA_dcrb_nov_mar_plot,h=6,w=8)
ggsave(here::here('test scripts','norCA_dcrb_nov_mar.png'),norCA_dcrb_nov_mar_plot,h=8,w=6)
ggsave(here::here('test scripts','cenCA_dcrb_apr_jul.png'),cenCA_dcrb_apr_jul_plot,h=6,w=8)
ggsave(here::here('test scripts','norCA_dcrb_apr_jul.png'),norCA_dcrb_apr_jul_plot,h=8,w=6)
```

## Coastwide 

```{r}
h_nov_mar_sf <- h_nov_mar %>% 
  left_join(grd,by="GRID5KM_ID") %>% 
  st_as_sf()

bbox <- st_bbox(h_nov_mar_sf)
h_nov_mar_plot <- ggplot()+
  geom_sf(data=BIA_mn,fill='gray80',col='black')+
  geom_sf(data=h_nov_mar_sf,aes(fill=z),col=NA,na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col=NA,fill='gray60')+
  scale_fill_viridis(na.value=NA,option="D",name="Humpback Abundance",breaks=seq(-2,3,by=0.5),limits=c(-2,3))+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  labs(title="Average Humpback Abundance\nNov-Mar 2009-2018")+
  theme(axis.text.x.bottom = element_text(angle=45),
          title=element_text(size=16))
h_nov_mar_plot
ggsave(here::here('test scripts','CA_mn_nov_mar.png'),h_nov_mar_plot,h=8,w=6)

# for DCRB, with z-score
dcrb_nov_mar_sf <- dcrb_nov_mar %>% 
  left_join(grd,by="GRID5KM_ID") %>% 
  st_as_sf()

bbox <- st_bbox(dcrb_nov_mar_sf)
dcrb_nov_mar_plot <- ggplot()+
  geom_sf(data=BIA_mn,fill='gray80',col='black')+
  geom_sf(data=dcrb_nov_mar_sf,aes(fill=z),col=NA,na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col=NA,fill='gray60')+
  scale_fill_viridis(na.value=NA,option="D",name="DCRB Pings\n(Z-score)")+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  labs(title="Dungeness Crab Pings\nNov-Mar 2009-2018")+
  theme(axis.text.x.bottom = element_text(angle=45),
          title=element_text(size=16))
dcrb_nov_mar_plot
ggsave(here::here('test scripts','CA_dcrb_nov_mar_z.png'),dcrb_nov_mar_plot,h=8,w=6)

# for DCRB, with percentiles
dcrb_nov_mar_sf <- dcrb_nov_mar %>%
  mutate(perc=percent_rank(pings)) %>% 
  left_join(grd,by="GRID5KM_ID") %>% 
  st_as_sf()

bbox <- st_bbox(dcrb_nov_mar_sf)
dcrb_nov_mar_plot <- ggplot()+
  geom_sf(data=BIA_mn,fill='gray80',col='black')+
  geom_sf(data=dcrb_nov_mar_sf,aes(fill=perc*100),col=NA,na.rm=T,alpha=0.8)+
  geom_sf(data=coaststates,col=NA,fill='gray60')+
  scale_fill_viridis(na.value=NA,option="D",name="DCRB Pings\n(Percentile)",breaks=seq(0,100,by=20),limits=c(0,100))+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  labs(title="Dungeness Crab Pings\nNov-Mar 2009-2018")+
  theme(axis.text.x.bottom = element_text(angle=45),
          title=element_text(size=16))
dcrb_nov_mar_plot
ggsave(here::here('test scripts','CA_dcrb_nov_mar_perc.png'),dcrb_nov_mar_plot,h=8,w=6)

```

## Maps with Regional Definitions

```{r,fig.height=8,fig.width=8}
# NorCA, CenCA polygons
norCA <- dcrb_nov_mar_sf %>% 
  filter(Region=="NorCA") %>% 
  ungroup() %>% 
  summarise() %>% 
  st_bbox()
cenCA <- dcrb_nov_mar_sf %>% 
  filter(Region=="CenCA") %>% 
  ungroup() %>% 
  summarise() %>% 
  st_bbox()

CAtop <- 1277869
CAsplit <- 922869
CAbot <- 362869
rgns <- tibble(hlines=c(CAtop,CAsplit,CAbot))
# norCA[1] <- norCA[1]-100000
# norCA %<>% 
#   st_as_sfc()%>% 
#   st_as_sf() %>% 
#   mutate(rgn="NorCA")
# 
# 
# cenCA[1] <- cenCA[1]-200000
# cenCA %<>% 
#   st_as_sfc() %>% 
#   st_as_sf() %>% 
#   mutate(rgn="CenCA")
# rgns <- st_sym_difference(norCA,cenCA)

# for DCRB, with percentiles
dcrb_nov_mar_sf <- dcrb_nov_mar %>%
  mutate(perc=percent_rank(pings)) %>% 
  left_join(grd,by="GRID5KM_ID") %>% 
  st_as_sf()

bbox <- st_bbox(dcrb_nov_mar_sf)
ltys <- c('Region'=2)
cols <- c('Humpback BIA'='red')
dcrb_nov_mar_plot <- ggplot()+
  geom_hline(data=rgns,aes(yintercept=hlines,linetype='Region'))+
  geom_sf(data=dcrb_nov_mar_sf,aes(fill=perc*100),col=NA,na.rm=T,alpha=0.8)+
  geom_sf(data=BIA_mn,fill=NA,aes(col='Humpback BIA'),size=1)+
  geom_sf(data=coaststates,col=NA,fill='gray60')+
  annotate("text",x=833000,y=600000,label="CenCA")+
  annotate("text",x=833000,y=1000000,label="NorCA")+
  scale_color_manual(values=cols,guide=guide_legend(title=NULL))+
  scale_linetype_manual(values=ltys,guide=guide_legend(title=NULL))+
  scale_fill_viridis(na.value=NA,option="D",name="DCRB Pings\n(Percentile)",breaks=seq(0,100,by=20),limits=c(0,100))+
  coord_sf(xlim=c(bbox[1],bbox[3]),ylim=c(bbox[2],bbox[4]))+
  labs(title="Dungeness Crab Pings\nNov-Mar 2009-2018",y="",x="")+
  theme(axis.text.x.bottom = element_text(angle=45),
          title=element_text(size=16),
        panel.grid.major=element_blank(),
        panel.grid.minor=element_blank(),
        legend.position = c(0.8,0.6),
        legend.background = element_rect(fill='white'),
        legend.spacing.y = unit(5,'mm'))
dcrb_nov_mar_plot
ggsave(here::here('test scripts','CA_dcrb_nov_mar_perc.png'),dcrb_nov_mar_plot,h=8,w=6)

```

