---
title: "Add Day of Season"
output: html_document
---

## Purpose

Add a day-of-season indicator for all fish tickets. These will be at the scale of year and port-group (instead of state, since different parts of each state may open at different times).

Currently (2/4/2020), we use a season start indicator derived from the fish ticket data itself, not from the official start dates. There are pros and cons of this approach.

```{r,message=F,warning=F}
library(here)
library(tidyverse)
library(magrittr)
library(lubridate)

# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=12,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=8,color="black"),
        panel.grid.major = element_line(color="gray50",linetype=3),
        panel.grid.minor = element_line(color="gray50",linetype=3))
theme_set(plot_theme)
```

## Import Fish Tickets

Import all fish ticket data and sort by tickets whose target (measured through revenue) is Dungeness Crab.

```{r}
fishtix <- purrr::map_df(2009:2019, function(yr){
  read_rds(paste0(here::here('data','processed','fish tickets'),"/",yr,"fishtix_vlengths.rds"))
})

# only Dungeness target
fishtix %<>% filter(TARGET_rev=="DCRB")
```

## Example Fishing Seasons for Different Ports

Test with an example from each state (arbitrarily chosen): 
Crescent City, 2008-2009 season
Coos Bay, OR, 2012-2013 season
Coastal Washington ports, 2011-12 season

```{r,fig.width=8,fig.height=6}
# general function to do a cumulative landings curve
plot_cumlandings <- function(state,season,savedir){
  stateabbr <- substr(state,1,1)
  subdat <- fishtix %>% 
    filter(removal_type_code %in% c("C","D")) %>%
    # add general indicator of crab season
    mutate(crab_season=case_when(
      month(date)>10 ~ paste0(year,"-",year+1),
      month(date)<=10 ~ paste0(year-1,"-",year))) %>% 
    filter(crab_season ==season,agency_code==stateabbr)%>%
    group_by(agency_code,port_group_code) %>% 
    select(agency_code,port_group_code,date,DCRB_lbs) %>% 
    arrange(date) %>% 
    mutate(cum_landings=cumsum(DCRB_lbs)) %>% 
    mutate(prop_tot=cum_landings/last(cum_landings)) %>% 
    ungroup() %>% 
    group_by(agency_code,port_group_code) %>% 
    mutate(startdate = first(date[prop_tot>=0.01]))
  
  # calculate date of 1% landings achievement
  # startdate <-  subdat %>% filter(prop_tot>=0.01) %>% slice(1) %>% pluck("date")
  
  out <-subdat %>% 
    ggplot()+
    geom_line(aes(date,prop_tot),col='gray50')+
    geom_point(aes(date,prop_tot),col='gray50')+
    geom_hline(yintercept=0.01,linetype=1,col='red')+
    geom_vline(aes(xintercept = startdate),linetype=1,col='red',size=1)+
    geom_text(aes(x=startdate,label=paste0("Start:",startdate)),y=0.8,hjust=-0.1,check_overlap=TRUE)+
    labs(title=paste0('Cumulative landings by Port Group\n',season," ",state),x="Date",y="Cumulative Proportion Total Landings")+
    facet_wrap(~port_group_code)
  fn <- paste0(here::here('fishing behavior','plots'),'/',state,str_replace_all(season,"-","_"),".png")
  ggsave(fn,out,h=6,w=8)
}
# plot and save
states <- c("California","Oregon","Washington")
seasons <- fishtix %>%
  mutate(crab_season=case_when(
      month(date)>10 ~ paste0(year,"-",year+1),
      month(date)<=10 ~ paste0(year-1,"-",year))) %>% 
  pluck('crab_season') %>% unique() %>% setdiff("2019-2020")

for(i in states) {
  for(k in seasons){
    plot_cumlandings(i,k)
  }
}

```

## Calculate One Percent Landings Achievement

These cumulative distribution suggest that a nice way might be to indicate the date of 1% landings achievement for each port group, since each season, no matter when it starts, begins with an intense derby that should be clear in the landings data. Using a 1 percent cutoff seems robust to the landings that 'dribble in' before the official season starts.

```{r}
one_percent_landings_achievement <- fishtix %>% 
  filter(removal_type_code %in% c("C","D")) %>%
  # add general indicator of crab season
  mutate(crab_season=case_when(
    month(date)>10 ~ paste0(year,"-",year+1),
    month(date)<=10 ~ paste0(year-1,"-",year))) %>% 
  group_by(agency_code,port_group_code,crab_season) %>% 
  arrange(date) %>% 
  mutate(cum_landings=cumsum(DCRB_lbs)) %>% 
  mutate(prop_tot=cum_landings/last(cum_landings)) %>% 
  filter(prop_tot>=0.01) %>% 
  summarise(start_of_season_oneperc=first(date,order_by = date)) %>% 
  ungroup()
```

## Apply to Fish Ticket Data

```{r}
fishtix_dos <- fishtix %>% 
  filter(removal_type_code %in% c("C","D")) %>%
  # add general indicator of crab season
  mutate(crab_season=case_when(
    month(date)>10 ~ paste0(year,"-",year+1),
    month(date)<=10 ~ paste0(year-1,"-",year))) %>% 
  
  # Join the one percent landings key
  left_join(one_percent_landings_achievement)

names(fishtix_dos)
```

Finally, add day of season to each ticket as a difference between the ticket's landing date and the start date of the season

```{r}
fishtix_dos %<>%
  ungroup() %>% 
  mutate(ticket_day_of_season=difftime(date,start_of_season_oneperc,units='days')) %>% 
  mutate(ticket_day_of_season=as.numeric(ticket_day_of_season))

glimpse(fishtix_dos %>% select(Rec_ID,port_group_code,date,crab_season,start_of_season_oneperc,ticket_day_of_season))
```

Overall distribution of tickets by day of season

```{r}
fishtix_dos %>% 
  ggplot(aes(ticket_day_of_season))+
  geom_density(fill='blue',alpha=0.5)
```

There's a little bit of weirdness at the end. Some seasons in some places seem like they may start before Nov 1...?

## Save 

Save a version of the fish tickets with day of season included.

```{r}
write_rds(fishtix_dos,here::here('data','processed','fish tickets','fish_tickets_w_dayofseason.rds'))
```

