---
title: "Matching VMS Data to Spatial Grid"
output: 
  html_document:
    toc: true
    toc_float: true
---

## Purpose

Matching the output VMS data to a multi-region grid to use for summarizing fishing effort

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r libs,message=FALSE,warning=FALSE}
library(tidyverse)
library(magrittr)
library(sf)
library(here)
library(knitr)
```


## Sample Year

Import files for an example year to match. We use the matched, filtered, but non-interpolated VMS data

```{r}
# clean, interpolated vms data
vms <- read_rds(here::here('data','processed','matched','interpolation','2012interpolated.rds'))

# 5x5 grid shapefile
grd <- read_sf(here::here('data','raw','grid','regions_master_final_lamb.shp'))
names(grd)

```

## Match Points

```{r}
# convert vms to spatial object (longitude/latitude)
vms_sf <- vms %>%
  st_as_sf(coords=c('LONGITUDE','LATITUDE'),crs=4326) %>% 
  # then, convert to planar projection to match the grid
  st_transform(st_crs(grd))

# do the join
pt <- proc.time()
vms_grd_match <- vms_sf %>%
  st_join(grd)
proc.time()-pt
```

## Check

Did the join work? How many unique 5km, 25km, and 50km cells are represented in the 2009 data?

```{r checks}
names(vms_grd_match)

length(unique(vms_grd_match$GRID5KM_ID))

length(unique(vms_grd_match$BAND_25KM))

length(unique(vms_grd_match$BAND_50KM))
```

## All Data

We now match all of the VMS points from all years. We process each year just as one above.

### Import

```{r import all}
# clear wrkspace
rm(list=setdiff(ls(),'grd'))

# load RData files
vms_all <- purrr::map_df(2009:2019, function(yr){
  read_rds(paste0(here::here('data','processed','matched','interpolation'),'/',yr,"interpolated.rds"))
})
```

### Match

```{r}
# convert VMS to spatial object
pt <- proc.time()
vms_all_sf <- vms_all %>%
  st_as_sf(coords=c('LONGITUDE','LATITUDE'),crs=4326) %>% 
  # then, convert to planar projection to match the grid
  st_transform(st_crs(grd))
x<-proc.time()-pt
```

Took `r round(x[3]/60,2)` minutes to do the conversion. Now for the join...

```{r join to grid}
# do the join
pt <- proc.time()
vms_all_grd_match <- vms_all_sf %>%
  st_join(grd)
x<-proc.time()-pt
```

The join took`r round(x[3]/60,2)` minutes.

### Add BIAs

We can add whale BIAs into the data as well. They have previously been matched to the 5km grid, courtesy of Sam Woodman.

```{r add BIAs}
BIA <- read_csv(here::here('data','raw','grid','Grid5km_BIA_overlap.csv'),col_types = cols())
BIA %<>% select(GRID5KM_ID,BIA_mn,BIA_bm)

vms_all_grd_match %<>% left_join(BIA,by='GRID5KM_ID')
```

### Save and Write

```{r}
# non-spatial version
glimpse(vms_all_grd_match)
vms_all_grd_match %>% 
  st_set_geometry(NULL) %>% 
  write_rds(here::here('data','processed','matched','interpolation','vms_all_interpolated_w_grd.rds'))
```






