---
title: "Check Port Buffer Zones, CA"
author: "M. Fisher"
date: "1/21/2023"
output: html_document
---


## Purpose

Verify that we aren't losing a ton of potential fishing data by using a 3km buffer zone around California ports of landing (instead of 1.5km).


```{r setup, include=FALSE}

library(tidyverse)
library(magrittr)
library(here)
library(sf)
library(geosphere)
library(lubridate)
library(rgdal)
library(rgeos)
library(rnaturalearth)

# ggplot theme
plot_theme <-   theme_minimal()+
  theme(text=element_text(family="sans",size=12,color="black"),
        legend.text = element_text(size=14),
        axis.title=element_text(family="sans",size=14,color="black"),
        axis.text=element_text(family="sans",size=12,color="black"),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank())

theme_set(plot_theme)

knitr::opts_chunk$set(echo = FALSE)
```


## Data
VMS data from 2010
```{r}
vms_wfilters <- read_rds(here::here('data','processed','matched','filtering','2010matched_unfiltered.rds'))
```

port coordinates
```{r}
portlist_coords <- read_csv(here::here('data','raw','port_coords_fromBlake_edited.csv'),col_types='cddd') %>% 
  select(port_code,Lon,Lat) %>% 
  set_names(c('port_code','portlon','portlat'))
```
<br>

Filter for California ports, with > 0 Dungeness crab landings for commercial use.
```{r}
vmsdat <- vms_wfilters %>% filter(agency_code=="C") %>% 
  filter(DCRB_lbs > 0) %>%
  filter(removal_type_code %in% c("C","D","U"))

vmsdat %<>%
  dplyr::select(-contains('pretrip')) %>%
  dplyr::select(-contains('posttrip'))
  
portdat <- portlist_coords %>% filter(port_code %in% vmsdat$pacfin_port_code)
```
<br>

Get rid of records on land / too fast / too far out to sea
```{r}
vmsdat_filtered1 <- vmsdat %>% 
  filter(keep_remove_bathy=="keep" & keep_remove_portdist == "keep" & keep_remove_speed=="keep") %>%
    dplyr::select(pacfin_port_code,Rec_ID, drvid,westcoastdate,
                  LATITUDE,LONGITUDE,AVG_SPEED,avg_speed_recalc,
                  in_port_3km, in_port_1p5km,keep_remove_not_intrip_3km, keep_remove_not_intrip_1p5km) %>%
  mutate(in_port_3km=ifelse(is.na(in_port_3km),0,1),
         in_port_1p5km=ifelse(is.na(in_port_1p5km),0,1))
```

## Proportion Removed

What proportion of VMS records per port are marked "remove" using the 3km v. 1.5km buffer zones?
```{r fig.width=12, fig.height=8}
n_by_port <- vmsdat_filtered1 %>%
  pivot_longer(cols=c(keep_remove_not_intrip_3km, keep_remove_not_intrip_1p5km), names_to="buffer_size", values_to="to_remove") %>%
  mutate(buffer_size=ifelse(buffer_size=="keep_remove_not_intrip_3km","3km","1.5km")) %>%
  group_by(pacfin_port_code,buffer_size) %>%
           summarise(totalVMS=n())
plotdat <- vmsdat_filtered1 %>%
  pivot_longer(cols=c(keep_remove_not_intrip_3km, keep_remove_not_intrip_1p5km), names_to="buffer_size", values_to="to_remove") %>%
  mutate(buffer_size=ifelse(buffer_size=="keep_remove_not_intrip_3km","3km","1.5km")) %>%
  group_by(pacfin_port_code,buffer_size,to_remove) %>%
    summarise(nVMS=n()) %>%
  left_join(n_by_port) %>%
  mutate(propVMS=nVMS/totalVMS)

ggplot(plotdat, aes(x=buffer_size,y=propVMS,fill=to_remove)) +
  geom_col() +
  facet_grid(cols=vars(pacfin_port_code)) + 
  scale_fill_manual(values=c("aquamarine3","coral3")) +
  theme(axis.text.x=element_text(angle=90,hjust=1,vjust=0.5))
```

The largest discrepancies seem to be: MNT, OSM, kind of BRG

## Map Removals

Map the points marked **keep** for 1.5km buffer, that were **removed** for the 3km buffer. The fill color of the points represents avg speed.
```{r}
# keep only records that aren't kept for both buffers
diff.removals <- vmsdat_filtered1 %>%
  filter((keep_remove_not_intrip_3km=="keep" & keep_remove_not_intrip_1p5km=="remove") | (keep_remove_not_intrip_3km=="remove" & keep_remove_not_intrip_1p5km=="keep"))

mnt.diff <- diff.removals %>% filter(pacfin_port_code=="MNT") %>%
  dplyr::select(Rec_ID, drvid, pacfin_port_code) %>% distinct() %>%
  left_join(vmsdat_filtered1)
```
<br>


```{r}
for(r in unique(mnt.diff$Rec_ID)){
plot3k <-   ggplot() +
    geom_point(data=filter(portlist_coords, port_code==unique(mnt.diff$pacfin_port_code)), aes(x=portlon, y=portlat), size=3, col="red") +
    geom_point(data=mnt.diff %>% filter(Rec_ID==r), aes(x=LONGITUDE, y=LATITUDE,col=keep_remove_not_intrip_3km), size = 2, alpha = 0.7) + 
  ggtitle("3km") +
    xlim(c(-123,-121))+ ylim(c(36.5,37)) + theme(legend.position="none")
plot1p5k <-   ggplot() +
    geom_point(data=filter(portlist_coords, port_code==unique(mnt.diff$pacfin_port_code)), aes(x=portlon, y=portlat), size=3, col="red") +
    geom_point(data=mnt.diff %>% filter(Rec_ID==r), aes(x=LONGITUDE, y=LATITUDE,col=keep_remove_not_intrip_1p5km), size = 2, alpha = 0.7) + 
  ggtitle("1.5km") +
    xlim(c(-123,-121))+ ylim(c(36.5,37)) + theme(legend.title=element_blank())
plot_grid(list(plot3k, plot1p5k), ncol=2, rel_width=c(0.8,1.0))
}


```



## Check Speeds

### in buffer zone

These are records within the buffer zones that have average speeds > 0. 
```{r}
for(p in portdat$port_code){
  plotdat <- vmsdat_filtered1 %>% filter(pacfin_port_code==p) %>% 
    filter(!is.na(avg_speed_recalc) & avg_speed_recalc > 0) %>%
    dplyr::select(-keep_remove_not_intrip_3km, -keep_remove_not_intrip_1p5km) %>%
    pivot_longer(cols=c(in_port_3km, in_port_1p5km), names_to="buffer_size",values_to="in_port_boolean") %>%
    mutate(buffer_size=ifelse(buffer_size=="in_port_3km","3km","1.5km")) %>%
    filter(in_port_boolean==1) 

  
  ggplot(plotdat,aes(x=avg_speed_recalc, fill=buffer_size))+
    geom_histogram() +
    facet_grid(rows=vars(buffer_size),scales="free_x") +
    labs(y="VMS records") + theme(legend.position="none")
}
```
<br>

### records marked remove

These are records that have been marked as "in port" *and* have a calculated speed of less than 2 knots (1 m/s)
```{r}
for(p in portdat$port_code){
  p="SF"
  plotdat <- vmsdat_filtered1 %>% filter(pacfin_port_code==p) %>% filter(!is.na(avg_speed_recalc)) %>%
    dplyr::select(-c(in_port_3km, in_port_1p5km)) %>%
    pivot_longer(cols=c(keep_remove_not_intrip_3km,keep_remove_not_intrip_1p5km), names_to="buffer_size", values_to="to_remove") %>%
    mutate(buffer_size=ifelse(buffer_size=="keep_remove_not_intrip_3km","3km","1.5km")) %>%
    filter(to_remove=="remove") %>%
    filter(avg_speed_recalc>0)

  
  ggplot(plotdat,aes(x=avg_speed_recalc, fill=buffer_size))+
    geom_histogram() +
    facet_grid(rows=vars(buffer_size),scales="free_x") +
    labs(y="VMS records")
}
```


## Cross-ref ports

Sometimes, the vessel is marked as "in port" for a different port of landing than where the fish ticket is recorded. how frequently does this happen for records marked "remove"?
