---
title: "QC - compare processed fish tix"
output: html_document
---
Date written: 1/21/2023

By: M. Fisher

## Purpose

Compare a prior version of the processed fish tickets (with / without vessel lengths) to a newer version. 

**2023-01-21:** compare process tickets from Owen (Feb 20, 2020) with newly processed tickets (file: 2006_2016_compFT.csv; most updated version of script 01)



## Set up

```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(magrittr)
library(knitr)
knitr::opts_chunk$set(echo = TRUE)
```

User inputs
```{r}
process_year <- seq(2009,2015)
oldtix_dir   <- 'data/processed/fish tickets compFT'
newtix_dir   <- 'data/processed/fish tickets'
```
<br>

## Data

All processed fish tickets should be (1) .rds format, and (2) named by default by scripts 01 & 02.

```{r read_new}
for(y in process_year){
  tmptix   <- read_rds(here(newtix_dir, paste0(y,"fishtix.rds")))
  tmptix.v <- read_rds(here(newtix_dir, paste0(y, "fishtix_vlengths.rds")))
  if(y==process_year[1]){
    newtix <- tmptix
    newtix.v <- tmptix.v
  } else{
    newtix %<>% bind_rows(tmptix)
    newtix.v %<>% bind_rows(tmptix.v)
  }
}
```

```{r read_old}
for(y in process_year){
  tmptix   <- read_rds(here(oldtix_dir, paste0(y,"fishtix.rds")))
  tmptix.v <- read_rds(here(oldtix_dir, paste0(y, "fishtix_vlengths.rds")))
  if(y==process_year[1]){
    oldtix <- tmptix
    oldtix.v <- tmptix.v
  } else{
    oldtix %<>% bind_rows(tmptix)
    oldtix.v %<>% bind_rows(tmptix.v)
  }
}
```
<br>

## Compare tickets

What columns in the older version are missing in the newer version?
```{r}
old_missing <- colnames(oldtix)[which(!(colnames(oldtix) %in% colnames(newtix)))]; old_missing
oldv_missing <- colnames(oldtix.v)[which(!(colnames(oldtix.v) %in% colnames(newtix.v)))]; oldv_missing
```

What columns in the newer version are missing in the older version?
```{r}
new_missing <- colnames(newtix)[which(!(colnames(newtix) %in% colnames(oldtix)))]; new_missing
newv_missing <- colnames(newtix.v)[which(!(colnames(newtix.v) %in% colnames(oldtix.v)))]; newv_missing
```


Which records are different in the older v. newer versions?
```{r}
diff.records <- anti_join(oldtix,newtix, by=all_of(colnames(newtix)))
diff.records.v <- anti_join(oldtix.v,newtix.v, by=all_of(colnames(newtix.v)))
```

```{r}
dim(diff.records)
dim(diff.records.v)
```




